### 2019-02-25 docs/manuscript/main.tex
[genesis]

%*******************************************************************************
\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\graphicspath{{../}} %%%REPOSITORY PATH

\title{Strengths and Weaknesses of Recurrent Quantification Analysis 
in the context of Human-Humanoid Interaction}
\author[1,*]{Miguel Xochicale}
\author[1]{Chirs Baber}
\affil[1]{University of Birmingham, 
	School of Engineering, 
	Birmingham, 
	B15 2TT, 
	UK}
\affil[*]{map479@bham.ac.uk}

%\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Movement variability is defined as the variations that occur in motor
performance across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly when traditional
methods in time domain and frequency domain fail to detect tiny modulations in
frequency or phase for time series. We therefore consider methodologies from
nonlinear dynamics such as reconstructed state space (RSS), uniform time-delay
embedding (UTDE), recurrence plots (RPs) and recurrence quantification analysis
(RQA) metrics. These algorithms require time series measured with sensors that
provide well sampled data with little noise. However, for this study, we are
interested in the analysis of data collected through wearable inertial sensors
(IMUs) and its post-processing effects in each of the nonlinear dynamics
methodologies. Twenty right-handed healthy participants (mean and standard
deviation (SD) age of mean=19.8 (SD=1.39)) performed an experiment in
human-humanoid imitation activities (H-H-I). Then IMUs were attached to the
participants and the humanoid robot performed simple vertical and horizontal
arm movements in normal and faster speed. With four window lengths and three
levels of smoothed time series, we found visual differences in the patterns of
RSSs and RPs. Then using metrics of RQA, we find out that the type of movements
and the level of smoothness affects those metrics. In particular, we found that
entropy values from RQA were well distributed and presenting variation in all
the conditions for time series. Hence, we demonstrated the potential of
nonlinear techniques to quantify human movement variability in the context of
H-H-I can enhance the development of better diagnostic tools for various
applications rehabilitation, sport science or for new forms of human-robot
interaction.
%280
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}
Human movement variability is a complex system where not only multiple
joints and limbs are involved for a specific task in a determined environment
but also how we process the external information with all of our available senses 
and use our experiences \cite{davids2003}. 
Recent studies in human motion recognition have revealed the possibility to 
estimate features from lower dimensions signals to distinguish differences between 
styles of pedalling motion \cite{Quintana-Duque2012, Quintana-Duque2016}, 
gait identification \cite{sama2013, frank2010} 
or pattern recognition from biological signals \cite{gomezgarcia2014}.
The lower dimension signals from biological signals are time series 
of $1-$dimension in $\mathbb{R}$ which commonly are 
noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
Hence, we are hypothesising that nonlinear dynamics can be used to 
objectively quantify such human movement variability of lower
dimension signals \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
For instance, Bradley et al. 2015 \cite{bradley2015} reviewed methods for
nonlinear time series analysis based on the appropriate estimation of the 
embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
reconstruct the state space,
where an $n$-dimensional reconstructed state space using $1-$dimensional 
time series,
can preserve the topological properties of an unknown $M$-dimensional 
state space \cite{takens1981}.
Also, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
which show black and white dots as recurrences in a given $n$-dimensional system, 
and Recurrence Quantification Analysis (RQA) metrics can pick out important 
directions and statistics in RPs.
In general, RPs and RQA provide a more intuitive meaning of the time series,
for instance, RQA is quantitatively and qualitatively independent of 
embedding dimension (also verified experimentally \cite{iwanski1998}).
However, the estimation of embedding parameters and the selection of the 
right parameters to perform RQA is still an open problem.
Bradley et al. 2015 \cite{bradley2015} pointed out that
there is no general technique that can be used to compute the embedding
parameters since the time series are system-dependent which means that
once computing the values for embedding parameters, these may only work for
one purpose (e.g., prediction) and may not work well for another purpose
(e.g., computing dynamical invariants).
Additionally, the methodologies of nonlinear dynamics for 
computing the embedding parameters e.g., autocorrelation, mutual information, 
and nearest neighbour require data which is well 
sampled and with little noise \cite{garland2016}
or require purely deterministic signals \cite{kantz2003}.
Similarly, these nonlinear analysis methods for computing the embedding 
parameters can break down
with real-world datasets which have generally different length, 
different values of accuracy and precision (rounding errors due to finite 
precision of the measurement apparatus which include frequency 
acquisition \cite{frank2010}),
and data may be contaminated with different sources of noise
\cite{garland2016}.
It is surprising that even with the previous constraints with regard to 
the quality of data, and the problem with the estimation of embedding parameters,
the use of nonlinear dynamics have proven to be helpful to understand and 
characterise dynamics of time series 
\cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
gomezgarcia2014, gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.

%Another point to consider with time series analysis using nonlinear dynamics
%is the appropriate use of different post-processing techniques such as 
%interpolation, filtering or normalisation.
%With that in mind, it can be said that there is little research and 
%understanding on the effects for post-processing techniques
%in the interpretation of reconstructed state spaces, RPs and RQA.

For this work, we are interested in exploring and investigating  
the effects of different features of time series 
(e.g. levels of smoothness, window data length,
structures of time series based on types of movements, 
types of sensors, participants and velocities) to estimate 
embedded parameters for uniform time-delay embedding, 
recurrence plots and recurrence quantification analysis.
Hence, we conducted an experiment with twenty right-handed 
healthy participants in the context of human-humanoid imitation 
activities where participants were asked to 
imitate simple arm movements performed by a humanoid robot.
The primary aim of this work is to explore the following questions:
\begin{itemize}
\item What are the effects on three nonlinear methods 
	(reconstructed state space with uniform time-delay embedding, 
	Recurrence Plots, and Recurrence Quantification Analysis), 
	for different embedding parameters, different recurrence thresholds 
	and different characteristics of time series 
	(structure of the signal, window length and smoothness of the signal)?, and 
\item What are the strengths and weaknesses of Recurrence
	Quantification Analysis for the previous conditions
	of nonlinear analysis methods?
\end{itemize}

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Results}
To give insight to the research questions, 
we considered various conditions of time series collected for this work 
(see Experiment section for more details) which are described as follows
\begin{itemize}
\item Three levels of smoothness for the normalised data 
	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
	lengths (29 and 159) with the same polynomial degree 
	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
\item four velocities arm movement activities: 
	horizontal normal (HN), horizontal faster (HF), 
	vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
	10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}

%\subsection*{Time series}
After the data collection, raw time series were windowed, normalised and 
smoothed. Then, due to space limitations and to have simple visualisation, 
we only present 10-sec (500 samples) window length time series for 
three participants (p01, p01 and p03) performing horizontal 
arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
    	\caption{ 
	{\bf Time series for horizontal and vertical arm movements.}
		(A/D) raw-normalised (sg0zmuv), 
		(B/E) normalised-smoothed 1 (sg1zmuv) and
		(C/F) normalised-smoothed 2 (sg2zmuv).
		Time series are only for three participants 
		($p01$, $p02$, and $p03$) 
		for horizontal and vertical arm movements in normal 
		and faster velocity (HN, HF, VN, VF) 
		with the normalised GyroZ or GyroY axis 
		and with one sensor attached to the participant (HS01) 
		and other sensor attached to the robot (RS01).	
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:ts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

%\subsection*{Reconstructed State Spaces using Uniform Time-Delay Embedding}
\subsection*{Reconstructed State Spaces}
One of the main challenges of the implementation of 
Uniform Time-Delay Embedding to Reconstructed State Spaces is the 
selection of embedding parameters because each time series is unique in 
terms of its structure (modulation of amplitude, frequency and phase) 
\cite{ frank2010, sama2013, bradley2015}.
With that in mind, the problem for this work is not to compute individual 
embedding parameters for each of the time series but to deal with a 
selecting of two parameters that can represent all the time series. 
Our solution for that problem was to compute a sample mean over all 
values in each of the conditions of the time series 
%of Figs~\ref{fig:ts}
for minimum dimension values and 
%Figs~\ref{fig:caoami} 
for minimum delay values as follows.
%, resulting in an average minimum embedding parameters 
%of ($\overline{m_0}=6$, $\overline{\tau_0}=8$).

\subsubsection*{Minimum embedding parameters}
Minimum embedding parameters were firstly computed 
with False Nearest Neighbour and Average Mutual Information.
%for Uniform Time-Delay Embedding (UTDE),
%Reconstructed State Space (RSS), 
%Recurrence Plots (RPs) and 
%Recurrence Quantification Analysis (RQAs).
Figures \ref{fig:caoami}(A) show that minimum embedding values 
for sensor HS01 appear to show more variations as their 
interquartile range of the box plots are near to 1 
with three exceptions (HF sg0, VN sg0, and VF sg1).
Minimum embedding values appear to be constant for sensor RS01 
as their interquartile range 
in the box plots are near to 0.1 with the exception of 
two axis (HF sg0 and VF sg1).
Additionally, it can be seen in Figs \ref{fig:caoami}(A) a 
decrease of mean values (rhombus) in the box plots
as smoothness of time series increase.
Similarly, the first minimum values of AMI values are 
shown in the box plots of Figs \ref{fig:caoami}(B).
It can be seen that values for HS01 tend to be more spread as the smoothness 
of the time series is increasing 
(see the increase of both mean (rhombus) and interquartile range).
However, AMI values for RS01 do not show such increase in relation with
the increase of smoothness excepting for HF and VF
(Figs \ref{fig:caoami}(B)).

We also computed an overall minimum embedding parameters that represent 
participants, activities, sensors and levels of smoothness, using a sample 
mean of all values in Figs~\ref{fig:caoami}.
The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:caoami}(A) 
is $\overline{m}_0=6$ 
and the sample mean for minimum values of AMIs from Figs~\ref{fig:caoami}(B)
is $\overline{\tau}_0=8$.
%Hence, we use the average minimum embedding parameters 
%($\overline{m_0}=6$, $\overline{\tau_0}=8$) to compute 
%Reconstructed State Spaces (RSSs), Recurrence Plots (RPs) and
%Recurrence Quantification Analysis (RQA) metrics for human-humanoid activities.
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/caoami/pdf/fig2}
	\caption{
	{\bf Box plots of minimum embedding parameters for horizontal and vertical arm movements.} 
	Box plots of (A) minimum embedding dimensions 
	and (B) first minimum AMI values for 
	Horizontal Normal (HN), Horizontal Faster (HF),
	Vertical Normal (VN) and Vertical Faster (VF)
	with sensors attached to participants (HS01) and
	sensor attached to robot (RS01).
	Minimum embedding dimensions are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:caoami}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsubsection*{Uniform Time-Delay Embedding}
Uniform time-delay embedding were computed with
embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
the first three axis of the rotated data of the PCA are shown 
for the reconstructed state spaces in 
Figs~\ref{fig:rss_aHw10} for horizontal arm movements and 
Figs~\ref{fig:rss_aVw10} for vertical arm movements.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig3}
\caption{
	{\bf RSSs for horizontal arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters $m=6$, $\tau=8$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rss_aHw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig4}
    \caption{
	{\bf RSSs for vertical arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters $m=6$, $\tau=8$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rss_aVw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

Evidently, it is easy to observe by eye the differences in each of the
trajectories in the reconstructed state spaces 
(Figs~\ref{fig:rss_aHw10}, \ref{fig:rss_aVw10}), 
however one might be not objective when quantifying those differences 
since those observation might vary from person to person.
With that in mind, we tried to objectively quantify those differences 
using euclidean distances between the origin to each of the points in the 
trajectories, however these results created a suspicious metric, specially 
for trajectories which looked very messy.
With that in mind, we computed Recurrence Plots and 
Recurrence Quantification Analysis to objectively quantify 
the differences in each of the cases of the time series.

\subsection*{Recurrences Plots}
Considering the time series of Fig~\ref{fig:ts}, 
Recurrence Plots were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and
for vertical arm movements (Fig~\ref{fig:rp_aV}) 
using the average embedding parameters ($m=6$, $\tau=8$) 
and an recurrence threshold of $\epsilon=1$.
With regard to the selection of recurrence threshold,
Marwan et al. \cite{marwan2011} pointed out that choosing an appropriate 
recurrence threshold is crucial to get meaningful representations in RPs, 
however, for this work where quantifying movement variability is our aim,
we do not compute recurrence thresholds as as long as it is able to 
represent the dynamical transitions in each of the time series.

As similar as with the Reconstructed State Spaces, 
the differences in the RPs can be easily noticed by eye 
for different time series (Figs~\ref{fig:rp_aV} and ~\ref{fig:rp_aH}), 
which lead us to apply Recurrence Quantification Analysis 
to have an objective quantification of each of the time series.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aH}
\caption{
	{\bf RPs for horizontal arm movements.}	
	Recurrence plots were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rp_aH}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aV}
\caption{
	{\bf RPs for vertical arm movements.}	
	Recurrence plots were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:rp_aV}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
Considering the RPs for 20 participants performing four activities 
(HN, HF, VN and VF) with sensors attached to the human (HS01) and to the 
humanoid robot (RS01) and with the increase of smoothness of time series 
(sg0zmuvGyroZ, sg1zmuvGyroZ and sg2zmuvGyroZ), 
we hence compute four metrics of RQA metrics (REC, DET, RATIO and ENTR) with 
embedding parameters $m=6$, $\tau=8$ and recurrence threshold $\epsilon=1$.

%\newpage
\subsubsection*{REC values}
It can be seen in the box plots of Figs~\ref{fig:RQABP}(A) that REC values, 
representing the \% of black dots in the RPs, 
are more spread for HN and VN movements (higher interquartile range) 
than HF and VF movements (lower interquartile range) for HS01 sensor. 
In contrast, REC values for RS01 sensor present little variation 
(interquartile range of 0.01).
With regard to the increase of smoothness of time series 
(sg0, sg1 and sg2), REC values present little 
variation as the smoothness is increasing for time series from HS01 
(changes of mean values (rhombus)) while REC values are more affected with 
the smoothness for data from RS01 
(see the incremental changes of mean values (rhombus)).
%See Figs~\ref{fig:rec_aH} and \ref{fig:rec_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual REC values for each participant.

\subsubsection*{DET values}
Figs \ref{fig:RQABP}(B) illustrate DET values, 
representing predictability and organisation of the RPs, 
which change very little (interquartile range is around 0.1) 
for type of movement, level of smoothness or type of sensor.
%See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual DET values for each participant.

\subsubsection*{RATIO values}
Figs \ref{fig:RQABP}(C) present RATIO values, representing dynamic transitions, 
for horizontal and vertical movements.
It can be seen that RATIO values for HS01 sensor vary less 
for HN movements (interquartile range around 2)
than HF movements (interquartile range around 5).
%which is a similar behaviour of RATIO values for oRS01 sensor.
It can also be noticed a decrease of variation in RATIO values as the 
smoothness of the time series is increasing (grey rhombus).
%See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
%\ref{appendix:e:ep} 
%for more details about individual RATIO values for each participant.

\subsubsection*{ENTR values}
Figs \ref{fig:RQABP}(D) show ENTR values, representing the complexity of 
the structure the time series, for both horizontal and vertical movements.
ENTR values for HS01 sensor show more variation 
(interquartile range around 0.5)
than ENTR values for RS01 sensor which appear 
to be more constant (interquartile range 0.1).
It can also be said that the smoothness of time series affects
each of the axis by an increase of mean values (see gray rhombos).
%See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
%\ref{appendix:e:ep}
%for more details about individual ENTR values for each participant.

%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa-bp}
    \caption
	[Box plots for RQA values]{
	{\bf Box plots for RQA values.}
	Box plots of (A) REC, (B) DET, (C) RATIO, and (D) ENTR values 
	for 20 participants performing HN, HF, VN and VF movements
	with sensors HS01, RS01 and three smoothed-normalised  
	time series (sg0, sg1 and sg2).
	RQA values were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available from \cite{srep2019}.
        }
    \label{fig:RQABP}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

%
%
%\subsection*{RQA metrics with different embedding parameters, 
%	recurrence thresholds, window lengths, levels of smoothness, 
%	and time series structures.}
%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}). We also computed 3D surfaces of RQA metrics 
%for different sensors and different activities 
%(Fig.~\ref{fig:topo_sensoractivities}). RQA metrics are also affected by 
%the window length where for example four window lengths of 100, 250, 500 
%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}). Three level of 
%smoothness were computed for RQA metrics showing smoothed 3D surfaces and 
%the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
%Similarly, 3D surfaces of RQA metrics were also computed for three 
%participants (Fig.~\ref{fig:topo_participants}).
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/rqas}
%%\includegraphics[width=1.0\textwidth]{rqas}
%    \caption{
%	{\bf 3D surfaces for RQA metrics.}
%	3D surfaces for REC, DET, RATIO and ENTR values with increasing 
%	pair embedding parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) 
%	and recurrence thresholds (  $ 0.2 \ge \epsilon \le 3 $).
%	RQA metrics values for time series of participant p01 using HS01 sensor, 
%	HN activity and sg0zmuvGyroZ axis and 500 samples window length.
%        R code to reproduce the figure is available from \cite{hwum2018}.
%	}
%\label{fig:topo_rqas}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/sa}
%%\includegraphics[width=1.0\textwidth]{sa}
%    \caption{
%	{\bf 3D surfaces of RQA metrics for sensors and activities.}
%	3D surfaces with increasing embedding parameters and recurrence 
%	thresholds are for HS01 and RS01 sensors of HN, HF, VN and VF 
%	activities. RQA metrics values are for time series of 
%	participant p01 for sensors (HS01 and RS01), 
%	activities (HN, HF, VN and VF) and for 	sg0zmuvGyroZ axis with 
%	500 samples window length. 
%	R code to reproduce the figure is available from \cite{hwum2018}.
%       }
%\label{fig:topo_sensoractivities}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/w}
%%\includegraphics[width=1.0\textwidth]{w}
%    \caption{
%	{\bf 3D surfaces for RQAs metrics with four window lengths.}
%	3D surfaces of RQAs metric values with increasing embedding 
%	parameters and recurrence thresholds are for four window lengths 
%	(w100, w250, w500 and  w750).
%	RQA metrics values are for time series of participant p01 
%	using HS01 sensor, HN activity and sg0zmuvGyroZ axis.
%	R code to reproduce the figure is available from \cite{hwum2018}.
%        }
%\label{fig:topo_sensoractivities}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/s}
%%\includegraphics[width=1.0\textwidth]{s}
%    \caption{
%	{\bf 3D surfaces for RQA metrics with three levels of smoothness.}
%       	3D surfaces of RQAs metric values with increasing embedding 
%	parameters and recurrence thresholds are for three levels of 
%	smoothness (sg0zmuvGyroZ, sg1zmuvGyroZ and sg1zmuvGyroZ).
%	RQA metrics values are for time series of participant p01 using 
%	HS01 sensor, HN activity and 500 samples window length.
%	R code to reproduce the figure is available from \cite{hwum2018}.
% }
%\label{fig:topo_smoothness}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/p}
%%\includegraphics[width=1.0\textwidth]{p}
%    \caption{
%	{\bf 3D surfaces for RQA metrics with three participants.}
%       	3D surfaces of RQAs metric values for participants p01, p02 and p03 
%	with increasing embedding parameters and recurrence thresholds.
%	RQA metrics values are for time series of HS01 sensor, 
%	HN activity and 500 samples window length.
%	R code to reproduce the figure is available from \cite{hwum2018}.
% }
%\label{fig:topo_participants}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%
%
%
%
%
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\section*{Discussion}
%%The Discussion should be succinct and must not contain subheadings.
%
%It is evidently that time series from different sources (participants, 
%movements, axis type, window length or levels of smoothness) 
%presents visual differences for embedding parameters and therefore for RRSs. 
%For which, the selection of embedding parameters was our first challenge 
%where we computed embedding parameters for each time series and then computed 
%a sample mean over all time series in order to get two embedding parameters 
%to compute all RSSs with its corresponded type of movement. Then we found 
%that the quantification of variability with regard to the shape of the 
%trajectories in RSSs requires more investigation since our original 
%proposed method base on euclidean metric failed to quantify those 
%trajectories. Specially, for trajectories which were not well unfolded. 
%With that in mind, we proceed to take advantage of four RQA metrics 
%(REC, DET, RATIO and ENTR) in order to avoid any subjective interpretations 
%or personal bias with regard to the evolution of the trajectories in RSSs.
%
%
%\subsection*{RQA metrics with fixed parameters}
%Considering that RQA metrics were computed with fixed embedding parameters 
%($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
%the following. REC values, which represents the \% of black points in the RPs, 
%were more affected with and increase in normal speed movements (HN and VN) 
%than faster movements (HF and VF) for the sensor attached to the participants 
%(HS01). Such decrease of REC values from normal speed to faster speed 
%movements is also presented in data from sensor attached to the robot (RS01), 
%and little can be said with regard to the dynamics of the time series coming 
%from RS01. Similarly, DET values, representing predictability and 
%organisation in the RPs, present little variation in the any of the time 
%series where little can be said. In contrast, RATIO values, which represent 
%dynamic transitions, were more variable for faster movements (HF and VF) 
%than normal speed movements (HN and VN) with sensors attached to the 
%participants (HS01). For data coming from sensors attached to the robot 
%(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
%more than values coming from vertical movmentes (VN, VF).
%With that, it can be said that RATIO values can represent a bit better
%than REC or DET metrics for the variability of imitation activities in 
%each of the conditions for time series.
%In the same way, ENTR values for HN were higher than values for HF
%and ENTR values varied more for sensor attached to participants 
%than ENTR values for sensors of the robot. It is evidently that 
%the higher the entropy the more complex the dynamics are, 
%however, ENTR values for HN appear a bit higher than HF values, 
%for which we believe this happens because of the structure the time series
%which appear more complex for HN than  HF movements which presented a 
%more consistence repetition.
%
%We observed that some RQA metrics are affected by the smoothness of data.
%For which, we also explored the effect of smoothness of raw-normalised data 
%where, for example, REC and DET values were not affected by the smoothness 
%of data since these seemed to be constants. However, for RATIO values, 
%the effect of smoothness can be noticed with a slight decrease of amplitude 
%in any of the time series conditions which is also presented with ENTR values.
%
%
%\subsection*{RQA metrics with different parameters}
%Patterns in RPs and metrics for RQA are independent of embedding dimension 
%parameters \cite{iwanski1998}, however, that is not the case when using 
%different recurrence thresholds. Such changes of recurrence threshold values 
%can modify the patterns in RPs and therefore the values of RQA metrics.
%We therefore computed 3D surfaces to explore the sensibility and robustness of 
%embedding parameters and recurrence threshold in RQA  metrics. Following the 
%same methodology of computing 3D surfaces, we also considered variation of 
%window length size to present RQA metrics dependencies with embedding 
%parameters, recurrence thresholds and window length size.
%
%
%
%\section*{Conclusions}
%We conclude that using a different level of smoothness for time series help us 
%to visualise and to quantify the variation of movements between participants 
%using RSSs, RPs and RQA. Also, it is important to mention that some RQA's 
%metrics (e.g. DET and  ENTR) are more robust to the effect of smoothness 
%of time series.
%
%Althought our our experiment is limited to twenty healthy right-handed 
%participants of a range age of mean 19.8 SD=1.39, RQA metrics can quantify 
%human movement variability. With that in mind, we conclude that 
%quantification of human-humanoid imitation activities is possible for 
%participants of different ages, state of health and anthropomorphic features.
%
%
%%
%%Such understanding and measurement of movement variability using
%%cheap wearable inertial sensors lead us to have a more intuitive selection of parameters
%%to reconstruct the state spaces and to create meaningful interpretations
%%of the recurrence plots and the results of the metrics with recurrence quantification 
%%analysis. 
%%
%
%In general, activity type, window length and structure of the time series 
%affects the values of the metrics of RQA for which certain RQA metrics 
%are better to describe determined type of movement.
%Using determined RQA metrics depends on what one want to quantify, for instance, 
%one can find predicability, organisation of the RPs, dynamics transitions, 
%or complexity and determinism.
%
%
%%
%%However, we believe that further investigation is required to find the 
%%right balance between the level of smoothness of the signal
%%and defining what is the aim of RQA where 
%%its representations using RSS, RP and RQA.
%%Specially, where the level of smoothness does not affect 
%%the variation of each of the movements quantification. 
%%%We can also conclude that finding the right balance between 
%%%smoothness and the raw data to capture movement variability is 
%%%a still a problem that has many avenues for exploration.
%%
%
%Similarly, such differences in time series created differences in each of the 
%RQA metrics, for instance, RATIO and ENTR are helpful to distinguish 
%differences in any of the categories of the time series (sensor, activity, 
%level of smoothness and number of participant), however for certain time series 
%(data from the sensor attached to the robot) seemed to have little variations 
%between each of the participants. The latter phenomena was in a way evidently
%as robot degrees of freedom did not allow it to move with a wide range of 
%variability. 
%
%
%
%\section*{Future Work}
%
%\subsection*{Inertial Sensors}
%To have more fundamental understating of nature of signals collected through 
%inertial sensors in the context of human-robot interaction, we are considering 
%to apply derivates to the acceleration data. We can then explore the jerkiness 
%of movements and therefore the nature of arm movements which typically have 
%minimum jerk \cite{flash1985}, its relationship with different body 
%parts \cite{devries1982, mori2012} or the application of higher derivatives 
%of displacement with respect time such as snap, crackle and 
%pop \cite{eager2016}.
%
%\subsection*{RQA}
%Having presented our results with RQA metrics, we believe that further 
%investigation is required to have more robust metrics. For example, 
%Marwan et al. \cite{marwan2007, marwan2015} reviewed different aspects to 
%compute RPs using different criteria for neighbours, different norms 
%($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or different methods 
%to select the recurrence threshold $\epsilon$, such as using certain 
%percentage of the signal \cite{letellier2006}, the amount of noise or 
%using a factor based on the standard deviation of the observational noise 
%among many others \cite{marwan2007}.
%
%
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\section*{Materials and methods}
%\subsection*{State Space Reconstruction}
%The method of state space reconstruction was originally proposed by 
%\cite{packard1980} and formalised by \cite{takens1981}. Since then, different 
%investigations and disciplines have benefited from the use of the method of 
%state space reconstruction \cite{aguirre2009, stergiou2011, frank2010, 
%sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
%based on uniform time-delay embedding methodology which is a simple 
%matrix implementation that can reconstruct an unknown $d-$dimensional 
%manifold $M$ from a scalar time series (e.g. one-dimensional 
%time series in $\mathbb{R}$). A manifold, in this context, is a 
%multidimensional curved surface within a space (e.g. a saddle) 
%\cite{guastello-gregson2011}.
%
%The use of a scalar time series is the main advantage of the method of state 
%space reconstruction which in essence preserve dynamic invariants such as 
%correlation dimension, fractal dimension, Lyaponov exponents, Kolmogorov-Sinai 
%entropy and detrended fluctuation analysis \cite{bradley2015, 
%Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
%However, selecting appropriate embedding parameters which are sued to apply the 
%state space reconstruction is still an open challenge for which we present 
%introductions for the methodologies to compute such embedding parameters 
%With that in mind, in the following subsections, we describe in more detail 
%the state space reconstruction theorem (RSSs), uniform time-delay embedding 
%theorem (UTDE), false nearest neighbours (FNN), average mutual 
%information (AMI) and other methodologies for state space reconstruction.
%
%\subsection*{State Space Reconstruction Theorem}
%Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
%uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:ssr}
%  s(t)=f^t [s(0)],
%\end{equation}
%%%********************************[EQUATION]************************************
%where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
%$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
%unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
%function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
%iteration of $f$ that corresponds to an initial position 
%$s(0) \in M $ \cite{takens1981}.
%%$f$ is a smooth dynamical system, where smooth means a 
%%continuously differentiable system
%%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
%Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
%can be obtained with
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:measurement}
%  x(t)=h[s(t)],
%\end{equation}
%%%********************************[EQUATION]************************************
%where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
%Reconstructed state space can then be described as an $n-$dimensional state 
%space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
%$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
%is the uniform time-delay embedding with a dimension embedding $m$
%and delay embedding $\tau$ and
%$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
%of dimensionality (e.g. Principal Component Analysis, 
%Singular Value Decomposition, etc) being $n \leq m$.
%With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
%defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
%$\boldsymbol{X}(t) = \Phi(s(t))$,
%where $\Phi$ is a diffeomorphic map \cite{takens1981}
%whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
%dimension of $M$ \cite{garland2016}.
%% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
%% is called diffeomorphic if it is one-to-one correspondence and its inverse
%% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
%Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
%space then a composition of functions represented with $F^t$ is induced on the
%reconstructed state space determined:
%%by true smooth dynamical system $f^t$ and $\Phi$:
%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:st}
%  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
%\end{equation}
%%%********************************[EQUATION]************************************
%With this in mind, an embedding is defined as "a smooth one-to-one 
%coordinate transformation with a smooth inverse" and the total reconstruction 
%map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
%Fig~\ref{fig:ssr} illustrates the state space reconstruction.
%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/rss/casdagli1991/figure1/rss_v04}
%%\includegraphics[width=1.0\textwidth]{rss_v04}
%    \caption{
%	{\bf State space reconstruction methodology.}
%	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
%	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
%    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
%    	The time-delay embedding represented as the $\Phi$, maps the original
%    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
%	time-delay embedding matrix $\boldsymbol{X}(t)$.
%    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
%	a new state $y(t)$ of dimensions $n < m$.
%    	(A) $M-$dimensional manifold representing the state space 
%	(e.g. Lorenz system);
%    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
%   	(C) $m-$dimensional reconstructed state space with 
%    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
%    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
%	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
%	where $\Phi$ is the delay reconstruction map and 
%	$\Psi$ is the coordinate transformation map.
%	This figure is adapted from 
%   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}
%	and R code to reproduce it is available from \cite{hwum2018}.
%    }
%    \label{fig:ssr}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%
%
%\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
%Frank et al. and Sama et al. refer to the state space reconstruction
%as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
%However, we consider the term "uniform time-delay embedding"
%as more descriptive and appropriate terminology for our work.
%
%The uniform time-delay embedding is represented as a matrix of uniform delayed 
%copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
%the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
%samples of $\{ \boldsymbol{x}_n \}$.
%$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
%The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
%and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
%where $i$ goes from $0,1, \dots, (m-1)$.
%Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
%information of unobserved state variables and encapsulates the information of
%the delayed copies of the available time series in the uniform time-delay 
%embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
%$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:tde}
%\boldsymbol{X}^{m}_{\tau}  =
%\begin{pmatrix}
%\boldsymbol{ \tilde{x} }_n \\
%\boldsymbol{ \tilde{x} }_{n-\tau} \\
%\boldsymbol{ \tilde{x} }_{n-2\tau} \\
%\vdots \\
%\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
%\end{pmatrix}^\intercal, 
%\end{equation}
%%%********************************[EQUATION]************************************
%where $m$ is the embedding dimension, $\tau$ is the embedding delay and
%$ ^\intercal$ denotes the transpose.
%$m$ and $\tau$ are known as embedding parameters.
%The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
%$N-(m-1)\tau$ rows and $m$ columns and 
%$N-(m-1)\tau$ defines the length of each delayed copy 
%of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
%%For further details and explicit examples of uniform time-delay 
%%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
%
%
%
%
%\subsection*{Estimation of Embedding Parameters}
%The estimation of the embedding parameters ($m$ and $\tau$) 
%is a fundamental step for the state space reconstruction with the use
%of uniform time-delay embedding method.
%With this in mind, we review two of the most common algorithms,
%which will be used in our work, to compute the embedding
%parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%*******************************************************************************
%\subsubsection*{False Nearest Neighbours}
%To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
%used the method of false neighbours which can be understood as follows:
%on the one hand, when the embedding dimension is too small to unfold the 
%attractor "not all points that lie close each other will be neighbours and 
%some points appear as neighbours because of the attractor has been projected 
%down into an smaller space", on the other hand, when increasing the embedding 
%dimension "points that are near to each other in the sufficient
%embedding dimension should remain close as the dimension increase from $m$ 
%to $m+1$ \cite{krakovska2015}".
%From a mathematical point of view, the state space reconstruction theorem is 
%done when the attractor is unfolded with either the minimum embedding 
%dimension, $m_0$, or any other embedding dimension value where 
%$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
%leads to excessive computations \cite{bradley2015}. With this in mind, 
%Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
%where only the time-series and one delay embedding value are necessary 
%to select the minimum embedding dimension. Cao's algorithm is based 
%on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:e}
%  \begin{aligned}
%	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
%	 &=
%       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
%       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
%            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
%  \end{aligned}
%\end{equation}
%%%********************************[EQUATION]************************************
%where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
%embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
%From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
%and $\tau$ for which $E_1(m)$ is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:e1}
%E_1(m) = \frac{ E(m+1) } { E(m)}.
%\end{equation}
%%%********************************[EQUATION]************************************
%$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
%in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
%As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
%than some $m_0$, if the time series comes from a multidimensional state space
%then $m_0 + 1$ is the minimum dimension".
%Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
%stochastic signals. $E_2(m)$ is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:e2}
%E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
%\end{equation}
%%%********************************[EQUATION]************************************
%where
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:ee}
%E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
%|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
%\end{equation}
%%%********************************[EQUATION]************************************
%For instance, when the signal comes from random noise (values that are 
%independent from each other), all $E_2(m)$ values are approximately equal 
%to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
%not constant for all $m$ (e.g. $E_2(m) \neq 1$).
%
%As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
%series: the solution for the $x$ variable of the Lorenz system 
%(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
%and a variance of one (Fig~\ref{fig:e1e2}F).
%We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
%The $E_1(m)$ values for the chaotic time series appear to be constant
%after the dimension is equal to six.
%The determination of six is given that any value of $m$ can be used as they 
%are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
%$E_2(m)$ values, for chaotic time series, are different to one 
%(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
%chaotic time series the minimum embedding dimension the time series 
%comes from a deterministic signal. With regard to the noise time series,  
%$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
%which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
%$E_1(m)$ values then indicate the minimum embedding dimension of the 
%noisy time series is thirteen, however all of the $E_2(m)$ values are 
%approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
%concluded that noise time series is a stochastic signal.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/falseNearestNeighbours/cao97/fig_v04}
%%\includegraphics[width=1.0\textwidth]{fig_v04}
%    \caption{
%	{\bf Minimum dimension embedding values with Cao's method.} 
%	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
%	with variations of $\tau$ values from one to twenty
%	for (E) chaotic and (F) random time series.
%	R code to reproduce the figure is available from \cite{hwum2018}.
%        }
%    \label{fig:e1e2}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
%computed but also a variation of $\tau$ from 1 to 20 is presented. 
%The purpose of such variation for $\tau$ is to show its independence with
%regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
%(Fig~\ref{fig:e1e2}A,B,C, and D). 
%However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
%a new threshold where $m$ values appear to be constant in $E_1 (m)$.
%In the case of the given examples and reported results, we defined such 
%threshold as 0.05. Further investigation is required for the selection of the 
%threshold in the $E_1(m)$, as the selection of the threshold in this work is 
%base on no particular method but visual inspection.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%*******************************************************************************
%\subsubsection*{Average Mutual Information}
%When selecting the delay dimension parameter, $\tau$, one can consider the 
%following two cases:
%(i) when $\tau$ is too small, the elements of time-delay embedding will be along
%the bisectrix of the phase space and the reconstruction is generally not 
%satisfactory, 
%(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
%time-delay embedding will become spread and uncorrelated which makes 
%recovering the underlying attractor more difficult if not 
%impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
%With regard to the algorithms to compute $\tau$, 
%Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
%function in which the first zero crossing is considered as the minimum delay 
%embedding parameter. However, the autocorrelation function is a linear 
%statistic over which the Average Mutual Information (AMI) algorithm is 
%preferred because the AMI takes into account the nonlinear dynamical 
%correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
%algorithm is described below to estimate the minimum delay embedding 
%parameter, \texorpdfstring{$\tau_0$}{T}.
%
%To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
%and then a probability distribution of data is computed \cite{kantz2003}.
%AMI is therefore denoted by $I(\tau)$ which is the average mutual 
%information between the original time series, $x(n)$, and the delayed 
%time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
%AMI is defined by
%%%********************************[EQUATION]************************************
%\begin{equation}\label{eq:ami}
%I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
%\end{equation}
%%%********************************[EQUATION]************************************
%Probabilities are defined as follows:
%$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
%the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
%the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
%that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
%in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
%For small $\tau$, AMI will be large and it will then decrease more or 
%less rapidly. As $\tau$ increase and goes to a large limit, 
%$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
%factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
%to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
%where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
%where the redundancy is the least" \cite{kantz2003}.
%
%For example, we compute the AMI for two time series:
%A) the $x$ solution of the Lorenz system, and
%B) a noise time series using a normal distribution with mean zero and 
%standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
%concluded that the amount of knowledge for any noise time series is zero 
%for which the first minimum embedding parameter is $\tau_0=1$. 
%On the contrary, the first minimum of the AMI for the chaotic time series 
%is $\tau_0=17$ which is the value that maximize the independence 
%between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
%\cite{bradley2015}.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/averageMutualInformation/example/ami_plots/figure_v04}
%%\includegraphics[width=1.0\textwidth]{figure_v04}
%    \caption{
%	{\bf Minimum delay embedding values with AMI's method.} 
%    	(A, B) AMI values where its first minimum value in the curve
%	is the minimum time delay embedding ($\tau_0$), 
%	for (C) a chaotic and (D) noise time series.
%	R code to reproduce the figure is available from \cite{hwum2018}.
%        }
%    \label{fig:amis}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%Similarly as Cao's algorithm negatives, AMI's algorithm is not an
%exception for negatives, which are worthwhile to mention for further 
%investigations.
%For instance, (i) is not clear why the choose of the first minimum of the AMI 
%is the minimum delay embedding parameter \cite{kantz2003} and 
%(ii) the probability distribution of the AMI function is computed
%with the use of histograms which depends on a heuristic choice of number of bins
%for which AMI depends on partitioning \cite{garcia2005e71}.
%
%
%
%
%
%
%%*******************************************************************************
%\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
%\subsection*{Recurrence Plots}
%Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
%in conservative systems, however such discovery was not put into practice 
%until the development of faster computers \cite{marwan2007},
%for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
%where recurrences in the dynamics of a system can be visualised using 
%Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
%propose a tool, called Recurrence Plot (RP), that provides insights into 
%high-dimensional dynamical systems where trajectories are very difficult to 
%visualise. Therefore, "RP helps us to investigate the 
%$m-$dimensional phase space trajectories through a two-dimensional 
%representation of its recurrences" \cite{marwan2015}.
%Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
%to the methodologies of the state space reconstruction and other dynamic 
%invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
%the recurrences of the trajectories in the phase space can provide 
%important clues to characterise natural process that present, for
%instance, periodicities (as Milankovitch cycles) or irregular cycles 
%(as El Ni\~no Southern Oscillation). 
%Such recurrences can not only be presented visually using Recurrence Plots (RP) 
%but also be quantified with Recurrence Quantification metrics, which leads 
%to applications of these in various areas such as economy, physiology, 
%neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
%
%For the creation of a recurrence plot based on time series 
%$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
%reconstruction with uniform time-delay embedding 
%$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
%\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
%where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
%and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
%The recurrence plot is therefore a two-dimensional $N \times N$ square 
%matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
%whenever $X(i)$ is sufficiently close to $X(j)$: 
%%%********************************[EQUATION]************************************
%\begin{equation}
%\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
%\end{equation}
%%%********************************[EQUATION]************************************
%where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
%$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
%function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
%(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
%RP is also characterised with a line of identity (LOI) which is a  
%diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/rqa/figure11/v02/figure11-v02}
%%\includegraphics[width=1.0\textwidth]{figure11-v02}
%    \caption{
%	{\bf Recurrence Plots.} 
%	(A) State space of the Lorenz system with controlling parameters 
%	($\rho=28, \sigma=10, \beta=8/3$). 
%	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
%	(black circle) of a given point at $i$ is a recurrent point and is 
%	represented as a black dot in the recurrence plot at location 
%	$(i, j)$ or white otherwise.
%	(B) Recurrence plot using the 
%	three components of the Lorenz system and the RP with no embeddings 
%	and threshold $\epsilon=5$.
%	This figure is adapted from \cite{marwan2015} and R code to reproduce 
%	it is available from \cite{hwum2018}.
%	}
%    \label{fig:mrp}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\subsection*{Structures of Recurrence Plots}
%Pattern formations in the RPs can be designated either 
%as topology for large-scale patterns or texture for small-scale patterns.
%In the case of topology, the following pattern formations are commonly presented:
%(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
%uniformly distributed noise (Fig~\ref{fig:rp2}A), 
%(ii) periodic and quasi-periodic systems where diagonal lines and 
%checkerboard structures represent oscillating systems, e.g., sinusoidal 
%signals (Fig~\ref{fig:rp2}B), 
%(iii) drift where paling or darkening recurrence points away from 
%the LOI is caused by drifting systems, 
%e.g., logistic map (Fig~\ref{fig:rp2}C), and
%(iv) disrupted where recurrence points are presented white areas or 
%bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
%(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
%Texture patterns in RPs can be categorised as:
%(i) single or isolated recurrence points that represent rare occurring states, 
%and do not persist for any time or fluctuate heavily,
%(ii) dots forming diagonal lines where the length of the small-scale parallel 
%lines in the diagonal are related to the ratio of determinism or predictability 
%in the dynamics of the system, and
%(iii) dots forming vertical and horizontal lines where the length of the 
%lines represent a time length where a state does not change or change very 
%slowly and these patterns formation represent discontinuities in the signal, 
%and (iv) dots clustering to inscribe rectangular regions which are related 
%to laminar states or singularities \cite{marwan2015}.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/rqa/figure12/v02/figure12-v02}
%%\includegraphics[width=1.0\textwidth]{figure12-v02}
%    \caption{
%	{\bf Patterns in Recurrence Plots.} 
%	Time-series with its respective recurrence plot for:
%	(A) uniformly distributed noise,
%	(B) super-positionet harmonic oscillation 
%	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
%	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
%	with a linearly increase term ($0.01*i$), and
%	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
%	This figure is adapted from \cite{marwan2015} and R code 
%	to reproduce it is available from \cite{hwum2018}.
%	}
%    \label{fig:rp2}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------
%
%Although, each of the previous pattern descriptions of the structures in the 
%RP offer an idea of the characteristics of dynamical systems, 
%these might be misinterpreted and conclusions might tend to be subjective 
%as these require the interpretation of a particular researcher(s).
%Because of that, recurrence quantification analyis (RQA) offer objective 
%methodologies to quantify such visual characteristics of previous 
%recurrent pattern structures in the RP \cite{zbilut1992}.
%
%%*******************************************************************************
%\subsection*{Recurrence Quantifications Analysis (RQA)}
%Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
%the density of recurrence points in RPs, then histograms of lengths for 
%diagonal lines in RPs were studied by \cite{trulla1996} which were the 
%introduction to the term recurrence quantification analysis (RQA) 
%\cite{marwan2008}. RQA has been applied in many fields such as life science, 
%engineering, physics, and others \cite{marwan2008}. Particularly in human 
%movement to investigate noise and complexity of postural control 
%\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
%coordination \cite{duran2017}. The success of RQA is not only due to its 
%simple algorithmic implementation but also to its capacity to detect tiny 
%modulations in frequency or phase which are not detectable using standard 
%methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
%RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
%dimension which is verified experimentally by \cite{iwanski1998}.
%RQA metrics comprehend percentage of recurrence, percentage of determinism, 
%ratio, Shannon entropy of the frequency distributions of the line lengths,
%maximal line length and divergence, trend and laminariy 
%\cite{marwan2007, marwan2015}. For this work, we considered only four 
%RQA metrics, due to its consistency with our preliminary experiments, 
%which are described below. Such metrics are computed the nonlinearTseries 
%R package \cite{nonlinearTseries2016}.
%
%\subsubsection*{REC values}
%The percentage of recurrence (REC) is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}
%REC(\epsilon,N) = 
%		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
%		\mathbf{R}^{m}_{i,j}(\epsilon),
%\end{equation}
%%%********************************[EQUATION]************************************
%which enumerates the black dots in the RP excluding the line of identity.
%REC is a measure of the relative density of recurrence points in the sparse 
%matrix \cite{marwan2015}.
%%REC is computed as follow with the nonlinearTseries package 
%% \cite{nonlinearTseries} 
%%  hist = getHistograms(neighs, ntakens, lmin, vmin)
%%  # calculate the number of recurrence points from the recurrence rate. The
%%  # recurrence rate counts the number of points at every distance in a concrete
%%  # side of the main diagonal.
%%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
%%  # add the maindiagonal
%%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
%%  # calculate the recurrence rate dividing the number of recurrent points at a
%%  # given distance by all points that could be at that distance
%%  recurrence_rate_vector = 
%%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
%%  # percentage of recurrent points
%%  REC = (numberRecurrencePoints) / ntakens ^ 2
%
%\subsubsection*{DET values} 
%The percent determinism (DET) is defined as the fraction of recurrence points
%that form diagonal lines and it is determined by
%%%********************************[EQUATION]************************************
%\begin{equation}
%DET=
%	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
%	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
%\end{equation}
%%%********************************[EQUATION]************************************
%where 
%%%********************************[EQUATION]************************************
%\begin{equation}
%H_D(l) = 
%	\sum^{N}_{i,j=1} 
%	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
%	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
%	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
%\end{equation}
%%%********************************[EQUATION]************************************
%is the histogram of the lengths of the diagonal structures in the RP.
%DET can be interpreted as the predictability of the system for periodic signals 
%which, in essence, have longer diagonal lines than the short diagonals lines
%for chaotic signals or absent diagonal lines for stochastic signals 
%\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
%the organisation of points in RPs  \cite{iwanski1998}. 
%%percent determinism (DET) is computed as follow with the nonlinearTseries 
%%package \cite{nonlinearTseries}  
%% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%%                                       lmin = 2, lDiagonalHistogram,
%%                                       recurrence_rate_vector, maxDistanceMD) {
%%  #begin parameter computations
%%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
%%  DET = num / numberRecurrencePoints
%
%
%\subsubsection*{RATIO values}
%RATIO is defined as the ratio between DET and REC and it is calculated from 
%the frequency distributions of the lengths of the diagonal lines.
%RATIO is useful to discover dynamic transitions \cite{marwan2015}.
%%  diagP = calculateDiagonalParameters(
%%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
%%    recurrence_rate_vector, maxDistanceMD
%%  )
%% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%%                                       lmin = 2, lDiagonalHistogram,
%%                                       recurrence_rate_vector, maxDistanceMD) {
%%  #begin parameter computations
%%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
%%  DET = num / numberRecurrencePoints
%% 
%%
%%    RATIO = diagP$DET / REC
%%
%
%\subsubsection*{ENT values}
%ENT is the Shannon entropy of the frequency distribution of the diagonal 
%line lengths and it is defined as
%%%********************************[EQUATION]************************************
%\begin{equation}
%ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with \quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
%\end{equation}
%%%********************************[EQUATION]************************************
%ENT reflects the complexity of the deterministic structure in the system.
%For instance, for uncorrelated noise or oscillations, 
%the value of ENT is rather small and indicates low complexity of the system,
%therefore "the higher the ENT is the more complex the dynamics are" 
%\cite{marwan2007, marwan2015}.
%%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
%%
%%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
%%                                       lmin = 2, lDiagonalHistogram,
%%                                       recurrence_rate_vector, maxDistanceMD) {
%%
%%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
%%  diff_0 = which(pl > 0)
%%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
%
% 
%\subsection*{Sensitivity and robustness of RPs and RQA.}
%RP and RQA are a very young field in nonlinear dynamics and many questions 
%are still open, for instance, different parameters for window length size 
%of the time series, embedding parameters or recurrence threshold can 
%generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
%
%The selection of recurrence threshold, $\epsilon$, can depend on the system 
%that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
%require to be very small, for trajectory reconstruction $\epsilon$ requires 
%to have a large thresholds or when studying dynamical transition 
%there is little importance about the selection of the threshold 
%\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
%the recurrence threshold  should be five times larger 
%than the standard deviation of the observational noise
%or the use of diagonal structures within the RP is suggested in order
%to find the optimal recurrence threshold for (quasi-)periodic process 
%\cite{marwan2011}.
%
%Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
%of choosing the right embedding parameters to perform RQA for which 
%many experiments have to be performed using different parameters in order 
%to have a better intuition of the nature of the time series and how 
%this is represented by using RQA.
%
%With that in mind, this work explores the sensitivity and robustness 
%of the window size of time series, embedding parameters for RSS with UTDE 
%and recurrence threshold for RP and RQA in order to gain a better insight 
%into the underlying time series collected from inertial sensors in the 
%context of human-humanoid imitation activities.
%
%
%
%
%
%
%
%%*******************************************************************************
%%*******************************************************************************
%%*******************************************************************************
%\section*{Experiment} \label{sec:experiment}
%We conducted an experiment in the context of human-humanoid imitation (HHI) 
%activities where participants were asked to imitate simple horizontal and 
%vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
%Such simple movements were repeated ten times for the participant 
%who copied NAO's arm movements in a face-to-face imitation activity.
%Also, wearable inertial measurement unit (IMU) sensors were attached 
%to the right hand of the participant and to the left hand of the robot 
%(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
%IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
%accelerometer, gyroscope and magnetometer sensors and quaternions 
%\cite{Comotti2014}. 
%%A further description of the NeMEMSi IMU sensors 
%%is given in \nameref{S1_AppendixB}.
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%  \centering
%\includegraphics[width=1.0\textwidth]{figures/experiment/hri/hri-v01}
%%\includegraphics[width=1.0\textwidth]{hri-v01}
%    \caption{
%	{\bf Human-humanoid imitation activities.} 
%    		Face-to-face human-humanoid imitation (HHI) activities for 
%		(A) HHI of horizontal arm movement, 
%		(B) Humanoid horizontal arm movement,
%		(C) HHI of vertical arm movement, and 
%		(D) Humanoid vertical arm movement.
%        }
%    \label{fig:hri}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%\subsection*{Participants}
%Twenty-three participants,
%from now on defined as $pN$ where $N$ is the number of participant, were 
%invited to do the experiment. However, data for three participants were 
%not used because the instructions for $p01$, who was the only left-handed,
%were mistakenly given in a way that movements were performed different
%from what had been planned, and for participants $p13$ and $p16$ 
%data were corrupted because bluetooth communications problems with the 
%sensors. With that in mind, data for twenty participants were analysed in 
%this work.
%
%Of the twenty participants, all of them are right-handed healthy participants 
%of whom four are females and sixteen are males with a mean and standard 
%deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
%informed consent forms prior to participation in the experiment.
%
%\subsection*{Human-humanoid imitation activities}
%For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
%two of which were attached to the right hand of the participant and the 
%other two to the left hand of the humanoid robot.
%Then, each participant was asked to imitate repetitions of simple horizontal
%and vertical arm movements performed by the humanoid robot in the following 
%conditions:
%(i) ten repetitions of horizontal arm movement at normal (HN) 
%	and faster (HF) speed (Figure~\ref{fig:hri} A), and
%(ii) ten repetitions of vertical arm movement at normal (VN) 
%	and faster (VF) speed (Figure~\ref{fig:hri} C).
%The normal and faster speed of arm movements is defined by the duration 
%in number of samples of one repetition of NAO's arm movements. 
%We select NAO's arm movements duration to distinguish between normal and 
%faster arm movements as NAO's movements have less variation between 
%repetition to repetition. The duration for one repetition of the horizontal 
%arm movement at normal speed, HN, is about 5 seconds considering 
%that each repetition last around 250 samples.
%For horizontal arm movement at faster speed, HF, each repetition were performed 
%in around 2 seconds which correspond to 90 samples of data.
%The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
%which is around 300 samples of data. For vertical arm movement at 
%faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
%to 120 samples of data.
%To visualise the distinction between normal and faster speed for horizontal 
%and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
%for axes Z and Y of the gyroscope sensors with four window lengths: 
%2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
%and 15-sec (750-samples).
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht] 
%\centering
%\includegraphics[width=1.0\textwidth]{figures/experiment/speedtimeseries/v01/sts01}
%%\includegraphics[width=1.0\textwidth]{sts01}
%    \caption{
%	{\bf Time series duration of horizontal and vertical arm movements.} 
%    		Time series of smoothed data from gyroscope sensor 
%		for different speed arm movements performed by NAO: 
%		(A) Horizontal Normal arm movement (HN), 
%		(B) Horizontal Faster arm movement (HF),
%		(C) Vertical Normal arm movement (VN) and 
%		(D) Vertical Faster arm movement (VF).
%		Additionally, 
%		(A) shows window sizes for 2-seconds (100 samples), 
%		5-seconds (250 samples), 10-seconds (500 samples) 
%		and 15-seconds (750 samples)
%		which are also presented in (B), (C) and (D).
%		R code to reproduce figure is available \cite{hwum2018}.
%        }
%	\label{fig:sts}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------
%
%
%
%
%\subsection*{Data from Inertial Measurement Units} 
%	\label{sec:experiment:subsec:imu}
%\subsubsection*{Raw data}
%Considering the work of \cite{shoaib2016} which provided evidence of an 
%improvement in recognition activities when combining data from 
%accelerometer and gyroscope. We focus our analysis from data of the 
%accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
%leave the data of the magnetometer and quaternions for further investigation 
%because of their possible variations with regard to magnetic disturbances.
%
%Data from the accelerometer is defined by triaxial time series 
%$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
%(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
%which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
%representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
%Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
%its respective axes subscripts $x,y,z$, where $n$ is the sample index 
%and $N$ is the same maximum length of all axes for the time series.
%Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
%%%---------------------------------(EQUATION)----------------------------------
%\begin{equation}\label{eq:A}
%\boldsymbol{A} =
%\begin{pmatrix}
%  A_x(n) \\
%  A_y(n) \\
%  A_z(n)
%\end{pmatrix}
%=
%\begin{pmatrix}
% a_x(1),a_x(2),\dots,a_x(N) \\
% a_y(1),a_y(2),\dots,a_y(N) \\
% a_z(1),a_z(2),\dots,a_z(N) 
%\end{pmatrix},
%\end{equation}
%%%---------------------------------(EQUATION)-----------------------------------
%and 
%%%---------------------------------(EQUATION)-----------------------------------
%\begin{equation}\label{eq:G}
%\boldsymbol{G} =
%\begin{pmatrix}
% G_x(n) \\
% G_y(n) \\
% G_z(n)
%\end{pmatrix}
%=
%\begin{pmatrix}
% g_x(1),g_x(2),\dots,g_x(N) \\
% g_y(1),g_y(2),\dots,g_y(N) \\
% g_z(1),g_z(2),\dots,g_z(N) 
%\end{pmatrix}.
%\end{equation}
%%%---------------------------------(EQUATION)-----------------------------------
%
%
%
%
%\subsubsection*{Postprocessing data}
%After the collection of raw data from four NeMEMsi sensors,
%time synchronisation alignment and interpolation were performed 
%in order to create time series with the same length and synchronised time.
%We refer the reader to \cite{Comotti2014} for further
%details about the time synchronisation process.
%
%\subsubsection*{Data normalization}
%Data is normalised to have zero mean and unit variance 
%using sample mean and sample standard deviation.
%The sample mean and sample standard deviation using $x(n)$ is given by
%%%---------------------------------(EQUATION)----------------------------------
%\begin{equation}\label{eq:ms}
%\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad  \sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
%\end{equation}
%%%---------------------------------(EQUATION)----------------------------------
%and the normalised data, $\hat{x}(n)$, is computed as follows
%%%---------------------------------(EQUATION)----------------------------------
%\begin{equation}\label{eq:normalization}
%\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
%\end{equation}
%%%---------------------------------(EQUATION)----------------------------------
%
%\subsubsection*{Smoothing data}
%Commonly, a low-pass filter is the method either to capture the low 
%frequencies that represent \%99 of the human body energy or to get 
%the gravitational and body motion components of 
%accelerations \cite{anguita2013}. However, for this work the elimination of 
%certain range of frequencies is not the main focus but the conservation of 
%the structure in the time series in terms of the width and heights where, 
%for instance, Savitzky-Golay filter can help to accomplish 
%such task \cite{press1992}. Savitzky-Golay filter is based on the 
%principle of moving window averaging which preserves the area under 
%the curve (the zeroth moment) and its mean position in time 
%(the first moment) but the line width (the second moment) is violated 
%and that results, for example, in the case of spectrometric data
%where a narrow spectral line is presented with reduced height and width. 
%With that in mind, the aim of Savitzky-Golay filtering is to find filter 
%coefficients $c_n$ that preserve higher momentums which are based on local 
%least-square polynomial approximations \cite{savitzkygolay1964, 
%press1992, schafer2011}.
%Therefore, Savitzky-Golay coefficients are therefore computed using 
%an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
%\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
%$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
%Smoothed signal is represented with a tilde over the original 
%signal: $\tilde{x}(n)$.
%
%\subsubsection*{Window size data}
%With regard to the window size, \cite{shoaib2016} investigated its effects 
%using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
%and combination of inertial sensors (accelerometer, gyroscope and linear 
%acceleration sensor) to improve the activity recognition performance for 
%repetitive activities (walking, jogging and biking) and less repetitive 
%activities (smoking, eating, giving a talk or drinking a coffee).
%With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
%increase of window length improve the recognition of complex activities 
%because these requires a large window length to learn the repetitive 
%motion patterns. Particularly, one of the recommendations is to use large 
%window size to recognise less repetitive activities which mainly involve 
%random hand gestures. Therefore, for the four activities 
%(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
%we select only four window sizes for analysis: 2-s window (100 samples), 
%5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
%(Fig~\ref{fig:sts}).
%
%
%

%% As a guideline references should be limited to 60 (this is not 
%% strictly enforced).
%\bibliography{sample}
%\bibliography{references}
\bibliography{../references/references}

\section*{Acknowledgements}

This work was funded by the University of Birmingham and the 
Mexican National Council of Science and Technology as part of 
Miguel Xochicale's PhD degree. I would like to thank to Dr. Dolores 
Columba Perez Flores and Prof. Martin J Russell for their helpful 
comments to polish the use of the language of mathematics and to 
Constantino Antonio Garcia Martinez et al. for developing the R 
package nonlinearTseries that help to accelerate the 
analysis of the nonlinear time series in this work.

\section*{Author contributions statement}
Contribution for this work of 
Miguel Xochicale (MX) and Chris Baber (CB) is as follows:
\begin{description}
\item[Conceptualisation] MX, CB
\item[Data Curation] MX
\item[Formal Analysis] MX
\item[Funding Acquisition] MX, CB
\item[Investigation] MX
\item[Methodology] MX
\item[Project Administration] MX
\item[Resources] CB
\item[Software] MX
\item[Supervision] CB
\item[Validation] MX
\item[Verification] MX
\item[Writing - Original Draft Preparation] MX
\item[Writing - Review] MX, CB
\item[Writing - Editing] MX
\end{description}

%
%\section*{Additional information}
%
%To include, in this order: \textbf{Accession codes} 
%(where applicable); \textbf{Competing interests} (mandatory statement). 
%
%The corresponding author is responsible for submitting a 
%\href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on 
%behalf of all authors of the paper. This statement must be included 
%in the submitted article file.
%
%\begin{figure}[ht]
%\centering
%\includegraphics[width=\linewidth]{stream}
%\caption{Legend (350 words max). Example legend text.}
%\label{fig:stream}
%\end{figure}

%\begin{table}[ht]
%\centering
%\begin{tabular}{|l|l|l|}
%\hline
%Condition & n & p \\
%\hline
%A & 5 & 0.1 \\
%\hline
%B & 10 & 0.01 \\
%\hline
%\end{tabular}
%\caption{\label{tab:example}Legend (350 words max). Example legend text.}
%\end{table}
%
%Figures and tables can be referenced in LaTeX using the ref command, 
%e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}

### 2019-02-25 code/anthropometrics/rscript/computing_statistics.R

@@ -1,0 +1,66 @@
+###############################################################################	
+#
+#  Computing age statistics of twenty participants.
+#
+#
+#
+#
+# Written by Miguel P Xochicale [http://mxochicale.github.io]
+#
+# If you see any errors or have any questions
+# please create an issue at https://github.com/mxochicale/phd-thesis-code-data/issues
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Statistics for 20 participants 
+	# (2) Filtering particpants (NOT USED)
+
+
+library('data.table') #
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+setwd('../')
+main_path <- getwd()
+data_path <- paste(main_path,'/data', sep='')
+
+
+
+################################################################################
+# (1) Statistics for 20 participants 
+
+datafilename <- 'data_20p.csv'
+#datafilename <- 'data_06p.csv'
+
+full_anthropometric_path_and_datafilename <- paste( data_path, '/', datafilename, sep='')
+pNNanthropometric <- fread(full_anthropometric_path_and_datafilename)
+
+all_pNNanthropometric <- pNNanthropometric[, head(.SD, .N), ] # replace .N for the number of participants
+summary(all_pNNanthropometric)
+
+m <- mean(all_pNNanthropometric$Age)
+sd <- sd(all_pNNanthropometric$Age)
+message('mean age: ', m, '   SD age: ' , sd)
+
+
+#################################################################################
+## (2) Filtering particpants (NOT USED)
+#males_pNNanthropometric <- all_pNNanthropometric[Gender == 'Male' &  Handeness == 'Right', !c("MusicTraining") ]
+#summary(males_pNNanthropometric)
+#
+#females_pNNanthropometric <- all_pNNanthropometric[Gender == 'Female' &  Handeness == 'Right', !c("MusicTraining") ]
+#summary(females_pNNanthropometric)
+#
+#mf_pNNanthropometric <- all_pNNanthropometric[Gender == c('Male','Female') &  Handeness == 'Right', !c("MusicTraining") ]
+#summary(mf_pNNanthropometric)
+#
+
+
+
+
+
+
+#############################################
+setwd(r_scripts_path) ## go back to the r-script source path

### 2019-02-25 code/creation_of_curated_timeseries/code_raw2aligned/A_mscripts/A_hri_rawData_TO_TimeAlignedDataForSeparateActivities_p01_to_p22_octave_linux.m

@@ -1,0 +1,2889 @@
+%
+% octave --no-gui
+% A_hri_rawData_TO_TimeAlignedDataForSeparateActivities_p01_to_p22_octave_linux
+%
+% sensors_timealignment_different_length_p16() function is used to correct the
+% time idex for participant 16
+%
+%
+% Written by Miguel P Xochicale [http://mxochicale.github.io]
+%
+% If you see any errors or have any questions
+% please create an issue at https://github.com/mxochicale/phd-thesis-code-data/issues
+%
+
+
+
+format long;
+clear; % Clear all variables and functions from memory
+
+m_scripts_path = pwd;
+
+source("MyFunctions_octave_linux.m");
+% two cd.. to go the main path
+
+cd ..;
+cd ..;
+main_path = pwd;
+cd ..;
+cd ..;
+cd ..;
+cd ..;
+user_path = pwd;
+
+
+data_path='/data_raw_p01-p22'
+datapath=strcat(main_path, data_path);
+
+tmpdatapath='/tmp/phdtmpdata';
+outcomes_path = strcat(user_path, tmpdatapath, '/time_aligned_different_length_data_for_separate_activities_p01_to_p22_octave_linux','/');
+
+outcomes_path_char= char(outcomes_path);
+if ~exist(outcomes_path_char,'dir');
+  {
+    myk =  cstrcat('system(',' ''mkdir', ' -p ',outcomes_path_char,''')');
+    eval( sprintf(myk) );
+  }
+end
+
+
+p01_data_path = strcat(datapath,'/p01/p01-HRI/');
+[p01_s01m, p01_s02m, p01_s03m, p01_s04m, p01_max_length_sensors] = cvs2arrays(p01_data_path);
+
+p02_data_path = strcat(datapath,'/p02/p02-HRI/');
+[p02_s01m, p02_s02m, p02_s03m, p02_s04m, p02_max_length_sensors] = cvs2arrays(p02_data_path);
+
[…]

### 2019-02-25 code/creation_of_curated_timeseries/code_raw2aligned/A_mscripts/test_hri_readrawdata.m

@@ -1,0 +1,175 @@
+%
+
+% USAGE
+% octave --no-gui
+% readrawdataHRI
+
+%  M-Script to read data files for SIX participant from \data_P01-P22.
+%  Data is separated by using labels00.txt in four activitities (a1,...,a4).
+%
+
+% Written by Miguel P Xochicale [http://mxochicale.github.io]
+%
+% If you see any errors or have any questions
+% please create an issue at https://github.com/mxochicale/phd-thesis-code-data/issues
+%
+
+
+format long;
+clear; % Clear all variables and functions from memory
+
+
+m_scripts_path = pwd;
+source("MyFunctions_octave_linux.m");
+
+cd ..;
+cd ..;
+main_path = pwd;
+
+data_path='/data_raw_p01-p22';
+datapath=strcat(main_path, data_path);
+
+
+p01_data_path = strcat(datapath,'/p01/p01-HRI/');
+[p01_s01m, p01_s02m, p01_s03m, p01_s04m, p01_max_length_sensors] = cvs2arrays(p01_data_path);
+
+%p02_data_path = strcat(datapath,'/p02/p02-HRI/');
+%[p02_s01m, p02_s02m, p02_s03m, p02_s04m, p02_max_length_sensors] = cvs2arrays(p02_data_path);
+%
+%p03_data_path = strcat(datapath,'/p03/p03-HRI/');
+%[p03_s01m, p03_s02m, p03_s03m, p03_s04m, p03_max_length_sensors] = cvs2arrays(p03_data_path);
+%
+%p04_data_path = strcat(datapath,'/p04/p04-HRI/');
+%[p04_s01m, p04_s02m, p04_s03m, p04_s04m, p04_max_length_sensors] = cvs2arrays(p04_data_path);
+%
+%p05_data_path = strcat(datapath,'/p05/p05-HRI/');
+%[p05_s01m, p05_s02m, p05_s03m, p05_s04m, p05_max_length_sensors] = cvs2arrays(p05_data_path);
+%
[…]

### 2019-02-25 code/creation_of_curated_timeseries/code_raw2aligned/B_rscripts/BA_creating_TidiedInterpolatedDATA_p01p22.R

@@ -1,0 +1,384 @@
+#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
+#
+#
+# FileName:           BA_creating_TidiedInterpolatedDATA_p01p22.R
+# FileDescription:
+#
+# Variables:
+#     participantsNN number
+#     aNN: activities
+#     sNN: sensors
+#
+# NOTE: Before running the current script, make sure that you have run:
+#           octave --no-gui
+#           /mscripts/rawData_TO_TimeAlignedDataForSeparateActivities16_p01_to_p22_octave_linux
+#
+
+
+#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
+# Defining participants, activities and sensors to read
+participantsNN <- c(1:22)
+#participantsNN <- c(1:12,14:15,17:22) # Discarding p13 and p16
+
+
+aNN <- c(1,2,3,4) # available activities 1,2,3,4
+sNN <- c(1,2,3,4)# available sensors 1,2,3,4
+
+participant_number <- c("p01_","p02_", "p03_", "p04_", "p05_","p06_","p07_","p08_","p09_","p10_","p11_","p12_","p13_","p14_","p15_","p16_","p17_","p18_","p19_","p20_","p21_","p22_")
+#participant_index <-  c(4,      6,      8,      10,     12,     14,   16,     18,   20,    22,    24,    26,    28,    30,    32,    34,    36,    38,    40,    42,    44,    46)
+#participant_index <-  c(2,      3,      4,      5,     6,     7,    8,     9,   10,    11,    12,    13,    14,    15,    16,    17,    18,    19,    20,    21,    22,    23)
+participant_index <-  c(3,      5,      7,      9,     11,     13,   15,     17,   19,    21,    23,    25,    27,    29,    31,    33,    35,    37,    39,    41,    43,    45)
+
+
+activity_number <- c("a01_","a02_","a03_","a04_")
+sensor_number <- c("s01_","s02_","s03_","s04_")
+
+
+
+
+
+#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
+# Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(akima) # for interpolation function aspline()
+
+
+options(digits=15)
+
+
+
[…]

### 2019-02-25 code/creation_of_curated_timeseries/code_raw2aligned/B_rscripts/BB_preprocessed_TidiedInterpolatedDATA_p01p22-v06.R

@@ -1,0 +1,150 @@
+###############################################################################	
+#
+# Create TidiedInterpolatedTHENpreprocessedData-v03.datatable
+# 
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading raw tidied interpolated data
+	# (3) Filtering out variables from the data.table
+		### (3.1) Selecting Participants
+		### (3.5) Adding/Deleting postprocessing vectors
+	# (4) Creating and Changing to Preprosseced Data Path
+	# (5) Writing Data
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../../")
+github_repository_path <- getwd()
+
+home_path <- Sys.getenv("HOME") 
+setwd(home_path)
+
+
+feature_path <- '/data/timeseries'
+outcomes_data_path <- paste(github_repository_path, feature_path, sep="")
+
+tidinterdata <- '/tmp/phdtmpdata/outcomes/preProcessedDataTable_p01_to_p22'
+preprossededata_path <- paste(home_path, tidinterdata, sep="")
+setwd(file.path(preprossededata_path))
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+################################################################################
+# (2) Reading raw tidied  interpolated data
+data <- fread("TidiedInterpolatedData.datatable", header=TRUE)
+
+################################################################################
+# (3) Filtering out variables from the data.table
+xdata <- data
+
+### Renaming Activity Variables
+xdata$Activity <- gsub('a01', 'HN', xdata$Activity)
+xdata$Activity <- gsub('a02', 'VN', xdata$Activity)
+xdata$Activity <- gsub('a03', 'HF', xdata$Activity)
+xdata$Activity <- gsub('a04', 'VF', xdata$Activity)
+
+## Renaming Sensor Variables
+xdata$Sensor <- gsub('s01', 'HS01', xdata$Sensor)
+xdata$Sensor <- gsub('s02', 'HS02', xdata$Sensor)
+xdata$Sensor <- gsub('s03', 'RS01', xdata$Sensor)
+xdata$Sensor <- gsub('s04', 'RS02', xdata$Sensor)
+
+
[…]

### 2019-02-25 code/dependencies/A_install-R.sh

@@ -1,0 +1,48 @@
+#!/bin/bash  
+
+##USAGE:
+#chmod +x file
+#./file
+
+
+
+sudo apt-key adv –keyserver keyserver.ubuntu.com –recv-keys E084DAB9
+##or
+#gpg --keyserver keyserver.ubuntu.com --recv-key E084DAB9
+#gpg -a --export E084DAB9 | sudo apt-key add -
+
+
+
+#SELECT UBUNTU VERSION
+
+# Basic format of next line deb https://<my.favorite.cran.mirror>/bin/linux/ubuntu <enter your ubuntu version>/
+#sudo add-apt-repository 'deb https://cran.ma.imperial.ac.uk/bin/linux/ubuntu precise/'
+
+## Ubuntu 12.04: precise
+#sudo add-apt-repository 'deb https://cloud.r-project.org/bin/linux/ubuntu precise/'
+
+## Ubuntu 14.04: trusty
+#sudo add-apt-repository 'deb https://cloud.r-project.org/bin/linux/ubuntu precise/'
+
+## Ubuntu 16.04: xenial
+sudo add-apt-repository 'deb-src https://cloud.r-project.org/bin/linux/ubuntu xenial'
+
+
+
+sudo apt-get update
+
+
+## DEPENDENCIES
+sudo apt-get update
+sudo apt-get --yes install gfortran
+sudo apt-get --yes install libblas-dev libatlas-base-dev liblapack-dev libatlas-base-dev libncurses5-dev
+sudo apt-get --yes install cdbs
+sudo apt-get --yes install r-base r-base-dev
+
+
+## devtools dependencies
+sudo apt-get --yes install libssl-dev libcurl4-openssl-dev
+sudo apt-get --yes install libxml2-dev # for xml2 calling load_all()
+sudo apt-get --yes install r-cran-rgl # rgl for nonlinearTseries
+
+

### 2019-02-25 code/rscripts/A_dataset/A_dataset.R

@@ -1,0 +1,209 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading raw tidied interpolated data
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/data/dataset'
+
+### Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, feature_path, sep="")
+### Raw Data Path
+rawdata_path <- paste(github_repo_path,'/data/timeseries',sep='')
+setwd(file.path(rawdata_path))
+#
+#
+#################################################################################
+## (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+source( paste(github_repo_path,'/code/rfunctions/ollin_cencah.R',sep='') )
+#
+#
+
+
+################################################################################
+# (2) Reading raw tidied  interpolated data
+xdata <- fread("TidiedInterpolatedRawData-v06.dt", header=TRUE)
+# add time variable
+xdata[, Time:=Sample/50, ]
+# add one sample to start at 0 instead of 0.2
+xdata[, Time:=Time-0.02, ]
+# reorder 
[…]

### 2019-02-25 code/rscripts/B_timeseries/B_timeseries.R

@@ -1,0 +1,309 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+#setwd("../")
+#github_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/data/dataset'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path,"/docs/figures/timeseries/src", sep="")
+
+
+## Data Path
+data_path <-  paste(github_repo_path, feature_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(ggplot2)
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+####  Deleting axis and leave only the ones for 
+#data <- data[, !c('Timestamp', 'MagnX', 'MagnY', 'MagnZ', 'qW', 'qX', 'qY', 'qZ') , with=FALSE]
+#       
+#"AccX" , "AccY", "AccZ", "GyroX", "GyroY", "GyroZ", "sg0zmuvAccX", "sg0zmuvAccY", "sg0zmuvAccZ",
+#"sg0zmuvGyroX", "sg0zmuvGyroY", "sg0zmuvGyroZ",
+#"sg1AccX", "sg1AccY", "sg1AccZ",
+#"sg1GyroX", "sg1GyroY", "sg1GyroZ", 
[…]

### 2019-02-25 code/rscripts/C_cao/B_cao.R

@@ -1,0 +1,597 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+	
+	# (4) Filtering data.table variables
+		# (4.1) Windowing data
+		# (4.2) Filtering variables
+			# (4.2.1) Activities Selection
+			# (4.2.2) Sensor Selection
+			# (4.2.3) Participant Selection
+			# (4.2.4) Axis Selection
+
+	# (5) Average Mutual Information 
+		# (5.1) Parameter selections
+		# (5.2) MAIN LOOP for CAOS 
+
+	# (6) Creating Outcome Data Path
+	# (7)  Writing Data
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+home_path <- Sys.getenv("HOME") 
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+feature_path <- '/data/utde'
+dataset_path <- '/data/dataset'
+
+## Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+
+################################################################################
+# (2) Reading data
+
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+################################################################################
+# (4) Filtering out variables in xdata data.table
+
+
+data <- data[,.(
+	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY,
[…]

### 2019-02-25 code/rscripts/C_cao/CA_cao_mm_H.R

@@ -1,0 +1,374 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+
+	# (6) Plot E values
+		# (6.1) main if for:
+			# (6.1.0) Creating paths for plots
+			# (6.1.1) E1 and E2 values
+			# (6.1.2) Minimum Embedding Dimensions Plots
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+setwd("../")
+github_path <- getwd()
+
+##VERSION 
+version <- '00'
+dataset_path <- '/data/utde'
+
+## Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, dataset_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+################################################################################
+# (2) Reading data
+
+
+#acts <- 'H'
+#windowl<-'w2'
+#wdata <- fread('MED-w2.dt', header=TRUE)
+
+
+#acts <- 'H'
+#windowl<-'w5'
+#wdata <- fread('MED-w5.dt', header=TRUE)
+
+acts <- 'H'
+windowl<-'w10'
+wdata <- fread('MED-w10.dt', header=TRUE)
+
+#acts <- 'H'
+#windowl<-'w15'
+#wdata <- fread('MED-w15.dt', header=TRUE)
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+	## (4.2.1) Activities Selection
+
+	ACPSAmdim <- NULL
+
+
+	activities <- c('HN','HF')
+
+	#########################################################
+	for (activity_k in 1:length(activities) ) {
+
+	activityk <- activities[activity_k]
+	message(activityk)
+	awdata <- wdata
+
+	if (activityk == 'HN' ) {
[…]

### 2019-02-25 code/rscripts/C_cao/CB_cao_mm_V.R

@@ -1,0 +1,378 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+
+	# (6) Plot E values
+		# (6.1) main if for:
+			# (6.1.0) Creating paths for plots
+			# (6.1.1) E1 and E2 values
+			# (6.1.2) Minimum Embedding Dimensions Plots
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+setwd("../")
+github_path <- getwd()
+
+##VERSION 
+version <- '00'
+dataset_path <- '/data/utde'
+
+## Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, dataset_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+################################################################################
+# (2) Reading data
+
+
+#acts <- 'V'
+#windowl<-'w2'
+#wdata <- fread('MED-w2.dt', header=TRUE)
+
+
+#acts <- 'V'
+#windowl<-'w5'
+#wdata <- fread('MED-w5.dt', header=TRUE)
+
+acts <- 'V'
+windowl<-'w10'
+wdata <- fread('MED-w10.dt', header=TRUE)
+
+#acts <- 'V'
+#windowl<-'w15'
+#wdata <- fread('MED-w15.dt', header=TRUE)
+
+
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+	## (4.2.1) Activities Selection
+
+	ACPSAmdim <- NULL
+
+
+	activities <- c('VN','VF')
+
+	#########################################################
+	for (activity_k in 1:length(activities) ) {
+
+	activityk <- activities[activity_k]
+	message(activityk)
+	awdata <- wdata
+
+	if (activityk == 'HN' ) {
[…]

### 2019-02-25 code/rscripts/C_cao/EA_plotcao_avMED_H.R

@@ -1,0 +1,315 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+
+	# (6) Plot E values
+		# (6.1) main if for:
+			# (6.1.0) Creating paths for plots
+			# (6.1.1) E1 and E2 values
+			# (6.1.2) Minimum Embedding Dimensions Plots
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+plots_path <- '/docs/figures/caoami/src'
+dataset_path <- '/data/utde'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, plots_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(ggplot2)
+library(data.table) # for manipulating data
+library(RColorBrewer) #for brewer.pal
+
+################################################################################
+# (2) Reading data
+
+
+
+
+#
+#windowl<-'w2'
+#aMED <- fread('aMED-H-w2.dt', header=TRUE)
+
+
+#windowl<-'w5'
+#aMED <- fread('aMED-H-w5.dt', header=TRUE)
+
+
+windowl<-'w10'
+aMED <- fread('aMED-H-w10.dt', header=TRUE)
+
+#windowl<-'w15'
+#aMED <- fread('aMED-H-w15.dt', header=TRUE)
+
+
+
+
+
+
+
+
+#################################################################################
+#### (6) Plot E values
+
+
+# (6.1.1)  Creating paths for plots
+if (file.exists(outcomes_plot_path)){
[…]

### 2019-02-25 code/rscripts/C_cao/EB_plotcao_avMED_V.R

@@ -1,0 +1,328 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+
+	# (6) Plot E values
+		# (6.1) main if for:
+			# (6.1.0) Creating paths for plots
+			# (6.1.1) E1 and E2 values
+			# (6.1.2) Minimum Embedding Dimensions Plots
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+plots_path <- '/docs/figures/caoami/src'
+dataset_path <- '/data/utde'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, plots_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(ggplot2)
+library(data.table) # for manipulating data
+library(RColorBrewer) #for brewer.pal
+
+################################################################################
+# (2) Reading data
+
+
+
+
+
+#windowl<-'w2'
+#aMED <- fread('aMED-V-w2.dt', header=TRUE)
+#
+#
+#windowl<-'w5'
+#aMED <- fread('aMED-V-w5.dt', header=TRUE)
+
+
+windowl<-'w10'
+aMED <- fread('aMED-V-w10.dt', header=TRUE)
+
+
+#windowl<-'w15'
+#aMED <- fread('aMED-V-w15.dt', header=TRUE)
+
+
+
+
+
+
+
+
+#################################################################################
+#### (6) Plot E values
+
+
+# (6.1.1)  Creating paths for plots
+if (file.exists(outcomes_plot_path)){
[…]

### 2019-02-25 code/rscripts/D_ami/B_ami.R

@@ -1,0 +1,441 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+	
+	# (4) Filtering data.table variables
+		# (4.1) Windowing data
+		# (4.2) Filtering variables
+			# (4.2.1) Sensor Selection
+			# (4.2.2) Activities Selection
+			# (4.2.3) Participant Selection
+			# (4.2.4) Axis Selection
+
+	# (5) Average Mutual Information 
+		# (5.1) Plot Avarage Mutual Information 
+		# (5.2) Creating and Changing to Preprosseced Data Path
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+homepath <- Sys.getenv("HOME")
+#message( 'homepath: ',   homepath )
+main_home_path <- '/quetzalcoalt'
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/data/utde'
+dataset_path <- '/data/dataset'
+
+## Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+#################################################################################
+## (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+library(devtools)
+load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+################################################################################
[…]

### 2019-02-25 code/rscripts/D_ami/CA_plotamiH.R

@@ -1,0 +1,318 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+	
+	# (4) Filtering data.table variables
+		# (4.1) Windowing data
+		# (4.2) Filtering variables
+			# (4.2.1) Sensor Selection
+			# (4.2.2) Activities Selection
+			# (4.2.3) Participant Selection
+			# (4.2.4) Axis Selection
+
+	# (5) Average Mutual Information 
+		# (5.1) Plot Avarage Mutual Information 
+		# (5.2) Creating and Changing to Preprosseced Data Path
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/caoami/src'
+dataset_path <- '/data/utde'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+library(ggplot2)
+library(RColorBrewer) #for brewer.pal
+
+
+###############################################################################
+## (2) Reading data
+
+
+noparticipants <- 20
+
+
+#AMI <- fread('AMI-w2.dt', header=TRUE)
+#MTD <- fread('MTD-w2.dt', header=TRUE)
+#windowl<-'w2'
+#
+
+#
+#AMI <- fread('AMI-w5.dt', header=TRUE)
+#MTD <- fread('MTD-w5.dt', header=TRUE)
+#windowl<-'w5'
+#
+#
+AMI <- fread('AMI-w10.dt', header=TRUE)
[…]

### 2019-02-25 code/rscripts/D_ami/CB_plotamiV.R

@@ -1,0 +1,314 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+	
+	# (4) Filtering data.table variables
+		# (4.1) Windowing data
+		# (4.2) Filtering variables
+			# (4.2.1) Sensor Selection
+			# (4.2.2) Activities Selection
+			# (4.2.3) Participant Selection
+			# (4.2.4) Axis Selection
+
+	# (5) Average Mutual Information 
+		# (5.1) Plot Avarage Mutual Information 
+		# (5.2) Creating and Changing to Preprosseced Data Path
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/caoami/src'
+dataset_path <- '/data/utde'
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+library(ggplot2)
+library(RColorBrewer) #for brewer.pal
+
+
+################################################################################
+# (2) Reading data
+
+
+noparticipants <- 20
+
+#
+#AMI <- fread('AMI-w2.dt', header=TRUE)
+#MTD <- fread('MTD-w2.dt', header=TRUE)
+#windowl<-'w2'
+#
+#
+#
+#AMI <- fread('AMI-w5.dt', header=TRUE)
+#MTD <- fread('MTD-w5.dt', header=TRUE)
+#windowl<-'w5'
+#
+
+AMI <- fread('AMI-w10.dt', header=TRUE)
[…]

### 2019-02-25 code/rscripts/E_rss/BA_mean_caoH.R

@@ -1,0 +1,263 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+
+	# (6) Plot E values
+		# (6.1) main if for:
+			# (6.1.0) Creating paths for plots
+			# (6.1.1) E1 and E2 values
+			# (6.1.2) Minimum Embedding Dimensions Plots
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+dataset_path <- '/data/utde'
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(ggplot2)
+library(data.table) # for manipulating data
+library(RColorBrewer) #for brewer.pal
+
+################################################################################
+# (2) Reading data
+
+
+
+
+#
+#windowl<-'w2'
+#aMED <- fread('aMED-H-w2.dt', header=TRUE)
+#MED <- fread('MED-w2.dt', header=TRUE)
+#EE <- fread('EE-w2.dt', header=TRUE)
+#
+
+#windowl<-'w5'
+#aMED <- fread('aMED-H-w5.dt', header=TRUE)
+#MED <- fread('MED-w5.dt', header=TRUE)
+#EE <- fread('EE-w5.dt', header=TRUE)
+#
+
+windowl<-'w10'
+aMED <- fread('aMED-H-w10.dt', header=TRUE)
+MED <- fread('MED-w10.dt', header=TRUE)
+EE <- fread('EE-w10.dt', header=TRUE)
+
+
+#windowl<-'w15'
[…]

### 2019-02-25 code/rscripts/E_rss/BB_mean_caoV.R

@@ -1,0 +1,276 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+
+	# (6) Plot E values
+		# (6.1) main if for:
+			# (6.1.0) Creating paths for plots
+			# (6.1.1) E1 and E2 values
+			# (6.1.2) Minimum Embedding Dimensions Plots
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+dataset_path <- '/data/utde'
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(ggplot2)
+library(data.table) # for manipulating data
+library(RColorBrewer) #for brewer.pal
+
+################################################################################
+# (2) Reading data
+
+
+
+
+##
+#windowl<-'w2'
+#aMED <- fread('aMED-V-w2.dt', header=TRUE)
+#MED <- fread('MED-w2.dt', header=TRUE)
+#EE <- fread('EE-w2.dt', header=TRUE)
+#
+#
+#windowl<-'w5'
+#aMED <- fread('aMED-V-w5.dt', header=TRUE)
+#MED <- fread('MED-w5.dt', header=TRUE)
+#EE <- fread('EE-w5.dt', header=TRUE)
+#
+
+windowl<-'w10'
+aMED <- fread('aMED-V-w10.dt', header=TRUE)
+MED <- fread('MED-w10.dt', header=TRUE)
+EE <- fread('EE-w10.dt', header=TRUE)
+
+
+#windowl<-'w15'
[…]

### 2019-02-25 code/rscripts/E_rss/BC_mean_amiH.R

@@ -1,0 +1,154 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+	
+	# (4) Filtering data.table variables
+		# (4.1) Windowing data
+		# (4.2) Filtering variables
+			# (4.2.1) Sensor Selection
+			# (4.2.2) Activities Selection
+			# (4.2.3) Participant Selection
+			# (4.2.4) Axis Selection
+
+	# (5) Average Mutual Information 
+		# (5.1) Plot Avarage Mutual Information 
+		# (5.2) Creating and Changing to Preprosseced Data Path
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+dataset_path <- '/data/utde'
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+
+library(ggplot2)
+library(RColorBrewer) #for brewer.pal
+
+
+################################################################################
+# (2) Reading data
+
+
+noparticipants <- 20
+
+
+#AMI <- fread('AMI-w2.dt', header=TRUE)
+#MTD <- fread('MTD-w2.dt', header=TRUE)
+#windowl<-'w2'
+#
+
+#
+#AMI <- fread('AMI-w5.dt', header=TRUE)
+#MTD <- fread('MTD-w5.dt', header=TRUE)
+#windowl<-'w5'
+#
+
+AMI <- fread('AMI-w10.dt', header=TRUE)
+MTD <- fread('MTD-w10.dt', header=TRUE)
+windowl<-'w10'
+
+
+#AMI <- fread('AMI-w15.dt', header=TRUE)
[…]

### 2019-02-25 code/rscripts/E_rss/BD_mean_amiV.R

@@ -1,0 +1,136 @@
+###############################################################################	
+#
+# 
+# 
+#
+#
+#
+#
+###############################################################################	
+	# TABLE OF CONTENT:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading data
+	
+	# (4) Filtering data.table variables
+		# (4.1) Windowing data
+		# (4.2) Filtering variables
+			# (4.2.1) Sensor Selection
+			# (4.2.2) Activities Selection
+			# (4.2.3) Participant Selection
+			# (4.2.4) Axis Selection
+
+	# (5) Average Mutual Information 
+		# (5.1) Plot Avarage Mutual Information 
+		# (5.2) Creating and Changing to Preprosseced Data Path
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+##VERSION 
+version <- '00'
+dataset_path <- '/data/utde'
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(ggplot2)
+library(RColorBrewer) #for brewer.pal
+
+
+################################################################################
+# (2) Reading data
+
+
+noparticipants <- 20
+
+
+#AMI <- fread('AMI-w2.dt', header=TRUE)
+#MTD <- fread('MTD-w2.dt', header=TRUE)
+#windowl<-'w2'
+
+#AMI <- fread('AMI-w5.dt', header=TRUE)
+#MTD <- fread('MTD-w5.dt', header=TRUE)
+#windowl<-'w5'
+#
+
+AMI <- fread('AMI-w10.dt', header=TRUE)
+MTD <- fread('MTD-w10.dt', header=TRUE)
+windowl<-'w10'
+
+#AMI <- fread('AMI-w15.dt', header=TRUE)
[…]

### 2019-02-25 code/rscripts/E_rss/C_rss.R

@@ -1,0 +1,435 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rss/src'
+dataset_path <- '/data/dataset'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+source( paste(github_repo_path,'/code/rfunctions/ollin_cencah.R',sep='') )
+
+
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+
+
+
+
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+### (4.1) Windowing Data [xdata[,.SD[1:2],by=.(Participant,Activity,Sensor)]]
+
+
+###########################
+###### one window lenght
+#windowsl <- c(100)
+#windowsn <- c('w2')
+#dimensions <- c(4)
+#delays <- c(5)
+#
+
+##########################
+##### one window lenght
+windowsl <- c(500)
+windowsn <- c('w10')
+dimensions <- c(6)
+delays <- c(8)
+
+
+
+############################
[…]

### 2019-02-25 code/rscripts/F_rps/Ba_timeseries_aH.R

@@ -1,0 +1,320 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rps/src'
+dataset_path <- '/data/dataset'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+
+# axis for horizontal movments
+data <- data[,.(
+	sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
+	), by=. (Participant,Activity,Sensor,Sample,Time)]
+
+
+setkey(data, Participant)
+pNN <- c('p01')
+#pNN <- 		c(	'p01', 'p02', 'p03', 'p04', 'p05', 'p06', 'p07', 'p08', 'p09', 'p10',
[…]

### 2019-02-25 code/rscripts/F_rps/Bb_timeseries_aV.R

@@ -1,0 +1,315 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rps/src'
+dataset_path <- '/data/dataset'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+# axis for horizontal movments
+data <- data[,.(
+	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY
+	), by=. (Participant,Activity,Sensor,Sample,Time)]
+
+
+setkey(data, Participant)
+pNN <- c('p01')
+#pNN <- 		c(	'p01', 'p02', 'p03', 'p04', 'p05', 'p06', 'p07', 'p08', 'p09', 'p10',
+#			'p11', 'p12', 'p13', 'p14', 'p15', 'p16', 'p17', 'p18', 'p19', 'p20')	
+data <- data[.(pNN)]
+
[…]

### 2019-02-25 code/rscripts/F_rps/Ca_rp_aH.R

@@ -1,0 +1,559 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rps/src'
+feature_plot_path <- '/rp_plots/H'
+dataset_path <- '/data/dataset'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+#r_scripts_path <- getwd()
+#setwd("../../../../../")
+#github_repo_path <- getwd()
+#setwd("../")
+#github_path <- getwd()
+#
+#
+###VERSION 
+#version <- '05'
+#feature_path <- '/rqa/hri'
+#feature_plot_path <- '/rp_plots/H'
+#	
+#
+### Outcomes Plot Path
+#outcomes_plot_path <- paste(github_path,"/phd-thesis/figs/results", feature_path, '/v', version,sep="")
+#
+### Data Path
+#data_path <- paste(github_repo_path,'/data-outputs', feature_path, '/v', version, sep="")
+#setwd(file.path(data_path))
+#
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
[…]

### 2019-02-25 code/rscripts/F_rps/Cb_rp_aV.R

@@ -1,0 +1,527 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rps/src'
+feature_plot_path <- '/rp_plots/V'
+dataset_path <- '/data/dataset'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
+
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(ggplot2)
+library(RColorBrewer)
+
+library(devtools)
+load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+
+
+
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+
+# axis for horizontal movments
+data <- data[,.(
+	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY
+	), by=. (Participant,Activity,Sensor,Sample,Time)]
+
+
[…]

### 2019-02-25 code/rscripts/F_rps/D_rqa.R

@@ -1,0 +1,516 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/data/rqa'
+dataset_path <- '/data/dataset'
+
+
+## Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(ggplot2)
+library(RColorBrewer)
+
+library(devtools)
+load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+
+
+
+
+
+
+
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
[…]

### 2019-02-25 code/rscripts/F_rps/Ea_rqa_plots_aH.R

@@ -1,0 +1,370 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+##VERSION 
+version <- '00'
+feature_plot_path <- '/docs/figures/rqa/src/H'
+#feature_path <- '/docs/figures/rps/src'
+dataset_path <- '/data/rqa'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_plot_path, sep="")
+
+## RQA Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+#rqa_path <- paste(data_path, '/rqa', sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(ggplot2)
+library(RColorBrewer)
+
+library(devtools)
+load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+
+
+
+################################################################################
+# (2) Reading data
+
+#windowksecs <- c('w2')
[…]

### 2019-02-25 code/rscripts/F_rps/Eb_rqa_plots_aV.R

@@ -1,0 +1,377 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+	# (3) Creating paths
+	# (4) Selecting Variables in data.table
+		# (4.1)	Selecting Participants
+	# (5) Adding vectors
+		# (5.1) Deleting some Magnetomer and quaternion data
+		# (5.2) zero mean and unit variance
+		# (5.3) Savitzky-Golay filter
+	# (6) Selecting Axis after postprocessing
+	# (7) Creating preprocessed data path
+	# (8) Writing data.table object to a file
+
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+##VERSION 
+version <- '00'
+feature_plot_path <- '/docs/figures/rqa/src/V'
+#feature_path <- '/docs/figures/rps/src'
+dataset_path <- '/data/rqa'
+
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path, feature_plot_path, sep="")
+
+## RQA Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+#rqa_path <- paste(data_path, '/rqa', sep="")
+setwd(file.path(data_path))
+
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(ggplot2)
+library(RColorBrewer)
+
+library(devtools)
+load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+
+
+
+
+
+################################################################################
+# (2) Reading data
+
+#windowksecs <- c('w2')
[…]

### 2019-02-25 code/rscripts/G_3Drqa/Fa_3Drqa_datatable.R

@@ -1,0 +1,547 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/data/rqa'
+dataset_path <- '/data/dataset'
+
+## Outcomes Data Path
+outcomes_data_path <- paste(github_repo_path, feature_path,  sep="")
+### Outcomes Plot Path
+#outcomes_plot_path <- paste(github_repo_path,'/hwum-manuscript/figures/results', feature_path, '/v', version, sep="")
+## Data Path
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(nonlinearTseries)
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+library(ggplot2)
+library(RColorBrewer)
+
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+
+
+data <- data[,.(
+	#sg0zmuvGyroZ
+	#sg0zmuvGyroZ, sg1zmuvGyroZ
+	#sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
+	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY, sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
+	), by=. (Participant,Activity,Sensor,Sample,Time)]
+
+
+
+
+
+################################################################################
+################################################################################
+################################################################################
[…]

### 2019-02-25 code/rscripts/G_3Drqa/Fb_3Drqa_plots.R

@@ -1,0 +1,358 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rqa/src/3drqa'
+data_path <- '/data/rqa'
+#feature_plot_path <- '/rqa_plots/topologies'
+
+## Outcomes Data Path
+#outcomes_data_path <- paste(github_repo_path,tput', feature_path, '/v', version, sep="")
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, data_path, sep="")
+setwd(file.path(data_path))
+
+
+#rqa_path <- paste(outcomes_data_path, '/rqa', sep="")
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(plot3D)
+
+
+
+################################################################################
+# (2) Reading data
+#
+#file_ext <- paste('RQAs_p01w100.dt',sep='')
+#W100 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w250.dt',sep='')
+#W250 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w500.dt',sep='')
+#W500 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w750.dt',sep='')
+#W750 <- fread( file_ext, header=TRUE)
+#
+#W <- rbind(W100, W250, W500, W750)
+#
+
+#file_ext <- paste('RQAs_p02w500.dt',sep='')
+#W <- fread( file_ext, header=TRUE)
+#
+
+file_ext <- paste('RQAs_p03w500.dt',sep='')
+W <- fread( file_ext, header=TRUE)
+
+
+
+################################################################################
+################################################################################
+################################################################################
[…]

### 2019-03-22 docs/manuscript/main.tex
[first draft]

@@
 \begin{abstract}
-Movement variability is defined as the variations that occur in motor
-performance across multiple repetitions of a task and such behaviour is an
+Human movement variability occurs in motor performance 
+across multiple repetitions of a task and such behaviour is an
 inherent feature within and between each persons' movement. Quantifying
@@
 methods in time domain and frequency domain fail to detect tiny modulations in
-frequency or phase for time series. We therefore consider methodologies from
-nonlinear dynamics such as reconstructed state space (RSS), uniform time-delay
-embedding (UTDE), recurrence plots (RPs) and recurrence quantification analysis
-(RQA) metrics. These algorithms require time series measured with sensors that
-provide well sampled data with little noise. However, for this study, we are
-interested in the analysis of data collected through wearable inertial sensors
-(IMUs) and its post-processing effects in each of the nonlinear dynamics
-methodologies. Twenty right-handed healthy participants (mean and standard
-deviation (SD) age of mean=19.8 (SD=1.39)) performed an experiment in
-human-humanoid imitation activities (H-H-I). Then IMUs were attached to the
-participants and the humanoid robot performed simple vertical and horizontal
-arm movements in normal and faster speed. With four window lengths and three
-levels of smoothed time series, we found visual differences in the patterns of
-RSSs and RPs. Then using metrics of RQA, we find out that the type of movements
-and the level of smoothness affects those metrics. In particular, we found that
-entropy values from RQA were well distributed and presenting variation in all
-the conditions for time series. Hence, we demonstrated the potential of
-nonlinear techniques to quantify human movement variability in the context of
-H-H-I can enhance the development of better diagnostic tools for various
-applications rehabilitation, sport science or for new forms of human-robot
-interaction.
-%280
+frequency or phase for time series. 
+For this work, we hence investigate methodologies from nonlinear dynamics such as 
+reconstructed state space (RSS), uniform time-delay embedding (UTDE), 
+recurrence plots (RPs) and recurrence quantification analysis (RQA) metrics. 
+Particularly, we are interested in the weaknesses and robustness 
+of nonlinear dynamics tools from raw and post-processed data of 
+wearable inertial sensors (IMUs). 
+In the reported experiment, twenty right-handed healthy participants 
+imitated simple vertical and horizontal arm movements in normal 
+and faster speed from an humanoid robot.
+With four window lengths and three levels of smoothed time series, 
+we found visual differences in the patterns of RSSs and RPs. 
+Then using metrics of RQA, we find out that the type of movements
+and the level of smoothness affects those metrics. 
+Our contribution for this work is that entropy values from RQA were 
+well distributed and presenting variation in all the conditions for time series. 
+This work have then the potential to enhance the development of 
+better diagnostic tools for various applications rehabilitation, 
+sport science or for new forms of human-robot interaction.
 \end{abstract}
+%226
 
@@
 \subsection*{Recurrences Plots}
-Considering the time series of Fig~\ref{fig:ts}, 
-Recurrence Plots were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and
-for vertical arm movements (Fig~\ref{fig:rp_aV}) 
+Recurrence Plots (RP) were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and 
+vertical arm movements (Fig~\ref{fig:rp_aV}) 
 using the average embedding parameters ($m=6$, $\tau=8$) 
 and an recurrence threshold of $\epsilon=1$.
-With regard to the selection of recurrence threshold,
-Marwan et al. \cite{marwan2011} pointed out that choosing an appropriate 
-recurrence threshold is crucial to get meaningful representations in RPs, 
-however, for this work where quantifying movement variability is our aim,
-we do not compute recurrence thresholds as as long as it is able to 
-represent the dynamical transitions in each of the time series.
-
-As similar as with the Reconstructed State Spaces, 
-the differences in the RPs can be easily noticed by eye 
-for different time series (Figs~\ref{fig:rp_aV} and ~\ref{fig:rp_aH}), 
-which lead us to apply Recurrence Quantification Analysis 
-to have an objective quantification of each of the time series.
+
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
-Considering the RPs for 20 participants performing four activities 
-(HN, HF, VN and VF) with sensors attached to the human (HS01) and to the 
-humanoid robot (RS01) and with the increase of smoothness of time series 
-(sg0zmuvGyroZ, sg1zmuvGyroZ and sg2zmuvGyroZ), 
-we hence compute four metrics of RQA metrics (REC, DET, RATIO and ENTR) with 
-embedding parameters $m=6$, $\tau=8$ and recurrence threshold $\epsilon=1$.
-
-%\newpage
+%Considering the RPs for 20 participants performing four activities 
+%(HN, HF, VN and VF) with sensors attached to the human and 
+%to the humanoid robot (HS01, RS01) and the increase of smoothness 
+%of time series (sg0zmuvGyroZ, sg1zmuvGyroZ and sg2zmuvGyroZ),
+Similarly to RP, we computed four RQA metrics (REC, DET, RATIO and ENTR) 
+with embedding parameters $m=6$, $\tau=8$ and 
+recurrence threshold $\epsilon=1$.
+
 \subsubsection*{REC values}
@@
 for type of movement, level of smoothness or type of sensor.
+There is also increase of DET values as the smoothness of the signal increase 
+(see the incremental changes of mean values (rhombus)).
+It can be noted that the interquartile range for faster movements
+(HF and VF) with no smoothing (sg0) is lower than the other
+levels of smoothness (sg1 and sg2).
 %See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
@@
 %which is a similar behaviour of RATIO values for oRS01 sensor.
-It can also be noticed a decrease of variation in RATIO values as the 
-smoothness of the time series is increasing (grey rhombus).
+For faster movements (HF and VF), it can be noted a decrease of 
+RATIO values as the smoothness of the time series is increasing (rhombus).
+For normal movements (HN and VN), the decrease of RATIO values
+is less evident than faster movements (see rhombus).
 %See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
@@
 \subsubsection*{ENTR values}
-Figs \ref{fig:RQABP}(D) show ENTR values, representing the complexity of 
-the structure the time series, for both horizontal and vertical movements.
-ENTR values for HS01 sensor show more variation 
+Figs \ref{fig:RQABP}(D) show ENTR values, Shannon entropy values 
+which represent the complexity of 
+the structure the time series, 
+for type of movement, level of smoothness or type of sensor.
+ENTR values for HS01 sensor show a higher variation  
 (interquartile range around 0.5)
@@
 to be more constant (interquartile range 0.1).
-It can also be said that the smoothness of time series affects
-each of the axis by an increase of mean values (see gray rhombos).
+It can also be noted the increase of smoothness of time series 
+also is presented with an increase of ENTR values (see gray rhombus).
 %See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
@@
 
-%
-%
-%\subsection*{RQA metrics with different embedding parameters, 
+\subsection*{3D RQA ENTR}
+Choosing an appropriate recurrence threshold is crucial to get 
+meaningful representations in RP and RQA, however, as previously shown, 
+one can selected the default recurrence threshold of $\epsilon=1$, 
+as long as it is able to represent the dynamical transitions 
+in each of the time series \cite{marwan2011}. %%Marwan et al. \cite{marwan2011} 
+However, to show how that the selection of recurrence threshold affects RQA
+values, 
+we computed ENTR values of RQA metrics for different embedding parameters
+$[1,10]= \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
+$[1,10]= \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $
+with an increase of 1
+and different recurrence thresholds 
+$[0.2,3]= \{ \epsilon \in \mathbb{R} | 0.2 \le \epsilon \le 3.0  \} $
+with increments of 0.1.
+For easily visualisation, Figs \ref{fig:RQA-IND} only show 3D RQA for 
+three recurrence thresholds $\epsilon=1, 2, 3$ with three levels of 
+smoothness (sg0, sg1, sg2). It can be noted in Figs \ref{fig:RQA-IND} 
+that the increase of recurrence threshold is associated to the increase 
+of ENTR values in any of the levels of smoothness (see values of the ENTR bar).
+One can also see that the maximum values of ENTR 
+for $\epsilon=1$ and sg0 are for embedding dimensions of 2 
+and then as $\epsilon$ increases the maximum values of ENTR are for
+an associated increase of in the embedding dimension.
+It can also be noted that the increase of level of smoothness (from sg0 to sg2) 
+is associated with both the increase of ENTR values and its smoothed 3D shape.
+
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}
+\centering
+\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa_epsilons}
+    \caption
+	[3D RQA ENTR values]{
+	{\bf 3D RQA ENTR values.}
+	RQA ENTR values 
+	for different embedding parameters
+	$[1,10]= \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
+	$[1,10]= \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
+	with an increase of 1
+	and different recurrence thresholds $\epsilon=1, 2, 3$.
+	RQA ENTR values are for $p03$, sensor HS01, of a window size of 10-secs (500 samples).
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+        }
+    \label{fig:RQA-IND}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+%of ENTR and plotted 3D surfaces using an unitary 
+%and a decimal increase of 0.1 for recurrence thresholds 
+% (Fig.~\ref{fig:topo_rqas}). 
+%Hence, Fig.~\ref{fig:topo_rqas}(A) shows an increase for 
+%REC values, the percentage of black docs in the RP, 
+%as the recurrence threshold increase,
+%while the variation for embedding parameters creates little decrease 
+%of REC values as the embedding dimensions.
+% increase and even a more slighter 
+%decrements of REC values for the increase of $\tau$.
+%For the 3D surface of DET values (Fig.~\ref{fig:topo_rqas}(B)), 
+%representing 
+%predictability and organisation of the RPs, one can be noted a plateau
+%for DET values near to 1 for embedding dimension parameters of less 
+%than 5 and recurrence threshold values of greater than 2 (red surface). 
+%It can also be observed that the increases of delay embedding made 
+%the DET values increase so as to make an cascade effect in the surface 
+%along with the increase of dimension embedding $m$.
+%For RATIO values, representing dynamic transitions, 
+%Fig.~\ref{fig:topo_rqas} shows that the 3D surface present 
+%a plateau (blue surface) of RATIO values 
+%near to zero for recurrence thresholds greater than 1.0, while 
+%fluctuations are more evident for recurrence thresholds of less than 1.0,
+%particularly it can also be noted an increase in the fluctuations of 
+%RATIO values as the embedding dimension is increasing. 
+%For ENTR values in Fig.~\ref{fig:topo_rqas}(D), 
+%representing the complexity of the structures in time series, 
+%one can be noted that the increase of recurrence threshold is, 
+%not strictly proportional to the increase of ENTR values. 
+%It can also be observed in Fig.~\ref{fig:topo_rqas}(D)
+%that the increase of delay embeddings affects 
+%little the ENTR values for embedding dimensions of 1, while 
+%for higher values of embedding dimensions 
+%there is a decrease of ENTR values, and 
+%there is a decrease of ENTR values as delay dimension value is increasing.
+%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}
+%\centering
+%\includegraphics[width=1.0\textwidth]{rqas}
+%    \caption
+%	[3D surfaces for RQA metrics]{
+%	{\bf 3D surfaces for RQA metrics.}
+%	3D surfaces for (A) REC, (B) DET, (C) RATIO and (D) ENTR values 
+%	with increasing pair of embedding parameters 
+%	($0 \ge m \le 10$, $0 \ge \tau \le 10$) 
+%	and recurrence thresholds (  $ 0.2 \ge \epsilon \le 3 $).
+%	RQA metrics are computed with the time series of participant $p01$ using 
+%	HS01 sensor, HN activity, sg0zmuvGyroZ axis and 500 samples 
+%	for window size length.
+%	Code and data to reproduce the figure is available from \cite{srep2019}.
+%	}
+%\label{fig:topo_rqas}
+%\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+%
+
+\subsection*{Other effects on 3D RQA ENTR values}
+% metrics with different embedding parameters, 
 %	recurrence thresholds, window lengths, levels of smoothness, 
 %	and time series structures.}
+
 %Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
@@
 %of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
-%(Fig.~\ref{fig:topo_rqas}). We also computed 3D surfaces of RQA metrics 
-%for different sensors and different activities 
-%(Fig.~\ref{fig:topo_sensoractivities}). RQA metrics are also affected by 
+%(Fig.~\ref{fig:topo_rqas}). 
+We also computed 3D surfaces of RQA metrics for different sensors 
+and different activities (Fig~\ref{fig:3dRQAENTR_sensoractivities}).
+From the maximum values of ENTR (lateral bars), one can see 
+a decrease of ENTR values for activities going from normal to 
+faster velocity and from human sensor (HS01) to robot sensor (RS01) 
+(See Fig~\ref{fig:3dRQAENTR_sensoractivities}). 
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa_sensors_activities}
+    \caption{
+	{\bf 3D surfaces of RQA ENTR metrics for sensors and activities.}
+	RQA ENTR values with embedding parameters
+	$[1,10]= \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
+	$[1,10]= \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
+	for different activities and sensors. 
+	RQA ENTR values are for $p03$, sg0 and window size of 10-secs (500 samples).
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+       }
+\label{fig:3dRQAENTR_sensoractivities}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+%RQA metrics are also affected by 
 %the window length where for example four window lengths of 100, 250, 500 
-%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}). Three level of 
-%smoothness were computed for RQA metrics showing smoothed 3D surfaces and 
-%the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
+%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}).
+
 %Similarly, 3D surfaces of RQA metrics were also computed for three 
 %participants (Fig.~\ref{fig:topo_participants}).
+
+%X Three level of 
+%X smoothness were computed for RQA metrics showing smoothed 3D surfaces and 
+%X the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
+
 %%%---------------------------------(FIGURE)-------------------------------------
@@
 %%%---------------------------------(FIGURE)------------------------------------
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/sa}
-%%\includegraphics[width=1.0\textwidth]{sa}
-%    \caption{
-%	{\bf 3D surfaces of RQA metrics for sensors and activities.}
-%	3D surfaces with increasing embedding parameters and recurrence 
-%	thresholds are for HS01 and RS01 sensors of HN, HF, VN and VF 
-%	activities. RQA metrics values are for time series of 
-%	participant p01 for sensors (HS01 and RS01), 
-%	activities (HN, HF, VN and VF) and for 	sg0zmuvGyroZ axis with 
-%	500 samples window length. 
-%	R code to reproduce the figure is available from \cite{hwum2018}.
-%       }
-%\label{fig:topo_sensoractivities}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
+
 %%%---------------------------------(FIGURE)-------------------------------------
@@
 %
-%%*******************************************************************************
-%%*******************************************************************************
-%%*******************************************************************************
-%\section*{Discussion}
-%%The Discussion should be succinct and must not contain subheadings.
-%
-%It is evidently that time series from different sources (participants, 
-%movements, axis type, window length or levels of smoothness) 
-%presents visual differences for embedding parameters and therefore for RRSs. 
-%For which, the selection of embedding parameters was our first challenge 
-%where we computed embedding parameters for each time series and then computed 
-%a sample mean over all time series in order to get two embedding parameters 
-%to compute all RSSs with its corresponded type of movement. Then we found 
-%that the quantification of variability with regard to the shape of the 
-%trajectories in RSSs requires more investigation since our original 
-%proposed method base on euclidean metric failed to quantify those 
-%trajectories. Specially, for trajectories which were not well unfolded. 
-%With that in mind, we proceed to take advantage of four RQA metrics 
-%(REC, DET, RATIO and ENTR) in order to avoid any subjective interpretations 
-%or personal bias with regard to the evolution of the trajectories in RSSs.
-%
-%
-%\subsection*{RQA metrics with fixed parameters}
-%Considering that RQA metrics were computed with fixed embedding parameters 
-%($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
-%the following. REC values, which represents the \% of black points in the RPs, 
-%were more affected with and increase in normal speed movements (HN and VN) 
-%than faster movements (HF and VF) for the sensor attached to the participants 
-%(HS01). Such decrease of REC values from normal speed to faster speed 
-%movements is also presented in data from sensor attached to the robot (RS01), 
-%and little can be said with regard to the dynamics of the time series coming 
-%from RS01. Similarly, DET values, representing predictability and 
-%organisation in the RPs, present little variation in the any of the time 
-%series where little can be said. In contrast, RATIO values, which represent 
-%dynamic transitions, were more variable for faster movements (HF and VF) 
-%than normal speed movements (HN and VN) with sensors attached to the 
-%participants (HS01). For data coming from sensors attached to the robot 
-%(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
-%more than values coming from vertical movmentes (VN, VF).
-%With that, it can be said that RATIO values can represent a bit better
-%than REC or DET metrics for the variability of imitation activities in 
-%each of the conditions for time series.
-%In the same way, ENTR values for HN were higher than values for HF
-%and ENTR values varied more for sensor attached to participants 
-%than ENTR values for sensors of the robot. It is evidently that 
-%the higher the entropy the more complex the dynamics are, 
-%however, ENTR values for HN appear a bit higher than HF values, 
-%for which we believe this happens because of the structure the time series
-%which appear more complex for HN than  HF movements which presented a 
-%more consistence repetition.
-%
-%We observed that some RQA metrics are affected by the smoothness of data.
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\section*{Discussion}
+%The Discussion should be succinct and must not contain subheadings.
+
+%DISCUSSION
+%Similarly as with the Reconstructed State Spaces, 
+%the differences in the RPs can be easily noticed by eye 
+%for different time series (Figs~\ref{fig:rp_aV} and ~\ref{fig:rp_aH}), 
+%which lead us to apply Recurrence Quantification Analysis (RQA) 
+%in order to have an objective quantification of each of the time series.
+%That said, 
+%%Considering the time series of Fig~\ref{fig:ts}, 
+%
+
+It is evidently that time series from different sources 
+(participants, movements, axis type, window length or levels of smoothness) 
+presents visual differences for embedding parameters and therefore for RRSs. 
+For which, the selection of embedding parameters was our first challenge 
+where we computed embedding parameters for each time series 
+(Fig \label{fig:caoami}) and then computed a sample mean over 
+all time series in order to get two embedding parameters 
+to compute all RRSs (Figs \label{fig:rss_aHw10} and \label{fig:rss_aVw10}). 
+%with its corresponded type of movement. 
+Then we found that the quantification of variability with regard to 
+the shape of the trajectories in RSSs requires more investigation 
+since our original proposed method base on euclidean metric failed 
+to quantify those trajectories. Specially, for trajectories which 
+were not well unfolded. 
+With that in mind, we proceed to apply 
+Recurrence Quantification Analysis (RQA) metrics 
+(REC, DET, RATIO and ENTR) in order to avoid any subjective interpretations 
+or personal bias with regard to the representation of the trajectories in RSSs.
+
+\subsection*{RQA metrics with fixed parameters}
+Considering that RQA metrics were computed with fixed embedding parameters 
+($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
+the following. REC values, which represents the \% of black points in the RPs, 
+were more affected with and increase in normal speed movements (HN and VN) 
+than faster movements (HF and VF) for the sensor attached to the participants 
+(HS01). Such decrease of REC values from normal speed to faster speed 
+movements is also presented in data from sensor attached to the robot (RS01), 
+and little can be said with regard to the dynamics of the time series coming 
+from RS01 (Fig \ref{fig:RQABP}A).
+Similarly, DET values, representing predictability and 
+organisation in the RPs, present little variation in the any of the time 
+series where little can be said (Fig \ref{fig:RQABP}B).
+In contrast, RATIO values, which represent 
+dynamic transitions, were more variable for faster movements (HF and VF) 
+than normal speed movements (HN and VN) with sensors attached to the 
+participants (HS01). For data coming from sensors attached to the robot 
+(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
+more than values coming from vertical movmentes (VN, VF) 
+(Fig \ref{fig:RQABP}C).
+With that, it can be said that RATIO values can represent a bit better
+than REC or DET metrics for the variability of imitation activities in 
+each of the conditions for time series.
+Similarly, ENTR values for HN were higher than values for HF
+and ENTR values varied more for sensor attached to participants 
+than ENTR values for sensors of the robot. It is evidently that 
+the higher the entropy the more complex the dynamics are, 
+however, ENTR values for HN appear a bit higher than HF values, 
+for which we believe this happens because of the structure the time series
+which appear more complex for HN than  HF movements which presented a 
+more consistence repetition (Fig \ref{fig:RQABP}D).
+
+Additionally, we observed that some RQA metrics are affected by the 
+smoothness of data. 
 %For which, we also explored the effect of smoothness of raw-normalised data 
-%where, for example, REC and DET values were not affected by the smoothness 
-%of data since these seemed to be constants. However, for RATIO values, 
-%the effect of smoothness can be noticed with a slight decrease of amplitude 
-%in any of the time series conditions which is also presented with ENTR values.
-%
-%
-%\subsection*{RQA metrics with different parameters}
-%Patterns in RPs and metrics for RQA are independent of embedding dimension 
-%parameters \cite{iwanski1998}, however, that is not the case when using 
-%different recurrence thresholds. Such changes of recurrence threshold values 
-%can modify the patterns in RPs and therefore the values of RQA metrics.
-%We therefore computed 3D surfaces to explore the sensibility and robustness of 
-%embedding parameters and recurrence threshold in RQA  metrics. Following the 
-%same methodology of computing 3D surfaces, we also considered variation of 
-%window length size to present RQA metrics dependencies with embedding 
-%parameters, recurrence thresholds and window length size.
-%
-%
-%
-%\section*{Conclusions}
-%We conclude that using a different level of smoothness for time series help us 
-%to visualise and to quantify the variation of movements between participants 
-%using RSSs, RPs and RQA. Also, it is important to mention that some RQA's 
-%metrics (e.g. DET and  ENTR) are more robust to the effect of smoothness 
-%of time series.
-%
-%Althought our our experiment is limited to twenty healthy right-handed 
-%participants of a range age of mean 19.8 SD=1.39, RQA metrics can quantify 
-%human movement variability. With that in mind, we conclude that 
+%where, 
+For instance, REC and DET values were not completely affected by the 
+smoothness of time-series since these RQA values seemed to be constants. 
+However, for RATIO values, 
+the effect of smoothness can be noticed with a slight decrease of amplitude 
+in any of the time series conditions which is also presented with ENTR values.
+
+\subsection*{RQA metrics with different parameters}
+Iwansky et al. \cite{iwanski1998} stated that patterns in RPs and 
+metrics for RQA are independent of embedding dimension parameters, 
+however, that is not the case when using 
+different recurrence thresholds. Such changes of recurrence threshold values 
+can modify the patterns in RPs and therefore the values of RQA metrics.
+We therefore computed 3D surfaces to explore the sensibility and robustness of 
+embedding parameters and recurrence threshold in RQA  metrics. Following the 
+same methodology of computing 3D surfaces, we also considered variation of 
+window length size to present RQA metrics dependencies with embedding 
+parameters, recurrence thresholds and window length size.
+
+\section*{Conclusions}
+Generally, we can conclude that using a different level of 
+smoothness for time series help us to visualise and to quantify 
+the variation of movements 
+between participants using RSSs, RPs and RQA. 
+It is important to mention that some RQA's metrics 
+(e.g. DET and ENTR) are more robust to the effect of 
+smoothness of time series.
+However, using determined RQA metrics will be depend on 
+what one want to quantify, for instance, one can compute 
+predicability, organisation,  
+dynamics transitions, or complexity and determinism of the 
+input time series. However, RQA metrics show certain constrains with 
+regard to activity type, window length and structure of the time series.
+%Similarly, such differences in time series created differences in each of the 
+%RQA metrics, 
+For instance, RATIO and ENTR are helpful to distinguish 
+differences in any of the categories of the time series (sensor, activity, 
+level of smoothness and number of participant), however for certain time series
+such as time series from the sensor attached to the robot seemed to have 
+little variations between each of the participants. 
+The latter phenomena was in a way evidently as robot degrees of freedom 
+did not allow it to move with a wide range of variability.
+
+Additionally, we can point out that even thought our experiment is 
+limited to twenty healthy right-handed participants of a range age 
+of mean 19.8 SD=1.39, RQA metrics show the potential to quantify 
+human movement variability in the context of 
+human-humanoid imitation activities.
+
+%With that in mind, we conclude that 
 %quantification of human-humanoid imitation activities is possible for 
 %participants of different ages, state of health and anthropomorphic features.
-%
-%
-%%
-%%Such understanding and measurement of movement variability using
-%%cheap wearable inertial sensors lead us to have a more intuitive selection of parameters
-%%to reconstruct the state spaces and to create meaningful interpretations
-%%of the recurrence plots and the results of the metrics with recurrence quantification 
-%%analysis. 
-%%
-%
-%In general, activity type, window length and structure of the time series 
-%affects the values of the metrics of RQA for which certain RQA metrics 
-%are better to describe determined type of movement.
-%Using determined RQA metrics depends on what one want to quantify, for instance, 
-%one can find predicability, organisation of the RPs, dynamics transitions, 
-%or complexity and determinism.
-%
-%
-%%
-%%However, we believe that further investigation is required to find the 
-%%right balance between the level of smoothness of the signal
-%%and defining what is the aim of RQA where 
-%%its representations using RSS, RP and RQA.
-%%Specially, where the level of smoothness does not affect 
-%%the variation of each of the movements quantification. 
-%%%We can also conclude that finding the right balance between 
-%%%smoothness and the raw data to capture movement variability is 
-%%%a still a problem that has many avenues for exploration.
-%%
-%
-%Similarly, such differences in time series created differences in each of the 
-%RQA metrics, for instance, RATIO and ENTR are helpful to distinguish 
-%differences in any of the categories of the time series (sensor, activity, 
-%level of smoothness and number of participant), however for certain time series 
-%(data from the sensor attached to the robot) seemed to have little variations 
-%between each of the participants. The latter phenomena was in a way evidently
-%as robot degrees of freedom did not allow it to move with a wide range of 
-%variability. 
-%
-%
-%
-%\section*{Future Work}
-%
-%\subsection*{Inertial Sensors}
-%To have more fundamental understating of nature of signals collected through 
-%inertial sensors in the context of human-robot interaction, we are considering 
-%to apply derivates to the acceleration data. We can then explore the jerkiness 
-%of movements and therefore the nature of arm movements which typically have 
-%minimum jerk \cite{flash1985}, its relationship with different body 
-%parts \cite{devries1982, mori2012} or the application of higher derivatives 
-%of displacement with respect time such as snap, crackle and 
-%pop \cite{eager2016}.
-%
-%\subsection*{RQA}
-%Having presented our results with RQA metrics, we believe that further 
-%investigation is required to have more robust metrics. For example, 
-%Marwan et al. \cite{marwan2007, marwan2015} reviewed different aspects to 
-%compute RPs using different criteria for neighbours, different norms 
-%($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or different methods 
-%to select the recurrence threshold $\epsilon$, such as using certain 
-%percentage of the signal \cite{letellier2006}, the amount of noise or 
-%using a factor based on the standard deviation of the observational noise 
-%among many others \cite{marwan2007}.
-%
-%
-%
-%
-%
-%
-%
-%%*******************************************************************************
-%%*******************************************************************************
-%%*******************************************************************************
-%\section*{Materials and methods}
-%\subsection*{State Space Reconstruction}
-%The method of state space reconstruction was originally proposed by 
-%\cite{packard1980} and formalised by \cite{takens1981}. Since then, different 
-%investigations and disciplines have benefited from the use of the method of 
-%state space reconstruction \cite{aguirre2009, stergiou2011, frank2010, 
-%sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
-%based on uniform time-delay embedding methodology which is a simple 
-%matrix implementation that can reconstruct an unknown $d-$dimensional 
-%manifold $M$ from a scalar time series (e.g. one-dimensional 
-%time series in $\mathbb{R}$). A manifold, in this context, is a 
-%multidimensional curved surface within a space (e.g. a saddle) 
-%\cite{guastello-gregson2011}.
-%
-%The use of a scalar time series is the main advantage of the method of state 
-%space reconstruction which in essence preserve dynamic invariants such as 
-%correlation dimension, fractal dimension, Lyaponov exponents, Kolmogorov-Sinai 
-%entropy and detrended fluctuation analysis \cite{bradley2015, 
-%Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
-%However, selecting appropriate embedding parameters which are sued to apply the 
-%state space reconstruction is still an open challenge for which we present 
-%introductions for the methodologies to compute such embedding parameters 
-%With that in mind, in the following subsections, we describe in more detail 
-%the state space reconstruction theorem (RSSs), uniform time-delay embedding 
-%theorem (UTDE), false nearest neighbours (FNN), average mutual 
-%information (AMI) and other methodologies for state space reconstruction.
-%
-%\subsection*{State Space Reconstruction Theorem}
-%Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
-%uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:ssr}
-%  s(t)=f^t [s(0)],
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
-%$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
-%unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
-%function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
-%iteration of $f$ that corresponds to an initial position 
-%$s(0) \in M $ \cite{takens1981}.
-%%$f$ is a smooth dynamical system, where smooth means a 
-%%continuously differentiable system
-%%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
-%Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
-%can be obtained with
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:measurement}
-%  x(t)=h[s(t)],
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
-%Reconstructed state space can then be described as an $n-$dimensional state 
-%space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
-%$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
-%is the uniform time-delay embedding with a dimension embedding $m$
-%and delay embedding $\tau$ and
-%$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
-%of dimensionality (e.g. Principal Component Analysis, 
-%Singular Value Decomposition, etc) being $n \leq m$.
-%With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
-%defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
-%$\boldsymbol{X}(t) = \Phi(s(t))$,
-%where $\Phi$ is a diffeomorphic map \cite{takens1981}
-%whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
-%dimension of $M$ \cite{garland2016}.
-%% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
-%% is called diffeomorphic if it is one-to-one correspondence and its inverse
-%% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
-%Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
-%space then a composition of functions represented with $F^t$ is induced on the
-%reconstructed state space determined:
-%%by true smooth dynamical system $f^t$ and $\Phi$:
-%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:st}
-%  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%With this in mind, an embedding is defined as "a smooth one-to-one 
-%coordinate transformation with a smooth inverse" and the total reconstruction 
-%map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
-%Fig~\ref{fig:ssr} illustrates the state space reconstruction.
-%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/rss/casdagli1991/figure1/rss_v04}
-%%\includegraphics[width=1.0\textwidth]{rss_v04}
-%    \caption{
-%	{\bf State space reconstruction methodology.}
-%	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
-%	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
-%    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
-%    	The time-delay embedding represented as the $\Phi$, maps the original
-%    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
-%	time-delay embedding matrix $\boldsymbol{X}(t)$.
-%    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
-%	a new state $y(t)$ of dimensions $n < m$.
-%    	(A) $M-$dimensional manifold representing the state space 
-%	(e.g. Lorenz system);
-%    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
-%   	(C) $m-$dimensional reconstructed state space with 
-%    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
-%    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
-%	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
-%	where $\Phi$ is the delay reconstruction map and 
-%	$\Psi$ is the coordinate transformation map.
-%	This figure is adapted from 
-%   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}
-%	and R code to reproduce it is available from \cite{hwum2018}.
-%    }
-%    \label{fig:ssr}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
-%
-%
-%
-%\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
-%Frank et al. and Sama et al. refer to the state space reconstruction
-%as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
-%However, we consider the term "uniform time-delay embedding"
-%as more descriptive and appropriate terminology for our work.
-%
-%The uniform time-delay embedding is represented as a matrix of uniform delayed 
-%copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
-%the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
-%samples of $\{ \boldsymbol{x}_n \}$.
-%$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
-%The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
-%and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
-%where $i$ goes from $0,1, \dots, (m-1)$.
-%Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
-%information of unobserved state variables and encapsulates the information of
-%the delayed copies of the available time series in the uniform time-delay 
-%embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
-%$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:tde}
-%\boldsymbol{X}^{m}_{\tau}  =
-%\begin{pmatrix}
-%\boldsymbol{ \tilde{x} }_n \\
-%\boldsymbol{ \tilde{x} }_{n-\tau} \\
-%\boldsymbol{ \tilde{x} }_{n-2\tau} \\
-%\vdots \\
-%\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
-%\end{pmatrix}^\intercal, 
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where $m$ is the embedding dimension, $\tau$ is the embedding delay and
-%$ ^\intercal$ denotes the transpose.
-%$m$ and $\tau$ are known as embedding parameters.
-%The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
-%$N-(m-1)\tau$ rows and $m$ columns and 
-%$N-(m-1)\tau$ defines the length of each delayed copy 
-%of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
-%%For further details and explicit examples of uniform time-delay 
-%%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
-%
-%
-%
-%
-%\subsection*{Estimation of Embedding Parameters}
-%The estimation of the embedding parameters ($m$ and $\tau$) 
-%is a fundamental step for the state space reconstruction with the use
-%of uniform time-delay embedding method.
-%With this in mind, we review two of the most common algorithms,
-%which will be used in our work, to compute the embedding
-%parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%*******************************************************************************
-%\subsubsection*{False Nearest Neighbours}
-%To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
-%used the method of false neighbours which can be understood as follows:
-%on the one hand, when the embedding dimension is too small to unfold the 
-%attractor "not all points that lie close each other will be neighbours and 
-%some points appear as neighbours because of the attractor has been projected 
-%down into an smaller space", on the other hand, when increasing the embedding 
-%dimension "points that are near to each other in the sufficient
-%embedding dimension should remain close as the dimension increase from $m$ 
-%to $m+1$ \cite{krakovska2015}".
-%From a mathematical point of view, the state space reconstruction theorem is 
-%done when the attractor is unfolded with either the minimum embedding 
-%dimension, $m_0$, or any other embedding dimension value where 
-%$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
-%leads to excessive computations \cite{bradley2015}. With this in mind, 
-%Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
-%where only the time-series and one delay embedding value are necessary 
-%to select the minimum embedding dimension. Cao's algorithm is based 
-%on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:e}
-%  \begin{aligned}
-%	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
-%	 &=
-%       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
-%       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
-%            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
-%  \end{aligned}
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
-%embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
-%From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
-%and $\tau$ for which $E_1(m)$ is defined as
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:e1}
-%E_1(m) = \frac{ E(m+1) } { E(m)}.
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
-%in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
-%As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
-%than some $m_0$, if the time series comes from a multidimensional state space
-%then $m_0 + 1$ is the minimum dimension".
-%Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
-%stochastic signals. $E_2(m)$ is defined as
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:e2}
-%E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:ee}
-%E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
-%|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%For instance, when the signal comes from random noise (values that are 
-%independent from each other), all $E_2(m)$ values are approximately equal 
-%to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
-%not constant for all $m$ (e.g. $E_2(m) \neq 1$).
-%
-%As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
-%series: the solution for the $x$ variable of the Lorenz system 
-%(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
-%and a variance of one (Fig~\ref{fig:e1e2}F).
-%We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
-%The $E_1(m)$ values for the chaotic time series appear to be constant
-%after the dimension is equal to six.
-%The determination of six is given that any value of $m$ can be used as they 
-%are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
-%$E_2(m)$ values, for chaotic time series, are different to one 
-%(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
-%chaotic time series the minimum embedding dimension the time series 
-%comes from a deterministic signal. With regard to the noise time series,  
-%$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-%which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
-%$E_1(m)$ values then indicate the minimum embedding dimension of the 
-%noisy time series is thirteen, however all of the $E_2(m)$ values are 
-%approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
-%concluded that noise time series is a stochastic signal.
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/falseNearestNeighbours/cao97/fig_v04}
-%%\includegraphics[width=1.0\textwidth]{fig_v04}
-%    \caption{
-%	{\bf Minimum dimension embedding values with Cao's method.} 
-%	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
-%	with variations of $\tau$ values from one to twenty
-%	for (E) chaotic and (F) random time series.
-%	R code to reproduce the figure is available from \cite{hwum2018}.
-%        }
-%    \label{fig:e1e2}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
-%
-%It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
-%computed but also a variation of $\tau$ from 1 to 20 is presented. 
-%The purpose of such variation for $\tau$ is to show its independence with
-%regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-%(Fig~\ref{fig:e1e2}A,B,C, and D). 
-%However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
-%a new threshold where $m$ values appear to be constant in $E_1 (m)$.
-%In the case of the given examples and reported results, we defined such 
-%threshold as 0.05. Further investigation is required for the selection of the 
-%threshold in the $E_1(m)$, as the selection of the threshold in this work is 
-%base on no particular method but visual inspection.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%*******************************************************************************
-%\subsubsection*{Average Mutual Information}
-%When selecting the delay dimension parameter, $\tau$, one can consider the 
-%following two cases:
-%(i) when $\tau$ is too small, the elements of time-delay embedding will be along
-%the bisectrix of the phase space and the reconstruction is generally not 
-%satisfactory, 
-%(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
-%time-delay embedding will become spread and uncorrelated which makes 
-%recovering the underlying attractor more difficult if not 
-%impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
-%With regard to the algorithms to compute $\tau$, 
-%Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
-%function in which the first zero crossing is considered as the minimum delay 
-%embedding parameter. However, the autocorrelation function is a linear 
-%statistic over which the Average Mutual Information (AMI) algorithm is 
-%preferred because the AMI takes into account the nonlinear dynamical 
-%correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
-%algorithm is described below to estimate the minimum delay embedding 
-%parameter, \texorpdfstring{$\tau_0$}{T}.
-%
-%To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
-%and then a probability distribution of data is computed \cite{kantz2003}.
-%AMI is therefore denoted by $I(\tau)$ which is the average mutual 
-%information between the original time series, $x(n)$, and the delayed 
-%time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
-%AMI is defined by
-%%%********************************[EQUATION]************************************
-%\begin{equation}\label{eq:ami}
-%I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%Probabilities are defined as follows:
-%$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
-%the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
-%the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
-%that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
-%in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
-%For small $\tau$, AMI will be large and it will then decrease more or 
-%less rapidly. As $\tau$ increase and goes to a large limit, 
-%$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
-%factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
-%to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
-%where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
-%where the redundancy is the least" \cite{kantz2003}.
-%
-%For example, we compute the AMI for two time series:
-%A) the $x$ solution of the Lorenz system, and
-%B) a noise time series using a normal distribution with mean zero and 
-%standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
-%concluded that the amount of knowledge for any noise time series is zero 
-%for which the first minimum embedding parameter is $\tau_0=1$. 
-%On the contrary, the first minimum of the AMI for the chaotic time series 
-%is $\tau_0=17$ which is the value that maximize the independence 
-%between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
-%\cite{bradley2015}.
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/averageMutualInformation/example/ami_plots/figure_v04}
-%%\includegraphics[width=1.0\textwidth]{figure_v04}
-%    \caption{
-%	{\bf Minimum delay embedding values with AMI's method.} 
-%    	(A, B) AMI values where its first minimum value in the curve
-%	is the minimum time delay embedding ($\tau_0$), 
-%	for (C) a chaotic and (D) noise time series.
-%	R code to reproduce the figure is available from \cite{hwum2018}.
-%        }
-%    \label{fig:amis}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
-%Similarly as Cao's algorithm negatives, AMI's algorithm is not an
-%exception for negatives, which are worthwhile to mention for further 
-%investigations.
-%For instance, (i) is not clear why the choose of the first minimum of the AMI 
-%is the minimum delay embedding parameter \cite{kantz2003} and 
-%(ii) the probability distribution of the AMI function is computed
-%with the use of histograms which depends on a heuristic choice of number of bins
-%for which AMI depends on partitioning \cite{garcia2005e71}.
-%
-%
-%
-%
-%
-%
-%%*******************************************************************************
-%\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
-%\subsection*{Recurrence Plots}
-%Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
-%in conservative systems, however such discovery was not put into practice 
-%until the development of faster computers \cite{marwan2007},
-%for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
-%where recurrences in the dynamics of a system can be visualised using 
-%Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
-%propose a tool, called Recurrence Plot (RP), that provides insights into 
-%high-dimensional dynamical systems where trajectories are very difficult to 
-%visualise. Therefore, "RP helps us to investigate the 
-%$m-$dimensional phase space trajectories through a two-dimensional 
-%representation of its recurrences" \cite{marwan2015}.
-%Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
-%to the methodologies of the state space reconstruction and other dynamic 
-%invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
-%the recurrences of the trajectories in the phase space can provide 
-%important clues to characterise natural process that present, for
-%instance, periodicities (as Milankovitch cycles) or irregular cycles 
-%(as El Ni\~no Southern Oscillation). 
-%Such recurrences can not only be presented visually using Recurrence Plots (RP) 
-%but also be quantified with Recurrence Quantification metrics, which leads 
-%to applications of these in various areas such as economy, physiology, 
-%neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
-%
-%For the creation of a recurrence plot based on time series 
-%$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
-%reconstruction with uniform time-delay embedding 
-%$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
-%\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
-%where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
-%and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
-%The recurrence plot is therefore a two-dimensional $N \times N$ square 
-%matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
-%whenever $X(i)$ is sufficiently close to $X(j)$: 
-%%%********************************[EQUATION]************************************
-%\begin{equation}
-%\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
-%$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
-%function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-%(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
-%RP is also characterised with a line of identity (LOI) which is a  
-%diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
-%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/rqa/figure11/v02/figure11-v02}
-%%\includegraphics[width=1.0\textwidth]{figure11-v02}
-%    \caption{
-%	{\bf Recurrence Plots.} 
-%	(A) State space of the Lorenz system with controlling parameters 
-%	($\rho=28, \sigma=10, \beta=8/3$). 
-%	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
-%	(black circle) of a given point at $i$ is a recurrent point and is 
-%	represented as a black dot in the recurrence plot at location 
-%	$(i, j)$ or white otherwise.
-%	(B) Recurrence plot using the 
-%	three components of the Lorenz system and the RP with no embeddings 
-%	and threshold $\epsilon=5$.
-%	This figure is adapted from \cite{marwan2015} and R code to reproduce 
-%	it is available from \cite{hwum2018}.
-%	}
-%    \label{fig:mrp}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
-%
-%
-%
-%
-%
-%%*******************************************************************************
-%%*******************************************************************************
-%%*******************************************************************************
-%\subsection*{Structures of Recurrence Plots}
-%Pattern formations in the RPs can be designated either 
-%as topology for large-scale patterns or texture for small-scale patterns.
-%In the case of topology, the following pattern formations are commonly presented:
-%(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-%uniformly distributed noise (Fig~\ref{fig:rp2}A), 
-%(ii) periodic and quasi-periodic systems where diagonal lines and 
-%checkerboard structures represent oscillating systems, e.g., sinusoidal 
-%signals (Fig~\ref{fig:rp2}B), 
-%(iii) drift where paling or darkening recurrence points away from 
-%the LOI is caused by drifting systems, 
-%e.g., logistic map (Fig~\ref{fig:rp2}C), and
-%(iv) disrupted where recurrence points are presented white areas or 
-%bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-%(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
-%Texture patterns in RPs can be categorised as:
-%(i) single or isolated recurrence points that represent rare occurring states, 
-%and do not persist for any time or fluctuate heavily,
-%(ii) dots forming diagonal lines where the length of the small-scale parallel 
-%lines in the diagonal are related to the ratio of determinism or predictability 
-%in the dynamics of the system, and
-%(iii) dots forming vertical and horizontal lines where the length of the 
-%lines represent a time length where a state does not change or change very 
-%slowly and these patterns formation represent discontinuities in the signal, 
-%and (iv) dots clustering to inscribe rectangular regions which are related 
-%to laminar states or singularities \cite{marwan2015}.
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/rqa/figure12/v02/figure12-v02}
-%%\includegraphics[width=1.0\textwidth]{figure12-v02}
-%    \caption{
-%	{\bf Patterns in Recurrence Plots.} 
-%	Time-series with its respective recurrence plot for:
-%	(A) uniformly distributed noise,
-%	(B) super-positionet harmonic oscillation 
-%	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-%	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-%	with a linearly increase term ($0.01*i$), and
-%	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-%	This figure is adapted from \cite{marwan2015} and R code 
-%	to reproduce it is available from \cite{hwum2018}.
-%	}
-%    \label{fig:rp2}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
-%
-%Although, each of the previous pattern descriptions of the structures in the 
-%RP offer an idea of the characteristics of dynamical systems, 
-%these might be misinterpreted and conclusions might tend to be subjective 
-%as these require the interpretation of a particular researcher(s).
-%Because of that, recurrence quantification analyis (RQA) offer objective 
-%methodologies to quantify such visual characteristics of previous 
-%recurrent pattern structures in the RP \cite{zbilut1992}.
-%
-%%*******************************************************************************
-%\subsection*{Recurrence Quantifications Analysis (RQA)}
-%Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
-%the density of recurrence points in RPs, then histograms of lengths for 
-%diagonal lines in RPs were studied by \cite{trulla1996} which were the 
-%introduction to the term recurrence quantification analysis (RQA) 
-%\cite{marwan2008}. RQA has been applied in many fields such as life science, 
-%engineering, physics, and others \cite{marwan2008}. Particularly in human 
-%movement to investigate noise and complexity of postural control 
-%\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
-%coordination \cite{duran2017}. The success of RQA is not only due to its 
-%simple algorithmic implementation but also to its capacity to detect tiny 
-%modulations in frequency or phase which are not detectable using standard 
-%methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
-%RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
-%dimension which is verified experimentally by \cite{iwanski1998}.
-%RQA metrics comprehend percentage of recurrence, percentage of determinism, 
-%ratio, Shannon entropy of the frequency distributions of the line lengths,
-%maximal line length and divergence, trend and laminariy 
-%\cite{marwan2007, marwan2015}. For this work, we considered only four 
-%RQA metrics, due to its consistency with our preliminary experiments, 
-%which are described below. Such metrics are computed the nonlinearTseries 
-%R package \cite{nonlinearTseries2016}.
-%
-%\subsubsection*{REC values}
-%The percentage of recurrence (REC) is defined as
-%%%********************************[EQUATION]************************************
-%\begin{equation}
-%REC(\epsilon,N) = 
-%		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
-%		\mathbf{R}^{m}_{i,j}(\epsilon),
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%which enumerates the black dots in the RP excluding the line of identity.
-%REC is a measure of the relative density of recurrence points in the sparse 
-%matrix \cite{marwan2015}.
-%%REC is computed as follow with the nonlinearTseries package 
-%% \cite{nonlinearTseries} 
-%%  hist = getHistograms(neighs, ntakens, lmin, vmin)
-%%  # calculate the number of recurrence points from the recurrence rate. The
-%%  # recurrence rate counts the number of points at every distance in a concrete
-%%  # side of the main diagonal.
-%%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
-%%  # add the maindiagonal
-%%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
-%%  # calculate the recurrence rate dividing the number of recurrent points at a
-%%  # given distance by all points that could be at that distance
-%%  recurrence_rate_vector = 
-%%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
-%%  # percentage of recurrent points
-%%  REC = (numberRecurrencePoints) / ntakens ^ 2
-%
-%\subsubsection*{DET values} 
-%The percent determinism (DET) is defined as the fraction of recurrence points
-%that form diagonal lines and it is determined by
-%%%********************************[EQUATION]************************************
-%\begin{equation}
-%DET=
-%	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
-%	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%where 
-%%%********************************[EQUATION]************************************
-%\begin{equation}
-%H_D(l) = 
-%	\sum^{N}_{i,j=1} 
-%	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
-%	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
-%	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%is the histogram of the lengths of the diagonal structures in the RP.
-%DET can be interpreted as the predictability of the system for periodic signals 
-%which, in essence, have longer diagonal lines than the short diagonals lines
-%for chaotic signals or absent diagonal lines for stochastic signals 
-%\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
-%the organisation of points in RPs  \cite{iwanski1998}. 
-%%percent determinism (DET) is computed as follow with the nonlinearTseries 
-%%package \cite{nonlinearTseries}  
-%% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%%                                       lmin = 2, lDiagonalHistogram,
-%%                                       recurrence_rate_vector, maxDistanceMD) {
-%%  #begin parameter computations
-%%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
-%%  DET = num / numberRecurrencePoints
-%
-%
-%\subsubsection*{RATIO values}
-%RATIO is defined as the ratio between DET and REC and it is calculated from 
-%the frequency distributions of the lengths of the diagonal lines.
-%RATIO is useful to discover dynamic transitions \cite{marwan2015}.
-%%  diagP = calculateDiagonalParameters(
-%%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
-%%    recurrence_rate_vector, maxDistanceMD
-%%  )
-%% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%%                                       lmin = 2, lDiagonalHistogram,
-%%                                       recurrence_rate_vector, maxDistanceMD) {
-%%  #begin parameter computations
-%%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
-%%  DET = num / numberRecurrencePoints
-%% 
-%%
-%%    RATIO = diagP$DET / REC
-%%
-%
-%\subsubsection*{ENT values}
-%ENT is the Shannon entropy of the frequency distribution of the diagonal 
-%line lengths and it is defined as
-%%%********************************[EQUATION]************************************
-%\begin{equation}
-%ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with \quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
-%\end{equation}
-%%%********************************[EQUATION]************************************
-%ENT reflects the complexity of the deterministic structure in the system.
-%For instance, for uncorrelated noise or oscillations, 
-%the value of ENT is rather small and indicates low complexity of the system,
-%therefore "the higher the ENT is the more complex the dynamics are" 
-%\cite{marwan2007, marwan2015}.
-%%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
-%%
-%%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%%                                       lmin = 2, lDiagonalHistogram,
-%%                                       recurrence_rate_vector, maxDistanceMD) {
-%%
-%%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
-%%  diff_0 = which(pl > 0)
-%%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
-%
-% 
-%\subsection*{Sensitivity and robustness of RPs and RQA.}
-%RP and RQA are a very young field in nonlinear dynamics and many questions 
-%are still open, for instance, different parameters for window length size 
-%of the time series, embedding parameters or recurrence threshold can 
-%generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
-%
-%The selection of recurrence threshold, $\epsilon$, can depend on the system 
-%that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
-%require to be very small, for trajectory reconstruction $\epsilon$ requires 
-%to have a large thresholds or when studying dynamical transition 
-%there is little importance about the selection of the threshold 
-%\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
-%the recurrence threshold  should be five times larger 
-%than the standard deviation of the observational noise
-%or the use of diagonal structures within the RP is suggested in order
-%to find the optimal recurrence threshold for (quasi-)periodic process 
-%\cite{marwan2011}.
-%
-%Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
-%of choosing the right embedding parameters to perform RQA for which 
-%many experiments have to be performed using different parameters in order 
-%to have a better intuition of the nature of the time series and how 
-%this is represented by using RQA.
-%
-%With that in mind, this work explores the sensitivity and robustness 
-%of the window size of time series, embedding parameters for RSS with UTDE 
-%and recurrence threshold for RP and RQA in order to gain a better insight 
-%into the underlying time series collected from inertial sensors in the 
-%context of human-humanoid imitation activities.
-%
-%
-%
-%
-%
-%
-%
-%%*******************************************************************************
-%%*******************************************************************************
-%%*******************************************************************************
-%\section*{Experiment} \label{sec:experiment}
-%We conducted an experiment in the context of human-humanoid imitation (HHI) 
-%activities where participants were asked to imitate simple horizontal and 
-%vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
-%Such simple movements were repeated ten times for the participant 
-%who copied NAO's arm movements in a face-to-face imitation activity.
-%Also, wearable inertial measurement unit (IMU) sensors were attached 
-%to the right hand of the participant and to the left hand of the robot 
-%(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
-%IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
-%accelerometer, gyroscope and magnetometer sensors and quaternions 
-%\cite{Comotti2014}. 
-%%A further description of the NeMEMSi IMU sensors 
-%%is given in \nameref{S1_AppendixB}.
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%  \centering
-%\includegraphics[width=1.0\textwidth]{figures/experiment/hri/hri-v01}
-%%\includegraphics[width=1.0\textwidth]{hri-v01}
-%    \caption{
-%	{\bf Human-humanoid imitation activities.} 
-%    		Face-to-face human-humanoid imitation (HHI) activities for 
-%		(A) HHI of horizontal arm movement, 
-%		(B) Humanoid horizontal arm movement,
-%		(C) HHI of vertical arm movement, and 
-%		(D) Humanoid vertical arm movement.
-%        }
-%    \label{fig:hri}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
-%
-%\subsection*{Participants}
-%Twenty-three participants,
-%from now on defined as $pN$ where $N$ is the number of participant, were 
-%invited to do the experiment. However, data for three participants were 
-%not used because the instructions for $p01$, who was the only left-handed,
-%were mistakenly given in a way that movements were performed different
-%from what had been planned, and for participants $p13$ and $p16$ 
-%data were corrupted because bluetooth communications problems with the 
-%sensors. With that in mind, data for twenty participants were analysed in 
-%this work.
-%
-%Of the twenty participants, all of them are right-handed healthy participants 
-%of whom four are females and sixteen are males with a mean and standard 
-%deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
-%informed consent forms prior to participation in the experiment.
-%
-%\subsection*{Human-humanoid imitation activities}
-%For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
-%two of which were attached to the right hand of the participant and the 
-%other two to the left hand of the humanoid robot.
-%Then, each participant was asked to imitate repetitions of simple horizontal
-%and vertical arm movements performed by the humanoid robot in the following 
-%conditions:
-%(i) ten repetitions of horizontal arm movement at normal (HN) 
-%	and faster (HF) speed (Figure~\ref{fig:hri} A), and
-%(ii) ten repetitions of vertical arm movement at normal (VN) 
-%	and faster (VF) speed (Figure~\ref{fig:hri} C).
-%The normal and faster speed of arm movements is defined by the duration 
-%in number of samples of one repetition of NAO's arm movements. 
-%We select NAO's arm movements duration to distinguish between normal and 
-%faster arm movements as NAO's movements have less variation between 
-%repetition to repetition. The duration for one repetition of the horizontal 
-%arm movement at normal speed, HN, is about 5 seconds considering 
-%that each repetition last around 250 samples.
-%For horizontal arm movement at faster speed, HF, each repetition were performed 
-%in around 2 seconds which correspond to 90 samples of data.
-%The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
-%which is around 300 samples of data. For vertical arm movement at 
-%faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
-%to 120 samples of data.
-%To visualise the distinction between normal and faster speed for horizontal 
-%and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
-%for axes Z and Y of the gyroscope sensors with four window lengths: 
-%2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
-%and 15-sec (750-samples).
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht] 
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/experiment/speedtimeseries/v01/sts01}
-%%\includegraphics[width=1.0\textwidth]{sts01}
-%    \caption{
-%	{\bf Time series duration of horizontal and vertical arm movements.} 
-%    		Time series of smoothed data from gyroscope sensor 
-%		for different speed arm movements performed by NAO: 
-%		(A) Horizontal Normal arm movement (HN), 
-%		(B) Horizontal Faster arm movement (HF),
-%		(C) Vertical Normal arm movement (VN) and 
-%		(D) Vertical Faster arm movement (VF).
-%		Additionally, 
-%		(A) shows window sizes for 2-seconds (100 samples), 
-%		5-seconds (250 samples), 10-seconds (500 samples) 
-%		and 15-seconds (750 samples)
-%		which are also presented in (B), (C) and (D).
-%		R code to reproduce figure is available \cite{hwum2018}.
-%        }
-%	\label{fig:sts}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
-%
-%
-%
-%
-%\subsection*{Data from Inertial Measurement Units} 
-%	\label{sec:experiment:subsec:imu}
-%\subsubsection*{Raw data}
-%Considering the work of \cite{shoaib2016} which provided evidence of an 
-%improvement in recognition activities when combining data from 
-%accelerometer and gyroscope. We focus our analysis from data of the 
-%accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
-%leave the data of the magnetometer and quaternions for further investigation 
-%because of their possible variations with regard to magnetic disturbances.
-%
-%Data from the accelerometer is defined by triaxial time series 
-%$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
-%(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
-%which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
-%representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
-%Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
-%its respective axes subscripts $x,y,z$, where $n$ is the sample index 
-%and $N$ is the same maximum length of all axes for the time series.
-%Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
-%%%---------------------------------(EQUATION)----------------------------------
-%\begin{equation}\label{eq:A}
-%\boldsymbol{A} =
-%\begin{pmatrix}
-%  A_x(n) \\
-%  A_y(n) \\
-%  A_z(n)
-%\end{pmatrix}
-%=
-%\begin{pmatrix}
-% a_x(1),a_x(2),\dots,a_x(N) \\
-% a_y(1),a_y(2),\dots,a_y(N) \\
-% a_z(1),a_z(2),\dots,a_z(N) 
-%\end{pmatrix},
-%\end{equation}
-%%%---------------------------------(EQUATION)-----------------------------------
-%and 
-%%%---------------------------------(EQUATION)-----------------------------------
-%\begin{equation}\label{eq:G}
-%\boldsymbol{G} =
-%\begin{pmatrix}
-% G_x(n) \\
-% G_y(n) \\
-% G_z(n)
-%\end{pmatrix}
-%=
-%\begin{pmatrix}
-% g_x(1),g_x(2),\dots,g_x(N) \\
-% g_y(1),g_y(2),\dots,g_y(N) \\
-% g_z(1),g_z(2),\dots,g_z(N) 
-%\end{pmatrix}.
-%\end{equation}
-%%%---------------------------------(EQUATION)-----------------------------------
-%
-%
-%
-%
-%\subsubsection*{Postprocessing data}
-%After the collection of raw data from four NeMEMsi sensors,
-%time synchronisation alignment and interpolation were performed 
-%in order to create time series with the same length and synchronised time.
-%We refer the reader to \cite{Comotti2014} for further
-%details about the time synchronisation process.
-%
-%\subsubsection*{Data normalization}
-%Data is normalised to have zero mean and unit variance 
-%using sample mean and sample standard deviation.
-%The sample mean and sample standard deviation using $x(n)$ is given by
-%%%---------------------------------(EQUATION)----------------------------------
-%\begin{equation}\label{eq:ms}
-%\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad  \sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
-%\end{equation}
-%%%---------------------------------(EQUATION)----------------------------------
-%and the normalised data, $\hat{x}(n)$, is computed as follows
-%%%---------------------------------(EQUATION)----------------------------------
-%\begin{equation}\label{eq:normalization}
-%\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
-%\end{equation}
-%%%---------------------------------(EQUATION)----------------------------------
-%
-%\subsubsection*{Smoothing data}
-%Commonly, a low-pass filter is the method either to capture the low 
-%frequencies that represent \%99 of the human body energy or to get 
-%the gravitational and body motion components of 
-%accelerations \cite{anguita2013}. However, for this work the elimination of 
-%certain range of frequencies is not the main focus but the conservation of 
-%the structure in the time series in terms of the width and heights where, 
-%for instance, Savitzky-Golay filter can help to accomplish 
-%such task \cite{press1992}. Savitzky-Golay filter is based on the 
-%principle of moving window averaging which preserves the area under 
-%the curve (the zeroth moment) and its mean position in time 
-%(the first moment) but the line width (the second moment) is violated 
-%and that results, for example, in the case of spectrometric data
-%where a narrow spectral line is presented with reduced height and width. 
-%With that in mind, the aim of Savitzky-Golay filtering is to find filter 
-%coefficients $c_n$ that preserve higher momentums which are based on local 
-%least-square polynomial approximations \cite{savitzkygolay1964, 
-%press1992, schafer2011}.
-%Therefore, Savitzky-Golay coefficients are therefore computed using 
-%an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
-%\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
-%$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
-%Smoothed signal is represented with a tilde over the original 
-%signal: $\tilde{x}(n)$.
-%
-%\subsubsection*{Window size data}
-%With regard to the window size, \cite{shoaib2016} investigated its effects 
-%using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
-%and combination of inertial sensors (accelerometer, gyroscope and linear 
-%acceleration sensor) to improve the activity recognition performance for 
-%repetitive activities (walking, jogging and biking) and less repetitive 
-%activities (smoking, eating, giving a talk or drinking a coffee).
-%With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
-%increase of window length improve the recognition of complex activities 
-%because these requires a large window length to learn the repetitive 
-%motion patterns. Particularly, one of the recommendations is to use large 
-%window size to recognise less repetitive activities which mainly involve 
-%random hand gestures. Therefore, for the four activities 
-%(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
-%we select only four window sizes for analysis: 2-s window (100 samples), 
-%5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
-%(Fig~\ref{fig:sts}).
-%
-%
-%
+
+%Such understanding and measurement of movement variability using
+%cheap wearable inertial sensors lead us to have a more intuitive selection of parameters
+%to reconstruct the state spaces and to create meaningful interpretations
+%of the recurrence plots and the results of the metrics with recurrence quantification 
+%analysis. 
+%
+
+%
+%However, we believe that further investigation is required to find the 
+%right balance between the level of smoothness of the signal
+%and defining what is the aim of RQA where 
+%its representations using RSS, RP and RQA.
+%Specially, where the level of smoothness does not affect 
+%the variation of each of the movements quantification. 
+%%We can also conclude that finding the right balance between 
+%%smoothness and the raw data to capture movement variability is 
+%%a still a problem that has many avenues for exploration.
+%
+
+\section*{Future Work}
+
+\subsection*{Inertial Sensors}
+To have more fundamental understating of nature of signals collected through 
+inertial sensors in the context of human-robot interaction, we are considering 
+to apply derivates to the acceleration data. We can then explore the jerkiness 
+of movements and therefore the nature of arm movements which typically have 
+minimum jerk \cite{flash1985}, its relationship with different body 
+parts \cite{devries1982, mori2012} or the application of higher derivatives 
+of displacement with respect time such as snap, crackle and pop \cite{eager2016}.
+
+\subsection*{RQA ENTR}
+Having shown that RQA ENTR metrics are robust against different data
+time series, we believe that further investigation is required to be done. 
+For example, Marwan et al. \cite{marwan2007, marwan2015} reviewed 
+different aspects to compute RPs using different criteria for neighbours, 
+different norms ($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or 
+different methods to select the recurrence threshold $\epsilon$, 
+such as using certain percentage of the signal \cite{letellier2006}, 
+the amount of noise or using a factor based on the standard deviation 
+of the observational noise among many others \cite{marwan2007}.
+
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\section*{Methods}
+\subsection*{State Space Reconstruction}
+The method of state space reconstruction was originally proposed by 
+\cite{packard1980} and formalised by \cite{takens1981}. Since then, different 
+investigations and disciplines have benefited from the use of the method of 
+state space reconstruction \cite{aguirre2009, stergiou2011, frank2010, 
+sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
+based on uniform time-delay embedding methodology which is a simple 
+matrix implementation that can reconstruct an unknown $d-$dimensional 
+manifold $M$ from a scalar time series (e.g. one-dimensional 
+time series in $\mathbb{R}$). A manifold, in this context, is a 
+multidimensional curved surface within a space (e.g. a saddle) 
+\cite{guastello-gregson2011}.
+
+The use of a scalar time series is the main advantage of the method of state 
+space reconstruction which in essence preserve dynamic invariants such as 
+correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
+entropy and detrended fluctuation analysis \cite{bradley2015, 
+Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
+However, selecting appropriate embedding parameters which are sued to apply the 
+state space reconstruction is still an open challenge for which we present 
+introductions for the methodologies to compute such embedding parameters 
+With that in mind, in the following subsections, we describe in more detail 
+the state space reconstruction theorem (RSSs), uniform time-delay embedding 
+theorem (UTDE), false nearest neighbours (FNN) and average mutual 
+information (AMI). 
+%and other methodologies for state space reconstruction.
+
+\subsection*{State Space Reconstruction Theorem}
+Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
+uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:ssr}
+  s(t)=f^t [s(0)],
+\end{equation}
+%%********************************[EQUATION]************************************
+where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
+$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
+unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
+function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
+iteration of $f$ that corresponds to an initial position 
+$s(0) \in M $ \cite{takens1981}.
+%$f$ is a smooth dynamical system, where smooth means a 
+%continuously differentiable system
+%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
+Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
+can be obtained with
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:measurement}
+  x(t)=h[s(t)],
+\end{equation}
+%%********************************[EQUATION]************************************
+where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
+Reconstructed state space can then be described as an $n-$dimensional state 
+space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
+$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
+is the uniform time-delay embedding with a dimension embedding $m$
+and delay embedding $\tau$ and
+$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
+of dimensionality (e.g. Principal Component Analysis, 
+Singular Value Decomposition, etc) being $n \leq m$.
+With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
+defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
+$\boldsymbol{X}(t) = \Phi(s(t))$,
+where $\Phi$ is a diffeomorphic map \cite{takens1981}
+whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
+dimension of $M$ \cite{garland2016}.
+% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
+% is called diffeomorphic if it is one-to-one correspondence and its inverse
+% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
+Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
+space then a composition of functions represented with $F^t$ is induced on the
+reconstructed state space determined:
+%by true smooth dynamical system $f^t$ and $\Phi$:
+%********************************[EQUATION]************************************
+\begin{equation}\label{eq:st}
+  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
+\end{equation}
+%%********************************[EQUATION]************************************
+With this in mind, an embedding is defined as "a smooth one-to-one 
+coordinate transformation with a smooth inverse" and the total reconstruction 
+map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
+Fig~\ref{fig:ssr} illustrates the state space reconstruction.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
+    \caption{
+	{\bf State space reconstruction methodology.}
+	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
+	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
+    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
+    	The time-delay embedding represented as the $\Phi$, maps the original
+    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
+	time-delay embedding matrix $\boldsymbol{X}(t)$.
+    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
+	a new state $y(t)$ of dimensions $n < m$.
+    	(A) $M-$dimensional manifold representing the state space 
+	(e.g. Lorenz system);
+    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
+   	(C) $m-$dimensional reconstructed state space with 
+    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
+    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
+	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
+	where $\Phi$ is the delay reconstruction map and 
+	$\Psi$ is the coordinate transformation map.
+	This figure is adapted from 
+   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+    }
+    \label{fig:ssr}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
+Frank et al. and Sama et al. refer to the state space reconstruction
+as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
+However, we consider the term "uniform time-delay embedding"
+as more descriptive and appropriate terminology for our work.
+Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed 
+copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
+the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
+samples of $\{ \boldsymbol{x}_n \}$.
+$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
+The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
+and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
+where $i$ goes from $0,1, \dots, (m-1)$.
+Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
+information of unobserved state variables and encapsulates the information of
+the delayed copies of the available time series in the uniform time-delay 
+embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
+$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:tde}
+\boldsymbol{X}^{m}_{\tau}  =
+\begin{pmatrix}
+\boldsymbol{ \tilde{x} }_n \\
+\boldsymbol{ \tilde{x} }_{n-\tau} \\
+\boldsymbol{ \tilde{x} }_{n-2\tau} \\
+\vdots \\
+\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
+\end{pmatrix}^\intercal, 
+\end{equation}
+%%********************************[EQUATION]************************************
+where $m$ is the embedding dimension, $\tau$ is the embedding delay and
+$ ^\intercal$ denotes the transpose.
+$m$ and $\tau$ are known as embedding parameters.
+The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
+$N-(m-1)\tau$ rows and $m$ columns and 
+$N-(m-1)\tau$ defines the length of each delayed copy 
+of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
+%For further details and explicit examples of uniform time-delay 
+%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
+
+\subsection*{Estimation of Embedding Parameters}
+The estimation of the embedding parameters ($m$ and $\tau$) 
+is a fundamental step for the state space reconstruction with the use
+of uniform time-delay embedding method.
+With this in mind, we review two of the most common algorithms,
+which will be used in our work, to compute the embedding
+parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%*******************************************************************************
+\subsubsection*{False Nearest Neighbours}
+To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
+used the method of false neighbours which can be understood as follows:
+on the one hand, when the embedding dimension is too small to unfold the 
+attractor "not all points that lie close each other will be neighbours and 
+some points appear as neighbours because of the attractor has been projected 
+down into an smaller space", on the other hand, when increasing the embedding 
+dimension "points that are near to each other in the sufficient
+embedding dimension should remain close as the dimension increase from $m$ 
+to $m+1$ \cite{krakovska2015}".
+From a mathematical point of view, the state space reconstruction theorem is 
+done when the attractor is unfolded with either the minimum embedding 
+dimension, $m_0$, or any other embedding dimension value where 
+$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
+leads to excessive computations \cite{bradley2015}. With this in mind, 
+Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
+where only the time-series and one delay embedding value are necessary 
+to select the minimum embedding dimension. Cao's algorithm is based 
+on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:e}
+  \begin{aligned}
+	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
+	 &=
+       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
+       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
+            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
+  \end{aligned}
+\end{equation}
+%%********************************[EQUATION]************************************
+where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
+embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
+From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
+and $\tau$ for which $E_1(m)$ is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:e1}
+E_1(m) = \frac{ E(m+1) } { E(m)}.
+\end{equation}
+%%********************************[EQUATION]************************************
+$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
+in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
+As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
+than some $m_0$, if the time series comes from a multidimensional state space
+then $m_0 + 1$ is the minimum dimension".
+Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
+stochastic signals. $E_2(m)$ is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:e2}
+E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
+\end{equation}
+%%********************************[EQUATION]************************************
+where
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:ee}
+E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
+|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
+\end{equation}
+%%********************************[EQUATION]************************************
+For instance, when the signal comes from random noise (values that are 
+independent from each other), all $E_2(m)$ values are approximately equal 
+to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
+not constant for all $m$ (e.g. $E_2(m) \neq 1$).
+
+As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
+series: the solution for the $x$ variable of the Lorenz system 
+(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
+and a variance of one (Fig~\ref{fig:e1e2}F).
+We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
+The $E_1(m)$ values for the chaotic time series appear to be constant
+after the dimension is equal to six.
+The determination of six is given that any value of $m$ can be used as they 
+are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
+$E_2(m)$ values, for chaotic time series, are different to one 
+(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
+chaotic time series the minimum embedding dimension the time series 
+comes from a deterministic signal. With regard to the noise time series,  
+$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
+which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
+$E_1(m)$ values then indicate the minimum embedding dimension of the 
+noisy time series is thirteen, however all of the $E_2(m)$ values are 
+approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
+concluded that noise time series is a stochastic signal.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
+    \caption{
+	{\bf Minimum dimension embedding values with Cao's method.} 
+	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
+	with variations of $\tau$ values from one to twenty
+	for (E) chaotic and (F) random time series.
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+        }
+    \label{fig:e1e2}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
+computed but also a variation of $\tau$ from 1 to 20 is presented. 
+The purpose of such variation for $\tau$ is to show its independence with
+regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
+(Fig~\ref{fig:e1e2}A,B,C, and D). 
+However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
+a new threshold where $m$ values appear to be constant in $E_1 (m)$.
+In the case of the given examples and reported results, we defined such 
+threshold as 0.05. Further investigation is required for the selection of the 
+threshold in the $E_1(m)$, as the selection of the threshold in this work is 
+base on no particular method but visual inspection.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%*******************************************************************************
+\subsubsection*{Average Mutual Information}
+When selecting the delay dimension parameter, $\tau$, one can consider the 
+following two cases:
+(i) when $\tau$ is too small, the elements of time-delay embedding will be along
+the bisectrix of the phase space and the reconstruction is generally not 
+satisfactory, 
+(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
+time-delay embedding will become spread and uncorrelated which makes 
+recovering the underlying attractor more difficult if not 
+impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
+With regard to the algorithms to compute $\tau$, 
+Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
+function in which the first zero crossing is considered as the minimum delay 
+embedding parameter. However, the autocorrelation function is a linear 
+statistic over which the Average Mutual Information (AMI) algorithm is 
+preferred because the AMI takes into account the nonlinear dynamical 
+correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
+algorithm is described below to estimate the minimum delay embedding 
+parameter, \texorpdfstring{$\tau_0$}{T}.
+
+To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
+and then a probability distribution of data is computed \cite{kantz2003}.
+AMI is therefore denoted by $I(\tau)$ which is the average mutual 
+information between the original time series, $x(n)$, and the delayed 
+time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
+AMI is defined by
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:ami}
+I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
+\end{equation}
+%%********************************[EQUATION]************************************
+Probabilities are defined as follows:
+$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
+the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
+the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
+that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
+in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
+For small $\tau$, AMI will be large and it will then decrease more or 
+less rapidly. As $\tau$ increase and goes to a large limit, 
+$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
+factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
+to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
+where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
+where the redundancy is the least" \cite{kantz2003}.
+
+For example, we compute the AMI for two time series:
+A) the $x$ solution of the Lorenz system, and
+B) a noise time series using a normal distribution with mean zero and 
+standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
+concluded that the amount of knowledge for any noise time series is zero 
+for which the first minimum embedding parameter is $\tau_0=1$. 
+On the contrary, the first minimum of the AMI for the chaotic time series 
+is $\tau_0=17$ which is the value that maximize the independence 
+between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
+\cite{bradley2015}.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
+    \caption{
+	{\bf Minimum delay embedding values with AMI's method.} 
+    	(A, B) AMI values where its first minimum value in the curve
+	is the minimum time delay embedding ($\tau_0$), 
+	for (C) a chaotic and (D) noise time series.
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+        }
+    \label{fig:amis}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+Similarly as Cao's algorithm negatives, AMI's algorithm is not an
+exception for negatives, which are worthwhile to mention for further 
+investigations.
+For instance, (i) is not clear why the choose of the first minimum of the AMI 
+is the minimum delay embedding parameter \cite{kantz2003} and 
+(ii) the probability distribution of the AMI function is computed
+with the use of histograms which depends on a heuristic choice of number of bins
+for which AMI depends on partitioning \cite{garcia2005e71}.
+
+%*******************************************************************************
+\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
+\subsection*{Recurrence Plots}
+Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
+in conservative systems, however such discovery was not put into practice 
+until the development of faster computers \cite{marwan2007},
+for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
+where recurrences in the dynamics of a system can be visualised using 
+Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
+propose a tool, called Recurrence Plot (RP), that provides insights into 
+high-dimensional dynamical systems where trajectories are very difficult to 
+visualise. Therefore, "RP helps us to investigate the 
+$m-$dimensional phase space trajectories through a two-dimensional 
+representation of its recurrences" \cite{marwan2015}.
+Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
+to the methodologies of the state space reconstruction and other dynamic 
+invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
+the recurrences of the trajectories in the phase space can provide 
+important clues to characterise natural process that present, for
+instance, periodicities (as Milankovitch cycles) or irregular cycles 
+(as El Ni\~no Southern Oscillation). 
+Such recurrences can not only be presented visually using Recurrence Plots (RP) 
+but also be quantified with Recurrence Quantification metrics, which leads 
+to applications of these in various areas such as economy, physiology, 
+neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
+
+For the creation of a recurrence plot based on time series 
+$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
+reconstruction with uniform time-delay embedding 
+$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
+\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
+where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
+and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
+The recurrence plot is therefore a two-dimensional $N \times N$ square 
+matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
+whenever $X(i)$ is sufficiently close to $X(j)$: 
+%%********************************[EQUATION]************************************
+\begin{equation}
+\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
+\end{equation}
+%%********************************[EQUATION]************************************
+where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
+$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
+function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
+(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
+RP is also characterised with a line of identity (LOI) which is a  
+diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
+%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
+    \caption{
+	{\bf Recurrence Plots.} 
+	(A) State space of the Lorenz system with controlling parameters 
+	($\rho=28, \sigma=10, \beta=8/3$). 
+	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
+	(black circle) of a given point at $i$ is a recurrent point and is 
+	represented as a black dot in the recurrence plot at location 
+	$(i, j)$ or white otherwise.
+	(B) Recurrence plot using the 
+	three components of the Lorenz system and the RP with no embeddings 
+	and threshold $\epsilon=5$.
+	This figure is adapted from \cite{marwan2015}.
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+	}
+    \label{fig:mrp}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\subsection*{Structures of Recurrence Plots}
+Pattern formations in the RPs can be designated either 
+as topology for large-scale patterns or texture for small-scale patterns.
+In the case of topology, the following pattern formations are commonly presented:
+(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
+uniformly distributed noise (Fig~\ref{fig:rp2}A), 
+(ii) periodic and quasi-periodic systems where diagonal lines and 
+checkerboard structures represent oscillating systems, e.g., sinusoidal 
+signals (Fig~\ref{fig:rp2}B), 
+(iii) drift where paling or darkening recurrence points away from 
+the LOI is caused by drifting systems, 
+e.g., logistic map (Fig~\ref{fig:rp2}C), and
+(iv) disrupted where recurrence points are presented white areas or 
+bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
+(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
+Texture patterns in RPs can be categorised as:
+(i) single or isolated recurrence points that represent rare occurring states, 
+and do not persist for any time or fluctuate heavily,
+(ii) dots forming diagonal lines where the length of the small-scale parallel 
+lines in the diagonal are related to the ratio of determinism or predictability 
+in the dynamics of the system, and
+(iii) dots forming vertical and horizontal lines where the length of the 
+lines represent a time length where a state does not change or change very 
+slowly and these patterns formation represent discontinuities in the signal, 
+and (iv) dots clustering to inscribe rectangular regions which are related 
+to laminar states or singularities \cite{marwan2015}.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
+    \caption{
+	{\bf Patterns in Recurrence Plots.} 
+	Time-series with its respective recurrence plot for:
+	(A) uniformly distributed noise,
+	(B) super-positionet harmonic oscillation 
+	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
+	with a linearly increase term ($0.01*i$), and
+	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	This figure is adapted from \cite{marwan2015}.
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+	}
+    \label{fig:rp2}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+Although, each of the previous pattern descriptions of the structures in the 
+RP offer an idea of the characteristics of dynamical systems, 
+these might be misinterpreted and conclusions might tend to be subjective 
+as these require the interpretation of a particular researcher(s).
+Because of that, recurrence quantification analyis (RQA) offer objective 
+methodologies to quantify such visual characteristics of previous 
+recurrent pattern structures in the RP \cite{zbilut1992}.
+
+%*******************************************************************************
+\subsection*{Recurrence Quantifications Analysis (RQA)}
+Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
+the density of recurrence points in RPs, then histograms of lengths for 
+diagonal lines in RPs were studied by \cite{trulla1996} which were the 
+introduction to the term recurrence quantification analysis (RQA) 
+\cite{marwan2008}. RQA has been applied in many fields such as life science, 
+engineering, physics, and others \cite{marwan2008}. Particularly in human 
+movement to investigate noise and complexity of postural control 
+\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
+coordination \cite{duran2017}. The success of RQA is not only due to its 
+simple algorithmic implementation but also to its capacity to detect tiny 
+modulations in frequency or phase which are not detectable using standard 
+methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
+RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
+dimension which is verified experimentally by \cite{iwanski1998}.
+RQA metrics comprehend percentage of recurrence, percentage of determinism, 
+ratio, Shannon entropy of the frequency distributions of the line lengths,
+maximal line length and divergence, trend and laminariy 
+\cite{marwan2007, marwan2015}. For this work, we considered only four 
+RQA metrics, due to its consistency with our preliminary experiments, 
+which are described below. Such metrics are computed the nonlinearTseries 
+R package \cite{nonlinearTseries2016}.
+
+\subsubsection*{REC values}
+The percentage of recurrence (REC) is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}
+REC(\epsilon,N) = 
+		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
+		\mathbf{R}^{m}_{i,j}(\epsilon),
+\end{equation}
+%%********************************[EQUATION]************************************
+which enumerates the black dots in the RP excluding the line of identity.
+REC is a measure of the relative density of recurrence points in the sparse 
+matrix \cite{marwan2015}.
+%REC is computed as follow with the nonlinearTseries package 
+% \cite{nonlinearTseries} 
+%  hist = getHistograms(neighs, ntakens, lmin, vmin)
+%  # calculate the number of recurrence points from the recurrence rate. The
+%  # recurrence rate counts the number of points at every distance in a concrete
+%  # side of the main diagonal.
+%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
+%  # add the maindiagonal
+%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
+%  # calculate the recurrence rate dividing the number of recurrent points at a
+%  # given distance by all points that could be at that distance
+%  recurrence_rate_vector = 
+%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
+%  # percentage of recurrent points
+%  REC = (numberRecurrencePoints) / ntakens ^ 2
+
+\subsubsection*{DET values} 
+The percent determinism (DET) is defined as the fraction of recurrence points
+that form diagonal lines and it is determined by
+%%********************************[EQUATION]************************************
+\begin{equation}
+DET=
+	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
+	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
+\end{equation}
+%%********************************[EQUATION]************************************
+where 
+%%********************************[EQUATION]************************************
+\begin{equation}
+H_D(l) = 
+	\sum^{N}_{i,j=1} 
+	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
+	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
+	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
+\end{equation}
+%%********************************[EQUATION]************************************
+is the histogram of the lengths of the diagonal structures in the RP.
+DET can be interpreted as the predictability of the system for periodic signals 
+which, in essence, have longer diagonal lines than the short diagonals lines
+for chaotic signals or absent diagonal lines for stochastic signals 
+\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
+the organisation of points in RPs  \cite{iwanski1998}. 
+%percent determinism (DET) is computed as follow with the nonlinearTseries 
+%package \cite{nonlinearTseries}  
+% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
+%                                       lmin = 2, lDiagonalHistogram,
+%                                       recurrence_rate_vector, maxDistanceMD) {
+%  #begin parameter computations
+%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
+%  DET = num / numberRecurrencePoints
+
+\subsubsection*{RATIO values}
+RATIO is defined as the ratio between DET and REC and it is calculated from 
+the frequency distributions of the lengths of the diagonal lines.
+RATIO is useful to discover dynamic transitions \cite{marwan2015}.
+%  diagP = calculateDiagonalParameters(
+%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
+%    recurrence_rate_vector, maxDistanceMD
+%  )
+% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
+%                                       lmin = 2, lDiagonalHistogram,
+%                                       recurrence_rate_vector, maxDistanceMD) {
+%  #begin parameter computations
+%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
+%  DET = num / numberRecurrencePoints
+% 
+%
+%    RATIO = diagP$DET / REC
+%
+
+\subsubsection*{ENT values}
+ENT is the Shannon entropy of the frequency distribution of the diagonal 
+line lengths and it is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}
+ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with 
+	\quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
+\end{equation}
+%%********************************[EQUATION]************************************
+ENT reflects the complexity of the deterministic structure in the system.
+For instance, for uncorrelated noise or oscillations, 
+the value of ENT is rather small and indicates low complexity of the system,
+therefore "the higher the ENT is the more complex the dynamics are" 
+\cite{marwan2007, marwan2015}.
+%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
+%
+%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
+%                                       lmin = 2, lDiagonalHistogram,
+%                                       recurrence_rate_vector, maxDistanceMD) {
+%
+%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
+%  diff_0 = which(pl > 0)
+%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
+
+\subsection*{Sensitivity and robustness of RPs and RQA.}
+RP and RQA are a very young field in nonlinear dynamics and many questions 
+are still open, for instance, different parameters for window length size 
+of the time series, embedding parameters or recurrence threshold can 
+generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
+
+The selection of recurrence threshold, $\epsilon$, can depend on the system 
+that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
+require to be very small, for trajectory reconstruction $\epsilon$ requires 
+to have a large thresholds or when studying dynamical transition 
+there is little importance about the selection of the threshold 
+\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
+the recurrence threshold  should be five times larger 
+than the standard deviation of the observational noise
+or the use of diagonal structures within the RP is suggested in order
+to find the optimal recurrence threshold for (quasi-)periodic process 
+\cite{marwan2011}.
+
+Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
+of choosing the right embedding parameters to perform RQA for which 
+many experiments have to be performed using different parameters in order 
+to have a better intuition of the nature of the time series and how 
+this is represented by using RQA.
+
+With that in mind, this work explores the sensitivity and robustness 
+of the window size of time series, embedding parameters for RSS with UTDE 
+and recurrence threshold for RP and RQA in order to gain a better insight 
+into the underlying time series collected from inertial sensors in the 
+context of human-humanoid imitation activities.
+
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\section*{Experiment} \label{sec:experiment}
+We conducted an experiment in the context of human-humanoid imitation (HHI) 
+activities where participants were asked to imitate simple horizontal and 
+vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
+Such simple movements were repeated ten times for the participant 
+who copied NAO's arm movements in a face-to-face imitation activity.
+Also, wearable inertial measurement unit (IMU) sensors were attached 
+to the right hand of the participant and to the left hand of the robot 
+(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
+IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
+accelerometer, gyroscope and magnetometer sensors and quaternions 
+\cite{Comotti2014}. 
+%A further description of the NeMEMSi IMU sensors 
+%is given in \nameref{S1_AppendixB}.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+  \centering
+\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
+    \caption{
+	{\bf Human-humanoid imitation activities.} 
+    		Face-to-face human-humanoid imitation (HHI) activities for 
+		(A) HHI of horizontal arm movement, 
+		(B) Humanoid horizontal arm movement,
+		(C) HHI of vertical arm movement, and 
+		(D) Humanoid vertical arm movement.
+        }
+    \label{fig:hri}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+\subsection*{Participants}
+Twenty-three participants,
+from now on defined as $pN$ where $N$ is the number of participant, were 
+invited to do the experiment. However, data for three participants were 
+not used because the instructions for $p01$, who was the only left-handed,
+were mistakenly given in a way that movements were performed different
+from what had been planned, and for participants $p13$ and $p16$ 
+data were corrupted because bluetooth communications problems with the 
+sensors. With that in mind, data for twenty participants were analysed in 
+this work.
+
+Of the twenty participants, all of them are right-handed healthy participants 
+of whom four are females and sixteen are males with a mean and standard 
+deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
+informed consent forms prior to participation in the experiment.
+
+\subsection*{Human-humanoid imitation activities}
+For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
+two of which were attached to the right hand of the participant and the 
+other two to the left hand of the humanoid robot.
+Then, each participant was asked to imitate repetitions of simple horizontal
+and vertical arm movements performed by the humanoid robot in the following 
+conditions:
+(i) ten repetitions of horizontal arm movement at normal (HN) 
+	and faster (HF) speed (Figure~\ref{fig:hri} A), and
+(ii) ten repetitions of vertical arm movement at normal (VN) 
+	and faster (VF) speed (Figure~\ref{fig:hri} C).
+The normal and faster speed of arm movements is defined by the duration 
+in number of samples of one repetition of NAO's arm movements. 
+We select NAO's arm movements duration to distinguish between normal and 
+faster arm movements as NAO's movements have less variation between 
+repetition to repetition. The duration for one repetition of the horizontal 
+arm movement at normal speed, HN, is about 5 seconds considering 
+that each repetition last around 250 samples.
+For horizontal arm movement at faster speed, HF, each repetition were performed 
+in around 2 seconds which correspond to 90 samples of data.
+The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
+which is around 300 samples of data. For vertical arm movement at 
+faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
+to 120 samples of data.
+To visualise the distinction between normal and faster speed for horizontal 
+and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
+for axes Z and Y of the gyroscope sensors with four window lengths: 
+2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
+and 15-sec (750-samples).
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht] 
+\centering
+\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
+    \caption{
+	{\bf Time series duration of horizontal and vertical arm movements.} 
+    		Time series of smoothed data from gyroscope sensor 
+		for different speed arm movements performed by NAO: 
+		(A) Horizontal Normal arm movement (HN), 
+		(B) Horizontal Faster arm movement (HF),
+		(C) Vertical Normal arm movement (VN) and 
+		(D) Vertical Faster arm movement (VF).
+		Additionally, 
+		(A) shows window sizes for 2-seconds (100 samples), 
+		5-seconds (250 samples), 10-seconds (500 samples) 
+		and 15-seconds (750 samples)
+		which are also presented in (B), (C) and (D).
+		Code and data to reproduce the figure is available from \cite{srep2019}.
+        }
+	\label{fig:sts}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+\subsection*{Data from Inertial Measurement Units} 
+	\label{sec:experiment:subsec:imu}
+\subsubsection*{Raw data}
+Considering the work of \cite{shoaib2016} which provided evidence of an 
+improvement in recognition activities when combining data from 
+accelerometer and gyroscope. We focus our analysis from data of the 
+accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
+leave the data of the magnetometer and quaternions for further investigation 
+because of their possible variations with regard to magnetic disturbances.
+
+Data from the accelerometer is defined by triaxial time series 
+$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
+(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
+which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
+representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
+Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
+its respective axes subscripts $x,y,z$, where $n$ is the sample index 
+and $N$ is the same maximum length of all axes for the time series.
+Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
+%%---------------------------------(EQUATION)----------------------------------
+\begin{equation}\label{eq:A}
+\boldsymbol{A} =
+\begin{pmatrix}
+  A_x(n) \\
+  A_y(n) \\
+  A_z(n)
+\end{pmatrix}
+=
+\begin{pmatrix}
+ a_x(1),a_x(2),\dots,a_x(N) \\
+ a_y(1),a_y(2),\dots,a_y(N) \\
+ a_z(1),a_z(2),\dots,a_z(N) 
+\end{pmatrix},
+\end{equation}
+%%---------------------------------(EQUATION)-----------------------------------
+and 
+%%---------------------------------(EQUATION)-----------------------------------
+\begin{equation}\label{eq:G}
+\boldsymbol{G} =
+\begin{pmatrix}
+ G_x(n) \\
+ G_y(n) \\
+ G_z(n)
+\end{pmatrix}
+=
+\begin{pmatrix}
+ g_x(1),g_x(2),\dots,g_x(N) \\
+ g_y(1),g_y(2),\dots,g_y(N) \\
+ g_z(1),g_z(2),\dots,g_z(N) 
+\end{pmatrix}.
+\end{equation}
+%%---------------------------------(EQUATION)-----------------------------------
+
+\subsubsection*{Postprocessing data}
+After the collection of raw data from four NeMEMsi sensors,
+time synchronisation alignment and interpolation were performed 
+in order to create time series with the same length and synchronised time.
+We refer the reader to \cite{Comotti2014} for further
+details about the time synchronisation process.
+
+\subsubsection*{Data normalization}
+Data is normalised to have zero mean and unit variance 
+using sample mean and sample standard deviation.
+The sample mean and sample standard deviation using $x(n)$ is given by
+%%---------------------------------(EQUATION)----------------------------------
+\begin{equation}\label{eq:ms}
+\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad 
+	\sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
+\end{equation}
+%%---------------------------------(EQUATION)----------------------------------
+and the normalised data, $\hat{x}(n)$, is computed as follows
+%%---------------------------------(EQUATION)----------------------------------
+\begin{equation}\label{eq:normalization}
+\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
+\end{equation}
+%%---------------------------------(EQUATION)----------------------------------
+
+\subsubsection*{Smoothing data}
+Commonly, a low-pass filter is the method either to capture the low 
+frequencies that represent \%99 of the human body energy or to get 
+the gravitational and body motion components of 
+accelerations \cite{anguita2013}. However, for this work the elimination of 
+certain range of frequencies is not the main focus but the conservation of 
+the structure in the time series in terms of the width and heights where, 
+for instance, Savitzky-Golay filter can help to accomplish 
+such task \cite{press1992}. Savitzky-Golay filter is based on the 
+principle of moving window averaging which preserves the area under 
+the curve (the zeroth moment) and its mean position in time 
+(the first moment) but the line width (the second moment) is violated 
+and that results, for example, in the case of spectrometric data
+where a narrow spectral line is presented with reduced height and width. 
+With that in mind, the aim of Savitzky-Golay filtering is to find filter 
+coefficients $c_n$ that preserve higher momentums which are based on local 
+least-square polynomial approximations \cite{savitzkygolay1964, 
+press1992, schafer2011}.
+Therefore, Savitzky-Golay coefficients are therefore computed using 
+an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
+\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
+$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
+Smoothed signal is represented with a tilde over the original 
+signal: $\tilde{x}(n)$.
+
+\subsubsection*{Window size data}
+With regard to the window size, \cite{shoaib2016} investigated its effects 
+using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
+and combination of inertial sensors (accelerometer, gyroscope and linear 
+acceleration sensor) to improve the activity recognition performance for 
+repetitive activities (walking, jogging and biking) and less repetitive 
+activities (smoking, eating, giving a talk or drinking a coffee).
+With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
+increase of window length improve the recognition of complex activities 
+because these requires a large window length to learn the repetitive 
+motion patterns. Particularly, one of the recommendations is to use large 
+window size to recognise less repetitive activities which mainly involve 
+random hand gestures. Therefore, for the four activities 
+(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
+we select only four window sizes for analysis: 2-s window (100 samples), 
+5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
+(Fig~\ref{fig:sts}).
 
@@
 \section*{Author contributions statement}
-Contribution for this work of 
-Miguel Xochicale (MX) and Chris Baber (CB) is as follows:
+Contributions for this work of 
+Miguel Xochicale (MX) and Chris Baber (CB) are as follows:
 \begin{description}

### 2019-03-22 code/rscripts/G_3Drqa/0_rqa_test.R

@@ -1,0 +1,598 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+##VERSION 
+version <- '00'
+#feature_path <- '/rqa'
+dataset_path <- '/data/dataset'
+
+## Outcomes Data Path
+#outcomes_data_path <- paste(github_path,'/hwum-dataset/data-output', feature_path, '/v', version, sep="")
+
+## Outcomes Plot Path
+#outcomes_plot_path <- paste(github_path,"/phd-thesis/figs/results", feature_path, '/v', version,sep="")
+
+## Data Path
+#data_path <- paste(github_path,'/hwum-dataset/data-output', feature_path, '/v', version, sep="")
+data_path <- paste(github_repo_path, dataset_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(signal)# for butterworth filter and sgolay
+library(ggplot2)
+library(RColorBrewer)
+
+library(devtools)
+load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
+
+
+################################################################################
+# (2) Reading data
+file_ext <- paste('xdata_v', version, '.dt',sep='')
+#file_ext <- paste('xdata_v01', '.dt',sep='')
+data <- fread( file_ext, header=TRUE)
+
+data <- data[,.(
+	#sg0zmuvGyroZ
+	sg0zmuvGyroZ, sg1zmuvGyroZ
+	#sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
+	), by=. (Participant,Activity,Sensor,Sample,Time)]
+
+
+
+
+
+################################################################################
[…]

### 2019-03-22 code/rscripts/G_3Drqa/B_3Drqa_plots_thesis.R

@@ -1,0 +1,367 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rqa/src/3drqa_thesis'
+data_path <- '/data/rqa'
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, data_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(plot3D)
+
+
+
+################################################################################
+# (2) Reading data
+#
+#file_ext <- paste('RQAs_p01w100.dt',sep='')
+#W100 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w250.dt',sep='')
+#W250 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w500.dt',sep='')
+#W500 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w750.dt',sep='')
+#W750 <- fread( file_ext, header=TRUE)
+#
+#W <- rbind(W100, W250, W500, W750)
+#
+
+#file_ext <- paste('RQAs_p02w500.dt',sep='')
+#W <- fread( file_ext, header=TRUE)
+#
+
+file_ext <- paste('RQAs_p03w500.dt',sep='')
+W <- fread( file_ext, header=TRUE)
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+## (x.x.x) RQA Metric Selection
+rqas <- c('ENTR')
+#rqas <- c('REC','DET', 'RATIO', 'ENTR')
+
+
+#########################################################
[…]

### 2019-03-22 code/rscripts/G_3Drqa/C_3Drqa_plots_epsilons.R

@@ -1,0 +1,389 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rqa/src/3drqa_epsilons'
+data_path <- '/data/rqa'
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, data_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(plot3D)
+
+
+
+################################################################################
+# (2) Reading data
+#
+#file_ext <- paste('RQAs_p01w100.dt',sep='')
+#W100 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w250.dt',sep='')
+#W250 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w500.dt',sep='')
+#W500 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w750.dt',sep='')
+#W750 <- fread( file_ext, header=TRUE)
+#
+#W <- rbind(W100, W250, W500, W750)
+#
+
+#file_ext <- paste('RQAs_p02w500.dt',sep='')
+#W <- fread( file_ext, header=TRUE)
+#
+
+file_ext <- paste('RQAs_p03w500.dt',sep='')
+W <- fread( file_ext, header=TRUE)
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+## (x.x.x) RQA Metric Selection
+rqas <- c('ENTR')
+#rqas <- c('REC','DET', 'RATIO', 'ENTR')
+
+
[…]

### 2019-03-22 code/rscripts/G_3Drqa/D_3Drqa_plots_sensors_activities.R

@@ -1,0 +1,388 @@
+###############################################################################	
+#
+# 
+#
+#
+#
+#
+#
+###############################################################################	
+	# OUTLINE:
+ 	# (0) Definifing paths 
+	# (1) Loading libraries and functions
+	# (2) Reading 
+
+
+#################
+# Start the clock!
+start.time <- Sys.time()
+
+
+
+
+################################################################################
+# (0) Defining paths for main_path, r_scripts_path, ..., etc.
+
+homepath <- Sys.getenv("HOME")
+main_home_path <- '/quetzalcoalt'
+r_scripts_path <- getwd()
+setwd("../../../")
+github_repo_path <- getwd()
+
+
+
+
+
+##VERSION 
+version <- '00'
+feature_path <- '/docs/figures/rqa/src/3drqa_sensors_activities'
+data_path <- '/data/rqa'
+
+## Outcomes Plot Path
+outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
+## Data Path
+data_path <- paste(github_repo_path, data_path, sep="")
+setwd(file.path(data_path))
+
+
+
+################################################################################
+# (1) Loading Functions and Libraries and Setting up digits
+library(data.table) # for manipulating data
+library(plot3D)
+
+
+
+################################################################################
+# (2) Reading data
+#
+#file_ext <- paste('RQAs_p01w100.dt',sep='')
+#W100 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w250.dt',sep='')
+#W250 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w500.dt',sep='')
+#W500 <- fread( file_ext, header=TRUE)
+#
+#file_ext <- paste('RQAs_p01w750.dt',sep='')
+#W750 <- fread( file_ext, header=TRUE)
+#
+#W <- rbind(W100, W250, W500, W750)
+#
+
+#file_ext <- paste('RQAs_p02w500.dt',sep='')
+#W <- fread( file_ext, header=TRUE)
+#
+
+file_ext <- paste('RQAs_p03w500.dt',sep='')
+W <- fread( file_ext, header=TRUE)
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+## (x.x.x) RQA Metric Selection
+rqas <- c('ENTR')
+#rqas <- c('REC','DET', 'RATIO', 'ENTR')
+
+
[…]

### 2019-03-22 code/rscripts/G_3Drqa/F0_rqa_test.R

@@ -1,598 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-
-
-homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-
-
-##VERSION 
-version <- '00'
-#feature_path <- '/rqa'
-dataset_path <- '/data/dataset'
-
-## Outcomes Data Path
-#outcomes_data_path <- paste(github_path,'/hwum-dataset/data-output', feature_path, '/v', version, sep="")
-
-## Outcomes Plot Path
-#outcomes_plot_path <- paste(github_path,"/phd-thesis/figs/results", feature_path, '/v', version,sep="")
-
-## Data Path
-#data_path <- paste(github_path,'/hwum-dataset/data-output', feature_path, '/v', version, sep="")
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(ggplot2)
-library(RColorBrewer)
-
-library(devtools)
-load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
-source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
-
-
-################################################################################
-# (2) Reading data
-file_ext <- paste('xdata_v', version, '.dt',sep='')
-#file_ext <- paste('xdata_v01', '.dt',sep='')
-data <- fread( file_ext, header=TRUE)
-
-data <- data[,.(
-	#sg0zmuvGyroZ
-	sg0zmuvGyroZ, sg1zmuvGyroZ
-	#sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
-	), by=. (Participant,Activity,Sensor,Sample,Time)]
-
-
-
-
-
-################################################################################
[…]

### 2019-03-22 code/rscripts/G_3Drqa/Fa_3Drqa_datatable.R

@@ -1,547 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/data/rqa'
-dataset_path <- '/data/dataset'
-
-## Outcomes Data Path
-outcomes_data_path <- paste(github_repo_path, feature_path,  sep="")
-### Outcomes Plot Path
-#outcomes_plot_path <- paste(github_repo_path,'/hwum-manuscript/figures/results', feature_path, '/v', version, sep="")
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(nonlinearTseries)
-source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
-
-library(ggplot2)
-library(RColorBrewer)
-
-
-################################################################################
-# (2) Reading data
-file_ext <- paste('xdata_v', version, '.dt',sep='')
-data <- fread( file_ext, header=TRUE)
-
-
-
-data <- data[,.(
-	#sg0zmuvGyroZ
-	#sg0zmuvGyroZ, sg1zmuvGyroZ
-	#sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
-	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY, sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
-	), by=. (Participant,Activity,Sensor,Sample,Time)]
-
-
-
-
-
-################################################################################
-################################################################################
-################################################################################
[…]

### 2019-03-22 code/rscripts/G_3Drqa/Fb_3Drqa_plots.R

@@ -1,358 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-
-homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-
-
-
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/docs/figures/rqa/src/3drqa'
-data_path <- '/data/rqa'
-#feature_plot_path <- '/rqa_plots/topologies'
-
-## Outcomes Data Path
-#outcomes_data_path <- paste(github_repo_path,tput', feature_path, '/v', version, sep="")
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
-## Data Path
-data_path <- paste(github_repo_path, data_path, sep="")
-setwd(file.path(data_path))
-
-
-#rqa_path <- paste(outcomes_data_path, '/rqa', sep="")
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(plot3D)
-
-
-
-################################################################################
-# (2) Reading data
-#
-#file_ext <- paste('RQAs_p01w100.dt',sep='')
-#W100 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w250.dt',sep='')
-#W250 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w500.dt',sep='')
-#W500 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w750.dt',sep='')
-#W750 <- fread( file_ext, header=TRUE)
-#
-#W <- rbind(W100, W250, W500, W750)
-#
-
-#file_ext <- paste('RQAs_p02w500.dt',sep='')
-#W <- fread( file_ext, header=TRUE)
-#
-
-file_ext <- paste('RQAs_p03w500.dt',sep='')
-W <- fread( file_ext, header=TRUE)
-
-
-
-################################################################################
-################################################################################
-################################################################################
[…]

### 2019-03-22 docs/figures/low-res-pdf.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+#usage
+# low-pdf.sh filename QualityNumber(50/70/100) Density(100)
+
+#inputfile = $1
+#quality $2
+#density 150x150
+
+#path=$pwd
+#echo $path
+
+#inkscape --export-pdf $1.pdf $1.svg
+inkscape vector/$1.svg --export-pdf pdf/$1.pdf 
+
+convert -density $3x$3 -quality $2 -compress jpeg  pdf/$1.pdf pdf/output.pdf
+mv pdf/output.pdf pdf/$1.pdf
+
+

### 2019-04-29 docs/corrections/29-03-2019/logbook/logbook.tex
[docs/corrections/29-03-2019] part-01

\title{Logbook for corrections }
\author{Miguel Xochicale}
\date{ \today }

\documentclass[12pt]{article}
\usepackage[margin=0.99in]{geometry}
\usepackage{enumitem}

% for asterik in circle
\usepackage{amsmath,amssymb,graphicx,color}
\newcommand{\invcircledast}{%
  \mathbin{\vphantom{\circledast}\text{%
    \ooalign{\smash{\blackcircle}\cr
             \hidewidth\smash{\textcolor{white}{$*$}}\hidewidth\cr
            }%
  }}%
}

\begin{document}
\maketitle

\begin{abstract}
This document presents a log-book for the corrections
of \emph{.../corrections/29-04-2019/comments/scanned-draft.pdf}.
\end{abstract}

%\tableofcontents
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Title and authors}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\title{The Application of Recurrent Quantification Analysis 
	to Human-Humanoid Interaction}

	Chris Baber

	\end{verbatim}
	\textit{
	SORTED:  Mon Apr  1 17:26:42 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abstract}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$  Not a conclusion just an observation

	\begin{verbatim}

Then with the use of RQA metrics, we find out that the type of 
movements and the level of smoothness affects those metrics. 
Specifically, entropy values from RQA were well distributed and 
presented variation in all the conditions for time series. 
This work have then the potential to enhance the development of 
better diagnostic tools for various applications in rehabilitation, 
sport science or for new forms of human-robot interaction.

	\end{verbatim}
	\textit{
	SORTED:  Mon Apr  1 17:44:55 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	added recommendations 	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:08:38 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
	recommendations were added	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:40:38 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}
 	added recommendations	
	\end{verbatim}
	\textit{
	SORTED:  Thu Apr  4 12:46:47 BST 2019
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

\begin{enumerate}

\item  (pp. 1)  \_$\wedge$

	\begin{verbatim}

	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

\end{enumerate}

\end{document}

### 2019-04-29 docs/manuscript/main.tex
[docs/corrections/29-03-2019] part-01

@@
 
-\title{Strengths and Weaknesses of Recurrent Quantification Analysis 
-in the context of Human-Humanoid Interaction}
+\title{The Application of Recurrent Quantification Analysis 
+to Human-Humanoid Interaction}
 \author[1,*]{Miguel Xochicale}
-\author[1]{Chirs Baber}
+\author[1]{Chris Baber}
 \affil[1]{University of Birmingham, 
@@
 we found visual differences in the patterns of RSSs and RPs. 
-Then using metrics of RQA, we find out that the type of movements
-and the level of smoothness affects those metrics. 
-Our contribution for this work is that entropy values from RQA were 
-well distributed and presenting variation in all the conditions for time series. 
+Then with the use of RQA metrics, we find out that the type of 
+movements and the level of smoothness affects those metrics. 
+Specifically, entropy values from RQA were well distributed and 
+presented variation in all the conditions for time series. 
 This work have then the potential to enhance the development of 
-better diagnostic tools for various applications rehabilitation, 
+better diagnostic tools for various applications in rehabilitation, 
 sport science or for new forms of human-robot interaction.
@@
 \section*{Introduction}
-Human movement variability is a complex system where not only multiple
+Human movement requires a complex system where not only multiple
 joints and limbs are involved for a specific task in a determined environment
-but also how we process the external information with all of our available senses 
-and use our experiences \cite{davids2003}. 
-Recent studies in human motion recognition have revealed the possibility to 
-estimate features from lower dimensions signals to distinguish differences between 
+but also perception and action of movement affects the physical performance 
+\cite{davids2003}. 
+In contrast, variability in humanoid movement is usually very small,
+as a result of the control system used. This means that, while humanoids 
+solve the degrees of freedom problem through join design and algorithms,
+humans tend to have more fluid and flexible approach. Consequently, 
+one can see much variability in the performance of even the simplest task.
+Studies of human motion reveal the possibility to 
+estimate features from lower dimension signals to distinguish differences between 
 styles of pedalling motion \cite{Quintana-Duque2012, Quintana-Duque2016}, 
-gait identification \cite{sama2013, frank2010} 
-or pattern recognition from biological signals \cite{gomezgarcia2014}.
-The lower dimension signals from biological signals are time series 
+gait identification \cite{sama2013, frank2010},
+and detection of pathologies \cite{gomezgarcia2014}.
+The lower dimension signals from biological systems are time series 
 of $1-$dimension in $\mathbb{R}$ which commonly are 
 noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
-Hence, we are hypothesising that nonlinear dynamics can be used to 
-objectively quantify such human movement variability of lower
+Hence, nonlinear dynamics can be used to 
+objectively quantify variability of lower
 dimension signals \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
@@
 state space \cite{takens1981}.
-Also, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
+Bradley et al. 2015 \cite{bradley2015} also reviewed the use of 
 Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
 which show black and white dots as recurrences in a given $n$-dimensional system, 
-and Recurrence Quantification Analysis (RQA) metrics can pick out important 
-directions and statistics in RPs.
-In general, RPs and RQA provide a more intuitive meaning of the time series,
+and Recurrence Quantification Analysis (RQA) metrics compute statistics in RPs.
+In general, RPs and RQA provide an intuitive meaning of the time series,
 for instance, RQA is quantitatively and qualitatively independent of 
@@
 there is no general technique that can be used to compute the embedding
-parameters since the time series are system-dependent which means that
-once computing the values for embedding parameters, these may only work for
+parameters since time series are system-dependent which means that
+computing the values for embedding parameters may only work for
 one purpose (e.g., prediction) and may not work well for another purpose
@@
 or require purely deterministic signals \cite{kantz2003}.
-Similarly, these nonlinear analysis methods for computing the embedding 
+Thus, these nonlinear analysis methods for computing the embedding 
 parameters can break down
@@
 \section*{Results}
-To give insight to the research questions, 
-we considered various conditions of time series collected for this work 
-(see Experiment section for more details) which are described as follows
-\begin{itemize}
-\item Three levels of smoothness for the normalised data 
-	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
-	lengths (29 and 159) with the same polynomial degree 
-	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
-\item four velocities arm movement activities: 
-	horizontal normal (HN), horizontal faster (HF), 
-	vertical normal (VN), and vertical faster (VF), and
-\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
-	10-sec (500 samples) and 15-sec (750 samples).
-\end{itemize}
-
-%\subsection*{Time series}
-After the data collection, raw time series were windowed, normalised and 
-smoothed. Then, due to space limitations and to have simple visualisation, 
-we only present 10-sec (500 samples) window length time series for 
-three participants (p01, p01 and p03) performing horizontal 
-arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
-    	\caption{ 
-	{\bf Time series for horizontal and vertical arm movements.}
-		(A/D) raw-normalised (sg0zmuv), 
-		(B/E) normalised-smoothed 1 (sg1zmuv) and
-		(C/F) normalised-smoothed 2 (sg2zmuv).
-		Time series are only for three participants 
-		($p01$, $p02$, and $p03$) 
-		for horizontal and vertical arm movements in normal 
-		and faster velocity (HN, HF, VN, VF) 
-		with the normalised GyroZ or GyroY axis 
-		and with one sensor attached to the participant (HS01) 
-		and other sensor attached to the robot (RS01).	
-	Code and data to reproduce the figure is available from \cite{srep2019}.
-        }
-    \label{fig:ts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-%\subsection*{Reconstructed State Spaces using Uniform Time-Delay Embedding}
+
 \subsection*{Reconstructed State Spaces}
-One of the main challenges of the implementation of 
+One of the challenges of the implementation of 
 Uniform Time-Delay Embedding to Reconstructed State Spaces is the 
@@
 Figures \ref{fig:caoami}(A) show that minimum embedding values 
-for sensor HS01 appear to show more variations as their 
-interquartile range of the box plots are near to 1 
+for sensor HS01 appear to show more variations as 
+interquartile range is near to 1 
 with three exceptions (HF sg0, VN sg0, and VF sg1).
 Minimum embedding values appear to be constant for sensor RS01 
-as their interquartile range 
-in the box plots are near to 0.1 with the exception of 
+as interquartile range is near to 0.1 with the exception of 
 two axis (HF sg0 and VF sg1).
@@
 However, AMI values for RS01 do not show such increase in relation with
-the increase of smoothness excepting for HF and VF
+the increase of smoothness except for HF and VF
 (Figs \ref{fig:caoami}(B)).
@@
 
-Evidently, it is easy to observe by eye the differences in each of the
+It is easy to observe by eye the differences in each of the
 trajectories in the reconstructed state spaces 
@@
 \subsection*{Other effects on 3D RQA ENTR values}
-% metrics with different embedding parameters, 
-%	recurrence thresholds, window lengths, levels of smoothness, 
-%	and time series structures.}
 
@@
 
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/rqas}
-%%\includegraphics[width=1.0\textwidth]{rqas}
-%    \caption{
-%	{\bf 3D surfaces for RQA metrics.}
-%	3D surfaces for REC, DET, RATIO and ENTR values with increasing 
-%	pair embedding parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) 
-%	and recurrence thresholds (  $ 0.2 \ge \epsilon \le 3 $).
-%	RQA metrics values for time series of participant p01 using HS01 sensor, 
-%	HN activity and sg0zmuvGyroZ axis and 500 samples window length.
-%        R code to reproduce the figure is available from \cite{hwum2018}.
-%	}
-%\label{fig:topo_rqas}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
-
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/w}
-%%\includegraphics[width=1.0\textwidth]{w}
-%    \caption{
-%	{\bf 3D surfaces for RQAs metrics with four window lengths.}
-%	3D surfaces of RQAs metric values with increasing embedding 
-%	parameters and recurrence thresholds are for four window lengths 
-%	(w100, w250, w500 and  w750).
-%	RQA metrics values are for time series of participant p01 
-%	using HS01 sensor, HN activity and sg0zmuvGyroZ axis.
-%	R code to reproduce the figure is available from \cite{hwum2018}.
-%        }
-%\label{fig:topo_sensoractivities}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/s}
-%%\includegraphics[width=1.0\textwidth]{s}
-%    \caption{
-%	{\bf 3D surfaces for RQA metrics with three levels of smoothness.}
-%       	3D surfaces of RQAs metric values with increasing embedding 
-%	parameters and recurrence thresholds are for three levels of 
-%	smoothness (sg0zmuvGyroZ, sg1zmuvGyroZ and sg1zmuvGyroZ).
-%	RQA metrics values are for time series of participant p01 using 
-%	HS01 sensor, HN activity and 500 samples window length.
-%	R code to reproduce the figure is available from \cite{hwum2018}.
-% }
-%\label{fig:topo_smoothness}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
-%
-%
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/results/rqa/v03/rqa_plots/topologies/figs/p}
-%%\includegraphics[width=1.0\textwidth]{p}
-%    \caption{
-%	{\bf 3D surfaces for RQA metrics with three participants.}
-%       	3D surfaces of RQAs metric values for participants p01, p02 and p03 
-%	with increasing embedding parameters and recurrence thresholds.
-%	RQA metrics values are for time series of HS01 sensor, 
-%	HN activity and 500 samples window length.
-%	R code to reproduce the figure is available from \cite{hwum2018}.
-% }
-%\label{fig:topo_participants}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
-%
 %*******************************************************************************
@@
 
-It is evidently that time series from different sources 
+It is evident that time series from different sources 
 (participants, movements, axis type, window length or levels of smoothness) 
@@
 were not well unfolded. 
-With that in mind, we proceed to apply 
+With that in mind, we apply 
 Recurrence Quantification Analysis (RQA) metrics 
@@
 	\label{sec:experiment:subsec:imu}
+
+To give insight to the research questions, 
+we considered various conditions of time series collected for this work 
+(see Experiment section for more details) which are described as follows
+\begin{itemize}
+\item Three levels of smoothness for the normalised data 
+	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
+	lengths (29 and 159) with the same polynomial degree 
+	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
+\item four velocities arm movement activities: 
+	horizontal normal (HN), horizontal faster (HF), 
+	vertical normal (VN), and vertical faster (VF), and
+\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
+	10-sec (500 samples) and 15-sec (750 samples).
+\end{itemize}
+%\subsection*{Time series}
+After the data collection, raw time series were windowed, normalised and 
+smoothed. Then, due to space limitations and to have simple visualisation, 
+we only present 10-sec (500 samples) window length time series for 
+three participants (p01, p01 and p03) performing horizontal 
+arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
+    	\caption{ 
+	{\bf Time series for horizontal and vertical arm movements.}
+		(A/D) raw-normalised (sg0zmuv), 
+		(B/E) normalised-smoothed 1 (sg1zmuv) and
+		(C/F) normalised-smoothed 2 (sg2zmuv).
+		Time series are only for three participants 
+		($p01$, $p02$, and $p03$) 
+		for horizontal and vertical arm movements in normal 
+		and faster velocity (HN, HF, VN, VF) 
+		with the normalised GyroZ or GyroY axis 
+		and with one sensor attached to the participant (HS01) 
+		and other sensor attached to the robot (RS01).	
+	Code and data to reproduce the figure is available from \cite{srep2019}.
+        }
+    \label{fig:ts}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
 \subsubsection*{Raw data}

### 2019-09-05 docs/corrections/29-03-2019/logbook/logbook.tex
completed corrections29-03-2019

@@
-\title{Logbook for corrections }
+\title{
+	Logbook of corrections \\ 
+	for manuscript-draft-29-04-2019 
+}
 \author{Miguel Xochicale}
@@
 \begin{abstract}
-This document presents a log-book for the corrections
+This is a log book document for the corrections
 of \emph{.../corrections/29-04-2019/comments/scanned-draft.pdf}.
@@
 	recommendations were added
+	(i.e. types of signals were moved to methods section)	
 	\end{verbatim}
@@
 	\begin{verbatim}
+	Recommendations with the use of English language were added	
+	\end{verbatim}
+	\textit{
+	SORTED:  Tue  3 Sep 07:15:23 BST 2019
+	}
+	\\
 
+	\begin{verbatim}
+	You need a sentence or two here
+	to explain why this is claimed.
 	\end{verbatim}
@@
 	\begin{verbatim}
-
+	Corrections with the use of English	
 	\end{verbatim}
 	\textit{
-	SORTED:  
+	SORTED:  Tue  3 Sep 07:26:41 BST 2019
+	}
+	\\
+
+\end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiment}
+
+\begin{enumerate}
+
+\item  (pp. )  \_$\wedge$
+
+	\begin{verbatim}
+	Needs a statment of ethics 
+	Section of Ethics has been added. However I am not sure 
+	if an Appendix should be added for further informotation.
+	\end{verbatim}
+	\textit{
+	SORTED:  Thu  5 Sep 07:14:33 BST 2019
 	}

### 2019-09-05 docs/manuscript/main.tex
completed corrections29-03-2019

@@
 smoothness of time series.
-However, using determined RQA metrics will be depend on 
-what one want to quantify, for instance, one can compute 
+Choice of RQA metrics will be depend on 
+what to quantify, for instance, one can compute 
 predicability, organisation,  
@@
 regard to activity type, window length and structure of the time series.
-%Similarly, such differences in time series created differences in each of the 
-%RQA metrics, 
 For instance, RATIO and ENTR are helpful to distinguish 
 differences in any of the categories of the time series (sensor, activity, 
-level of smoothness and number of participant), however for certain time series
-such as time series from the sensor attached to the robot seemed to have 
-little variations between each of the participants. 
-The latter phenomena was in a way evidently as robot degrees of freedom 
+level of smoothness and number of participant), however time series
+data from the sensor attached to the robot seemed to have 
+little variations between as humanoid degrees of freedom 
 did not allow it to move with a wide range of variability.
 
-Additionally, we can point out that even thought our experiment is 
+Additionally, we can point out that even though our experiment is 
 limited to twenty healthy right-handed participants of a range age 
@@
 human-humanoid imitation activities.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%	You need a sentence or two here
+%	to explain why this is claimed.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection*{Inertial Sensors}
-To have more fundamental understating of nature of signals collected through 
-inertial sensors in the context of human-robot interaction, we are considering 
-to apply derivates to the acceleration data. We can then explore the jerkiness 
-of movements and therefore the nature of arm movements which typically have 
+To better understand the signals collected through 
+inertial sensors in the context of human-humanoid interaction, 
+we could apply derivates to the acceleration data. 
+We can then explore the jerkiness of movements and therefore 
+the nature of arm movements which typically have 
 minimum jerk \cite{flash1985}, its relationship with different body 
@@
 Having shown that RQA ENTR metrics are robust against different data
-time series, we believe that further investigation is required to be done. 
+time series, we believe that further investigation is required. 
 For example, Marwan et al. \cite{marwan2007, marwan2015} reviewed 
@@
 \subsection*{State Space Reconstruction}
-The method of state space reconstruction was originally proposed by 
-\cite{packard1980} and formalised by \cite{takens1981}. Since then, different 
-investigations and disciplines have benefited from the use of the method of 
-state space reconstruction \cite{aguirre2009, stergiou2011, frank2010, 
+The method of state space reconstruction \cite{packard1980} 
+\cite{takens1981} has been applied in many disciplines 
+\cite{aguirre2009, stergiou2011, frank2010, 
 sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
@@
 \cite{Comotti2014}.
+
+\subsection*{Ethics}
+The experiments of this work were conducted in November 2016 and participants
+confirmed reading and understanding the participant information sheet of the 
+experiments and were able to withdraw from the experiment at any time without giving
+any reason. The design of the experiments adhered to the University of X
+regulations, data were anonymised and videos were stored only on a 
+personal computer in accordance with the Data Protection Act 1998. 
+%Refer to Appendix D for further information about the ethics, 
+%online participation information sheets and experiment check list.
+
 %A further description of the NeMEMSi IMU sensors 

### 2020-07-02 docs/corrections/2020-01-14/report/report.tex
completion of corrections-2020-01-14 and updates manuscript2020-07-03

\documentclass[10pt]{article}
\usepackage[margin=15mm]{geometry}
\usepackage{enumitem}
\usepackage{hyperref}

\title{
Logbook of corrections 2020-01-14 
}
\author{Miguel Xochicale}
\date{ \today }

\begin{document}
\maketitle

\begin{abstract}
This logbook contains comments made by Chris Baber 
(\emph{.../corrections/2020-01-14/comments/README.md})
and their responses. 
However, on 4th of February 2020, I realised 
that before tackling such comments (in Section I),
I will be reviewing the whole manuscript and report 
any amends (this start in section II).
\end{abstract}

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Chris Baber}
\subsection{Summarise results}

\begin{enumerate}

\item

At the end of the results section (before the Discussion) you should include a paragraph that summarises the main results
	\begin{verbatim}

Tue 30 Jun 00:21:09 BST 2020
To summarise this section of results, it has been found that computing individual embedding parameters 
for a particular structure of time series is an already solved problem
\cite{frank2010, sama2013, bradley2015}.
However, the challenge is to find embedding parameters that represent a set of different time series,
in order to compute uniform time-delay embedding, recurrence plots and recurrence quantification analysis.
That said, it has been proposed the use of sample mean of the set of embedding parameters
to then noticed that the selection of recurrence threshold, $\epsilon$, is also an open problem.
Hence, variation of recurrence thresholds and embedding parameters 
has been presented to show the relationships to different datasets 
(participants, activities, windows lengths and sensors).

Thu  2 Jul 20:57:04 BST 2020
To summarise this section of results, it can be said that computing
embedding parameters for individual structure of time-series 
data is already a solved problem \cite{frank2010, sama2013, bradley2015}. 
However, it has been shown the challenge of finding embedding parameters 
for nonlinear dynamic tools that represent a set of different time-series data.
That said, we proposed the use of sample mean of the set of embedding parameters
for RSSs, RP and RQA to then noticed that the selection of recurrence 
threshold, $\epsilon$, is also an open problem.
For which, this work proposed the variation of recurrence thresholds 
and embedding parameters to show the relationships of these to different datasets 
(participants, activities, windows lengths and sensors).

	\end{verbatim}
	\textit{
	SORTED: \\ 
Tue 30 Jun 00:21:15 BST 2020 \\
Thu  2 Jul 20:57:15 BST 2020\\
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comments for the conclusion}

\begin{enumerate}

\item

It is still not clear what conclusions you can draw about the 
performance of the participants (in the experiment) or 
comparing human with humanoid. While I can see that the point 
of the paper is on the discussion of the different methods, 
and their usefulness, 
I still think it would be good 
to make observations on variability in the actual 
movement (what is the difference between human and humanoid?, 
do some people show more variability than others?) 
and then ask whether (as well as being sensitive to the data) 
some of the methods are more sensitive to variability 
in performance.
	\begin{verbatim}

Thu  2 Jul 22:41:43 BST 2020
\section*{Conclusions}
This work allows to conclude that the choice of 
nonlinear analysis tool (e.g., RSSs, RPs, RQA metrics) will depend 
on what one would like to quantify on the time-series 
data (e.g., predicability, organisation, dynamics transitions, 
or complexity and determinism).
Then, time-series data characteristics 
(e.g., window size length, level of smoothness) plays an important role
as well on the results that nonlinear analysis tool can provide.
Similarly, the results of the nonlinear analysis tools
are associated with the structure of the time-series data 
(e.g. frequency, amplitude), the position of the sensor 
and activity performed by either a robot or human being
(as degrees of freedom from the humanoid are far less than
human movement). 
That said, it has been shown that the use of different 
characteristics of the time-series data 
(e.g., sensor, activity, level of smoothness and participants)
has help us to visualise and to quantify 
with nonlinear tools the variation of 
movements of, in this work, human-humanoid activities.
However, some limitation of nonlinear tools are related 
to the computation of different parameters 
(e.g., recurrence thresholds, embedding parameters)
that reflect the dynamics of individual characteristics 
of activity type, window length and structure of the time series.
Specifically, the example of DET values which appear 
to be constants across sensors, activities and levels of smoothness, 
whereas REC and RATIO, as function of REC, values 
show variation for certain sensors and movements. 
To then found out that RQA ENTR values with different recurrence 
thresholds were appropriate to quantify 
the different changes and variations of the characteristics of 
time-series data.
Therefore, we can conclude that this work provides a good starting 
point and reference to the use of Shannon Entropy to quantify 
human-humanoid imitation activities 
that can then lead to interesting results on the quantification 
of movement variability of participants with different ages, 
state of health and anthropomorphic features.

	\end{verbatim}
	\textit{
	LOG: \\
	Thu  2 Jul 22:41:43 BST 2020 \\
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Comments and question 2}
%
%
%\begin{enumerate}
%
%\item 
%
%It is still not clear what conclusions you can draw about the 
%performance of the participants (in the experiment) or 
%comparing human with humanoid. While I can see that the point 
%of the paper is on the discussion of the different methods, 
%and their usefulness, 
%I still think it would be good 
%to make observations on variability in the actual 
%movement (what is the difference between human and humanoid?, 
%\textbf{do some people show more variability than others?)}
%and then ask whether (as well as being sensitive to the data) 
%some of the methods are more sensitive to variability 
%in performance.
%
%	\begin{verbatim}
%
%	
%	\end{verbatim}
%	\textit{
%	SORTED: 
%	}
%	\\
%
%
%\end{enumerate}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Comments and question 3}
%
%
%\begin{enumerate}
%
%\item 
%
%It is still not clear what conclusions you can draw about the 
%performance of the participants (in the experiment) or 
%comparing human with humanoid. While I can see that the point 
%of the paper is on the discussion of the different methods, 
%and their usefulness, 
%I still think it would be good 
%to make observations on variability in the actual 
%movement (what is the difference between human and humanoid?, 
%do some people show more variability than others?) 
%and 
%\textbf{
%then ask whether (as well as being sensitive to the data) 
%some of the methods are more sensitive to variability 
%in performance.}
%
%	\begin{verbatim}
%
%	
%	\end{verbatim}
%	\textit{
%	SORTED: 
%	}
%	\\
%
%
%\end{enumerate}
%
%

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Miguel Xochicale}
\subsection{Introduction}
\begin{enumerate}

\item  (pp. ) 
	Human movement requires a complex system where not only multiple
joints and limbs are involved for a specific task in a determined environment
but also perception and action of movement affects the physical performance (davids2003)

	\begin{verbatim}
Human movement requires a complex system where not only multiple
joints and limbs are involved for a specific task in a determined environment
but also perception and action of movement that affects such 
physical performance 
\cite{davids2003}.

	\end{verbatim}
	\textit{
	ADDED: Mon 10 Feb 07:08:01 GMT 2020
	}
	\\

\item  (pp. ) 
In contrast, variability in humanoid movement is usually very small,
as a result of its control system cite{gouaillier2009}.

	\begin{verbatim}
In contrast, variability in humanoid movement is usually very small,
as a result of mechanical and dynamic capabilities \cite{gouaillier2009}. 
\end{verbatim}
	\textit{
	ADDED: Mon 10 Feb 07:25:54 GMT 2020
	}
	\\

\item  (pp. ) 
humans tend to have more fluid and flexible approach. Consequently, 
one can see much variability in performance of 
even the simplest task.

\begin{verbatim}
humans tend to have more fluid and flexible approach. Consequently, 
one can see much variability in human movement performance of 
even the simplest task.
\end{verbatim}
	\textit{
	ADDED: Mon 10 Feb 07:30:31 GMT 2020
	}
	\\

\item  (pp. ) 
nonlinear time series analysis based on the appropriate estimation of the 
\begin{verbatim}
nonlinear time series analysis based on the estimation of the 
\end{verbatim}
	\textit{
	ADDED: Mon 10 Feb 16:32:51 GMT 2020
	}
	\\

\item  (pp. ) 
Bradley et al. 2015 cite{bradley2015} also reviewed the use of 
Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
which show black and white dots as recurrences in a given $n$-dimensional system, 
and Recurrence Quantification Analysis (RQA) metrics compute statistics in RPs.

\begin{verbatim}

Similarly, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
which show black and white dots as recurrences in a given $n$-dimensional system, 
and Recurrence Quantification Analysis (RQA) metrics that compute 
statistics in RPs.

\end{verbatim}
	\textit{
	ADDED: Mon 10 Feb 16:42:27 GMT 2020
	}
	\\

\item  (pp. ) 
Additionally, the methodologies of nonlinear dynamics for 
computing the embedding parameters e.g., autocorrelation, mutual information, 
and nearest neighbour require data which is well 
sampled and with little noise \ cite{garland2016}
or require purely deterministic signals \ cite{kantz2003}.

\begin{verbatim}
Additionally, methods of nonlinear dynamics for 
computing embedding parameters e.g., autocorrelation, mutual information, 
and nearest neighbour require data which is well 
sampled and with little noise \cite{garland2016}
or require signals that are purely deterministic \cite{kantz2003}.

\end{verbatim}
	\textit{
	AMENDED: Mon 10 Feb 16:55:30 GMT 2020
	}
	\\

\item  (pp. )

Thus, these nonlinear analysis methods for computing the embedding 
parameters can break down
with real-world datasets which have generally different length, 
different values of accuracy and precision (rounding errors due to finite 
precision of the measurement apparatus which include frequency 
acquisition \ cite{frank2010}),
and data may be contaminated with different sources of noise
\ cite{garland2016}.

\begin{verbatim}
Thus, these methods of nonlinear analysis for computing the embedding 
parameters can break down
with real-world datasets which have generally different length, 
different values of accuracy and precision (rounding errors due to finite 
precision of the measurement apparatus which include frequency 
acquisition \cite{frank2010}),
and contaminated data with different sources of noise
\cite{garland2016}.
\end{verbatim}
	\textit{
	AMENDED: 
Mon 10 Feb 17:00:32 GMT 2020
	}
	\\

\item  (pp. ) 
It is surprising that even with the previous constraints with regard 
to the quality of data, and the problem with the estimation of embedding parameters,
the use of nonlinear dynamics have proven to be helpful to understand and 
characterise dynamics of time series 
\
\begin{verbatim}
It is then surprising that even with the previous constraints with regard 
to the quality of data, and the problem with the estimation of 
embedding parameters,
the use of nonlinear dynamics have proven to be helpful to understand and 
to characterise dynamics of time series 
\
\end{verbatim}
	\textit{
	AMENDED: Mon 10 Feb 17:02:59 GMT 2020
	}
	\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results}
\begin{enumerate}

\item  (pp. ) 
One of the challenges of the implementation of 
Uniform Time-Delay Embedding to Reconstructed State Spaces is the 
selection of embedding parameters because each time series is unique in 
terms of its structure (modulation of amplitude, frequency and phase) 
\ cite{ frank2010, sama2013, bradley2015}.

\begin{verbatim}
One of the challenges of the implementation of 
uniform time-delay embedding to reconstruct the state spaces is the 
selection of embedding parameters because each time series is unique in 
terms of its structure (e.g. modulation of amplitude, frequency and phase) 
\cite{ frank2010, sama2013, bradley2015}.
\end{verbatim}
	\textit{
	AMENDED: 
Tue 11 Feb 07:35:10 GMT 2020
	}
	\\

\item  (pp. ) 
With that in mind, the problem for this work is not to compute individual 
embedding parameters for each of the time series but to deal with a 
selecting of two parameters that can represent all the time series. 
Our solution for that problem was to compute a sample mean over all 
values in each of the conditions of the time series 
%of Figs~\ref{fig:ts}
for minimum dimension values and 
%Figs~\ref{fig:caoami} 
for minimum delay values as follows.
\begin{verbatim}
With that in mind, the problem for this work is not to compute individual 
embedding parameters for each of the time series but to deal with a 
selection of two parameters that can represent all the time series. 
Our solution for such problem, as explained in below sections, 
was to compute a sample mean over all 
values in each of the conditions of the time series 
%of Figs~\ref{fig:ts}
for minimum dimension values and 
%Figs~\ref{fig:caoami} 
for minimum delay values.
%
\end{verbatim}
	\textit{
	AMENDED: 
Tue 11 Feb 03:47:57 GMT 2020
	}
	\\

\item  (pp. ) 
Minimum embedding parameters were firstly computed 
with False Nearest Neighbour and Average Mutual Information.
%
\begin{verbatim}
Minimum embedding parameters were firstly computed 
with the methods of False Nearest Neighbour and Average Mutual Information.
%
\end{verbatim}
	\textit{
	AMENDED: 
Wed 12 Feb 10:30:37 GMT 2020
	}
	\\

\item  (pp. )

	{\bf Box plots of minimum embedding parameters for horizontal and vertical arm movements.}

	Minimum embedding dimensions are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available from \cite{srep2019}.

\begin{verbatim}

	{\bf Box plots of minimum embedding parameters.}

	Minimum embedding parameters are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2019}.
        }

\end{verbatim}
	\textit{
	AMENDED: \\ 
Thu  5 Mar 06:46:20 GMT 2020 \\
Sat  7 Mar 15:35:36 GMT 2020
	}
	\\

\item  UPDATE REFERENCE AND AMEND LEGEND IN ALL FIGURES \\

	Code and data to reproduce the figure is available from \textbackslash cite{srep2019}.

\begin{verbatim}

	Code and data to reproduce the figure is available in \cite{srep2020}.

\end{verbatim}
	\textit{
	AMENDED: 
Tue 10 Mar 23:46:25 GMT 2020
	}
	\\

\item  Change title

\title{The Application of Recurrent Quantification Analysis 
to Human-Humanoid Interaction}

\begin{verbatim}

\title{Recurrent Quantification Analysis of Human-Humanoid Interaction Activities}

\end{verbatim}
	\textit{
	AMENDED: 
Tue 10 Mar 23:42:40 GMT 2020
	}
	\\

\item  (pp. )

ORIGINAL
\begin{verbatim}
Figures \ref{fig:cao_ami}(A) show that minimum embedding values 
for sensor HS01 appear to show more variations as 
interquartile range is near to 1 
with three exceptions (HF sg0, VN sg0, and VF sg1).
Minimum embedding values appear to be constant for sensor RS01 
as interquartile range is near to 0.1 with the exception of 
two axis (HF sg0 and VF sg1).
Additionally, it can be seen in Figs \ref{fig:cao_ami}(A) a 
decrease of mean values (rhombus) in the box plots
as smoothness of time series increase.
Similarly, the first minimum values of AMI values are 
shown in the box plots of Figs \ref{fig:cao_ami}(B).
It can be seen that values for HS01 tend to be more spread as the smoothness 
of the time series is increasing 
(see the increase of both mean (rhombus) and interquartile range).
However, AMI values for RS01 do not show such increase in relation with
the increase of smoothness except for HF and VF
(Figs \ref{fig:caoami}(B)).

We also computed an overall minimum embedding parameters that represent 
participants, activities, sensors and levels of smoothness, using a sample 
mean of all values in Figs~\ref{fig:cao_ami}.
The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:cao_ami}(A) 
is $\overline{m}_0=6$ 
and the sample mean for minimum values of AMIs from Figs~\ref{fig:cao_ami}(B)
is $\overline{\tau}_0=8$
\end{verbatim}

AMENDED:
\begin{verbatim}
Figures \ref{fig:cao_ami}(A) show box plots for the minimum embedding values 
for sensors HS01 and RS01. Minimum embedding vaues for HS01 presents more 
variations than RS01 as interquartile range for HS01 is near to 1 (with the exceptions of HF sg0, VN sg0, and VF sg1)
whereas interquartile range for RS01 is near to 0.1 (with the exception of two axis HF sg0 and VF sg1).
Additionally, Figures \ref{fig:cao_ami}(A) show
a decrease of mean values (rhombus) in the box plots
as smoothness of time series increase (see sg0, sg1, sg2).
Figures \ref{fig:cao_ami}(B) show box plots for the first minimum values of AMI values.
It can be seen that values for HS01 tend to be more spread as the smoothness 
of the time series is increasing 
(see the increase of both mean (rhombus) and interquartile range).
However, AMI values for RS01 do not show such increase in relation with
the increase of smoothness except for HF and VF
(Figs \ref{fig:cao_ami}(B)).

We also computed a sample mean for an overall minimum embedding parameters that 
represent all participants, activities, sensors and levels of smoothness.
The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:cao_ami}(A) 
is $\overline{m}_0=6$ 
and the sample mean for minimum values of AMIs from Figs~\ref{fig:cao_ami}(B)
is $\overline{\tau}_0=8$.
%
\end{verbatim}

\textit{
	LOG: 
Thu 19 Mar 21:58:23 GMT 2020
	}
	\\

\item  (pp. )

ORIGINAL
\begin{verbatim}
Uniform time-delay embedding were computed with
embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
the first three axis of the rotated data of the PCA are shown 
for the reconstructed state spaces in 
Figs~\ref{fig:rss_aHw10} for horizontal arm movements and 
Figs~\ref{fig:rss_aVw10} for vertical arm movements.

It is easy to observe by eye the differences in each of the
trajectories in the reconstructed state spaces 
(Figs~\ref{fig:rss_aHw10}, \ref{fig:rss_aVw10}), 
however one might be not objective when quantifying those differences 
since those observation might vary from person to person.
With that in mind, we tried to objectively quantify those differences 
using euclidean distances between the origin to each of the points in the 
trajectories, however these results created a suspicious metric, specially 
for trajectories which looked very messy.
With that in mind, we computed Recurrence Plots and 
Recurrence Quantification Analysis to objectively quantify 
the differences in each of the cases of the time series.

\end{verbatim}
AMENDED
\begin{verbatim}

Uniform time-delay embedding were computed with the overall 
embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
the first three axis of the rotated data of the PCA are shown 
for the reconstructed state spaces of horizontal arm movements
(Figs~\ref{fig:rss_aHw10}) and vertical arm movements 
(Figs~\ref{fig:rss_aVw10}).
%
One can observe by eye the differences in each of the
trajectories of the reconstructed state spaces 
(Figs~\ref{fig:rss_aHw10} and \ref{fig:rss_aVw10}), 
however an objective quantification is required.
That being said, we tried to quantify those differences 
using euclidean distances between the origin 
to each of the points in their trajectories, 
which results did not capture the dynamics of 
the trajectories specially 
for those that looked very messy.
Such ambiguities lead us to consider the use of 
Recurrence Plots and Recurrence Quantification Analysis 
to objectively quantify dynamics the capture activities 
using time series.

\end{verbatim}

	\textit{
LOG: \\
Thu 19 Mar 22:21:44 GMT 2020  \\
Mon 13 Apr 13:22:42 BST 2020
	}
	\\

\item  (pp. )

ORIGINAL
\begin{verbatim}
Recurrence Plots (RP) were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and 
vertical arm movements (Fig~\ref{fig:rp_aV}) 
using the average embedding parameters ($m=6$, $\tau=8$) 
and an recurrence threshold of $\epsilon=1$.

	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
\end{verbatim}

AMENDED
\begin{verbatim}
Recurrence Plots (RP) were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and 
vertical arm movements (Fig~\ref{fig:rp_aV}) 
using the average embedding parameters ($m=6$, $\tau=8$) 
and an recurrence threshold of $\epsilon=1$.

	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
\end{verbatim}

	\textit{
LOG: \\
 Mon 13 Apr 13:31:49 BST 2020
	}
	\\

\item  (pp. ) \\ 
ORIGINAL
\begin{verbatim}

Similarly to RP, we computed four RQA metrics (REC, DET, RATIO and ENTR) 
with embedding parameters $m=6$, $\tau=8$ and 
recurrence threshold $\epsilon=1$.

\end{verbatim}

AMENDED
\begin{verbatim}
Four Recurrence Quantification Analysis (RQA) metrics (REC, DET, RATIO and ENTR) 
were also computed with $\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.

\end{verbatim}

\textit{
LOG: \\ 
Mon 13 Apr 13:42:03 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

	[Box plots for RQA values]{
	{\bf Box plots for RQA values.}
	Box plots of (A) REC, (B) DET, (C) RATIO, and (D) ENTR values 
	for 20 participants performing HN, HF, VN and VF movements
	with sensors HS01, RS01 and three smoothed-normalised  
	time series (sg0, sg1 and sg2).
	RQA values were computed with 
	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
	Code and data to reproduce the figure is available in \cite{srep2020}.

\end{verbatim}

AMENDED
\begin{verbatim}

	[Box plots for RQA values]{
	{\bf Box plots for RQA metrics.}
	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of 
	20 participants performing HN, HF, VN and VF movements
	with sensors HS01, RS01 and three smoothed-normalised  
	time series (sg0, sg1 and sg2).
	RQA values were computed with 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$. 
	Code and data to reproduce the figure is available in \cite{srep2020}.

\end{verbatim}

\textit{
LOG: \\ 
Mon 13 Apr 13:53:42 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\subsubsection*{REC values}
It can be seen in the box plots of Figs~\ref{fig:RQABP}(A) that REC values, 
representing the \% of black dots in the RPs, 
are more spread for HN and VN movements (higher interquartile range) 
than HF and VF movements (lower interquartile range) for HS01 sensor. 
In contrast, REC values for RS01 sensor present little variation 
(interquartile range of 0.01).
With regard to the increase of smoothness of time series 
(sg0, sg1 and sg2), REC values present little 
variation as the smoothness is increasing for time series from HS01 
(changes of mean values (rhombus)) while REC values are more affected with 
the smoothness for data from RS01 
(see the incremental changes of mean values (rhombus)).
%See Figs~\ref{fig:rec_aH} and \ref{fig:rec_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual REC values for each participant.

\end{verbatim}

AMENDED
\begin{verbatim}
Figs~\ref{fig:RQABP}(A) show box plots of REC values 
that represent the \% of black dots in RPs.
It can be noted that REC values are more spread 
for HN and VN movements (higher interquartile range) than 
for HF and VF movements (lower interquartile range) for HS01 sensor. 
In contrast, REC values for RS01 sensor present little variation 
(interquartile range of 0.01).
Regarding the increase of smoothness for time series 
(sg0, sg1 and sg2), REC values present little 
variation as the smoothness is increasing for time series from HS01 
(changes of mean values (rhombus)) whereas REC values are 
more affected with the smoothness for data from RS01 
(see the incremental changes of mean values (rhombus)).
%See Figs~\ref{fig:rec_aH} and \ref{fig:rec_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual REC values for each participant.

\end{verbatim}

\textit{
LOG: \\ 
Mon 13 Apr 17:06:01 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\subsubsection*{DET values}
Figs \ref{fig:RQABP}(B) illustrate DET values, 
representing predictability and organisation of the RPs, 
which change very little (interquartile range is around 0.1) 
for type of movement, level of smoothness or type of sensor.
There is also increase of DET values as the smoothness of the signal increase 
(see the incremental changes of mean values (rhombus)).
It can be noted that the interquartile range for faster movements
(HF and VF) with no smoothing (sg0) is lower than the other
levels of smoothness (sg1 and sg2).
%See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual DET values for each participant.

\end{verbatim}

AMENDED
\begin{verbatim}

\subsubsection*{DET values}
Figs \ref{fig:RQABP}(B) illustrate DET values 
that represent the predictability and organisation of RPs.
Generally, it can be noted little change of DET values 
(interquartile range is around 0.1) 
for type of movement, type of sensor 
but the increase of DET values as the smoothness of 
the signal increase 
(see the incremental changes of mean values (rhombus)).
However, the interquartile range for faster movements
(HF and VF) with no smoothing (sg0) is lower than the other
levels of smoothness (sg1 and sg2).
%See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
%for more details about individual DET values for each participant.

\end{verbatim}

\textit{
LOG: \\ 
Mon 13 Apr 17:28:11 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}
\subsubsection*{RATIO values}
Figs \ref{fig:RQABP}(C) present RATIO values, representing dynamic transitions, 
for horizontal and vertical movements.
It can be seen that RATIO values for HS01 sensor vary less 
for HN movements (interquartile range around 2)
than HF movements (interquartile range around 5).
%which is a similar behaviour of RATIO values for oRS01 sensor.
For faster movements (HF and VF), it can be noted a decrease of 
RATIO values as the smoothness of the time series is increasing (rhombus).
For normal movements (HN and VN), the decrease of RATIO values
is less evident than faster movements (see rhombus).
%See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
%\ref{appendix:e:ep} 
%for more details about individual RATIO values for each participant.

\end{verbatim}

AMENDED
\begin{verbatim}

\subsubsection*{RATIO values}
Figs \ref{fig:RQABP}(C) illustrate RATIO values that represent the 
dynamic transitions in RPs. 
Generally, it can be seen that RATIO values for HS01 sensor vary less 
for HN movements (interquartile range around 2)
than HF movements (interquartile range around 5).
For faster movements (HF and VF), it can be noted a decrease of 
the mean values the smoothness of the time series 
is increasing (rhombus). However, for normal movements (HN and VN), 
the decrease of mean values is less evident than 
faster movements (rhombus).
%See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
%\ref{appendix:e:ep} 
%for more details about individual RATIO values for each participant.

\end{verbatim}

\textit{
LOG: \\ 
Mon 13 Apr 17:45:43 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}
\subsubsection*{ENTR values}
Figs \ref{fig:RQABP}(D) show ENTR values, Shannon entropy values 
which represent the complexity of 
the structure the time series, 
for type of movement, level of smoothness or type of sensor.
ENTR values for HS01 sensor show a higher variation  
(interquartile range around 0.5)
than ENTR values for RS01 sensor which appear 
to be more constant (interquartile range 0.1).
It can also be noted the increase of smoothness of time series 
also is presented with an increase of ENTR values (see gray rhombus).
%See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
%\ref{appendix:e:ep}
%for more details about individual ENTR values for each participant.

\end{verbatim}

AMENDED
\begin{verbatim}
\subsubsection*{ENTR values}
Figs \ref{fig:RQABP}(D) show ENTR values that represent 
Shannon entropy values in RPs. 
Generally, 
ENTR values for HS01 sensor show a higher variation  
(interquartile range around 0.5)
than ENTR values for RS01 sensor 
(interquartile range 0.1).
It can also be noted the increase of mean values 
for ENTR values as the smoothness of time series 
increase (rhombus) for type of movement, 
type of sensor and level of smoothness.  
%See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
%\ref{appendix:e:ep}
%for more details about individual ENTR values for each participant.

\end{verbatim}

\textit{
LOG: \\ 
Mon 13 Apr 18:00:49 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}
	RQA ENTR values 
	for different embedding parameters
	$[1,10]= \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
	$[1,10]= \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
	with an increase of 1
	and different recurrence thresholds $\epsilon=1, 2, 3$.
	RQA ENTR values are for $p03$, sensor HS01, of a window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.

\end{verbatim}

AMENDED
\begin{verbatim}
	[3D RQA ENTR values]{
	{\bf 3D RQA ENTR values.}
	RQA ENTR values 
	for different embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
	incrementing by one
	and three recurrence thresholds $\epsilon=1, 2, 3$.
	RQA ENTR values were computed with data from $p03$, sensor HS01, with 
	a window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.

\end{verbatim}

\textit{
LOG: \\ 
Tue 14 Apr 09:19:00 BST 2020 \\
Tue 14 Apr 11:04:36 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\subsection*{3D RQA ENTR}
Choosing an appropriate recurrence threshold is crucial to get 
meaningful representations in RP and RQA, however, as previously shown, 
one can selected the default recurrence threshold of $\epsilon=1$, 
as long as it is able to represent the dynamical transitions 
in each of the time series \cite{marwan2011}. %%Marwan et al. \cite{marwan2011} 
However, to show how that the selection of recurrence threshold affects RQA
values, 
we computed ENTR values of RQA metrics for different embedding parameters
$[1,10]= \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
$[1,10]= \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $
with an increase of 1
and different recurrence thresholds 
$[0.2,3]= \{ \epsilon \in \mathbb{R} | 0.2 \le \epsilon \le 3.0  \} $
with increments of 0.1.
For easily visualisation, Figs \ref{fig:RQA-IND} only show 3D RQA for 
three recurrence thresholds $\epsilon=1, 2, 3$ with three levels of 
smoothness (sg0, sg1, sg2). It can be noted in Figs \ref{fig:RQA-IND} 
that the increase of recurrence threshold is associated to the increase 
of ENTR values in any of the levels of smoothness (see values of the ENTR bar).
One can also see that the maximum values of ENTR 
for $\epsilon=1$ and sg0 are for embedding dimensions of 2 
and then as $\epsilon$ increases the maximum values of ENTR are for
an associated increase of in the embedding dimension.
It can also be noted that the increase of level of smoothness (from sg0 to sg2) 
is associated with both the increase of ENTR values and its smoothed 3D shape.

\end{verbatim}

AMENDED
\begin{verbatim}
\subsection*{3D RQA ENTR}
It has been shown that one can select a default recurrence 
threshold of $\epsilon=1$ to represent the dynamical transitions 
in each of the time series \cite{marwan2011}.
%Choosing an appropriate recurrence threshold is crucial so as to get 
%meaningful representations in RP and RQA, however, 
However, to show how that the selection of recurrence threshold affects RQA
values, 
we computed ENTR values of RQA metrics for different embedding parameters
$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by one, 
three recurrence thresholds $\epsilon=1, 2, 3$.
and three levels of smoothness (sg0, sg1, sg2). 
%different recurrence thresholds 
%$\{ \epsilon \in \mathbb{R} | 0.2 \le \epsilon \le 3.0  \} $
%incrementing by 0.1.
%For easily visualisation, Figs \ref{fig:RQA-IND} only show 3D RQA for 
%three recurrence thresholds $\epsilon=1, 2, 3$ 
Figs \ref{fig:RQA-IND} show the increase of recurrence threshold is 
associated to the increase of ENTR values in any of the levels 
of smoothness (see values of the ENTR bar).
Similarly, it can be noted that the increase of level of 
smoothness (from sg0 to sg2) is associated with both the increase 
of ENTR values and the smoothness of the 3D surfaces.
%One can also see in Figs \ref{fig:RQA-IND} that the maximum values of 
%ENTR for $\epsilon=1$ and sg0 are for embedding dimensions of 2 
%and then as $\epsilon$ increases the maximum values of ENTR are for
%an associated increase of in the embedding dimension.

% SECOND AMMENMEND
\subsection*{3D RQA ENTR}
It has been shown that one can select a default recurrence 
threshold of $\epsilon=1$ to represent the dynamical transitions of 
time series \cite{marwan2011}.
%Choosing an appropriate recurrence threshold is crucial so as to get 
%meaningful representations in RP and RQA, however, 
However, to show how that the selection of recurrence threshold affects RQA
values, 
we computed ENTR values of RQA metrics for different embedding parameters
$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by one,
with the consideration of three recurrence thresholds $\epsilon=1, 2, 3$ 
and three levels of smoothness (sg0, sg1, sg2). 
%different recurrence thresholds 
%$\{ \epsilon \in \mathbb{R} | 0.2 \le \epsilon \le 3.0  \} $
%incrementing by 0.1.
%For easily visualisation, Figs \ref{fig:RQA-IND} only show 3D RQA for 
%three recurrence thresholds $\epsilon=1, 2, 3$ 
That being said, figures \ref{fig:RQA-IND} show the increase of recurrence threshold is 
associated to the increase of ENTR values in any of the levels 
of smoothness (see values of the ENTR bar).
Similarly, it can be noted that the increase of level of 
smoothness (from sg0 to sg2) is associated with both the increase 
of ENTR values and the smoothness of the 3D surfaces.
%One can also see in Figs \ref{fig:RQA-IND} that the maximum values of 
%ENTR for $\epsilon=1$ and sg0 are for embedding dimensions of 2 
%and then as $\epsilon$ increases the maximum values of ENTR are for
%an associated increase of in the embedding dimension.

%THIRD AMENMEND
\subsection*{3D RQA ENTR}
Marwan stated that the computation of an optimal recurrence threshold
$\epsilon$ is still an open question as it depends on a particular problem and question \cite{marwan2011}.
That said, recurrence threshold $\epsilon$ to study of dynamical transitions can be 
of little importance because of the relative change of the RQA measures 
does not depend too much on it in a certain range \cite{marwan2011}
%Choosing an appropriate recurrence threshold is crucial so as to get 
%meaningful representations in RP and RQA, however, 
However, to show how that the selection of recurrence threshold affects RQA
values, 
we computed ENTR values of RQA metrics for different embedding parameters
$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by one,
with the consideration of three recurrence thresholds $\epsilon=1, 2, 3$ 
and three levels of smoothness (sg0, sg1, sg2). 
%different recurren

%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa_epsilons}
    \caption
	[3D RQA ENTR values]{
	{\bf 3D RQA ENTR values.}
	RQA ENTR values are
	for different embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
	incrementing by one and three recurrence thresholds $\epsilon=1, 2, 3$.
	RQA ENTR values were computed with data from $p03$, sensor HS01, with 
	a window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:RQA-IND}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\end{verbatim}

\textit{
LOG: \\ 
Tue 14 Apr 11:29:33 BST 2020\\
Sun 26 Apr 11:21:45 BST 2020\\
Thu 25 Jun 07:22:03 BST 2020\\
Thu 25 Jun 22:36:42 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\subsection*{Other effects on 3D RQA ENTR values}
%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}). 
We also computed 3D surfaces of RQA metrics for different sensors 
and different activities (Fig~\ref{fig:3dRQAENTR_sensoractivities}).
From the maximum values of ENTR (lateral bars), one can see 
a decrease of ENTR values for activities going from normal to 
faster velocity and from human sensor (HS01) to robot sensor (RS01) 
(See Fig~\ref{fig:3dRQAENTR_sensoractivities}). 
%%---------------------------------(FIGURE)-------------------------------------

	{\bf 3D surfaces of RQA ENTR metrics for sensors and activities.}
	RQA ENTR values with embedding parameters
	$[1,10]= \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$[1,10]= \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	for different activities and sensors. 
	RQA ENTR values are for $p03$, sg0 and window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }

\end{verbatim}

AMENDED
\begin{verbatim}

%%\subsection*{Other effects on 3D RQA ENTR values}
%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}). 
Similarly, Figs~\ref{fig:3dRQAENTR_sensoractivities} show 3D surfaces of 
ENTR values for different sensors and different activities. 
Maximum values of ENTR (lateral bars) in Figs~\ref{fig:3dRQAENTR_sensoractivities} 
decrease for activities from normal (HN, VN) to faster (HF, VF) velocity and 
from human sensor (HS01) to robot sensor (RS01).

	{\bf 3D surfaces of RQA ENTR metrics for sensors and activities.}
	RQA ENTR values with embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1 $ for different activities and sensors. 
	RQA ENTR values are for $p03$, sg0 and window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.

\end{verbatim}

\textit{
LOG: \\ 
Tue 14 Apr 11:55:28 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

% Create a "Supplementatry information"
% ~/srep2019/docs/supplementary-information/report
added: Sun 26 Apr 11:28:02 BST 2020
SORTED: Completion of new datasets, figures and report on Thu 25 Jun 00:50:15 BST 2020

However the manuscript should need a compresed version of such results to summarise
the following:
added: Thu 25 Jun 06:52:00 BST 2020

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EPSILON, SMOTHNESS (DONE)
%X Three level of 
%X smoothness were computed for RQA metrics showing smoothed 3D surfaces and 
%X the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%ACTIVITIES AND SENORS (DONE)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% COMPARE PARTICIPANTS 
%Similarly, 3D surfaces of RQA metrics were also computed for three 
%participants (Fig.~\ref{fig:topo_participants}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% COMPARE WINDOWS
%RQA metrics are also affected by 
%the window length where for example four window lengths of 100, 250, 500 
%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}).

Figures~\ref{fig:3dRQAENTR_sensoractivities} show 3D surface plots of 
ENTR values for different sensors and different activities. 
It can be seen that maximum values of ENTR (lateral bars) in 
Figs~\ref{fig:3dRQAENTR_sensoractivities} decrease from normal (HN, VN) to faster (HF, VF) velocities 
and from human sensor (HS01) to robot sensor (RS01).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
    \caption{
	{\bf 3D surfaces plots of RQA ENTR values for sensors and activities.}
	RQA ENTR values with embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1 $ for different activities 
	(Horizontal Normal, Horizontal Faster, Vertical Normal, and Vertical Faster) and 
	sensors (Human Sensor 01 and Robot Sensor 01). 
	RQA ENTR values were computed from data of $p03$, sg0 and 
	window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_sensoractivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

Figures~\ref{fig:3dRQAENTR_participantsactivities}(A) show the similarity of 3D surface plots in 
relationship with four activities and three participants. 
It can be noted the subtle differences for each of the participants. 
However, 3D surface plots in figures~\ref{fig:3dRQAENTR_participantsactivities}(B) show 
little change for each participant as the data from a sensor attached to the robot. 
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{figures/rqa/output/rqa-participants}
    \caption{
	{\bf 3D surface plots of RQA ENTR value for different participants, activities and sensors}
	RQA ENTR values with embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1 $ for different activities 
	(Horizontal Normal, Horizontal Faster, Vertical Normal, and Vertical Faster) 
	three participants (p01, p02, and p03) and two main 
	classifications for (A) Human Sensor 01 and (B) Robot Sensor 01.
	RQA ENTR values were computed from data with sg0 and window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_participantsactivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

Figures \ref{fig:3dRQAENTR_windowsactivities} show the effect on the 3D surface plots
as the window length of the time series increase (w100 (2-sec), w250 (5-sec), w500 (10-sec) and w750 (15-sec)).
It can be noted that the increase of number of samples improves the quality 
of the surface plots where 100 samples roughly capture the dynamics of the activity
and 250, 500 and 750 samples show similar representation of RQA Entropy values 
(being the surface plot with 750 samples the best test case).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-windows}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for windows and activities.}
	RQA ENTR values for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$, 
	$\epsilon = 1 $ for different 
	windows size (e.g. w100 (100 samples), w250 (250 samples),
	w500 (500 samples) and w750 (750 samples)).
	RQA ENTR values were computed from data of $p01$ and sg0.
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_windowsactivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

Mon 29 Jun 22:35:05 BST 2020
For instance, 
Figures \ref{fig:RQA-IND} show the increase of recurrence threshold is 
associated to the increase of ENTR values in any of the levels 
of smoothness (see values of the ENTR bar).
Similarly, it can be noted that the increase of level of 
smoothness (sg0, sg1 and sg2) is associated with the increase 
of ENTR values of the 3D surface plots.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-epsilons}
    \caption
	[3D surface plots of RQA ENTR values]{
	{\bf 
	3D surface plots of RQA ENTR values for different recurrence threshold and smoothness levels.}
	RQA ENTR values are
	for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
	incrementing by one and three recurrence thresholds $\epsilon=1, 2, 3$.
	RQA ENTR values were computed with data from $p03$, sensor HS01, with 
	a window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:RQA-IND}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
Figures~\ref{fig:3dRQAENTR_sensoractivities} show 3D surface plots of 
ENTR values for different sensors and different activities.
Figs~\ref{fig:3dRQAENTR_sensoractivities} show that RQA ENTR (lateral bars) 
decrease from normal (HN, VN) to faster (HF, VF) velocities from both 
human sensor (HS01) to robot sensor (RS01).
Also, it can be noted that RQA ENTR values decrease from human sensor (HS01) 
to robot sensor (RS01).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for different sensors and activities.}
	RQA ENTR values are for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1 $ considering four activities 
	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
	Vertical Faster (VF) and sensors Human Sensor 01 (HS01) and 
	Robot Sensor (RS01).
	RQA ENTR values were computed from data of $p03$, sg0 and 
	window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_sensoractivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Figures~\ref{fig:3dRQAENTR_participantsactivities} show the similarity of 3D surface plots in 
relationship with three participants (p01, p02 and p03)
performing four activities (HN, HF, VN, and VF) and two sensors (HS01, RS01).
Figures~\ref{fig:3dRQAENTR_participantsactivities}(A) 
present subtle differences for each of the participants (see RQA ENTR bar). 
Figures~\ref{fig:3dRQAENTR_participantsactivities}(B) show 
less change of RQA ENTR values for each participant as the data is 
from a sensor attached to the robot. 
Also, from data of sensors attached to the robot, it can be noted 
that changes are more notable for faster movements than normal movements. 
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{figures/rqa/output/rqa-participants}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for different participants, activities and sensors.}
	RQA ENTR values are for participants (p01, p02, and p03) 
	in the categories of 
	(A) Human Sensor 01 (HS01) and 
	(B) Robot Sensor 01 (RS01)
	considering embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
	with $\epsilon = 1$ and four activities 
	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
	Vertical Faster (VF).
	RQA ENTR values were computed from data of sg0 and window size of 10-secs (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_participantsactivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Figures \ref{fig:3dRQAENTR_windowsactivities} show the effect of the increase 
of window length of the time series (w100 (2-sec), w250 (5-sec), w500 (10-sec) and w750 (15-sec))
on the 3D surface plots. 
It can be noted that the increase of number of samples improves the quality 
of the surface plots where 100 samples poorly capture the dynamics of each activity.
However as the window length increases from 250, 500 to 750 samples, 
3D surface plots show a similar representation of RQA ENTR values 
(being the surface plot with 750 samples the best test case).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-windows}
    \caption{
	{\bf 3D surface plots of RQA ENTR values for different windows lengths and activities.}
	RQA ENTR values are for embedding parameters
	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$, 
	with $\epsilon = 1 $ considering four 
	windows lengths (e.g., w100 (100 samples), w250 (250 samples),
	w500 (500 samples) and w750 (750 samples)) and
	four activities 
	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
	Vertical Faster (VF).
	RQA ENTR values were computed from data of $p01$ and sg0.
	Code and data to reproduce the figure is available in \cite{srep2020}.
       }
\label{fig:3dRQAENTR_windowsactivities}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

\end{verbatim}

AMENDED
\begin{verbatim}

\end{verbatim}

\textit{
LOG: \\ 
Thu 25 Jun 06:48:57 BST 2020 \\
Fri 26 Jun 03:31:54 BST 2020 \\
Mon 29 Jun 22:35:15 BST 2020
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\end{verbatim}

AMENDED
\begin{verbatim}

\end{verbatim}

\textit{
LOG: \\ 
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\end{verbatim}

AMENDED
\begin{verbatim}

\end{verbatim}

\textit{
LOG: \\ 
}
\\

\item  (pp. ) \\
ORIGINAL
\begin{verbatim}

\end{verbatim}

AMENDED
\begin{verbatim}

\end{verbatim}

\textit{
LOG: \\ 
}
\\

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discussion}
\begin{enumerate}

\item  (pp. )

ORIGINAL
\begin{verbatim}

Tue 30 Jun 22:28:39 BST 2020 
%The Discussion should be succinct and must not contain subheadings.

%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}).

It is evident that time series from different sources 
(participants, movements, axis type, window length or levels of smoothness) 
presents visual differences for embedding parameters and therefore for RRSs. 
For which, the selection of embedding parameters was our first challenge 
where we computed embedding parameters for each time series 
(Fig \label{fig:caoami}) and then computed a sample mean over 
all time series in order to get two embedding parameters 
to compute all RRSs (Figs \label{fig:rss_aHw10} and \label{fig:rss_aVw10}). 
%with its corresponded type of movement. 
Then we found that the quantification of variability with regard to 
the shape of the trajectories in RSSs requires more investigation 
since our original proposed method base on euclidean metric failed 
to quantify those trajectories. Specially, for trajectories which 
were not well unfolded. 
With that in mind, we apply 
Recurrence Quantification Analysis (RQA) metrics 
(REC, DET, RATIO and ENTR) in order to avoid any subjective interpretations 
or personal bias with regard to the representation of the trajectories in RSSs.

\subsection*{RQA metrics with fixed parameters}
Considering that RQA metrics were computed with fixed embedding parameters 
($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
the following. REC values, which represents the \% of black points in the RPs, 
were more affected with and increase in normal speed movements (HN and VN) 
than faster movements (HF and VF) for the sensor attached to the participants 
(HS01). Such decrease of REC values from normal speed to faster speed 
movements is also presented in data from sensor attached to the robot (RS01), 
and little can be said with regard to the dynamics of the time series coming 
from RS01 (Fig \ref{fig:RQABP}A).
Similarly, DET values, representing predictability and 
organisation in the RPs, present little variation in the any of the time 
series where little can be said (Fig \ref{fig:RQABP}B).
In contrast, RATIO values, which represent 
dynamic transitions, were more variable for faster movements (HF and VF) 
than normal speed movements (HN and VN) with sensors attached to the 
participants (HS01). For data coming from sensors attached to the robot 
(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
more than values coming from vertical movmentes (VN, VF) 
(Fig \ref{fig:RQABP}C).
With that, it can be said that RATIO values can represent a bit better
than REC or DET metrics for the variability of imitation activities in 
each of the conditions for time series.
Similarly, ENTR values for HN were higher than values for HF
and ENTR values varied more for sensor attached to participants 
than ENTR values for sensors of the robot. It is evidently that 
the higher the entropy the more complex the dynamics are, 
however, ENTR values for HN appear a bit higher than HF values, 
for which we believe this happens because of the structure the time series
which appear more complex for HN than  HF movements which presented a 
more consistence repetition (Fig \ref{fig:RQABP}D).

Additionally, we observed that some RQA metrics are affected by the 
smoothness of data. 
%For which, we also explored the effect of smoothness of raw-normalised data 
%where, 
For instance, REC and DET values were not completely affected by the 
smoothness of time-series since these RQA values seemed to be constants. 
However, for RATIO values, 
the effect of smoothness can be noticed with a slight decrease of amplitude 
in any of the time series conditions which is also presented with ENTR values.

\subsection*{RQA metrics with different parameters}
Iwansky et al. \cite{iwanski1998} stated that patterns in RPs and 
metrics for RQA are independent of embedding dimension parameters, 
however, that is not the case when using 
different recurrence thresholds. Such changes of recurrence threshold values 
can modify the patterns in RPs and therefore the values of RQA metrics.
We therefore computed 3D surfaces to explore the sensibility and robustness of 
embedding parameters and recurrence threshold in RQA  metrics. Following the 
same methodology of computing 3D surfaces, we also considered variation of 
window length size to present RQA metrics dependencies with embedding 
parameters, recurrence thresholds and window length size.

\end{verbatim}
AMENDED

\begin{verbatim}

Wed  1 Jul 00:33:42 BST 2020 
%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Discussion}
Time series from different sources  
(e.g., participants, movements, axis type) as well as
different characteristis (e.g., sample rate, window length or levels of smoothness) 
result in different embedding parameters and therefore different 
values for recurrence plots and recuerence quantification analysis metrics.
That said, while embedding parameters for individual
time series were succesfully computed, it has been shown that the quantification 
of variability with regard to the shape of the trajectories in 
reconstructed state spaces require more investigation
as our initial results based on euclidean distances 
in the trajectories failed to quantify trajectories 
which were not well unfolded. 
Then, the use of Recurrence Quantification Analysis (RQA) metrics 
(e.g., REC, DET, RATIO and ENTR) helped to quantify 
the variablity of different sources of time series, to then found that 
ENTR metric along with different recurrence thresholds showed 
the variation of different sources of time series.

Thu  2 Jul 22:26:10 BST 2020
%*******************************************************************************
\section*{Discussion}
Time series from different sources  
(e.g., participants, movements, axis type) as well as
different characteristics (e.g., sample rate, window length or levels of smoothness) 
result in different embedding parameters and therefore different 
values for recurrence plots and recurrence quantification analysis metrics.
That said, while embedding parameters for individual
time series were successfully computed, the quantification 
of variability with regard to the shape of the trajectories in 
reconstructed state spaces require more investigation
as our initial, approach based on euclidean distances on the trajectories, 
failed to quantify trajectories which were not well unfolded. 
To then found out that Recurrence Plots along with 
Recurrence Quantification Analysis (RQA) metrics (e.g., REC, DET, RATIO and ENTR) 
and its variation of embedding parameters and recurrence thresholds 
helped to quantify the variability of different sources of time series.
However, further investigation is required to be done in order to 
have better intuition and meaningful interpretation of nonlinear 
analysis such as to find a right 
balance among (i) the level of smoothness of the signal, 
(ii) the selection of recurrence thresholds and (iii) 
the range of embedding parameters.

\

\end{verbatim}

	\textit{
	LOG: \\
	Tue 30 Jun 22:28:39 BST 2020 \\
	Wed  1 Jul 00:33:42 BST 2020 \\
	Thu  2 Jul 22:26:16 BST 2020 \\
	}
	\\

\end{enumerate}

\subsection{Conclusions}

\begin{enumerate}

\item  (pp. )

ORIGINAL
\begin{verbatim}

\section*{Conclusions}
Generally, we can conclude that using a different level of 
smoothness for time series help us to visualise and to quantify 
the variation of movements 
between participants using RSSs, RPs and RQA. 
It is important to mention that some RQA's metrics 
(e.g. DET and ENTR) are more robust to the effect of 
smoothness of time series.
Choice of RQA metrics will be depend on 
what to quantify, for instance, one can compute 
predicability, organisation,  
dynamics transitions, or complexity and determinism of the 
input time series. However, RQA metrics show certain constrains with 
regard to activity type, window length and structure of the time series.
For instance, RATIO and ENTR are helpful to distinguish 
differences in any of the categories of the time series (sensor, activity, 
level of smoothness and number of participant), however time series
data from the sensor attached to the robot seemed to have 
little variations between as humanoid degrees of freedom 
did not allow it to move with a wide range of variability.

Additionally, we can point out that even though our experiment is 
limited to twenty healthy right-handed participants of a range age 
of mean 19.8 SD=1.39, RQA metrics show the potential to quantify 
human movement variability in the context of 
human-humanoid imitation activities.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	You need a sentence or two here
%	to explain why this is claimed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Additionally, we observed that some RQA metrics are affected by the 
%smoothness of data. 
%%For which, we also explored the effect of smoothness of raw-normalised data 
%%where, 
%For instance, REC and DET values were not completely affected by the 
%smoothness of time-series since these RQA values seemed to be constants. 
%However, for RATIO values, 
%the effect of smoothness can be noticed with a slight decrease of amplitude 
%in any of the time series conditions which is also presented with ENTR values.

%With that in mind, we conclude that 
%quantification of human-humanoid imitation activities is possible for 
%participants of different ages, state of health and anthropomorphic features.

%Such understanding and measurement of movement variability using
%cheap wearable inertial sensors lead us to have a more intuitive selection of parameters
%to reconstruct the state spaces and to create meaningful interpretations
%of the recurrence plots and the results of the metrics with recurrence quantification 
%analysis. 
%

%
%However, we believe that further investigation is required to find the 
%right balance between the level of smoothness of the signal
%and defining what is the aim of RQA where 
%its representations using RSS, RP and RQA.
%Specially, where the level of smoothness does not affect 
%the variation of each of the movements quantification. 
%%We can also conclude that finding the right balance between 
%%smoothness and the raw data to capture movement variability is 
%%a still a problem that has many avenues for exploration.
%

%Hence, variation of recurrence thresholds and embedding parameters 
%has been presented to show the relationships to different datasets 
%(participants, activities, windows lengths and sensors).

\end{verbatim}
AMENDED
\begin{verbatim}

Thu  2 Jul 00:15:27 BST 2020
\section*{Conclusions}
This work allows to conclude that the choice of 
nonlinear analysis tool (e.g., RSSs, RPs, RQA metrics) will depends 
on what one would like to quantify on the time-series 
data (e.g., predicability, organisation, dynamics transitions, 
or complexity and determinism).
Then, time-series data characteristics 
(e.g., window size length, level of smoothness) plays an important role
as well on the results that nonlinear analysis tool can provide.
Similarly, the results of the nonlinear analysis tools
are associated with the structure of the time-series data 
(e.g. frequency, amplitude), the position of the sensor 
and activity performed by either a robot or human being
(as degrees of freedom from the humanoid are far less than
human movement). 
That said, it is unquestionable that the use of different 
characteristics of the time-series data 
(e.g., sensor, activity, level of smoothness and participants)
has help us to visualise and to quantify 
with nonlinear tools the variation of 
movements of our human-humanoid activities.
However, with some limitation of the nonlinear tools such as 
RSSs cannot unfold trajectories, RPs 
are dependent on recurrence thresholds
and RQA metrics show certain constrains with 
regard to activity type, window length and structure of the time series.
Specifically, DET values appear to be constants across sensors, 
activities and levels of smoothness, 
whereas REC and RATIO, as function of REC, values 
show variation for certain sensors and movements. To then 
found that RQA ENTR values with recurrence thresholds  
were appropriate to quantify 
the different changes and variations of the characteristics of 
time-series data.

Further investigation is required to be done in order to 
have more intuition and meaningful interpretation of nonlinear 
analysis tools as well as to find a right 
balance among (i) the level of smoothness of the signal, 
(ii) the selection of recurrence thresholds and (iii) 
the range of embedding parameters. 
Nonetheless, this work provides a good starting 
point to conclude that Shannon Entropy is an approach that can
lead to interesting results on the quantification of human-humanoid imitation activities 
for participants of different ages, state of health and anthropomorphic features.

Thu  2 Jul 22:41:43 BST 2020
\section*{Conclusions}
This work allows to conclude that the choice of 
nonlinear analysis tool (e.g., RSSs, RPs, RQA metrics) will depend 
on what one would like to quantify on the time-series 
data (e.g., predicability, organisation, dynamics transitions, 
or complexity and determinism).
Then, time-series data characteristics 
(e.g., window size length, level of smoothness) plays an important role
as well on the results that nonlinear analysis tool can provide.
Similarly, the results of the nonlinear analysis tools
are associated with the structure of the time-series data 
(e.g. frequency, amplitude), the position of the sensor 
and activity performed by either a robot or human being
(as degrees of freedom from the humanoid are far less than
human movement). 
That said, it has been shown that the use of different 
characteristics of the time-series data 
(e.g., sensor, activity, level of smoothness and participants)
has help us to visualise and to quantify 
with nonlinear tools the variation of 
movements of, in this work, human-humanoid activities.
However, some limitation of nonlinear tools are related 
to the computation of different parameters 
(e.g., recurrence thresholds, embedding parameters)
that reflect the dynamics of individual characteristics 
of activity type, window length and structure of the time series.
Specifically, the example of DET values which appear 
to be constants across sensors, activities and levels of smoothness, 
whereas REC and RATIO, as function of REC, values 
show variation for certain sensors and movements. 
To then found out that RQA ENTR values with different recurrence 
thresholds were appropriate to quantify 
the different changes and variations of the characteristics of 
time-series data.
Therefore, we can conclude that this work provides a good starting 
point and reference to the use of Shannon Entropy to quantify 
human-humanoid imitation activities 
that can then lead to interesting results on the quantification 
of movement variability of participants with different ages, 
state of health and anthropomorphic features.

\end{verbatim}

	\textit{
	LOG:\\ 
	Thu  2 Jul 00:15:27 BST 2020 \\
	Thu  2 Jul 22:41:48 BST 2020 \\
	}
	\\

\end{enumerate}

\subsection{Future work}

\begin{enumerate}

\item  (pp. )

ORIGINAL
\begin{verbatim}

\section*{Future Work}

\subsection*{Inertial Sensors}
To better understand the signals collected through 
inertial sensors in the context of human-humanoid interaction, 
we could apply derivates to the acceleration data. 
We can then explore the jerkiness of movements and therefore 
the nature of arm movements which typically have 
minimum jerk \cite{flash1985}, its relationship with different body 
parts \cite{devries1982, mori2012} or the application of higher derivatives 
of displacement with respect time such as snap, crackle and pop \cite{eager2016}.

\subsection*{RQA ENTR}
Having shown that RQA ENTR metrics are robust against different data
time series, we believe that further investigation is required. 
For example, Marwan et al. \cite{marwan2007, marwan2015} reviewed 
different aspects to compute RPs using different criteria for neighbours, 
different norms ($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or 
different methods to select the recurrence threshold $\epsilon$, 
such as using certain percentage of the signal \cite{letellier2006}, 
the amount of noise or using a factor based on the standard deviation 
of the observational noise among many others \cite{marwan2007}.

\end{verbatim}
AMENDED
\begin{verbatim}

Future work is not necessarily part of the manuscript in the guidelines 
for which I leave it out for coming publications
\end{verbatim}

	\textit{
	LOG:\\ 
	Wed  1 Jul 00:39:16 BST 2020
	}
	\\

\end{enumerate}

\subsection{Abstract}

\begin{enumerate}

\item  (pp. )

ORIGINAL
\begin{verbatim}

\begin{abstract}
Human movement variability occurs in motor performance 
across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly when traditional
methods in time domain and frequency domain fail to detect tiny modulations in
frequency or phase for time series. 
For this work, we hence investigate methodologies from nonlinear dynamics such as 
reconstructed state space (RSS), uniform time-delay embedding (UTDE), 
recurrence plots (RPs) and recurrence quantification analysis (RQA) metrics. 
Particularly, we are interested in the weaknesses and robustness 
of nonlinear dynamics tools from raw and post-processed data of 
wearable inertial sensors (IMUs). 
In the reported experiment, twenty right-handed healthy participants 
imitated simple vertical and horizontal arm movements in normal 
and faster speed from an humanoid robot.
With four window lengths and three levels of smoothed time series, 
we found visual differences in the patterns of RSSs and RPs. 
Then with the use of RQA metrics, we find out that the type of 
movements and the level of smoothness affects those metrics. 
Specifically, entropy values from RQA were well distributed and 
presented variation in all the conditions for time series. 
This work have then the potential to enhance the development of 
better diagnostic tools for various applications in rehabilitation, 
sport science or for new forms of human-robot interaction.
\end{abstract}
%Number of words: 226

\end{verbatim}
AMENDED
\begin{verbatim}

Thu  2 Jul 00:37:16 BST 2020
\begin{abstract}
Human movement variability occurs in motor performance 
across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly 
when methods in time domain, frequency domain or nonlinear dynamics 
can break down due to the real-world time-series datasets. 
For this work, we hence investigate methodologies from nonlinear dynamics such as 
reconstructed state space (RSS), uniform time-delay embedding (UTDE), 
recurrence plots (RPs) and recurrence quantification analysis (RQA) metrics
with real-world time-series data.
Particularly, we are interested in the weaknesses and robustness 
of nonlinear dynamics tools of raw and post-processed data of 
wearable inertial sensors. 
That said, twenty right-handed healthy participants 
imitated simple vertical and horizontal arm movements in normal 
and faster velocity from an humanoid robot in order to have 
four window lengths and three levels of smoothed time-series data,
to then found visual differences in the patterns with RSSs and RPs. 
To then find out that the the use of RQA metrics
can quantify activities, types of sensors, windows lenghts 
and the level of smoothness. 
Specifically, we can conclude that Shannon Entropy can lead to 
interesting results on the quantification of human-humanoid imitation activities
for participants of different ages, state of health and anthropomorphic features.
to then enhance the development of 
better diagnostic tools for various applications in rehabilitation, 
sport science or for new forms of human-robot interaction.
\end{abstract}

Thu  2 Jul 22:54:44 BST 2020
\begin{abstract}
Human movement variability occurs in motor performance 
across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly 
when methods in time domain, frequency domain or nonlinear dynamics 
can break down due to the real-world time-series datasets. 
For this work, we therefore investigate nonlinear dynamics methods 
such as reconstructed state space (RSSs), uniform time-delay embedding, 
recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
with real-world time-series data.
Particularly, we are interested in the weaknesses and robustness 
of nonlinear dynamics tools with the use of raw and post-processed 
data of wearable inertial sensors. 
That said, twenty right-handed healthy participants 
imitated simple vertical and horizontal arm movements in normal 
and faster velocity from an humanoid robot in order to have 
four window lengths and three levels of smoothed time-series data,
to then found visual differences in the patterns with RSSs and RPs. 
We then found out that RQA metrics
can quantify activities, types of sensors, windows lenghts 
and level of smoothness. 
Specifically, we can conclude that Shannon Entropy can lead to 
interesting results on the quantification of 
movement variablity for participants of different ages, 
state of health and anthropomorphic features.
to then potentially enhance the development of 
better diagnostic tools for applications in rehabilitation, 
sport science or for new forms of human-robot interaction.
\end{abstract}

Thu  2 Jul 23:03:20 BST 2020
\begin{abstract}
Human movement variability occurs in motor performance 
across multiple repetitions of a task and such behaviour is an
inherent feature within and between each persons' movement. Quantifying
movement variability is still an open problem, particularly 
when methods in time domain, frequency domain or nonlinear dynamics 
can break down due to the real-world time-series datasets. 
For this work, we therefore investigate nonlinear dynamics methods 
such as reconstructed state space (RSSs), uniform time-delay embedding, 
recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
with real-world time-series data.
Particularly, we are interested in the weaknesses and robustness 
of nonlinear dynamics tools with the use of raw and post-processed 
data of wearable inertial sensors. 
That said, twenty right-handed healthy participants 
imitated simple vertical and horizontal arm movements in normal 
and faster velocity from an humanoid robot in order to have 
four window lengths and three levels of smoothed time-series data,
to then found visual differences in the patterns with RSSs and RPs
and particulatly the computed differences with RQA metrics
that help us to quantify activities, types of sensors, windows lenghts 
and level of smoothness. 
Specifically, we can conclude that RQA ENTR, Shannon Entropy, 
can lead to interesting results on the quantification of 
movement variablity for participants of different ages, 
state of health and anthropomorphic features to 
then enhance the development of 
better diagnostic tools for applications in rehabilitation, 
sport science or for new forms of human-humanoid interaction.
\end{abstract}

\end{verbatim}

	\textit{
	LOG:\\ 
	Thu  2 Jul 00:37:21 BST 2020\\
	Thu  2 Jul 22:54:53 BST 2020\\
	Thu  2 Jul 23:03:20 BST 2020
	}
	\\

\end{enumerate}

\subsection{Title}

\begin{enumerate}

\item  (pp. )

ORIGINAL
\begin{verbatim}
\title{Recurrent Quantification Analysis of Human-Humanoid Interaction Activities}
\end{verbatim}

AMEND 
\begin{verbatim}
Thu  2 Jul 22:44:53 BST 2020
\title{Recurrent Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
\end{verbatim}

	\textit{
	LOG:\\ 
	Thu  2 Jul 22:44:53 BST 2020
	}
	\\

\end{enumerate}

\end{document}

### 2020-07-02 docs/manuscript/main.tex
completion of corrections-2020-01-14 and updates manuscript2020-07-03

@@
 \usepackage[T1]{fontenc}
-\graphicspath{{../}} %%%REPOSITORY PATH
-
-\title{The Application of Recurrent Quantification Analysis 
-to Human-Humanoid Interaction}
+\graphicspath{{../}} %goes to path: docs/
+
+\title{Recurrent Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
 \author[1,*]{Miguel Xochicale}
@@
 inherent feature within and between each persons' movement. Quantifying
-movement variability is still an open problem, particularly when traditional
-methods in time domain and frequency domain fail to detect tiny modulations in
-frequency or phase for time series. 
-For this work, we hence investigate methodologies from nonlinear dynamics such as 
-reconstructed state space (RSS), uniform time-delay embedding (UTDE), 
-recurrence plots (RPs) and recurrence quantification analysis (RQA) metrics. 
+movement variability is still an open problem, particularly 
+when methods in time domain, frequency domain or nonlinear dynamics 
+can break down due to the real-world time-series datasets. 
+For this work, we therefore investigate nonlinear dynamics methods 
+such as reconstructed state space (RSSs), uniform time-delay embedding, 
+recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
+with real-world time-series data.
 Particularly, we are interested in the weaknesses and robustness 
-of nonlinear dynamics tools from raw and post-processed data of 
-wearable inertial sensors (IMUs). 
-In the reported experiment, twenty right-handed healthy participants 
+of nonlinear dynamics tools with the use of raw and post-processed 
+data of wearable inertial sensors. 
+That said, twenty right-handed healthy participants 
 imitated simple vertical and horizontal arm movements in normal 
-and faster speed from an humanoid robot.
-With four window lengths and three levels of smoothed time series, 
-we found visual differences in the patterns of RSSs and RPs. 
-Then with the use of RQA metrics, we find out that the type of 
-movements and the level of smoothness affects those metrics. 
-Specifically, entropy values from RQA were well distributed and 
-presented variation in all the conditions for time series. 
-This work have then the potential to enhance the development of 
-better diagnostic tools for various applications in rehabilitation, 
-sport science or for new forms of human-robot interaction.
+and faster velocity from an humanoid robot in order to have 
+four window lengths and three levels of smoothed time-series data,
+to then found visual differences in the patterns with RSSs and RPs
+and particulatly the computed differences with RQA metrics
+that help us to quantify activities, types of sensors, windows lenghts 
+and level of smoothness. 
+Specifically, we can conclude that RQA ENTR, Shannon Entropy, 
+can lead to interesting results on the quantification of 
+movement variablity for participants of different ages, 
+state of health and anthropomorphic features to 
+then enhance the development of 
+better diagnostic tools for applications in rehabilitation, 
+sport science or for new forms of human-humanoid interaction.
 \end{abstract}
-%226
 
@@
 joints and limbs are involved for a specific task in a determined environment
-but also perception and action of movement affects the physical performance 
-\cite{davids2003}. 
+but also perception and action of movement that affects such 
+physical performance \cite{davids2003}. 
 In contrast, variability in humanoid movement is usually very small,
-as a result of the control system used. This means that, while humanoids 
+as a result of mechanical and dynamic capabilities \cite{gouaillier2009}. 
+This means that, while humanoids 
 solve the degrees of freedom problem through join design and algorithms,
 humans tend to have more fluid and flexible approach. Consequently, 
-one can see much variability in the performance of even the simplest task.
+one can see much variability in human movement performance of even the simplest task.
 Studies of human motion reveal the possibility to 
@@
 For instance, Bradley et al. 2015 \cite{bradley2015} reviewed methods for
-nonlinear time series analysis based on the appropriate estimation of the 
+nonlinear time series analysis based on the estimation of the 
 embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
@@
 state space \cite{takens1981}.
-Bradley et al. 2015 \cite{bradley2015} also reviewed the use of 
+Similarly, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
 Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
 which show black and white dots as recurrences in a given $n$-dimensional system, 
-and Recurrence Quantification Analysis (RQA) metrics compute statistics in RPs.
+and Recurrence Quantification Analysis (RQA) metrics that compute statistics in RPs.
 In general, RPs and RQA provide an intuitive meaning of the time series,
@@
 (e.g., computing dynamical invariants).
-Additionally, the methodologies of nonlinear dynamics for 
-computing the embedding parameters e.g., autocorrelation, mutual information, 
+Additionally, methods of nonlinear dynamics for 
+computing embedding parameters e.g., autocorrelation, mutual information, 
 and nearest neighbour require data which is well 
 sampled and with little noise \cite{garland2016}
-or require purely deterministic signals \cite{kantz2003}.
-Thus, these nonlinear analysis methods for computing the embedding 
+or require signals that are purely deterministic \cite{kantz2003}.
+Thus, these methods of nonlinear analysis for computing the embedding 
 parameters can break down
@@
 acquisition \cite{frank2010}),
-and data may be contaminated with different sources of noise
+and contaminated data with different sources of noise
 \cite{garland2016}.
-It is surprising that even with the previous constraints with regard to 
+It is then surprising that even with the previous constraints with regard to 
 the quality of data, and the problem with the estimation of embedding parameters,
 the use of nonlinear dynamics have proven to be helpful to understand and 
-characterise dynamics of time series 
+to characterise dynamics of time series 
 \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
@@
 One of the challenges of the implementation of 
-Uniform Time-Delay Embedding to Reconstructed State Spaces is the 
+uniform time-delay embedding to reconstruct the state spaces is the 
 selection of embedding parameters because each time series is unique in 
-terms of its structure (modulation of amplitude, frequency and phase) 
+terms of its structure (e.g. modulation of amplitude, frequency and phase) 
 \cite{ frank2010, sama2013, bradley2015}.
@@
 embedding parameters for each of the time series but to deal with a 
-selecting of two parameters that can represent all the time series. 
-Our solution for that problem was to compute a sample mean over all 
+selection of two parameters that can represent all the time series. 
+Our solution for such problem, as explained in below sections, 
+was to compute a sample mean over all 
 values in each of the conditions of the time series 
@@
 %Figs~\ref{fig:caoami} 
-for minimum delay values as follows.
+for minimum delay values.
 %, resulting in an average minimum embedding parameters 
@@
 Minimum embedding parameters were firstly computed 
-with False Nearest Neighbour and Average Mutual Information.
+with the methods of False Nearest Neighbour and Average Mutual Information.
 %for Uniform Time-Delay Embedding (UTDE),
@@
 %Recurrence Quantification Analysis (RQAs).
-Figures \ref{fig:caoami}(A) show that minimum embedding values 
-for sensor HS01 appear to show more variations as 
-interquartile range is near to 1 
-with three exceptions (HF sg0, VN sg0, and VF sg1).
-Minimum embedding values appear to be constant for sensor RS01 
-as interquartile range is near to 0.1 with the exception of 
-two axis (HF sg0 and VF sg1).
-Additionally, it can be seen in Figs \ref{fig:caoami}(A) a 
-decrease of mean values (rhombus) in the box plots
-as smoothness of time series increase.
-Similarly, the first minimum values of AMI values are 
-shown in the box plots of Figs \ref{fig:caoami}(B).
+Figures \ref{fig:cao_ami}(A) show box plots for the minimum embedding values 
+for sensors HS01 and RS01. Minimum embedding vaues for HS01 presents more 
+variations than RS01 as interquartile range for HS01 is near to 1 (with the exceptions of HF sg0, VN sg0, and VF sg1)
+whereas interquartile range for RS01 is near to 0.1 (with the exception of two axis HF sg0 and VF sg1).
+Additionally, Figures \ref{fig:cao_ami}(A) show
+a decrease of mean values (rhombus) in the box plots
+as smoothness of time series increase (see sg0, sg1, sg2).
+Figures \ref{fig:cao_ami}(B) show box plots for the first minimum values of AMI values.
 It can be seen that values for HS01 tend to be more spread as the smoothness 
@@
 the increase of smoothness except for HF and VF
-(Figs \ref{fig:caoami}(B)).
-
-We also computed an overall minimum embedding parameters that represent 
-participants, activities, sensors and levels of smoothness, using a sample 
-mean of all values in Figs~\ref{fig:caoami}.
-The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:caoami}(A) 
+(Figs \ref{fig:cao_ami}(B)).
+
+We also computed a sample mean for an overall minimum embedding parameters that 
+represent all participants, activities, sensors and levels of smoothness.
+The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:cao_ami}(A) 
 is $\overline{m}_0=6$ 
-and the sample mean for minimum values of AMIs from Figs~\ref{fig:caoami}(B)
+and the sample mean for minimum values of AMIs from Figs~\ref{fig:cao_ami}(B)
 is $\overline{\tau}_0=8$.
@@
 	\caption{
-	{\bf Box plots of minimum embedding parameters for horizontal and vertical arm movements.} 
+	{\bf Box plots of minimum embedding parameters.} 
 	Box plots of (A) minimum embedding dimensions 
@@
 	sensor attached to robot (RS01).
-	Minimum embedding dimensions are for twenty participants 
+	Minimum embedding parameters are for twenty participants 
 	($p01$ to $p20$) with three smoothed signals 
@@
 	and window length of 10-sec (500 samples).
-	Code and data to reproduce the figure is available from \cite{srep2019}.
-        }
-    \label{fig:caoami}
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:cao_ami}
 \end{figure}
@@
 \subsubsection*{Uniform Time-Delay Embedding}
-Uniform time-delay embedding were computed with
+Uniform time-delay embedding were computed with the overall 
 embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
 the first three axis of the rotated data of the PCA are shown 
-for the reconstructed state spaces in 
-Figs~\ref{fig:rss_aHw10} for horizontal arm movements and 
-Figs~\ref{fig:rss_aVw10} for vertical arm movements.
+for the reconstructed state spaces of horizontal arm movements
+(Figs~\ref{fig:rss_aHw10}) and vertical arm movements 
+(Figs~\ref{fig:rss_aVw10}).
 %%---------------------------------(FIGURE)-------------------------------------
@@
 	Reconstructed state spaces were computed with 
-	embedding parameters $m=6$, $\tau=8$.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	embedding parameters 
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 	Reconstructed state spaces were computed with 
-	embedding parameters $m=6$, $\tau=8$.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	embedding parameters
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 
-It is easy to observe by eye the differences in each of the
-trajectories in the reconstructed state spaces 
-(Figs~\ref{fig:rss_aHw10}, \ref{fig:rss_aVw10}), 
-however one might be not objective when quantifying those differences 
-since those observation might vary from person to person.
-With that in mind, we tried to objectively quantify those differences 
-using euclidean distances between the origin to each of the points in the 
-trajectories, however these results created a suspicious metric, specially 
-for trajectories which looked very messy.
-With that in mind, we computed Recurrence Plots and 
-Recurrence Quantification Analysis to objectively quantify 
-the differences in each of the cases of the time series.
+One can observe by eye the differences in each of the
+trajectories of the reconstructed state spaces 
+(Figs~\ref{fig:rss_aHw10} and \ref{fig:rss_aVw10}), 
+however an objective quantification is required.
+That being said, we tried to quantify those differences 
+using euclidean distances between the origin 
+to each of the points in their trajectories, 
+which results did not capture the dynamics of 
+the trajectories specially 
+for those that looked very messy.
+Such ambiguities lead us to consider the use of 
+Recurrence Plots and Recurrence Quantification Analysis 
+to objectively quantify dynamics the capture activities 
+using time series.
 
@@
 vertical arm movements (Fig~\ref{fig:rp_aV}) 
-using the average embedding parameters ($m=6$, $\tau=8$) 
+using the average embedding parameters 
+($\overline{m}_0=6$, $\overline{\tau}_0=8$) 
 and an recurrence threshold of $\epsilon=1$.
-
 %%---------------------------------(FIGURE)-------------------------------------
@@
 	Recurrence plots were computed with 
-	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	embedding parameters 
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 	Recurrence plots were computed with 
-	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 \subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
-%Considering the RPs for 20 participants performing four activities 
-%(HN, HF, VN and VF) with sensors attached to the human and 
-%to the humanoid robot (HS01, RS01) and the increase of smoothness 
-%of time series (sg0zmuvGyroZ, sg1zmuvGyroZ and sg2zmuvGyroZ),
-Similarly to RP, we computed four RQA metrics (REC, DET, RATIO and ENTR) 
-with embedding parameters $m=6$, $\tau=8$ and 
-recurrence threshold $\epsilon=1$.
+Four Recurrence Quantification Analysis (RQA) metrics (REC, DET, RATIO and ENTR) 
+were also computed with $\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
 
 \subsubsection*{REC values}
-It can be seen in the box plots of Figs~\ref{fig:RQABP}(A) that REC values, 
-representing the \% of black dots in the RPs, 
-are more spread for HN and VN movements (higher interquartile range) 
-than HF and VF movements (lower interquartile range) for HS01 sensor. 
+Figs~\ref{fig:RQABP}(A) show box plots of REC values 
+that represent the \% of black dots in RPs.
+It can be noted that REC values are more spread 
+for HN and VN movements (higher interquartile range) than 
+for HF and VF movements (lower interquartile range) for HS01 sensor. 
 In contrast, REC values for RS01 sensor present little variation 
 (interquartile range of 0.01).
-With regard to the increase of smoothness of time series 
+Regarding the increase of smoothness for time series 
 (sg0, sg1 and sg2), REC values present little 
 variation as the smoothness is increasing for time series from HS01 
-(changes of mean values (rhombus)) while REC values are more affected with 
-the smoothness for data from RS01 
+(changes of mean values (rhombus)) whereas REC values are 
+more affected with the smoothness for data from RS01 
 (see the incremental changes of mean values (rhombus)).
@@
 \subsubsection*{DET values}
-Figs \ref{fig:RQABP}(B) illustrate DET values, 
-representing predictability and organisation of the RPs, 
-which change very little (interquartile range is around 0.1) 
-for type of movement, level of smoothness or type of sensor.
-There is also increase of DET values as the smoothness of the signal increase 
+Figs \ref{fig:RQABP}(B) illustrate DET values 
+that represent the predictability and organisation of RPs.
+Generally, it can be noted little change of DET values 
+(interquartile range is around 0.1) 
+for type of movement, type of sensor 
+but the increase of DET values as the smoothness of 
+the signal increase 
 (see the incremental changes of mean values (rhombus)).
-It can be noted that the interquartile range for faster movements
+However, the interquartile range for faster movements
 (HF and VF) with no smoothing (sg0) is lower than the other
@@
 \subsubsection*{RATIO values}
-Figs \ref{fig:RQABP}(C) present RATIO values, representing dynamic transitions, 
-for horizontal and vertical movements.
-It can be seen that RATIO values for HS01 sensor vary less 
+Figs \ref{fig:RQABP}(C) illustrate RATIO values that represent the 
+dynamic transitions in RPs. 
+Generally, it can be seen that RATIO values for HS01 sensor vary less 
 for HN movements (interquartile range around 2)
 than HF movements (interquartile range around 5).
-%which is a similar behaviour of RATIO values for oRS01 sensor.
 For faster movements (HF and VF), it can be noted a decrease of 
-RATIO values as the smoothness of the time series is increasing (rhombus).
-For normal movements (HN and VN), the decrease of RATIO values
-is less evident than faster movements (see rhombus).
+the mean values the smoothness of the time series 
+is increasing (rhombus). However, for normal movements (HN and VN), 
+the decrease of mean values is less evident than 
+faster movements (rhombus).
 %See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
@@
 \subsubsection*{ENTR values}
-Figs \ref{fig:RQABP}(D) show ENTR values, Shannon entropy values 
-which represent the complexity of 
-the structure the time series, 
-for type of movement, level of smoothness or type of sensor.
+Figs \ref{fig:RQABP}(D) show ENTR values that represent 
+Shannon entropy values in RPs. 
+Generally, 
 ENTR values for HS01 sensor show a higher variation  
 (interquartile range around 0.5)
-than ENTR values for RS01 sensor which appear 
-to be more constant (interquartile range 0.1).
-It can also be noted the increase of smoothness of time series 
-also is presented with an increase of ENTR values (see gray rhombus).
+than ENTR values for RS01 sensor 
+(interquartile range 0.1).
+It can also be noted the increase of mean values 
+for ENTR values as the smoothness of time series 
+increase (rhombus) for type of movement, 
+type of sensor and level of smoothness.  
 %See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa-bp}
+\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-bp}
     \caption
 	[Box plots for RQA values]{
-	{\bf Box plots for RQA values.}
-	Box plots of (A) REC, (B) DET, (C) RATIO, and (D) ENTR values 
-	for 20 participants performing HN, HF, VN and VF movements
+	{\bf Box plots for RQA metrics.}
+	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of 
+	20 participants performing HN, HF, VN and VF movements
 	with sensors HS01, RS01 and three smoothed-normalised  
@@
 	RQA values were computed with 
-	embedding parameters $m=6$, $\tau=8$ and $\epsilon=1$.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$. 
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 
+%\subsection*{RQA metrics with fixed parameters}
+%Considering that RQA metrics were computed with fixed embedding parameters 
+%($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
+%the following. REC values, which represents the \% of black points in the RPs, 
+%were more affected with and increase in normal speed movements (HN and VN) 
+%than faster movements (HF and VF) for the sensor attached to the participants 
+%(HS01). Such decrease of REC values from normal speed to faster speed 
+%movements is also presented in data from sensor attached to the robot (RS01), 
+%and little can be said with regard to the dynamics of the time series coming 
+%from RS01 (Fig \ref{fig:RQABP}A).
+%Similarly, DET values, representing predictability and 
+%organisation in the RPs, present little variation in the any of the time 
+%series where little can be said (Fig \ref{fig:RQABP}B).
+%In contrast, RATIO values, which represent 
+%dynamic transitions, were more variable for faster movements (HF and VF) 
+%than normal speed movements (HN and VN) with sensors attached to the 
+%participants (HS01). For data coming from sensors attached to the robot 
+%(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
+%more than values coming from vertical movmentes (VN, VF) 
+%(Fig \ref{fig:RQABP}C).
+%With that, it can be said that RATIO values can represent a bit better
+%than REC or DET metrics for the variability of imitation activities in 
+%each of the conditions for time series.
+%Similarly, ENTR values for HN were higher than values for HF
+%and ENTR values varied more for sensor attached to participants 
+%than ENTR values for sensors of the robot. It is evidently that 
+%the higher the entropy the more complex the dynamics are, 
+%however, ENTR values for HN appear a bit higher than HF values, 
+%for which we believe this happens because of the structure the time series
+%which appear more complex for HN than  HF movements which presented a 
+%more consistence repetition (Fig \ref{fig:RQABP}D).
+
 \subsection*{3D RQA ENTR}
-Choosing an appropriate recurrence threshold is crucial to get 
-meaningful representations in RP and RQA, however, as previously shown, 
-one can selected the default recurrence threshold of $\epsilon=1$, 
-as long as it is able to represent the dynamical transitions 
-in each of the time series \cite{marwan2011}. %%Marwan et al. \cite{marwan2011} 
-However, to show how that the selection of recurrence threshold affects RQA
+To show how that the selection of recurrence threshold affects RQA
 values, 
 we computed ENTR values of RQA metrics for different embedding parameters
-$[1,10]= \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
-$[1,10]= \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $
-with an increase of 1
-and different recurrence thresholds 
-$[0.2,3]= \{ \epsilon \in \mathbb{R} | 0.2 \le \epsilon \le 3.0  \} $
-with increments of 0.1.
-For easily visualisation, Figs \ref{fig:RQA-IND} only show 3D RQA for 
-three recurrence thresholds $\epsilon=1, 2, 3$ with three levels of 
-smoothness (sg0, sg1, sg2). It can be noted in Figs \ref{fig:RQA-IND} 
-that the increase of recurrence threshold is associated to the increase 
-of ENTR values in any of the levels of smoothness (see values of the ENTR bar).
-One can also see that the maximum values of ENTR 
-for $\epsilon=1$ and sg0 are for embedding dimensions of 2 
-and then as $\epsilon$ increases the maximum values of ENTR are for
-an associated increase of in the embedding dimension.
-It can also be noted that the increase of level of smoothness (from sg0 to sg2) 
-is associated with both the increase of ENTR values and its smoothed 3D shape.
-
+$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
+$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by one,
+with the consideration of three recurrence thresholds $\epsilon=1, 2, 3$ 
+and three levels of smoothness (sg0, sg1, sg2). 
+For instance, 
+Figures \ref{fig:RQA-IND} show the increase of recurrence threshold is 
+associated to the increase of ENTR values in any of the levels 
+of smoothness (see values of the ENTR bar).
+Similarly, it can be noted that the increase of level of 
+smoothness (sg0, sg1 and sg2) is associated with the increase 
+of ENTR values of the 3D surface plots.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa_epsilons}
+\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-epsilons}
     \caption
-	[3D RQA ENTR values]{
-	{\bf 3D RQA ENTR values.}
-	RQA ENTR values 
-	for different embedding parameters
-	$[1,10]= \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
-	$[1,10]= \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
-	with an increase of 1
-	and different recurrence thresholds $\epsilon=1, 2, 3$.
-	RQA ENTR values are for $p03$, sensor HS01, of a window size of 10-secs (500 samples).
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	[3D surface plots of RQA ENTR values]{
+	{\bf 
+	3D surface plots of RQA ENTR values for different recurrence threshold and smoothness levels.}
+	RQA ENTR values are
+	for embedding parameters
+	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
+	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
+	incrementing by one and three recurrence thresholds $\epsilon=1, 2, 3$.
+	RQA ENTR values were computed with data from $p03$, sensor HS01, with 
+	a window size of 10-secs (500 samples).
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 %%---------------------------------(FIGURE)------------------------------------
-
-%of ENTR and plotted 3D surfaces using an unitary 
-%and a decimal increase of 0.1 for recurrence thresholds 
-% (Fig.~\ref{fig:topo_rqas}). 
-%Hence, Fig.~\ref{fig:topo_rqas}(A) shows an increase for 
-%REC values, the percentage of black docs in the RP, 
-%as the recurrence threshold increase,
-%while the variation for embedding parameters creates little decrease 
-%of REC values as the embedding dimensions.
-% increase and even a more slighter 
-%decrements of REC values for the increase of $\tau$.
-%For the 3D surface of DET values (Fig.~\ref{fig:topo_rqas}(B)), 
-%representing 
-%predictability and organisation of the RPs, one can be noted a plateau
-%for DET values near to 1 for embedding dimension parameters of less 
-%than 5 and recurrence threshold values of greater than 2 (red surface). 
-%It can also be observed that the increases of delay embedding made 
-%the DET values increase so as to make an cascade effect in the surface 
-%along with the increase of dimension embedding $m$.
-%For RATIO values, representing dynamic transitions, 
-%Fig.~\ref{fig:topo_rqas} shows that the 3D surface present 
-%a plateau (blue surface) of RATIO values 
-%near to zero for recurrence thresholds greater than 1.0, while 
-%fluctuations are more evident for recurrence thresholds of less than 1.0,
-%particularly it can also be noted an increase in the fluctuations of 
-%RATIO values as the embedding dimension is increasing. 
-%For ENTR values in Fig.~\ref{fig:topo_rqas}(D), 
-%representing the complexity of the structures in time series, 
-%one can be noted that the increase of recurrence threshold is, 
-%not strictly proportional to the increase of ENTR values. 
-%It can also be observed in Fig.~\ref{fig:topo_rqas}(D)
-%that the increase of delay embeddings affects 
-%little the ENTR values for embedding dimensions of 1, while 
-%for higher values of embedding dimensions 
-%there is a decrease of ENTR values, and 
-%there is a decrease of ENTR values as delay dimension value is increasing.
-%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}
-%\centering
-%\includegraphics[width=1.0\textwidth]{rqas}
-%    \caption
-%	[3D surfaces for RQA metrics]{
-%	{\bf 3D surfaces for RQA metrics.}
-%	3D surfaces for (A) REC, (B) DET, (C) RATIO and (D) ENTR values 
-%	with increasing pair of embedding parameters 
-%	($0 \ge m \le 10$, $0 \ge \tau \le 10$) 
-%	and recurrence thresholds (  $ 0.2 \ge \epsilon \le 3 $).
-%	RQA metrics are computed with the time series of participant $p01$ using 
-%	HS01 sensor, HN activity, sg0zmuvGyroZ axis and 500 samples 
-%	for window size length.
-%	Code and data to reproduce the figure is available from \cite{srep2019}.
-%	}
-%\label{fig:topo_rqas}
-%\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-%
-
-\subsection*{Other effects on 3D RQA ENTR values}
-
-%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
-%determining embedding parameters, their method consisted on creating 3D 
-%surfaces with RQA metrics with an increase of embedding parameters 
-%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
-%fluctuations and gradual changes in the 3D surfaces that provide information 
-%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
-%created 3D surfaces for visual selection of not only embedding parameters 
-%but also recurrence thresholds. Following same methodologies, we explored 
-%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
-%using 3D surfaces by an unitary increase of the pair embedding 
-%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
-%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
-%(Fig.~\ref{fig:topo_rqas}). 
-We also computed 3D surfaces of RQA metrics for different sensors 
-and different activities (Fig~\ref{fig:3dRQAENTR_sensoractivities}).
-From the maximum values of ENTR (lateral bars), one can see 
-a decrease of ENTR values for activities going from normal to 
-faster velocity and from human sensor (HS01) to robot sensor (RS01) 
-(See Fig~\ref{fig:3dRQAENTR_sensoractivities}). 
+Figures~\ref{fig:3dRQAENTR_sensoractivities} show 3D surface plots of 
+ENTR values for different sensors and different activities.
+Figs~\ref{fig:3dRQAENTR_sensoractivities} show that RQA ENTR (lateral bars) 
+decrease from normal (HN, VN) to faster (HF, VF) velocities from both 
+human sensor (HS01) to robot sensor (RS01).
+Also, it can be noted that RQA ENTR values decrease from human sensor (HS01) 
+to robot sensor (RS01).
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/pdf/rqa_sensors_activities}
+\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
     \caption{
-	{\bf 3D surfaces of RQA ENTR metrics for sensors and activities.}
-	RQA ENTR values with embedding parameters
-	$[1,10]= \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
-	$[1,10]= \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
-	for different activities and sensors. 
-	RQA ENTR values are for $p03$, sg0 and window size of 10-secs (500 samples).
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	{\bf 3D surface plots of RQA ENTR values for different sensors and activities.}
+	RQA ENTR values are for embedding parameters
+	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
+	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
+	with $\epsilon = 1 $ considering four activities 
+	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
+	Vertical Faster (VF) and sensors Human Sensor 01 (HS01) and 
+	Robot Sensor (RS01).
+	RQA ENTR values were computed from data of $p03$, sg0 and 
+	window size of 10-secs (500 samples).
+	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
@@
 %%---------------------------------(FIGURE)-------------------------------------
-
-%RQA metrics are also affected by 
-%the window length where for example four window lengths of 100, 250, 500 
-%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}).
-
-%Similarly, 3D surfaces of RQA metrics were also computed for three 
-%participants (Fig.~\ref{fig:topo_participants}).
-
-%X Three level of 
-%X smoothness were computed for RQA metrics showing smoothed 3D surfaces and 
-%X the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
+Figures~\ref{fig:3dRQAENTR_participantsactivities} show the similarity of 3D surface plots in 
+relationship with three participants (p01, p02 and p03)
+performing four activities (HN, HF, VN, and VF) and two sensors (HS01, RS01).
+Figures~\ref{fig:3dRQAENTR_participantsactivities}(A) 
+present subtle differences for each of the participants (see RQA ENTR bar). 
+Figures~\ref{fig:3dRQAENTR_participantsactivities}(B) show 
+less change of RQA ENTR values for each participant as the data is 
+from a sensor attached to the robot. 
+Also, from data of sensors attached to the robot, it can be noted 
+that changes are more notable for faster movements than normal movements. 
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.9\textwidth]{figures/rqa/output/rqa-participants}
+    \caption{
+	{\bf 3D surface plots of RQA ENTR values for different participants, activities and sensors.}
+	RQA ENTR values are for participants (p01, p02, and p03) 
+	in the categories of 
+	(A) Human Sensor 01 (HS01) and 
+	(B) Robot Sensor 01 (RS01)
+	considering embedding parameters
+	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
+	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
+	with $\epsilon = 1$ and four activities 
+	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
+	Vertical Faster (VF).
+	RQA ENTR values were computed from data of sg0 and window size of 10-secs (500 samples).
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+       }
+\label{fig:3dRQAENTR_participantsactivities}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+Figures \ref{fig:3dRQAENTR_windowsactivities} show the effect of the increase 
+of window length of the time series (w100 (2-sec), w250 (5-sec), w500 (10-sec) and w750 (15-sec))
+on the 3D surface plots. 
+It can be noted that the increase of number of samples improves the quality 
+of the surface plots where 100 samples poorly capture the dynamics of each activity.
+However as the window length increases from 250, 500 to 750 samples, 
+3D surface plots show a similar representation of RQA ENTR values 
+(being the surface plot with 750 samples the best test case).
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-windows}
+    \caption{
+	{\bf 3D surface plots of RQA ENTR values for different windows lengths and activities.}
+	RQA ENTR values are for embedding parameters
+	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
+	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$, 
+	with $\epsilon = 1 $ considering four 
+	windows lengths (e.g., w100 (100 samples), w250 (250 samples),
+	w500 (500 samples) and w750 (750 samples)) and
+	four activities 
+	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
+	Vertical Faster (VF).
+	RQA ENTR values were computed from data of $p01$ and sg0.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+       }
+\label{fig:3dRQAENTR_windowsactivities}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+To summarise this section of results, it can be said that computing
+embedding parameters for individual structure of time-series 
+data is already a solved problem \cite{frank2010, sama2013, bradley2015}. 
+However, it has been shown the challenge of finding embedding parameters 
+for nonlinear dynamic tools that represent a set of different time-series data.
+That said, we proposed the use of sample mean of the set of embedding parameters
+for RSSs, RP and RQA to then noticed that the selection of recurrence 
+threshold, $\epsilon$, is also an open problem.
+For which, this work proposed the variation of recurrence thresholds 
+and embedding parameters to show the relationships of these to different datasets 
+(participants, activities, windows lengths and sensors).
 
@@
 \section*{Discussion}
-%The Discussion should be succinct and must not contain subheadings.
-
-%DISCUSSION
-%Similarly as with the Reconstructed State Spaces, 
-%the differences in the RPs can be easily noticed by eye 
-%for different time series (Figs~\ref{fig:rp_aV} and ~\ref{fig:rp_aH}), 
-%which lead us to apply Recurrence Quantification Analysis (RQA) 
-%in order to have an objective quantification of each of the time series.
-%That said, 
-%%Considering the time series of Fig~\ref{fig:ts}, 
-%
-
-It is evident that time series from different sources 
-(participants, movements, axis type, window length or levels of smoothness) 
-presents visual differences for embedding parameters and therefore for RRSs. 
-For which, the selection of embedding parameters was our first challenge 
-where we computed embedding parameters for each time series 
-(Fig \label{fig:caoami}) and then computed a sample mean over 
-all time series in order to get two embedding parameters 
-to compute all RRSs (Figs \label{fig:rss_aHw10} and \label{fig:rss_aVw10}). 
-%with its corresponded type of movement. 
-Then we found that the quantification of variability with regard to 
-the shape of the trajectories in RSSs requires more investigation 
-since our original proposed method base on euclidean metric failed 
-to quantify those trajectories. Specially, for trajectories which 
-were not well unfolded. 
-With that in mind, we apply 
-Recurrence Quantification Analysis (RQA) metrics 
-(REC, DET, RATIO and ENTR) in order to avoid any subjective interpretations 
-or personal bias with regard to the representation of the trajectories in RSSs.
-
-\subsection*{RQA metrics with fixed parameters}
-Considering that RQA metrics were computed with fixed embedding parameters 
-($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
-the following. REC values, which represents the \% of black points in the RPs, 
-were more affected with and increase in normal speed movements (HN and VN) 
-than faster movements (HF and VF) for the sensor attached to the participants 
-(HS01). Such decrease of REC values from normal speed to faster speed 
-movements is also presented in data from sensor attached to the robot (RS01), 
-and little can be said with regard to the dynamics of the time series coming 
-from RS01 (Fig \ref{fig:RQABP}A).
-Similarly, DET values, representing predictability and 
-organisation in the RPs, present little variation in the any of the time 
-series where little can be said (Fig \ref{fig:RQABP}B).
-In contrast, RATIO values, which represent 
-dynamic transitions, were more variable for faster movements (HF and VF) 
-than normal speed movements (HN and VN) with sensors attached to the 
-participants (HS01). For data coming from sensors attached to the robot 
-(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
-more than values coming from vertical movmentes (VN, VF) 
-(Fig \ref{fig:RQABP}C).
-With that, it can be said that RATIO values can represent a bit better
-than REC or DET metrics for the variability of imitation activities in 
-each of the conditions for time series.
-Similarly, ENTR values for HN were higher than values for HF
-and ENTR values varied more for sensor attached to participants 
-than ENTR values for sensors of the robot. It is evidently that 
-the higher the entropy the more complex the dynamics are, 
-however, ENTR values for HN appear a bit higher than HF values, 
-for which we believe this happens because of the structure the time series
-which appear more complex for HN than  HF movements which presented a 
-more consistence repetition (Fig \ref{fig:RQABP}D).
-
-Additionally, we observed that some RQA metrics are affected by the 
-smoothness of data. 
-%For which, we also explored the effect of smoothness of raw-normalised data 
-%where, 
-For instance, REC and DET values were not completely affected by the 
-smoothness of time-series since these RQA values seemed to be constants. 
-However, for RATIO values, 
-the effect of smoothness can be noticed with a slight decrease of amplitude 
-in any of the time series conditions which is also presented with ENTR values.
-
-\subsection*{RQA metrics with different parameters}
-Iwansky et al. \cite{iwanski1998} stated that patterns in RPs and 
-metrics for RQA are independent of embedding dimension parameters, 
-however, that is not the case when using 
-different recurrence thresholds. Such changes of recurrence threshold values 
-can modify the patterns in RPs and therefore the values of RQA metrics.
-We therefore computed 3D surfaces to explore the sensibility and robustness of 
-embedding parameters and recurrence threshold in RQA  metrics. Following the 
-same methodology of computing 3D surfaces, we also considered variation of 
-window length size to present RQA metrics dependencies with embedding 
-parameters, recurrence thresholds and window length size.
+Time series from different sources  
+(e.g., participants, movements, axis type) as well as
+different characteristics (e.g., sample rate, window length or levels of smoothness) 
+result in different embedding parameters and therefore different 
+values for recurrence plots and recurrence quantification analysis metrics.
+That said, while embedding parameters for individual
+time series were successfully computed, the quantification 
+of variability with regard to the shape of the trajectories in 
+reconstructed state spaces require more investigation
+as our initial, approach based on euclidean distances on the trajectories, 
+failed to quantify trajectories which were not well unfolded. 
+To then found out that Recurrence Plots along with 
+Recurrence Quantification Analysis (RQA) metrics (e.g., REC, DET, RATIO and ENTR) 
+and its variation of embedding parameters and recurrence thresholds 
+helped to quantify the variability of different sources of time series.
+However, further investigation is required to be done in order to 
+have better intuition and meaningful interpretation of nonlinear 
+analysis such as to find a right 
+balance among (i) the level of smoothness of the signal, 
+(ii) the selection of recurrence thresholds and (iii) 
+the range of embedding parameters.
 
 \section*{Conclusions}
-Generally, we can conclude that using a different level of 
-smoothness for time series help us to visualise and to quantify 
-the variation of movements 
-between participants using RSSs, RPs and RQA. 
-It is important to mention that some RQA's metrics 
-(e.g. DET and ENTR) are more robust to the effect of 
-smoothness of time series.
-Choice of RQA metrics will be depend on 
-what to quantify, for instance, one can compute 
-predicability, organisation,  
-dynamics transitions, or complexity and determinism of the 
-input time series. However, RQA metrics show certain constrains with 
-regard to activity type, window length and structure of the time series.
-For instance, RATIO and ENTR are helpful to distinguish 
-differences in any of the categories of the time series (sensor, activity, 
-level of smoothness and number of participant), however time series
-data from the sensor attached to the robot seemed to have 
-little variations between as humanoid degrees of freedom 
-did not allow it to move with a wide range of variability.
-
-Additionally, we can point out that even though our experiment is 
-limited to twenty healthy right-handed participants of a range age 
-of mean 19.8 SD=1.39, RQA metrics show the potential to quantify 
-human movement variability in the context of 
-human-humanoid imitation activities.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%	You need a sentence or two here
-%	to explain why this is claimed.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%With that in mind, we conclude that 
-%quantification of human-humanoid imitation activities is possible for 
-%participants of different ages, state of health and anthropomorphic features.
-
-%Such understanding and measurement of movement variability using
-%cheap wearable inertial sensors lead us to have a more intuitive selection of parameters
-%to reconstruct the state spaces and to create meaningful interpretations
-%of the recurrence plots and the results of the metrics with recurrence quantification 
-%analysis. 
-%
-
-%
-%However, we believe that further investigation is required to find the 
-%right balance between the level of smoothness of the signal
-%and defining what is the aim of RQA where 
-%its representations using RSS, RP and RQA.
-%Specially, where the level of smoothness does not affect 
-%the variation of each of the movements quantification. 
-%%We can also conclude that finding the right balance between 
-%%smoothness and the raw data to capture movement variability is 
-%%a still a problem that has many avenues for exploration.
-%
-
-\section*{Future Work}
-
-\subsection*{Inertial Sensors}
-To better understand the signals collected through 
-inertial sensors in the context of human-humanoid interaction, 
-we could apply derivates to the acceleration data. 
-We can then explore the jerkiness of movements and therefore 
-the nature of arm movements which typically have 
-minimum jerk \cite{flash1985}, its relationship with different body 
-parts \cite{devries1982, mori2012} or the application of higher derivatives 
-of displacement with respect time such as snap, crackle and pop \cite{eager2016}.
-
-\subsection*{RQA ENTR}
-Having shown that RQA ENTR metrics are robust against different data
-time series, we believe that further investigation is required. 
-For example, Marwan et al. \cite{marwan2007, marwan2015} reviewed 
-different aspects to compute RPs using different criteria for neighbours, 
-different norms ($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or 
-different methods to select the recurrence threshold $\epsilon$, 
-such as using certain percentage of the signal \cite{letellier2006}, 
-the amount of noise or using a factor based on the standard deviation 
-of the observational noise among many others \cite{marwan2007}.
+This work allows to conclude that the choice of 
+nonlinear analysis tool (e.g., RSSs, RPs, RQA metrics) will depend 
+on what one would like to quantify on the time-series 
+data (e.g., predicability, organisation, dynamics transitions, 
+or complexity and determinism).
+Then, time-series data characteristics 
+(e.g., window size length, level of smoothness) plays an important role
+as well on the results that nonlinear analysis tool can provide.
+Similarly, the results of the nonlinear analysis tools
+are associated with the structure of the time-series data 
+(e.g. frequency, amplitude), the position of the sensor 
+and activity performed by either a robot or human being
+(as degrees of freedom from the humanoid are far less than
+human movement). 
+That said, it has been shown that the use of different 
+characteristics of the time-series data 
+(e.g., sensor, activity, level of smoothness and participants)
+has help us to visualise and to quantify 
+with nonlinear tools the variation of 
+movements of, in this work, human-humanoid activities.
+However, some limitation of nonlinear tools are related 
+to the computation of different parameters 
+(e.g., recurrence thresholds, embedding parameters)
+that reflect the dynamics of individual characteristics 
+of activity type, window length and structure of the time series.
+Specifically, the example of DET values which appear 
+to be constants across sensors, activities and levels of smoothness, 
+whereas REC and RATIO, as function of REC, values 
+show variation for certain sensors and movements. 
+To then found out that RQA ENTR values with different recurrence 
+thresholds were appropriate to quantify 
+the different changes and variations of the characteristics of 
+time-series data.
+Therefore, we can conclude that this work provides a good starting 
+point and reference to the use of Shannon Entropy to quantify 
+human-humanoid imitation activities 
+that can then lead to interesting results on the quantification 
+of movement variability of participants with different ages, 
+state of health and anthropomorphic features.
 
@@
    	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
     }
@@
 	for (E) chaotic and (F) random time series.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 	for (C) a chaotic and (D) noise time series.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
 	}
@@
 \cite{marwan2011}.
-
 Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
@@
 context of human-humanoid imitation activities.
+%Such changes of recurrence threshold values 
+%can modify the patterns in RPs and therefore the values of RQA metrics.
+%We therefore computed 3D surfaces to explore the sensibility and robustness of 
+%embedding parameters and recurrence threshold in RQA  metrics. Following the 
+%same methodology of computing 3D surfaces, we also considered variation of 
+%window length size to present RQA metrics dependencies with embedding 
+%parameters, recurrence thresholds and window length size.
+
+%Choosing an appropriate recurrence threshold is crucial so as to get 
+%meaningful representations in RP and RQA, however,
+
+%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
+%metrics for RQA are independent of embedding dimension parameters, 
+%however, that is not the case when using 
+%different recurrence thresholds.
+
+%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
+%determining embedding parameters, their method consisted on creating 3D 
+%surfaces with RQA metrics with an increase of embedding parameters 
+%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
+%fluctuations and gradual changes in the 3D surfaces that provide information 
+%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
+%created 3D surfaces for visual selection of not only embedding parameters 
+%but also recurrence thresholds. Following same methodologies, we explored 
+%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
+%using 3D surfaces by an unitary increase of the pair embedding 
+%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
+%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
+%(Fig.~\ref{fig:topo_rqas}).
 
@@
 		which are also presented in (B), (C) and (D).
-		Code and data to reproduce the figure is available from \cite{srep2019}.
+		Code and data to reproduce the figure is available in \cite{srep2020}.
         }
@@
 		and other sensor attached to the robot (RS01).	
-	Code and data to reproduce the figure is available from \cite{srep2019}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
         }

### 2020-07-02 docs/supplementary-information/report/report.tex
completion of corrections-2020-01-14 and updates manuscript2020-07-03

\documentclass[12pt]{article}
\usepackage[margin=15mm]{geometry}
\usepackage{graphicx}
\graphicspath{{../../}} %%%goes to the path: docs/

\title{
Supplementary Material
}
\author{Miguel Xochicale}
\date{ \today }

\begin{document}
\maketitle

\begin{abstract}
Section 1 shows surface plots of 3D RQA ENTR values for 3 participants (p01, p02 and p03).
Datasets are for 
(a) window size of 100, 250, 500 and 750 samples (w100, w250, w500, w750);
(b) smoothnes: sg0, sg1 and sg2;
(c) movement: Horizontal Normal (HN), Horizontal Faster (HF), Vertical Normal (VN) and Vertical Faster (VF); and 
(d) sensors: Sensor attached to Human (HS01), sensor attached to Robot (RS01).
\end{abstract}

\tableofcontents
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{3D RQA ENTR values}
Location of code, data and figures for 3D RQA ENTR values is shown below 
\begin{verbatim}
## Code
$HOME/srep2020/code/rscripts/G_3Drqa
`> source(  paste( getwd(), '/C_3Drqa_plots_epsilons.R', sep=''), echo=TRUE )`

## Data
$HOME/srep2019/data/rqa$ tree -s

## Figures
$HOME/srep2020/docs/figures/rqa/src/3drqa_epsilons$ tree -s
\end{verbatim}

Datasets for NN participants where NN is 01, 02 and 03 are:
\begin{verbatim}
RQAs_pNNw100.dt
RQAs_pNNw250.dt
RQAs_pNNw500.dt
RQAs_pNNw750.dt
\end{verbatim}

\subsection{Participant 01}
%X Three level of
%X smoothness were computed for RQA metrics showing smoothed 3D surfaces and
%X the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
%After data collection, raw time series were windowed, normalised and smoothed.

Figures \ref{fig-p01-H-w100} and \ref{fig-p01-V-w100} are for a window size of 100 samples.
Figures \ref{fig-p01-H-w250} and \ref{fig-p01-V-w250} are for a window size of 250 samples.
Figures \ref{fig-p01-H-w500} and \ref{fig-p01-V-w500} are for a window size of 500 samples.
Figures \ref{fig-p01-H-w750} and \ref{fig-p01-V-w750} are for a window size of 750 samples.

%we only present 10-sec (500 samples) window length time series for
%three participants (p01, p01 and p03) performing horizontal
%arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig-p01-w100}).

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics[scale=1.0]{figures/rqa/output/epsilons/rqa-epsilonsp01w100Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 01 performing horizontal movements for a window size of 100 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-H-w100}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics[scale=1.0]{figures/rqa/output/epsilons/rqa-epsilonsp01w100Vertical}
    	\caption{
	{\bf RQA-Entr for participant 01 performing vertical movements for a window size of 100 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-V-w100}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp01w250Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 01 performing horizontal movements for a window size of 250 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-H-w250}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp01w250Vertical}
    	\caption{
	{\bf RQA-Entr for participant 01 performing vertical movements for a window size of 250 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-V-w250}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp01w500Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 01 performing horizontal movements for a window size of 500 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-H-w500}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp01w500Vertical}
    	\caption{
	{\bf RQA-Entr for participant 01 performing vertical movements for a window size of 500 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-V-w500}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp01w750Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 01 performing horizontal movements for a window size of 750 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-H-w750}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp01w750Vertical}
    	\caption{
	{\bf RQA-Entr for participant 01 performing vertical movements for a window size of 750 samples.}
		%RQA-Entr surface plots are for participant $p01$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p01-V-w750}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
\subsection{Participant 02}
%X Three level of
%X smoothness were computed for RQA metrics showing smoothed 3D surfaces and
%X the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
%After data collection, raw time series were windowed, normalised and
%smoothed. Then, due to space limitations and to have simple visualisation,
Figures \ref{fig-p02-H-w100} and \ref{fig-p02-V-w100} are for a window size of 100 samples.
Figures \ref{fig-p02-H-w250} and \ref{fig-p02-V-w250} are for a window size of 250 samples.
Figures \ref{fig-p02-H-w500} and \ref{fig-p02-V-w500} are for a window size of 500 samples.
Figures \ref{fig-p02-H-w750} and \ref{fig-p02-V-w750} are for a window size of 750 samples.

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics[scale=1.0]{figures/rqa/output/epsilons/rqa-epsilonsp02w100Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 02 performing horizontal movements for a window size of 100 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-H-w100}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics[scale=1.0]{figures/rqa/output/epsilons/rqa-epsilonsp02w100Vertical}
    	\caption{
	{\bf RQA-Entr for participant 02 performing vertical movements for a window size of 100 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-V-w100}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp02w250Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 02 performing horizontal movements for a window size of 250 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-H-w250}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp02w250Vertical}
    	\caption{
	{\bf RQA-Entr for participant 02 performing vertical movements for a window size of 250 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-V-w250}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp02w500Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 02 performing horizontal movements for a window size of 500 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-H-w500}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp02w500Vertical}
    	\caption{
	{\bf RQA-Entr for participant 02 performing vertical movements for a window size of 500 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-V-w500}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp02w750Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 02 performing horizontal movements for a window size of 750 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-H-w750}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp02w750Vertical}
    	\caption{
	{\bf RQA-Entr for participant 02 performing vertical movements for a window size of 750 samples.}
		%RQA-Entr surface plots are for participant $p02$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p02-V-w750}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
\subsection{Participant 03}
%X Three level of
%X smoothness were computed for RQA metrics showing smoothed 3D surfaces and
%X the level of smoothness increase (Fig.~\ref{fig:topo_smoothness}).
%After data collection, raw time series were windowed, normalised and
%smoothed. Then, due to space limitations and to have simple visualisation,
Figures \ref{fig-p03-H-w100} and \ref{fig-p03-V-w100} are for a window size of 100 samples.
Figures \ref{fig-p03-H-w250} and \ref{fig-p03-V-w250} are for a window size of 250 samples.
Figures \ref{fig-p03-H-w500} and \ref{fig-p03-V-w500} are for a window size of 500 samples.
Figures \ref{fig-p03-H-w750} and \ref{fig-p03-V-w750} are for a window size of 750 samples.

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics[scale=1.0]{figures/rqa/output/epsilons/rqa-epsilonsp03w100Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 03 performing horizontal movements for a window size of 100 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-H-w100}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics[scale=1.0]{figures/rqa/output/epsilons/rqa-epsilonsp03w100Vertical}
    	\caption{
	{\bf RQA-Entr for participant 03 performing vertical movements for a window size of 100 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-V-w100}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp03w250Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 03 performing horizontal movements for a window size of 250 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-H-w250}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp03w250Vertical}
    	\caption{
	{\bf RQA-Entr for participant 03 performing vertical movements for a window size of 250 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-V-w250}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp03w500Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 03 performing horizontal movements for a window size of 500 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-H-w500}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp03w500Vertical}
    	\caption{
	{\bf RQA-Entr for participant 03 performing vertical movements for a window size of 500 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-V-w500}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

\newpage
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[ht!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp03w750Horizontal}
    	\caption{
	{\bf RQA-Entr for participant 03 performing horizontal movements for a window size of 750 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-H-w750}
\end{figure}
%---------------------------------(FIGURE)------------------------------------
%---------------------------------FIGURE)-------------------------------------
\begin{figure}[hb!]
\centering
\includegraphics{figures/rqa/output/epsilons/rqa-epsilonsp03w750Vertical}
    	\caption{
	{\bf RQA-Entr for participant 03 performing vertical movements for a window size of 750 samples.}
		%RQA-Entr surface plots are for participant $p03$
		%for horizontal and vertical arm movements in normal
		%and faster velocity (HN, HF, VN, VF)
		%with the normalised GyroZ or GyroY axis
		%((sg0) raw-normalised (sg0zmuv),
		%(sg1) normalised-smoothed 1 and
		%(sg1) normalised-smoothed 2)
		%and with one sensor attached to the participant (HS01)
		%and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig-p03-V-w750}
\end{figure}
%---------------------------------(FIGURE)------------------------------------

%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\subsection{Activities and sensors}
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Participants}
%%Similarly, 3D surfaces of RQA metrics were also computed for three
%%participants (Fig.~\ref{fig:topo_participants}).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Window length}
%%RQA metrics are also affected by
%%the window length where for example four window lengths of 100, 250, 500
%%and 750 samples (Fig.~\ref{fig:topo_sensoractivities}).
%

\clearpage
\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2020-07-02 code/rscripts/G_3Drqa/0_rqa_test.R

@@ -25,25 +25,13 @@ start.time <- Sys.time()
 
 
 homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
 r_scripts_path <- getwd()
 setwd("../../../")
 github_repo_path <- getwd()
 
-
-##VERSION 
 version <- '00'
-#feature_path <- '/rqa'
 dataset_path <- '/data/dataset'
 
-## Outcomes Data Path
-#outcomes_data_path <- paste(github_path,'/hwum-dataset/data-output', feature_path, '/v', version, sep="")
-
-## Outcomes Plot Path
-#outcomes_plot_path <- paste(github_path,"/phd-thesis/figs/results", feature_path, '/v', version,sep="")
-
-## Data Path
-#data_path <- paste(github_path,'/hwum-dataset/data-output', feature_path, '/v', version, sep="")
 data_path <- paste(github_repo_path, dataset_path, sep="")
 setwd(file.path(data_path))
 
@@ -57,14 +45,13 @@ library(ggplot2)
 library(RColorBrewer)
 
 library(devtools)
-load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+#load_all( paste(homepath, main_home_path, '/nonlinearTseries', sep='' ))
+library(nonlinearTseries)
 source( paste(github_repo_path,'/code/rfunctions/extra_rqa.R',sep='') )
-
 
 ################################################################################
 # (2) Reading data
 file_ext <- paste('xdata_v', version, '.dt',sep='')
-#file_ext <- paste('xdata_v01', '.dt',sep='')
 data <- fread( file_ext, header=TRUE)
 
 data <- data[,.(

### 2020-07-02 code/rscripts/G_3Drqa/A_3Drqa_datatable.R

@@ -23,7 +23,6 @@ start.time <- Sys.time()
 ################################################################################
 # (0) Defining paths for main_path, r_scripts_path, ..., etc.
 homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
 r_scripts_path <- getwd()
 setwd("../../../")
 github_repo_path <- getwd()
@@ -36,8 +35,6 @@ dataset_path <- '/data/dataset'
 
 ## Outcomes Data Path
 outcomes_data_path <- paste(github_repo_path, feature_path,  sep="")
-### Outcomes Plot Path
-#outcomes_plot_path <- paste(github_repo_path,'/hwum-manuscript/figures/results', feature_path, '/v', version, sep="")
 ## Data Path
 data_path <- paste(github_repo_path, dataset_path, sep="")
 setwd(file.path(data_path))
@@ -70,9 +67,6 @@ data <- data[,.(
 	), by=. (Participant,Activity,Sensor,Sample,Time)]
 
 
-
-
-
 ################################################################################
 ################################################################################
 ################################################################################
@@ -84,8 +78,8 @@ W<-NULL#rqas for all windows
 ##########################
 ##### one window lenght
 
-#windowsl <- c(100)
-#windowsn <- c('w2')
+windowsl <- c(100)
+windowsn <- c('w2')
 
 #windowsl <- c(250)
 #windowsn <- c('w5')
@@ -95,7 +89,8 @@ W<-NULL#rqas for all windows
 
 #windowsl <- c(750)
 #windowsn <- c('w15')
-#
+
+
 
 
 ############################
@@ -103,10 +98,8 @@ W<-NULL#rqas for all windows
 #windowsl <- c(100,250,500,750)
 #windowsn <- c('w2', 'w5', 'w10', 'w15')
 #
-windowsl <- c(100,250,500)
-windowsn <- c('w2', 'w5', 'w10')
-
-
+#windowsl <- c(100,250,500)
+#windowsn <- c('w2', 'w5', 'w10')
 
 
 
@@ -120,7 +113,7 @@ windowsn <- c('w2', 'w5', 'w10')
 #### w15, 15-second window (750 samples) ## 100 to 850
 
 
-for ( wk in 1:(length(windowsl)) ) {
+for ( wk in 1:(length(windowsn)) ) {
 
 xdata <- data
 
[…]

### 2020-07-02 code/rscripts/G_3Drqa/B_3Drqa_plots_thesis.R

@@ -1,367 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-
-homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-
-
-
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/docs/figures/rqa/src/3drqa_thesis'
-data_path <- '/data/rqa'
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
-## Data Path
-data_path <- paste(github_repo_path, data_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(plot3D)
-
-
-
-################################################################################
-# (2) Reading data
-#
-#file_ext <- paste('RQAs_p01w100.dt',sep='')
-#W100 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w250.dt',sep='')
-#W250 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w500.dt',sep='')
-#W500 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w750.dt',sep='')
-#W750 <- fread( file_ext, header=TRUE)
-#
-#W <- rbind(W100, W250, W500, W750)
-#
-
-#file_ext <- paste('RQAs_p02w500.dt',sep='')
-#W <- fread( file_ext, header=TRUE)
-#
-
-file_ext <- paste('RQAs_p03w500.dt',sep='')
-W <- fread( file_ext, header=TRUE)
-
-
-
-################################################################################
-################################################################################
-################################################################################
-################################################################################
-## (x.x.x) RQA Metric Selection
-rqas <- c('ENTR')
-#rqas <- c('REC','DET', 'RATIO', 'ENTR')
-
-
-#########################################################
[…]

### 2020-07-02 code/rscripts/G_3Drqa/C_3Drqa_plots_epsilons.R

@@ -24,17 +24,12 @@ start.time <- Sys.time()
 # (0) Defining paths for main_path, r_scripts_path, ..., etc.
 
 homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
 r_scripts_path <- getwd()
 setwd("../../../")
 github_repo_path <- getwd()
 
 
-
-
-
 ##VERSION 
-version <- '00'
 feature_path <- '/docs/figures/rqa/src/3drqa_epsilons'
 data_path <- '/data/rqa'
 
@@ -55,31 +50,26 @@ library(plot3D)
 
 ################################################################################
 # (2) Reading data
-#
+
+## Participant 01
 #file_ext <- paste('RQAs_p01w100.dt',sep='')
-#W100 <- fread( file_ext, header=TRUE)
-#
 #file_ext <- paste('RQAs_p01w250.dt',sep='')
-#W250 <- fread( file_ext, header=TRUE)
-#
 #file_ext <- paste('RQAs_p01w500.dt',sep='')
-#W500 <- fread( file_ext, header=TRUE)
-#
 #file_ext <- paste('RQAs_p01w750.dt',sep='')
-#W750 <- fread( file_ext, header=TRUE)
-#
-#W <- rbind(W100, W250, W500, W750)
-#
 
+## Participant 02
+#file_ext <- paste('RQAs_p02w100.dt',sep='')
+#file_ext <- paste('RQAs_p02w250.dt',sep='')
 #file_ext <- paste('RQAs_p02w500.dt',sep='')
-#W <- fread( file_ext, header=TRUE)
-#
+#file_ext <- paste('RQAs_p02w750.dt',sep='')
 
-file_ext <- paste('RQAs_p03w500.dt',sep='')
+## Participant 03
+#file_ext <- paste('RQAs_p03w100.dt',sep='')
+#file_ext <- paste('RQAs_p03w250.dt',sep='')
+#file_ext <- paste('RQAs_p03w500.dt',sep='')
+file_ext <- paste('RQAs_p03w750.dt',sep='')
+
 W <- fread( file_ext, header=TRUE)
-
-
-
 ################################################################################
 ################################################################################
 ################################################################################
@@ -87,7 +77,6 @@ W <- fread( file_ext, header=TRUE)
 ## (x.x.x) RQA Metric Selection
 rqas <- c('ENTR')
 #rqas <- c('REC','DET', 'RATIO', 'ENTR')
-
 
[…]

### 2020-07-02 code/rscripts/G_3Drqa/D_3Drqa_plots_sensors_activities.R

@@ -24,7 +24,6 @@ start.time <- Sys.time()
 # (0) Defining paths for main_path, r_scripts_path, ..., etc.
 
 homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
 r_scripts_path <- getwd()
 setwd("../../../")
 github_repo_path <- getwd()
@@ -118,8 +117,8 @@ Wkw <-  Wk[ Window %in% selectWindow ]
 ################################################################################
 ################################################################################
 	## (4.2.1) Activities Selection
-	#activities <- c('HN','HF')
-	activities <- c('VN','VF')
+	activities <- c('HN','HF')
+	#activities <- c('VN','VF')
 
 
 

### 2020-10-18 docs/corrections/2020-07-03/report/report.tex
adds corrections from 2020-07-03 and updates the manuscript 2020-10-08

--- 2020-07-02 docs/manuscript/main.tex
@@
-%*******************************************************************************
-\documentclass[fleqn,10pt]{wlscirep}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\graphicspath{{../}} %goes to path: docs/
-
-\title{Recurrent Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
-\author[1,*]{Miguel Xochicale}
-\author[1]{Chris Baber}
-\affil[1]{University of Birmingham, 
-	School of Engineering, 
-	Birmingham, 
-	B15 2TT, 
-	UK}
-\affil[*]{map479@bham.ac.uk}
-
-%\keywords{Keyword1, Keyword2, Keyword3}
+\documentclass[10pt]{article}
+\usepackage[margin=20mm]{geometry}
+\usepackage{enumitem}
+\usepackage{hyperref}
+
+\title{
+Logbook of corrections 2020-07-03 
+}
+\author{Miguel Xochicale}
+\date{ \today }
+
+\begin{document}
+\maketitle
 
 \begin{abstract}
-Human movement variability occurs in motor performance 
-across multiple repetitions of a task and such behaviour is an
-inherent feature within and between each persons' movement. Quantifying
-movement variability is still an open problem, particularly 
-when methods in time domain, frequency domain or nonlinear dynamics 
-can break down due to the real-world time-series datasets. 
-For this work, we therefore investigate nonlinear dynamics methods 
-such as reconstructed state space (RSSs), uniform time-delay embedding, 
-recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
-with real-world time-series data.
-Particularly, we are interested in the weaknesses and robustness 
-of nonlinear dynamics tools with the use of raw and post-processed 
-data of wearable inertial sensors. 
-That said, twenty right-handed healthy participants 
-imitated simple vertical and horizontal arm movements in normal 
-and faster velocity from an humanoid robot in order to have 
-four window lengths and three levels of smoothed time-series data,
-to then found visual differences in the patterns with RSSs and RPs
-and particulatly the computed differences with RQA metrics
-that help us to quantify activities, types of sensors, windows lenghts 
-and level of smoothness. 
-Specifically, we can conclude that RQA ENTR, Shannon Entropy, 
-can lead to interesting results on the quantification of 
-movement variablity for participants of different ages, 
-state of health and anthropomorphic features to 
-then enhance the development of 
-better diagnostic tools for applications in rehabilitation, 
-sport science or for new forms of human-humanoid interaction.
+This logbook contains the suggested changes by Chris Baber 
+presented in (\emph{.../corrections/2020-07-03/comments/README.md}),
+fused text and further changes to the manuscript.
 \end{abstract}
 
-\begin{document}
-
-\flushbottom
-\maketitle
-\thispagestyle{empty}
-
+\tableofcontents
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Comments}
+\subsection{Text replacement}
+\begin{enumerate}
+\item 
+"Perhaps you could replace your text with mine?  
+\begin{verbatim}
+Instead of replacing test, I will fuse text
+\end{verbatim}
+\textit{
+SORTED: \\ 
+Sun  4 Oct 01:11:27 BST 2020\\
+}
+\\
+\end{enumerate}
+
+\subsection{Experiment description}
+\begin{enumerate}
+\item 
+"although still believe that there needs to be a bit more information in your description of the experiment, in terms of what participants had to do and in terms of the data collection"
+\begin{verbatim}
+
+Comestic changes changes to the experiment sections were made.
+These include the subsections of particianpts, ethics, hhi activiites,
+data collection from imus, raw data, data normalization smoothing, windowding
+
+\end{verbatim}
+\textit{
+SORTED: \\ 
+Sun 18 Oct 04:19:38 BST 2020
+\\
+}
+\\
+\end{enumerate}
+
+\section{Introduction}
+
+\subsection{ORIGINAL TEXT}
+\begin{verbatim}
 \section*{Introduction}
@@
 	of nonlinear analysis methods?
-\end{itemize}
-
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
+
+\end{verbatim}
+
+\subsection{SUGGESTED TEXT}
+\begin{verbatim}
+The complexity of human movement arises from the balancing of consistency 
+and variability in solving the Degrees of Freedom problem (Bernstein, xx). 
+Each joint can be moved in more than one plane, leading to many possible 
+combinations which need to be reduced into order to perform a control action (Newell, xx). 
+However, too much reduction can result in highly consistent movement which lacks 
+sufficient variability to allow adaptation to situational demands. 
+Consequently, one can see much variability in human movement even the 
+simplest of movements.  Variability can be beneficial in supporting adaptability 
+and can reflect experience and expertise of the performance, e.g., in cycling.  
+Uncontrolled variability, due to disease, disfunction or inexperience, 
+can be detrimental to performance and plays a role in detection of pathologies. 
+While the analysis of movement of variability is becoming increasingly 
+popular as a diagnostic tool, there remain challenges in terms of defining 
+the most appropriate methods and parameters to apply. In part, 
+these challenges stem from the fact that the identification movement 
+variability requires analysis of signals which are time-series of 1-dimension 
+in R which are noisy, nonlinear and non-stationary.  
+Further problems arise from assigning a plausible locus of control to 
+the movements, e.g., in order to determine whether variability is the 
+result of deliberative control by the performer or whether it arises 
+from exogenous or endogenous disturbances.  For this paper, our focus 
+is on the analysis of signals; specifically, in terms of objectively 
+quantifying variability of lower dimension signals using time-series analysis.
+
+Methods for time-series analysis generally involve the estimation 
+of embedding parameters (m embedding dimension and T embedding delay) 
+which are used to reconstruct the n-dimensional state space of 
+the 1-dimension time-series.  Key to the selection of these 
+parameters is the need to ensure that the topological properties 
+of the (unknown) M-dimensional state space.   A popular approach 
+to the construction of these state spaces involves the use of 
+Recurrence Plots (RPs) which provide a graphical two-dimensional 
+map (using black and white dots) to indicate recurring patterns 
+in the n-dimensional system.  While RPs provide a human 
+interpretable picture of the system, these require further 
+analysis to allow the properties of that system to be quantified, 
+and so Recurrence Quantification Analysis (RQA) can be applied.  
+However, the estimation of the embedding parameters for 
+RQA remains an open problem (Bradley et al., 2015).
+
+There is no agreed method for estimating embedding parameters 
+(for RQA or for other nonlinear analysis methods) because time-series 
+are system-dependent, i.e., these rely on the initial conditions, 
+and on the configuration and behaviour of the system.  
+This means that, unless one holds all of the influencing 
+variables constant, embedding parameters computed for one 
+instance may not apply to another instance.   
+One could apply methods, such as autocorrelation, mutual information, 
+nearest neighbour (and we will apply these in this paper), 
+but these methods assume that the data are well sampled, with 
+little noise and (usually) that the signals are purely deterministic.  
+As such, these methods can break-down in the face of real-world datasets 
+which could have different length, different degrees of precision, or 
+different levels of contamination from exogenous and endogenous 
+sources of ‘noise’.  What is, perhaps, surprising is that 
+even subject to these problems, the methods are useful and 
+nonlinear dynamics approaches continue to tell us much about 
+human movement.
+
+For this paper, we explore the role of RP and RQA in the 
+analysis of simple human movement.  We compare horizontal and 
+vertical arm movements performed by neurotypical participants 
+who are copying these movements being made by a humanoid robot.  
+This provides an initial point of comparison 
+(between human movement, which we assume to have a degree of variability, 
+and the robot, which we assume to have a degree of consistency).   
+Additionally, our interest lies in the estimation of embedding parameters 
+in light of different features of the data, e.g., levels of smoothing, 
+window length, structure of time-series based on movement, 
+types of sensors, individual differences between participants and 
+the movements that they perform. Specifically, we ask what are 
+the effects of different embedding parameters, recurrence thresholds 
+and characteristics of time-series on nonlinear analysis 
+methods (i.e., reconstructed state space with 
+uniform time-delay embedding, Recurrence Plots and 
+Recurrence Quantification Analysis)?
+
+\end{verbatim}
+
+\subsection{FUSED TEXT}
+\begin{verbatim}
+
+\section*{Introduction}
+The complexity of human movement arises from the process of mastering 
+redundant (bio)mechanical degrees of freedom (DOF) to successfully 
+accomplish any given motor task (Bernstein, 1967).
+Such DOF are independent coordinates to uniquely 
+describe the configuartions of the system which provides flexibiity 
+and stability to adapt to a change enviroment conditions but 
+that leads many possible combinations (Newell and Vaillancourt, 2001).
+Hence, human movement complexity can be seen as a balance 
+between the required DOF "to generate a stable (persistent) 
+and flexible (variable) behavioural output in response to 
+changing intentions and dynamic environmental
+conditions" \cite{davids2003}. 
+Consequently, one can see much variability in human movement even the 
+simplest of movements.  
+While the analysis of movement of variability is becoming increasingly 
+popular as a diagnostic tool or skill performance evaluation, 
+there remain challenges in terms of defining 
+the most appropriate methods and parameters to apply. 
+In part, 
+these challenges stem from the fact that the identification movement 
+variability requires analysis of signals which are time-series of 
+of $1-$dimension in $\mathbb{R}$ which are 
+noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
+Further problems arise from assigning a plausible locus of control to 
+the movements, e.g., in order to determine whether variability is the 
+result of deliberative control by the performer or whether it arises 
+from exogenous or endogenous disturbances.  For this paper, our focus 
+is on the analysis of signals; specifically, in terms of objectively 
+quantifying variability of lower dimension signals using time-series analysis.
+
+Methods for nonlinear time series analysis generally involve the estimation of the 
+embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
+reconstruct the state space,
+where an $n$-dimensional is reconstructed state space using $1-$dimensional 
+time series \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
+frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
+Key to the selection of these parameters is the need to preserve 
+topological properties of an unknown $M$-dimensional state space \cite{takens1981}.
+An common approach to the construction of these state spaces 
+involves Recurrence Plots (RPs), a graphical representation of black and white dots, 
+which shows recurrent patterns of a $n$-dimensional system.
+While RPs provide a human
+interpretable picture of the system, these require further
+analysis to allow the properties of that system to be quantified,
+and so Recurrence Quantification Analysis (RQA) can be applied.
+However, the estimation of the embedding parameters for
+RQA remains an open problem (Bradley et al. 2015 \cite{bradley2015})
+There is no agreed method to estimate embedding parameters
+for RQA and other nonlinear methods \cite{bradley2015} 
+because time series are system-dependent, i.e., these rely
+on the initial conditions, and on the configuration and behaviour of the system.  
+This means that, unless one holds all of the influencing 
+variables constant, embedding parameters computed for one 
+instance may not apply to another instance.
+
+One could apply methods, such as autocorrelation, mutual information, 
+nearest neighbour (and we will apply these in this paper), 
+but these methods assume that the data are well sampled, with 
+little noise and (usually) that the signals are purely deterministic \cite{garland2016, kantz2003}.
+That said, these methods can break-down in the face of real-world datasets 
+which could have 
+different length, 
+different values of accuracy and precision (rounding errors due to finite 
+precision of the measurement apparatus which include frequency 
+acquisition \cite{frank2010}), or 
+different levels of contamination from exogenous and endogenous 
+sources of ‘noise’ \cite{garland2016}.  What is, perhaps, surprising is that 
+even subject to these problems, the methods are useful and 
+nonlinear dynamics approaches continue to tell us much about 
+human movement \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
+gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.
+
+For this paper, we explore the role of RPs and RQA in the 
+analysis of simple human movement. 
+We compare horizontal and 
+vertical arm movements performed by neurotypical participants 
+who are copying these movements being made by a humanoid robot.  
+This provides an initial point of comparison 
+(between human movement, which we assume to have a degree of variability, 
+and the robot, which we assume to have a degree of consistency).   
+Additionally, our interest lies in the estimation of embedding parameters 
+in light of different features of the data, e.g., levels of smoothing, 
+window length, structure of time-series based on movement, 
+types of sensors, individual differences between participants and 
+the movements that they perform. 
+Specifically, we ask what are 
+the effects of different embedding parameters, recurrence thresholds 
+and characteristics of time-series on nonlinear analysis 
+methods (i.e., reconstructed state space with 
+uniform time-delay embedding, RPs and RQA)?
+
+\end{verbatim}
+
+\textit{
+SORTED: \\ 
+Tue 11 Aug 18:37:21 BST 2020 \\
+Sun 18 Oct 04:27:51 BST 2020
+\\
+}
+
+\section{Results}
+\subsection{ORIGINAL TEXT}
+\begin{verbatim}
+
 \section*{Results}
@@
 
+\end{verbatim}
+
+\subsection{SUGGESTED TEXT}
+\begin{verbatim}
+
+Reconstructed State Spaces
+As noted in the Introduction, a challenge in the implantation of uniform time-delay embedding 
+arises from the selection of embedding parameters because of the uniqueness of each time series 
+in terms of its structure (e.g., modulation of amplitude, frequency, phase etc.). 
+With that in mind, the options are to either calculate embedding parameters for each unique 
+instance (which can make comparison challenging) or to find parameters which can apply to all 
+instances in the study.  We recognise that the latter approach is not without its problems, 
+but our approach is to compute sample mean over all values in each of the conditions of the 
+time series for minimum dimension and minimum delay values.
+
+Minimum embedding parameters
+Minimum embedding parameters were initially computed using False Nearest Neighbour (FNN) 
+and Average Mutual Information (AMI).   We used FNN to calculate a value for m 
+and AMI to calculate a value for X? Comment: You should provide a sentence here 
+to justify the use these different methods for m and X? 
+To illustrate this, figure (A) show box plots for minimum embedding values for sensors 
+on the wrist of the human (HS01) and the robot (RS01).   
+As we had assumed, minimum embedding values for HS01 show greater variation than 
+those for RS01 (as indicated by differences in interquartile ranges).  
+Additionally, figure (A) show a decrease in mean values (rhombus) in the box plots 
+as smoothness of the time-series increases (see sg0, sg1, sg2).  
+Figure (B) show box plots for Average Mutual Information (AMI).  
+One can see that, in contrast to RS01, minimum values for HS01 tend to spread 
+as the smoothness of the time-series increases.
+The sample mean for minimum value of embedding parameter, m derived from 
+FNN  (figures A) is 6 and  Tfrom AMI (figures B) is 8.
+
+Uniform Time-delay embedding
+Using the minimum embedding parameters (defined above), 
+the first three axis of the rotated Principal Components Analysis (PCA) 
+are shown for the reconstructed state spaces of horizontal (figures  ) 
+and vertical (figures ) arm movements.   While visual inspection of these figures 
+suggests differences in the trajectories of the reconstructed state spaces, 
+we require an objective quantification to determine the extent of the differences.  
+One approach could be use Euclidean distances from the origin for points in these 
+trajectories, but this proved inconclusive and was not able to capture 
+the dynamics of the trajectories. Consequently, we applied RP and RQA.
+
+Recurrence Plots
+Using the minimum embedding parameters and a recurrence threshold of 1 
+Comment: you should explain why you chose 1 here Recurrence Plots (RP) were 
+computed for horizontal (figures ) and vertical (figures ) arm movements.
+
+Recurrence Quantification Analysis
+Four Recurrence Quantification Analysis metrics (percentage recurrence, REC, 
+representing the percentage of black dots in RP; percentage determinism, DET, 
+representing the predictability of the RP; ratio of DET / REC, RATIO; 
+Shannon entropy, ENTR) were computed using the same parameters as for RP.
+
+REC values
+Figure (A) presents box plots of REC which show that REC values, 
+for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements 
+in Horizontal (HS) or Vertical (VS) than for Fast (i.e., 2 seconds per movement).  
+This suggests greater variation between participants for the Slow movements.  
+For RS01, there is little variation between Slow and Fast movement.  
+In terms of smoothness, there seems little effect of HS01 but RS01 values 
+do show affects of smoothness.
+
+DET values
+Figure (b) presents DET values and shows little difference for type of movement or performer.  
+However, DET values are affected by changes in smoothness of the signal, 
+particularly for Fast movement.
+
+RATIO values
+Figure (C) presents the ratio of DET / REC.  These values, for the human performer, 
+vary less for HN than for HF movement.  
+Additionally, smoothness leads to a decrease in mean values for Fast movements.
+
+ENTR values
+Figure (d) shows ENTR values are higher for the human performer than the robot, 
+and vary with the smoothness of the time-series.
+
+3D RQA ENTR
+As ENTR appeared to be have higher sensitivity than the other measures, 
+we explored the impact of different embedding parameters /{m.../}, /{.../} 
+incrementing by 1 each run, with recurrence thresholds e = 1,2,3 and levels of 
+smoothness (sg0, sg1, sg2).  
+Figure ( ) shows that increasing recurrence 
+threshold leads to an increase in ENTR regardless of level of smoothness.  
+Increasing level of smoothness will also increase ENTR (figure () ).  
+In terms of movement or performer, RQA ENTR decrease from Slow (HS, VS) to Fast (HF, VF) 
+for both human and robot.   In terms of individual differences, for human participants, 
+figure ( ) compares p01, p02 and p03 (for illustrative purposes) – 
+both between each other and in comparison with the more consistent performance of the robot.  
+In terms of window length (w100 (2s), w250 (5s), w500(10s), w750 (15s)), figure ( ) 
+shows that improvement in capture with number of samples, although this has less 
+of an effect on RQA ENTR.
+
+\end{verbatim}
+
+\subsection{FUSED TEXT}
+\begin{verbatim}
+
 %*******************************************************************************
 %*******************************************************************************
-%*******************************************************************************
+\section*{Results}
+
+\subsection*{Reconstructed State Spaces}
+As noted in the Introduction, a challenge in the implementation of uniform time-delay embedding 
+arises from the selection of embedding parameters because of the uniqueness of each time series 
+in terms of its structure (e.g., modulation of amplitude, frequency, phase etc.). 
+With that in mind, the options are to either calculate embedding parameters for each unique 
+instance (which can make comparison challenging) or to find parameters which can apply to all 
+instances in the study.
+We recognise that the latter approach is not without its problems, 
+but our approach is to compute sample mean over all values in each of the conditions of the 
+time series for minimum dimension and minimum delay values.
+
+\subsubsection*{Minimum embedding parameters}
+Minimum embedding parameters were initially computed using False Nearest Neighbour (FNN) 
+and Average Mutual Information (AMI).   We used FNN to calculate a value for $m$
+to unfold the attractors and AMI to calculate a value for $\tau$ and maxime the 
+information in the unfolded attractor.
+To illustrate this, figure \ref{fig:cao_ami} (A) show box plots for minimum embedding 
+values for sensors on the wrist of the human (HS01) and the robot (RS01).   
+As we had assumed, minimum embedding values for HS01 show greater variation than 
+those for RS01 (as indicated by differences in interquartile ranges).  
+Additionally, figure \ref{fig:cao_ami} (A) show a decrease in mean values (rhombus) in the box plots 
+as smoothness of the time-series increases (see sg0, sg1, sg2).  
+Figure \ref{fig:cao_ami} (B) show box plots for Average Mutual Information (AMI).  
+One can see that, in contrast to RS01, minimum values for HS01 tend to spread 
+as the smoothness of the time-series increases.
+The sample mean for minimum value of embedding parameter, 
+$m$ derived from FNN  (figures \ref{fig:cao_ami} A) is $\overline{m}_0=6$ 
+and $\tau$ from AMI (figures \ref{fig:cao_ami} B) is $\overline{\tau}_0=8$.
+
+%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/caoami/pdf/fig2}
+	\caption{
+	{\bf Box plots of minimum embedding parameters.} 
+	Box plots of (A) minimum embedding dimensions 
+	and (B) first minimum AMI values for 
+	Horizontal Normal (HN), Horizontal Faster (HF),
+	Vertical Normal (VN) and Vertical Faster (VF)
+	with sensors attached to participants (HS01) and
+	sensor attached to robot (RS01).
+	Minimum embedding parameters are for twenty participants 
+	($p01$ to $p20$) with three smoothed signals 
+	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
+	and window length of 10-sec (500 samples).
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:cao_ami}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+\subsubsection*{Uniform Time-Delay Embedding}
+Using the overall embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$), 
+the first three axis of the rotated Principal Components Analysis (PCA) 
+are shown for the reconstructed state spaces of horizontal (figures~\ref{fig:rss_aHw10}) 
+and vertical (figures~\ref{fig:rss_aVw10}) arm movements. 
+While visual inspection of these figures 
+suggests differences in the trajectories of the reconstructed state spaces, 
+we require an objective quantification to determine the extent of the differences.  
+One approach could be use Euclidean distances from the origin for points in these 
+trajectories, but this proved inconclusive and was not able to capture 
+the dynamics of the trajectories. Consequently, we applied 
+Recurrence Plots and Recurrence Quantification Analysis.
+
+\subsection*{Recurrences Plots}
+Using the average embedding parameters 
+($\overline{m}_0=6$, $\overline{\tau}_0=8$) 
+and an recurrence threshold of $\epsilon=1$.
+As our interest is for dynamical transitions, 
+there is little importance on the selection of $\epsilon$ which in this case
+is 1. 
+Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rp_aH}) and 
+vertical (figures~\ref{fig:rp_aV}) arm movements.
+
+\subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
+Four Recurrence Quantification Analysis metrics (
+percentage recurrence, REC, representing the percentage of black dots in RP; 
+percentage determinism, DET, representing the predictability of the RP; 
+ratio of DET / REC, RATIO; Shannon entropy, ENTR) 
+were computed using the same parameters as for RP.
+%\subsubsection*{REC values}
+Figure~\ref{fig:RQABP}(A) presents box plots of REC values, 
+for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements 
+in Horizontal (HS) or Vertical (VS) than for Fast (i.e., 2 seconds per movement).  
+This suggests greater variation between participants for the Slow movements.  
+For RS01, there is little variation between Slow and Fast movement
+(interquartile range of 0.01). 
+In terms of smoothness, there seems little effect of HS01 but RS01 values 
+do show affects of smoothness (see the incremental changes of mean values (rhombus)).
+%\subsubsection*{DET values}
+Figure \ref{fig:RQABP}(B) presents DET values and shows 
+little difference for type of movement or performer.  
+However, DET values are affected by changes in smoothness of the signal, 
+particularly for Fast movement.
+%\subsubsection*{RATIO values}
+Figure \ref{fig:RQABP}(C) presents the ratio of DET / REC. 
+These values, for the human performer, vary less for HN than for HF movement.  
+Additionally, smoothness leads to a decrease in mean values for Fast movements.
+%\subsubsection*{ENTR values}
+Figure \ref{fig:RQABP}(D) shows ENTR values are higher for the human 
+performer than the robot, and vary with the smoothness of the time-series.
+
+\subsection*{3D RQA ENTR}
+As ENTR appeared to be have higher sensitivity than the other measures, 
+we explored the impact of different embedding parameters 
+$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
+$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ 
+incrementing by 1 each run, with recurrence thresholds $\epsilon=1, 2, 3$ 
+and levels of smoothness (sg0, sg1, sg2).  
+Figure \ref{fig:RQA-IND} shows that increasing recurrence 
+threshold leads to an increase in ENTR regardless of level of smoothness.  
+Similarly, increasing level of smoothness will also increase ENTR (figure \ref{fig:RQA-IND}).
+In terms of movement or performer, RQA ENTR decrease from Slow (HS, VS) to Fast (HF, VF) 
+for both human (HS01) and (RS01) (Figures~\ref{fig:3dRQAENTR_sensoractivities}).
+In terms of individual differences, for human participants, 
+figure \ref{fig:3dRQAENTR_participantsactivities} compares p01, p02 and p03 (for illustrative purposes) --
+both between each other and in comparison with the more consistent performance of the robot.  
+In terms of window length (w100 (2s), w250 (5s), w500(10s), w750 (15s)), 
+figure \ref{fig:3dRQAENTR_windowsactivities} shows that improvement in 
+capture with number of samples, although this has less of an effect on RQA ENTR.
+
+\end{verbatim}
+
+\textit{
+SORTED: \\ 
+Tue  1 Sep 23:43:03 BST 2020\\
+Sat  3 Oct 14:54:05 BST 2020\\
+Sat  3 Oct 23:55:12 BST 2020 \\
+}
+\\
+
+\section{Discussion}
+
+\subsection{ORIGINAL TEXT}
+\begin{verbatim}
+
 \section*{Discussion}
@@
 the range of embedding parameters.
+
+\end{verbatim}
+
+\subsection{SUGGESTED TEXT}
+\begin{verbatim}
+
+While there are many approaches to estimating embedding parameters for nonlinear analysis, 
+these can be influenced by the structure of the time-series data.   
+We show that, for RSS, RP and RQA, the estimation of embedding parameters can be performed 
+using a sample mean which, together with recurrence threshold, can be shown to be 
+influenced by activity, performer, window length and smoothness of time-series.  
+It is known that time-series from different sources and with different characteristics 
+require different embedding parameters, and this can produce different RPs.  
+We have shown how RQA metrics can help to quantify movement variability.
+
+\end{verbatim}
+
+\subsection{FUSED TEXT}
+\begin{verbatim}
+
+\section*{Discussion}
+While there are many approaches to estimating embedding parameters for nonlinear analysis, 
+these can be influenced by the structure of the time-series data.   
+We show that, for RSS, RP and RQA, the estimation of embedding parameters can be performed 
+using a sample mean which, together with recurrence threshold, can be shown to be 
+influenced by activity, performer, window length and smoothness of time-series.  
+It is known that time-series from different sources and with different characteristics 
+require different embedding parameters, and this can produce different RSS, RPs and RQAs. 
+Although this work helps to understand the open problem of 
+finding right balance among (i) the level of smoothness of the signal, 
+(ii) the selection of recurrence thresholds and (iii) 
+the range of embedding parameters, 
+we have shown how RQA metrics can help to quantify movement variability.
+
+\end{verbatim}
+
+\textit{
+SORTED: \\ 
+Sun  4 Oct 00:52:04 BST 2020\\
+}
+\\
+
+\section{Conclusions}
+
+\subsection{ORIGINAL TEXT}
+\begin{verbatim}
 
@@
 
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
-\section*{Methods}
-\subsection*{State Space Reconstruction}
-The method of state space reconstruction \cite{packard1980} 
-\cite{takens1981} has been applied in many disciplines 
-\cite{aguirre2009, stergiou2011, frank2010, 
-sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
-based on uniform time-delay embedding methodology which is a simple 
-matrix implementation that can reconstruct an unknown $d-$dimensional 
-manifold $M$ from a scalar time series (e.g. one-dimensional 
-time series in $\mathbb{R}$). A manifold, in this context, is a 
-multidimensional curved surface within a space (e.g. a saddle) 
-\cite{guastello-gregson2011}.
-
-The use of a scalar time series is the main advantage of the method of state 
-space reconstruction which in essence preserve dynamic invariants such as 
-correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
-entropy and detrended fluctuation analysis \cite{bradley2015, 
-Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
-However, selecting appropriate embedding parameters which are sued to apply the 
-state space reconstruction is still an open challenge for which we present 
-introductions for the methodologies to compute such embedding parameters 
-With that in mind, in the following subsections, we describe in more detail 
-the state space reconstruction theorem (RSSs), uniform time-delay embedding 
-theorem (UTDE), false nearest neighbours (FNN) and average mutual 
-information (AMI). 
-%and other methodologies for state space reconstruction.
-
-\subsection*{State Space Reconstruction Theorem}
-Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
-uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:ssr}
-  s(t)=f^t [s(0)],
-\end{equation}
-%%********************************[EQUATION]************************************
-where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
-$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
-unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
-function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
-iteration of $f$ that corresponds to an initial position 
-$s(0) \in M $ \cite{takens1981}.
-%$f$ is a smooth dynamical system, where smooth means a 
-%continuously differentiable system
-%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
-Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
-can be obtained with
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:measurement}
-  x(t)=h[s(t)],
-\end{equation}
-%%********************************[EQUATION]************************************
-where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
-Reconstructed state space can then be described as an $n-$dimensional state 
-space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
-$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
-is the uniform time-delay embedding with a dimension embedding $m$
-and delay embedding $\tau$ and
-$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
-of dimensionality (e.g. Principal Component Analysis, 
-Singular Value Decomposition, etc) being $n \leq m$.
-With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
-defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
-$\boldsymbol{X}(t) = \Phi(s(t))$,
-where $\Phi$ is a diffeomorphic map \cite{takens1981}
-whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
-dimension of $M$ \cite{garland2016}.
-% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
-% is called diffeomorphic if it is one-to-one correspondence and its inverse
-% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
-Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
-space then a composition of functions represented with $F^t$ is induced on the
-reconstructed state space determined:
-%by true smooth dynamical system $f^t$ and $\Phi$:
-%********************************[EQUATION]************************************
-\begin{equation}\label{eq:st}
-  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
-\end{equation}
-%%********************************[EQUATION]************************************
-With this in mind, an embedding is defined as "a smooth one-to-one 
-coordinate transformation with a smooth inverse" and the total reconstruction 
-map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
-Fig~\ref{fig:ssr} illustrates the state space reconstruction.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
-    \caption{
-	{\bf State space reconstruction methodology.}
-	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
-	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
-    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
-    	The time-delay embedding represented as the $\Phi$, maps the original
-    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
-	time-delay embedding matrix $\boldsymbol{X}(t)$.
-    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
-	a new state $y(t)$ of dimensions $n < m$.
-    	(A) $M-$dimensional manifold representing the state space 
-	(e.g. Lorenz system);
-    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
-   	(C) $m-$dimensional reconstructed state space with 
-    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
-    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
-	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
-	where $\Phi$ is the delay reconstruction map and 
-	$\Psi$ is the coordinate transformation map.
-	This figure is adapted from 
-   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-    }
-    \label{fig:ssr}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
-Frank et al. and Sama et al. refer to the state space reconstruction
-as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
-However, we consider the term "uniform time-delay embedding"
-as more descriptive and appropriate terminology for our work.
-Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed 
-copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
-the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
-samples of $\{ \boldsymbol{x}_n \}$.
-$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
-The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
-and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
-where $i$ goes from $0,1, \dots, (m-1)$.
-Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
-information of unobserved state variables and encapsulates the information of
-the delayed copies of the available time series in the uniform time-delay 
-embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
-$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:tde}
-\boldsymbol{X}^{m}_{\tau}  =
-\begin{pmatrix}
-\boldsymbol{ \tilde{x} }_n \\
-\boldsymbol{ \tilde{x} }_{n-\tau} \\
-\boldsymbol{ \tilde{x} }_{n-2\tau} \\
-\vdots \\
-\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
-\end{pmatrix}^\intercal, 
-\end{equation}
-%%********************************[EQUATION]************************************
-where $m$ is the embedding dimension, $\tau$ is the embedding delay and
-$ ^\intercal$ denotes the transpose.
-$m$ and $\tau$ are known as embedding parameters.
-The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
-$N-(m-1)\tau$ rows and $m$ columns and 
-$N-(m-1)\tau$ defines the length of each delayed copy 
-of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
-%For further details and explicit examples of uniform time-delay 
-%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
-
-\subsection*{Estimation of Embedding Parameters}
-The estimation of the embedding parameters ($m$ and $\tau$) 
-is a fundamental step for the state space reconstruction with the use
-of uniform time-delay embedding method.
-With this in mind, we review two of the most common algorithms,
-which will be used in our work, to compute the embedding
-parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%*******************************************************************************
-\subsubsection*{False Nearest Neighbours}
-To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
-used the method of false neighbours which can be understood as follows:
-on the one hand, when the embedding dimension is too small to unfold the 
-attractor "not all points that lie close each other will be neighbours and 
-some points appear as neighbours because of the attractor has been projected 
-down into an smaller space", on the other hand, when increasing the embedding 
-dimension "points that are near to each other in the sufficient
-embedding dimension should remain close as the dimension increase from $m$ 
-to $m+1$ \cite{krakovska2015}".
-From a mathematical point of view, the state space reconstruction theorem is 
-done when the attractor is unfolded with either the minimum embedding 
-dimension, $m_0$, or any other embedding dimension value where 
-$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
-leads to excessive computations \cite{bradley2015}. With this in mind, 
-Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
-where only the time-series and one delay embedding value are necessary 
-to select the minimum embedding dimension. Cao's algorithm is based 
-on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:e}
-  \begin{aligned}
-	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
-	 &=
-       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
-       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
-            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
-  \end{aligned}
-\end{equation}
-%%********************************[EQUATION]************************************
-where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
-embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
-From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
-and $\tau$ for which $E_1(m)$ is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:e1}
-E_1(m) = \frac{ E(m+1) } { E(m)}.
-\end{equation}
-%%********************************[EQUATION]************************************
-$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
-in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
-As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
-than some $m_0$, if the time series comes from a multidimensional state space
-then $m_0 + 1$ is the minimum dimension".
-Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
-stochastic signals. $E_2(m)$ is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:e2}
-E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
-\end{equation}
-%%********************************[EQUATION]************************************
-where
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:ee}
-E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
-|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
-\end{equation}
-%%********************************[EQUATION]************************************
-For instance, when the signal comes from random noise (values that are 
-independent from each other), all $E_2(m)$ values are approximately equal 
-to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
-not constant for all $m$ (e.g. $E_2(m) \neq 1$).
-
-As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
-series: the solution for the $x$ variable of the Lorenz system 
-(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
-and a variance of one (Fig~\ref{fig:e1e2}F).
-We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
-The $E_1(m)$ values for the chaotic time series appear to be constant
-after the dimension is equal to six.
-The determination of six is given that any value of $m$ can be used as they 
-are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
-$E_2(m)$ values, for chaotic time series, are different to one 
-(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
-chaotic time series the minimum embedding dimension the time series 
-comes from a deterministic signal. With regard to the noise time series,  
-$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
-$E_1(m)$ values then indicate the minimum embedding dimension of the 
-noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
-concluded that noise time series is a stochastic signal.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
-    \caption{
-	{\bf Minimum dimension embedding values with Cao's method.} 
-	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
-	with variations of $\tau$ values from one to twenty
-	for (E) chaotic and (F) random time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:e1e2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
-computed but also a variation of $\tau$ from 1 to 20 is presented. 
-The purpose of such variation for $\tau$ is to show its independence with
-regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-(Fig~\ref{fig:e1e2}A,B,C, and D). 
-However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
-a new threshold where $m$ values appear to be constant in $E_1 (m)$.
-In the case of the given examples and reported results, we defined such 
-threshold as 0.05. Further investigation is required for the selection of the 
-threshold in the $E_1(m)$, as the selection of the threshold in this work is 
-base on no particular method but visual inspection.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%*******************************************************************************
-\subsubsection*{Average Mutual Information}
-When selecting the delay dimension parameter, $\tau$, one can consider the 
-following two cases:
-(i) when $\tau$ is too small, the elements of time-delay embedding will be along
-the bisectrix of the phase space and the reconstruction is generally not 
-satisfactory, 
-(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
-time-delay embedding will become spread and uncorrelated which makes 
-recovering the underlying attractor more difficult if not 
-impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
-With regard to the algorithms to compute $\tau$, 
-Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
-function in which the first zero crossing is considered as the minimum delay 
-embedding parameter. However, the autocorrelation function is a linear 
-statistic over which the Average Mutual Information (AMI) algorithm is 
-preferred because the AMI takes into account the nonlinear dynamical 
-correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
-algorithm is described below to estimate the minimum delay embedding 
-parameter, \texorpdfstring{$\tau_0$}{T}.
-
-To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
-and then a probability distribution of data is computed \cite{kantz2003}.
-AMI is therefore denoted by $I(\tau)$ which is the average mutual 
-information between the original time series, $x(n)$, and the delayed 
-time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
-AMI is defined by
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:ami}
-I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
-\end{equation}
-%%********************************[EQUATION]************************************
-Probabilities are defined as follows:
-$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
-the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
-the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
-that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
-in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
-For small $\tau$, AMI will be large and it will then decrease more or 
-less rapidly. As $\tau$ increase and goes to a large limit, 
-$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
-factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
-to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
-where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
-where the redundancy is the least" \cite{kantz2003}.
-
-For example, we compute the AMI for two time series:
-A) the $x$ solution of the Lorenz system, and
-B) a noise time series using a normal distribution with mean zero and 
-standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
-concluded that the amount of knowledge for any noise time series is zero 
-for which the first minimum embedding parameter is $\tau_0=1$. 
-On the contrary, the first minimum of the AMI for the chaotic time series 
-is $\tau_0=17$ which is the value that maximize the independence 
-between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
-\cite{bradley2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
-    \caption{
-	{\bf Minimum delay embedding values with AMI's method.} 
-    	(A, B) AMI values where its first minimum value in the curve
-	is the minimum time delay embedding ($\tau_0$), 
-	for (C) a chaotic and (D) noise time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:amis}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-Similarly as Cao's algorithm negatives, AMI's algorithm is not an
-exception for negatives, which are worthwhile to mention for further 
-investigations.
-For instance, (i) is not clear why the choose of the first minimum of the AMI 
-is the minimum delay embedding parameter \cite{kantz2003} and 
-(ii) the probability distribution of the AMI function is computed
-with the use of histograms which depends on a heuristic choice of number of bins
-for which AMI depends on partitioning \cite{garcia2005e71}.
-
-%*******************************************************************************
-\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
-\subsection*{Recurrence Plots}
-Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
-in conservative systems, however such discovery was not put into practice 
-until the development of faster computers \cite{marwan2007},
-for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
-where recurrences in the dynamics of a system can be visualised using 
-Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
-propose a tool, called Recurrence Plot (RP), that provides insights into 
-high-dimensional dynamical systems where trajectories are very difficult to 
-visualise. Therefore, "RP helps us to investigate the 
-$m-$dimensional phase space trajectories through a two-dimensional 
-representation of its recurrences" \cite{marwan2015}.
-Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
-to the methodologies of the state space reconstruction and other dynamic 
-invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
-the recurrences of the trajectories in the phase space can provide 
-important clues to characterise natural process that present, for
-instance, periodicities (as Milankovitch cycles) or irregular cycles 
-(as El Ni\~no Southern Oscillation). 
-Such recurrences can not only be presented visually using Recurrence Plots (RP) 
-but also be quantified with Recurrence Quantification metrics, which leads 
-to applications of these in various areas such as economy, physiology, 
-neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
-
-For the creation of a recurrence plot based on time series 
-$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
-reconstruction with uniform time-delay embedding 
-$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
-\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
-where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
-and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
-The recurrence plot is therefore a two-dimensional $N \times N$ square 
-matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
-whenever $X(i)$ is sufficiently close to $X(j)$: 
-%%********************************[EQUATION]************************************
-\begin{equation}
-\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
-\end{equation}
-%%********************************[EQUATION]************************************
-where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
-$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
-function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
-RP is also characterised with a line of identity (LOI) which is a  
-diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
-%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
-    \caption{
-	{\bf Recurrence Plots.} 
-	(A) State space of the Lorenz system with controlling parameters 
-	($\rho=28, \sigma=10, \beta=8/3$). 
-	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
-	(black circle) of a given point at $i$ is a recurrent point and is 
-	represented as a black dot in the recurrence plot at location 
-	$(i, j)$ or white otherwise.
-	(B) Recurrence plot using the 
-	three components of the Lorenz system and the RP with no embeddings 
-	and threshold $\epsilon=5$.
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:mrp}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
-\subsection*{Structures of Recurrence Plots}
-Pattern formations in the RPs can be designated either 
-as topology for large-scale patterns or texture for small-scale patterns.
-In the case of topology, the following pattern formations are commonly presented:
-(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-uniformly distributed noise (Fig~\ref{fig:rp2}A), 
-(ii) periodic and quasi-periodic systems where diagonal lines and 
-checkerboard structures represent oscillating systems, e.g., sinusoidal 
-signals (Fig~\ref{fig:rp2}B), 
-(iii) drift where paling or darkening recurrence points away from 
-the LOI is caused by drifting systems, 
-e.g., logistic map (Fig~\ref{fig:rp2}C), and
-(iv) disrupted where recurrence points are presented white areas or 
-bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
-Texture patterns in RPs can be categorised as:
-(i) single or isolated recurrence points that represent rare occurring states, 
-and do not persist for any time or fluctuate heavily,
-(ii) dots forming diagonal lines where the length of the small-scale parallel 
-lines in the diagonal are related to the ratio of determinism or predictability 
-in the dynamics of the system, and
-(iii) dots forming vertical and horizontal lines where the length of the 
-lines represent a time length where a state does not change or change very 
-slowly and these patterns formation represent discontinuities in the signal, 
-and (iv) dots clustering to inscribe rectangular regions which are related 
-to laminar states or singularities \cite{marwan2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
-    \caption{
-	{\bf Patterns in Recurrence Plots.} 
-	Time-series with its respective recurrence plot for:
-	(A) uniformly distributed noise,
-	(B) super-positionet harmonic oscillation 
-	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-	with a linearly increase term ($0.01*i$), and
-	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:rp2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-Although, each of the previous pattern descriptions of the structures in the 
-RP offer an idea of the characteristics of dynamical systems, 
-these might be misinterpreted and conclusions might tend to be subjective 
-as these require the interpretation of a particular researcher(s).
-Because of that, recurrence quantification analyis (RQA) offer objective 
-methodologies to quantify such visual characteristics of previous 
-recurrent pattern structures in the RP \cite{zbilut1992}.
-
-%*******************************************************************************
-\subsection*{Recurrence Quantifications Analysis (RQA)}
-Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
-the density of recurrence points in RPs, then histograms of lengths for 
-diagonal lines in RPs were studied by \cite{trulla1996} which were the 
-introduction to the term recurrence quantification analysis (RQA) 
-\cite{marwan2008}. RQA has been applied in many fields such as life science, 
-engineering, physics, and others \cite{marwan2008}. Particularly in human 
-movement to investigate noise and complexity of postural control 
-\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
-coordination \cite{duran2017}. The success of RQA is not only due to its 
-simple algorithmic implementation but also to its capacity to detect tiny 
-modulations in frequency or phase which are not detectable using standard 
-methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
-RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
-dimension which is verified experimentally by \cite{iwanski1998}.
-RQA metrics comprehend percentage of recurrence, percentage of determinism, 
-ratio, Shannon entropy of the frequency distributions of the line lengths,
-maximal line length and divergence, trend and laminariy 
-\cite{marwan2007, marwan2015}. For this work, we considered only four 
-RQA metrics, due to its consistency with our preliminary experiments, 
-which are described below. Such metrics are computed the nonlinearTseries 
-R package \cite{nonlinearTseries2016}.
-
-\subsubsection*{REC values}
-The percentage of recurrence (REC) is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}
-REC(\epsilon,N) = 
-		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
-		\mathbf{R}^{m}_{i,j}(\epsilon),
-\end{equation}
-%%********************************[EQUATION]************************************
-which enumerates the black dots in the RP excluding the line of identity.
-REC is a measure of the relative density of recurrence points in the sparse 
-matrix \cite{marwan2015}.
-%REC is computed as follow with the nonlinearTseries package 
-% \cite{nonlinearTseries} 
-%  hist = getHistograms(neighs, ntakens, lmin, vmin)
-%  # calculate the number of recurrence points from the recurrence rate. The
-%  # recurrence rate counts the number of points at every distance in a concrete
-%  # side of the main diagonal.
-%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
-%  # add the maindiagonal
-%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
-%  # calculate the recurrence rate dividing the number of recurrent points at a
-%  # given distance by all points that could be at that distance
-%  recurrence_rate_vector = 
-%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
-%  # percentage of recurrent points
-%  REC = (numberRecurrencePoints) / ntakens ^ 2
-
-\subsubsection*{DET values} 
-The percent determinism (DET) is defined as the fraction of recurrence points
-that form diagonal lines and it is determined by
-%%********************************[EQUATION]************************************
-\begin{equation}
-DET=
-	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
-	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
-\end{equation}
-%%********************************[EQUATION]************************************
-where 
-%%********************************[EQUATION]************************************
-\begin{equation}
-H_D(l) = 
-	\sum^{N}_{i,j=1} 
-	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
-	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
-	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
-\end{equation}
-%%********************************[EQUATION]************************************
-is the histogram of the lengths of the diagonal structures in the RP.
-DET can be interpreted as the predictability of the system for periodic signals 
-which, in essence, have longer diagonal lines than the short diagonals lines
-for chaotic signals or absent diagonal lines for stochastic signals 
-\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
-the organisation of points in RPs  \cite{iwanski1998}. 
-%percent determinism (DET) is computed as follow with the nonlinearTseries 
-%package \cite{nonlinearTseries}  
-% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%                                       lmin = 2, lDiagonalHistogram,
-%                                       recurrence_rate_vector, maxDistanceMD) {
-%  #begin parameter computations
-%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
-%  DET = num / numberRecurrencePoints
-
-\subsubsection*{RATIO values}
-RATIO is defined as the ratio between DET and REC and it is calculated from 
-the frequency distributions of the lengths of the diagonal lines.
-RATIO is useful to discover dynamic transitions \cite{marwan2015}.
-%  diagP = calculateDiagonalParameters(
-%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
-%    recurrence_rate_vector, maxDistanceMD
-%  )
-% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%                                       lmin = 2, lDiagonalHistogram,
-%                                       recurrence_rate_vector, maxDistanceMD) {
-%  #begin parameter computations
-%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
-%  DET = num / numberRecurrencePoints
-% 
-%
-%    RATIO = diagP$DET / REC
-%
-
-\subsubsection*{ENT values}
-ENT is the Shannon entropy of the frequency distribution of the diagonal 
-line lengths and it is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}
-ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with 
-	\quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
-\end{equation}
-%%********************************[EQUATION]************************************
-ENT reflects the complexity of the deterministic structure in the system.
-For instance, for uncorrelated noise or oscillations, 
-the value of ENT is rather small and indicates low complexity of the system,
-therefore "the higher the ENT is the more complex the dynamics are" 
-\cite{marwan2007, marwan2015}.
-%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
-%
-%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%                                       lmin = 2, lDiagonalHistogram,
-%                                       recurrence_rate_vector, maxDistanceMD) {
-%
-%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
-%  diff_0 = which(pl > 0)
-%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
-
-\subsection*{Sensitivity and robustness of RPs and RQA.}
-RP and RQA are a very young field in nonlinear dynamics and many questions 
-are still open, for instance, different parameters for window length size 
-of the time series, embedding parameters or recurrence threshold can 
-generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
-
-The selection of recurrence threshold, $\epsilon$, can depend on the system 
-that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
-require to be very small, for trajectory reconstruction $\epsilon$ requires 
-to have a large thresholds or when studying dynamical transition 
-there is little importance about the selection of the threshold 
-\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
-the recurrence threshold  should be five times larger 
-than the standard deviation of the observational noise
-or the use of diagonal structures within the RP is suggested in order
-to find the optimal recurrence threshold for (quasi-)periodic process 
-\cite{marwan2011}.
-Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
-of choosing the right embedding parameters to perform RQA for which 
-many experiments have to be performed using different parameters in order 
-to have a better intuition of the nature of the time series and how 
-this is represented by using RQA.
-
-With that in mind, this work explores the sensitivity and robustness 
-of the window size of time series, embedding parameters for RSS with UTDE 
-and recurrence threshold for RP and RQA in order to gain a better insight 
-into the underlying time series collected from inertial sensors in the 
-context of human-humanoid imitation activities.
-%Such changes of recurrence threshold values 
-%can modify the patterns in RPs and therefore the values of RQA metrics.
-%We therefore computed 3D surfaces to explore the sensibility and robustness of 
-%embedding parameters and recurrence threshold in RQA  metrics. Following the 
-%same methodology of computing 3D surfaces, we also considered variation of 
-%window length size to present RQA metrics dependencies with embedding 
-%parameters, recurrence thresholds and window length size.
-
-%Choosing an appropriate recurrence threshold is crucial so as to get 
-%meaningful representations in RP and RQA, however,
-
-%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
-%metrics for RQA are independent of embedding dimension parameters, 
-%however, that is not the case when using 
-%different recurrence thresholds.
-
-%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
-%determining embedding parameters, their method consisted on creating 3D 
-%surfaces with RQA metrics with an increase of embedding parameters 
-%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
-%fluctuations and gradual changes in the 3D surfaces that provide information 
-%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
-%created 3D surfaces for visual selection of not only embedding parameters 
-%but also recurrence thresholds. Following same methodologies, we explored 
-%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
-%using 3D surfaces by an unitary increase of the pair embedding 
-%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
-%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
-%(Fig.~\ref{fig:topo_rqas}).
-
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
-\section*{Experiment} \label{sec:experiment}
-We conducted an experiment in the context of human-humanoid imitation (HHI) 
-activities where participants were asked to imitate simple horizontal and 
-vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
-Such simple movements were repeated ten times for the participant 
-who copied NAO's arm movements in a face-to-face imitation activity.
-Also, wearable inertial measurement unit (IMU) sensors were attached 
-to the right hand of the participant and to the left hand of the robot 
-(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
-IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
-accelerometer, gyroscope and magnetometer sensors and quaternions 
-\cite{Comotti2014}.
-
-\subsection*{Ethics}
-The experiments of this work were conducted in November 2016 and participants
-confirmed reading and understanding the participant information sheet of the 
-experiments and were able to withdraw from the experiment at any time without giving
-any reason. The design of the experiments adhered to the University of X
-regulations, data were anonymised and videos were stored only on a 
-personal computer in accordance with the Data Protection Act 1998. 
-%Refer to Appendix D for further information about the ethics, 
-%online participation information sheets and experiment check list.
-
-%A further description of the NeMEMSi IMU sensors 
-%is given in \nameref{S1_AppendixB}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-  \centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
-    \caption{
-	{\bf Human-humanoid imitation activities.} 
-    		Face-to-face human-humanoid imitation (HHI) activities for 
-		(A) HHI of horizontal arm movement, 
-		(B) Humanoid horizontal arm movement,
-		(C) HHI of vertical arm movement, and 
-		(D) Humanoid vertical arm movement.
-        }
-    \label{fig:hri}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-\subsection*{Participants}
-Twenty-three participants,
-from now on defined as $pN$ where $N$ is the number of participant, were 
-invited to do the experiment. However, data for three participants were 
-not used because the instructions for $p01$, who was the only left-handed,
-were mistakenly given in a way that movements were performed different
-from what had been planned, and for participants $p13$ and $p16$ 
-data were corrupted because bluetooth communications problems with the 
-sensors. With that in mind, data for twenty participants were analysed in 
-this work.
-
-Of the twenty participants, all of them are right-handed healthy participants 
-of whom four are females and sixteen are males with a mean and standard 
-deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
-informed consent forms prior to participation in the experiment.
-
-\subsection*{Human-humanoid imitation activities}
-For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
-two of which were attached to the right hand of the participant and the 
-other two to the left hand of the humanoid robot.
-Then, each participant was asked to imitate repetitions of simple horizontal
-and vertical arm movements performed by the humanoid robot in the following 
-conditions:
-(i) ten repetitions of horizontal arm movement at normal (HN) 
-	and faster (HF) speed (Figure~\ref{fig:hri} A), and
-(ii) ten repetitions of vertical arm movement at normal (VN) 
-	and faster (VF) speed (Figure~\ref{fig:hri} C).
-The normal and faster speed of arm movements is defined by the duration 
-in number of samples of one repetition of NAO's arm movements. 
-We select NAO's arm movements duration to distinguish between normal and 
-faster arm movements as NAO's movements have less variation between 
-repetition to repetition. The duration for one repetition of the horizontal 
-arm movement at normal speed, HN, is about 5 seconds considering 
-that each repetition last around 250 samples.
-For horizontal arm movement at faster speed, HF, each repetition were performed 
-in around 2 seconds which correspond to 90 samples of data.
-The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
-which is around 300 samples of data. For vertical arm movement at 
-faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
-to 120 samples of data.
-To visualise the distinction between normal and faster speed for horizontal 
-and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
-for axes Z and Y of the gyroscope sensors with four window lengths: 
-2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
-and 15-sec (750-samples).
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht] 
-\centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
-    \caption{
-	{\bf Time series duration of horizontal and vertical arm movements.} 
-    		Time series of smoothed data from gyroscope sensor 
-		for different speed arm movements performed by NAO: 
-		(A) Horizontal Normal arm movement (HN), 
-		(B) Horizontal Faster arm movement (HF),
-		(C) Vertical Normal arm movement (VN) and 
-		(D) Vertical Faster arm movement (VF).
-		Additionally, 
-		(A) shows window sizes for 2-seconds (100 samples), 
-		5-seconds (250 samples), 10-seconds (500 samples) 
-		and 15-seconds (750 samples)
-		which are also presented in (B), (C) and (D).
-		Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-	\label{fig:sts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-\subsection*{Data from Inertial Measurement Units} 
-	\label{sec:experiment:subsec:imu}
-
-To give insight to the research questions, 
-we considered various conditions of time series collected for this work 
-(see Experiment section for more details) which are described as follows
-\begin{itemize}
-\item Three levels of smoothness for the normalised data 
-	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
-	lengths (29 and 159) with the same polynomial degree 
-	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
-\item four velocities arm movement activities: 
-	horizontal normal (HN), horizontal faster (HF), 
-	vertical normal (VN), and vertical faster (VF), and
-\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
-	10-sec (500 samples) and 15-sec (750 samples).
-\end{itemize}
-%\subsection*{Time series}
-After the data collection, raw time series were windowed, normalised and 
-smoothed. Then, due to space limitations and to have simple visualisation, 
-we only present 10-sec (500 samples) window length time series for 
-three participants (p01, p01 and p03) performing horizontal 
-arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
-    	\caption{ 
-	{\bf Time series for horizontal and vertical arm movements.}
-		(A/D) raw-normalised (sg0zmuv), 
-		(B/E) normalised-smoothed 1 (sg1zmuv) and
-		(C/F) normalised-smoothed 2 (sg2zmuv).
-		Time series are only for three participants 
-		($p01$, $p02$, and $p03$) 
-		for horizontal and vertical arm movements in normal 
-		and faster velocity (HN, HF, VN, VF) 
-		with the normalised GyroZ or GyroY axis 
-		and with one sensor attached to the participant (HS01) 
-		and other sensor attached to the robot (RS01).	
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:ts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-\subsubsection*{Raw data}
-Considering the work of \cite{shoaib2016} which provided evidence of an 
-improvement in recognition activities when combining data from 
-accelerometer and gyroscope. We focus our analysis from data of the 
-accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
-leave the data of the magnetometer and quaternions for further investigation 
-because of their possible variations with regard to magnetic disturbances.
-
-Data from the accelerometer is defined by triaxial time series 
-$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
-(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
-which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
-representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
-Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
-its respective axes subscripts $x,y,z$, where $n$ is the sample index 
-and $N$ is the same maximum length of all axes for the time series.
-Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
-%%---------------------------------(EQUATION)----------------------------------
-\begin{equation}\label{eq:A}
-\boldsymbol{A} =
-\begin{pmatrix}
-  A_x(n) \\
-  A_y(n) \\
-  A_z(n)
-\end{pmatrix}
-=
-\begin{pmatrix}
- a_x(1),a_x(2),\dots,a_x(N) \\
- a_y(1),a_y(2),\dots,a_y(N) \\
- a_z(1),a_z(2),\dots,a_z(N) 
-\end{pmatrix},
-\end{equation}
-%%---------------------------------(EQUATION)-----------------------------------
-and 
-%%---------------------------------(EQUATION)-----------------------------------
-\begin{equation}\label{eq:G}
-\boldsymbol{G} =
-\begin{pmatrix}
- G_x(n) \\
- G_y(n) \\
- G_z(n)
-\end{pmatrix}
-=
-\begin{pmatrix}
- g_x(1),g_x(2),\dots,g_x(N) \\
- g_y(1),g_y(2),\dots,g_y(N) \\
- g_z(1),g_z(2),\dots,g_z(N) 
-\end{pmatrix}.
-\end{equation}
-%%---------------------------------(EQUATION)-----------------------------------
-
-\subsubsection*{Postprocessing data}
-After the collection of raw data from four NeMEMsi sensors,
-time synchronisation alignment and interpolation were performed 
-in order to create time series with the same length and synchronised time.
-We refer the reader to \cite{Comotti2014} for further
-details about the time synchronisation process.
-
-\subsubsection*{Data normalization}
-Data is normalised to have zero mean and unit variance 
-using sample mean and sample standard deviation.
-The sample mean and sample standard deviation using $x(n)$ is given by
-%%---------------------------------(EQUATION)----------------------------------
-\begin{equation}\label{eq:ms}
-\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad 
-	\sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
-\end{equation}
-%%---------------------------------(EQUATION)----------------------------------
-and the normalised data, $\hat{x}(n)$, is computed as follows
-%%---------------------------------(EQUATION)----------------------------------
-\begin{equation}\label{eq:normalization}
-\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
-\end{equation}
-%%---------------------------------(EQUATION)----------------------------------
-
-\subsubsection*{Smoothing data}
-Commonly, a low-pass filter is the method either to capture the low 
-frequencies that represent \%99 of the human body energy or to get 
-the gravitational and body motion components of 
-accelerations \cite{anguita2013}. However, for this work the elimination of 
-certain range of frequencies is not the main focus but the conservation of 
-the structure in the time series in terms of the width and heights where, 
-for instance, Savitzky-Golay filter can help to accomplish 
-such task \cite{press1992}. Savitzky-Golay filter is based on the 
-principle of moving window averaging which preserves the area under 
-the curve (the zeroth moment) and its mean position in time 
-(the first moment) but the line width (the second moment) is violated 
-and that results, for example, in the case of spectrometric data
-where a narrow spectral line is presented with reduced height and width. 
-With that in mind, the aim of Savitzky-Golay filtering is to find filter 
-coefficients $c_n$ that preserve higher momentums which are based on local 
-least-square polynomial approximations \cite{savitzkygolay1964, 
-press1992, schafer2011}.
-Therefore, Savitzky-Golay coefficients are therefore computed using 
-an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
-\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
-$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
-Smoothed signal is represented with a tilde over the original 
-signal: $\tilde{x}(n)$.
-
-\subsubsection*{Window size data}
-With regard to the window size, \cite{shoaib2016} investigated its effects 
-using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
-and combination of inertial sensors (accelerometer, gyroscope and linear 
-acceleration sensor) to improve the activity recognition performance for 
-repetitive activities (walking, jogging and biking) and less repetitive 
-activities (smoking, eating, giving a talk or drinking a coffee).
-With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
-increase of window length improve the recognition of complex activities 
-because these requires a large window length to learn the repetitive 
-motion patterns. Particularly, one of the recommendations is to use large 
-window size to recognise less repetitive activities which mainly involve 
-random hand gestures. Therefore, for the four activities 
-(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
-we select only four window sizes for analysis: 2-s window (100 samples), 
-5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
-(Fig~\ref{fig:sts}).
-
-%% As a guideline references should be limited to 60 (this is not 
-%% strictly enforced).
-%\bibliography{sample}
-%\bibliography{references}
-\bibliography{../references/references}
-
-\section*{Acknowledgements}
-
-This work was funded by the University of Birmingham and the 
-Mexican National Council of Science and Technology as part of 
-Miguel Xochicale's PhD degree. I would like to thank to Dr. Dolores 
-Columba Perez Flores and Prof. Martin J Russell for their helpful 
-comments to polish the use of the language of mathematics and to 
-Constantino Antonio Garcia Martinez et al. for developing the R 
-package nonlinearTseries that help to accelerate the 
-analysis of the nonlinear time series in this work.
-
-\section*{Author contributions statement}
-Contributions for this work of 
-Miguel Xochicale (MX) and Chris Baber (CB) are as follows:
-\begin{description}
-\item[Conceptualisation] MX, CB
-\item[Data Curation] MX
-\item[Formal Analysis] MX
-\item[Funding Acquisition] MX, CB
-\item[Investigation] MX
-\item[Methodology] MX
-\item[Project Administration] MX
-\item[Resources] CB
-\item[Software] MX
-\item[Supervision] CB
-\item[Validation] MX
-\item[Verification] MX
-\item[Writing - Original Draft Preparation] MX
-\item[Writing - Review] MX, CB
-\item[Writing - Editing] MX
-\end{description}
-
-%
-%\section*{Additional information}
-%
-%To include, in this order: \textbf{Accession codes} 
-%(where applicable); \textbf{Competing interests} (mandatory statement). 
-%
-%The corresponding author is responsible for submitting a 
-%\href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on 
-%behalf of all authors of the paper. This statement must be included 
-%in the submitted article file.
-%
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=\linewidth]{stream}
-%\caption{Legend (350 words max). Example legend text.}
-%\label{fig:stream}
-%\end{figure}
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{|l|l|l|}
-%\hline
-%Condition & n & p \\
-%\hline
-%A & 5 & 0.1 \\
-%\hline
-%B & 10 & 0.01 \\
-%\hline
-%\end{tabular}
-%\caption{\label{tab:example}Legend (350 words max). Example legend text.}
-%\end{table}
-%
-%Figures and tables can be referenced in LaTeX using the ref command, 
-%e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.
+\end{verbatim}
+
+\subsection{SUGGESTED TEXT}
+\begin{verbatim}
+In this paper we show how the selection of nonlinear analysis metric 
+(i.e., RSS, RP, RQA metrics) depends on what question one wishes to 
+address with time-series data (e.g., predictability, organisation, dynamics, complexity).  
+Time-series data characteristics (e.g., window length, smoothness), 
+time-series structure (e.g., frequency, amplitude) and 
+data source (e.g., sensor placement, performance, movement) all influence the results. 
+That said, it has been shown that the use of different characteristics of the data and 
+their collection can help us visualise and quantify variability of movement using 
+methods of nonlinear analysis.   
+There remain limitations of nonlinear methods in 
+relation to the estimation of parameters (e.g., recurrence threshold, embedding parameters) 
+which reflect the dynamics of specific movement and performers, window length 
+and structure of the time-series. 
+We note that RQA DET seems to show low sensitivity 
+to these differences, whereas REC and RATIO (primarily as a result of REC) show 
+variation across performers and movements.   
+RQA ENTR, with different recurrence thresholds, 
+can quantify variation in the time-series data and offers the most appropriate means 
+for analysing variability in movement to allow us to analyse individual differences 
+between human performers.
+
+This metric could be applied to analyse human participants 
+who might vary in age, state of health, anthropometric features and capability to perform movement."
+
+\end{verbatim}
+
+\subsection{FUSED TEXT}
+\begin{verbatim}
+
+\section*{Conclusions}
+In this paper we show how the selection of nonlinear analysis tool  
+(i.e., RSS, RP, RQA metrics) depends on what question one wishes to 
+address with time-series data (e.g., predictability, organisation, dynamics, complexity).  
+Time-series data characteristics (e.g., window length, smoothness), 
+time-series structure (e.g., frequency, amplitude) and 
+data source (e.g., sensor placement, performance, movement) all influence 
+the results that nonlinear analysis methods can provide.
+That said, it has been shown that the use of different characteristics of the data and 
+their collection can help us visualise and quantify variability of movement using 
+methods of nonlinear analysis.   
+There remain limitations of nonlinear methods in 
+relation to the estimation of parameters (e.g., recurrence threshold, embedding parameters) 
+which reflect the dynamics of specific movement and performers, window length 
+and structure of the time-series. 
+We note that RQA DET seems to show low sensitivity 
+to these differences, whereas REC and RATIO (primarily as a result of REC) show 
+variation across performers and movements.   
+RQA ENTR, with different recurrence thresholds, 
+can quantify variation in the time-series data and offers the most appropriate means 
+for analysing variability in movement to allow us to analyse individual differences 
+between human performers.  
+To then found out that RQA ENTR values with different recurrence 
+thresholds were appropriate to quantify 
+the different changes and variations of the characteristics of 
+time-series data.
+Therefore, the use of Shannon Entropy could be applied 
+to analyse human participants who might vary in age, 
+state of health, anthropometric features and capability to perform movement.
+
+\end{verbatim}
+
+\textit{
+SORTED: \\ 
+Sun  4 Oct 01:10:47 BST 2020 \\
+}
+\\
+
+\section{Abstract}
+
+\subsection{ORIGINAL TEXT}
+\begin{verbatim}
+
+\begin{abstract}
+Human movement variability occurs in motor performance 
+across multiple repetitions of a task and such behaviour is an
+inherent feature within and between each persons' movement. Quantifying
+movement variability is still an open problem, particularly 
+when methods in time domain, frequency domain or nonlinear dynamics 
+can break down due to the real-world time-series datasets. 
+For this work, we therefore investigate nonlinear dynamics methods 
+such as reconstructed state space (RSSs), uniform time-delay embedding, 
+recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
+with real-world time-series data.
+Particularly, we are interested in the weaknesses and robustness 
+of nonlinear dynamics tools with the use of raw and post-processed 
+data of wearable inertial sensors. 
+That said, twenty right-handed healthy participants 
+imitated simple vertical and horizontal arm movements in normal 
+and faster velocity from an humanoid robot in order to have 
+four window lengths and three levels of smoothed time-series data,
+to then found visual differences in the patterns with RSSs and RPs
+and particulatly the computed differences with RQA metrics
+that help us to quantify activities, types of sensors, windows lenghts 
+and level of smoothness. 
+Specifically, we can conclude that RQA ENTR, Shannon Entropy, 
+can lead to interesting results on the quantification of 
+movement variablity for participants of different ages, 
+state of health and anthropomorphic features to 
+then enhance the development of 
+better diagnostic tools for applications in rehabilitation, 
+sport science or for new forms of human-humanoid interaction.
+\end{abstract}
+
+\end{verbatim}
+
+\subsection{UPDATED TEXT}
+\begin{verbatim}
+
+\begin{abstract}
+Human movement variability arises from the process of 
+mastering redundant (bio)mechanical degrees of freedom (DOF)
+to successfully accomplish any given motor task
+(from the most complex to even the simplest one) 
+where flexibility and stability of many possible joint 
+combinations helps to adapt to environment conditions. 
+While the analysis of movement of variability is becoming increasingly 
+popular as a diagnostic tool or skill performance evaluation, 
+there are remain challenges in terms of defining 
+the most appropriate methods and parameters to apply. 
+For this work, we therefore investigate nonlinear dynamics methods 
+such as reconstructed state space (RSSs), uniform time-delay embedding, 
+recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
+with real-world time-series data of wearable inertial sensors. 
+That said, twenty right-handed healthy participants 
+imitated simple vertical and horizontal arm movements in normal 
+and faster velocity from an humanoid robot.
+We applied nonlinear methods to four activities, 
+four window lengths and three levels of smoothed time-series data,
+to then found visual differences in the patterns of RSSs and RPs
+and statistical differences with RQA metrics.
+We therefore conclude that Shannon Entropy for RQA is a robust method 
+that help us to quantify activities, types of sensors, windows lengths 
+and level of smoothness.
+Hence this work might enhance the development of 
+better diagnostic tools for applications in 
+rehabilitation and sport science for skill performance
+or new forms of human-humanoid interaction for
+quantification of movement adaptations and motor pathologies.
+\end{abstract}
+
+\end{verbatim}
+
+\textit{
+SORTED: \\ 
+Sun 18 Oct 05:44:59 BST 2020 \\
+}
+\\
 

### 2020-10-18 docs/manuscript/main.tex
adds corrections from 2020-07-03 and updates the manuscript 2020-10-08

@@
 
-\title{Recurrent Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
+\title{Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities}
 \author[1,*]{Miguel Xochicale}
@@
 	UK}
-\affil[*]{map479@bham.ac.uk}
+\affil[*]{email@user.com}
 
@@
 \begin{abstract}
-Human movement variability occurs in motor performance 
-across multiple repetitions of a task and such behaviour is an
-inherent feature within and between each persons' movement. Quantifying
-movement variability is still an open problem, particularly 
-when methods in time domain, frequency domain or nonlinear dynamics 
-can break down due to the real-world time-series datasets. 
+Human movement variability arises from the process of 
+mastering redundant (bio)mechanical degrees of freedom (DOF)
+to successfully accomplish any given motor task
+(from the most complex to even the simplest one) 
+where flexibility and stability of many possible joint 
+combinations helps to adapt to environment conditions. 
+While the analysis of movement of variability is becoming increasingly 
+popular as a diagnostic tool or skill performance evaluation, 
+there are remain challenges in terms of defining 
+the most appropriate methods and parameters to apply. 
 For this work, we therefore investigate nonlinear dynamics methods 
@@
 recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
-with real-world time-series data.
-Particularly, we are interested in the weaknesses and robustness 
-of nonlinear dynamics tools with the use of raw and post-processed 
-data of wearable inertial sensors. 
+with real-world time-series data of wearable inertial sensors. 
 That said, twenty right-handed healthy participants 
 imitated simple vertical and horizontal arm movements in normal 
-and faster velocity from an humanoid robot in order to have 
+and faster velocity from an humanoid robot.
+We applied nonlinear methods to four activities, 
 four window lengths and three levels of smoothed time-series data,
-to then found visual differences in the patterns with RSSs and RPs
-and particulatly the computed differences with RQA metrics
-that help us to quantify activities, types of sensors, windows lenghts 
+to then found visual differences in the patterns of RSSs and RPs
+and statistical differences with RQA metrics.
+We therefore conclude that Shannon Entropy for RQA is a robust method 
+that help us to quantify activities, types of sensors, windows lengths 
 and level of smoothness.
-Specifically, we can conclude that RQA ENTR, Shannon Entropy, 
-can lead to interesting results on the quantification of 
-movement variablity for participants of different ages, 
-state of health and anthropomorphic features to 
-then enhance the development of 
-better diagnostic tools for applications in rehabilitation, 
-sport science or for new forms of human-humanoid interaction.
+Hence this work might enhance the development of 
+better diagnostic tools for applications in 
+rehabilitation and sport science for skill performance
+or new forms of human-humanoid interaction for
+quantification of movement adaptations and motor pathologies.
 \end{abstract}
@@
 \begin{document}
-
 \flushbottom
@@
 \section*{Introduction}
-Human movement requires a complex system where not only multiple
-joints and limbs are involved for a specific task in a determined environment
-but also perception and action of movement that affects such 
-physical performance \cite{davids2003}. 
-In contrast, variability in humanoid movement is usually very small,
-as a result of mechanical and dynamic capabilities \cite{gouaillier2009}. 
-This means that, while humanoids 
-solve the degrees of freedom problem through join design and algorithms,
-humans tend to have more fluid and flexible approach. Consequently, 
-one can see much variability in human movement performance of even the simplest task.
-Studies of human motion reveal the possibility to 
-estimate features from lower dimension signals to distinguish differences between 
-styles of pedalling motion \cite{Quintana-Duque2012, Quintana-Duque2016}, 
-gait identification \cite{sama2013, frank2010},
-and detection of pathologies \cite{gomezgarcia2014}.
-The lower dimension signals from biological systems are time series 
-of $1-$dimension in $\mathbb{R}$ which commonly are 
+The complexity of human movement arises from the process of mastering 
+redundant (bio)mechanical degrees of freedom (DOF) to successfully 
+accomplish any given motor task (Bernstein, 1967).
+Such DOF are independent coordinates to uniquely 
+describe the configurations of the system which provides flexibility 
+and stability to adapt to a change environment conditions but 
+that leads many possible combinations (Newell and Vaillancourt, 2001).
+Hence, human movement complexity can be seen as a balance 
+between the required DOF "to generate a stable (persistent) 
+and flexible (variable) behavioural output in response to 
+changing intentions and dynamic environmental
+conditions" \cite{davids2003}. 
+Consequently, one can see much variability in human movement even in the 
+simplest of movements.  
+While the analysis of movement of variability is becoming increasingly 
+popular as a diagnostic tool or skill performance evaluation, 
+there remain challenges in terms of defining 
+the most appropriate methods and parameters to apply. 
+In part, 
+these challenges stem from the fact that the identification movement 
+variability requires analysis of signals which are time-series of 
+of $1-$dimension in $\mathbb{R}$ which are 
 noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
-Hence, nonlinear dynamics can be used to 
-objectively quantify variability of lower
-dimension signals \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
-frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
-For instance, Bradley et al. 2015 \cite{bradley2015} reviewed methods for
-nonlinear time series analysis based on the estimation of the 
+Further problems arise from assigning a plausible locus of control to 
+the movements, e.g., in order to determine whether variability is the 
+result of deliberative control by the performer or whether it arises 
+from exogenous or endogenous disturbances.  For this paper, our focus 
+is on the analysis of signals; specifically, in terms of objectively 
+quantifying variability of lower dimension signals using time-series analysis.
+
+Methods for nonlinear time series analysis generally involve the estimation of the 
 embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
 reconstruct the state space,
-where an $n$-dimensional reconstructed state space using $1-$dimensional 
-time series,
-can preserve the topological properties of an unknown $M$-dimensional 
-state space \cite{takens1981}.
-Similarly, Bradley et al. 2015 \cite{bradley2015} reviewed the use of 
-Recurrence Plots (RPs), a graphical representation of a two-dimensional map 
-which show black and white dots as recurrences in a given $n$-dimensional system, 
-and Recurrence Quantification Analysis (RQA) metrics that compute statistics in RPs.
-In general, RPs and RQA provide an intuitive meaning of the time series,
-for instance, RQA is quantitatively and qualitatively independent of 
-embedding dimension (also verified experimentally \cite{iwanski1998}).
-However, the estimation of embedding parameters and the selection of the 
-right parameters to perform RQA is still an open problem.
-Bradley et al. 2015 \cite{bradley2015} pointed out that
-there is no general technique that can be used to compute the embedding
-parameters since time series are system-dependent which means that
-computing the values for embedding parameters may only work for
-one purpose (e.g., prediction) and may not work well for another purpose
-(e.g., computing dynamical invariants).
-Additionally, methods of nonlinear dynamics for 
-computing embedding parameters e.g., autocorrelation, mutual information, 
-and nearest neighbour require data which is well 
-sampled and with little noise \cite{garland2016}
-or require signals that are purely deterministic \cite{kantz2003}.
-Thus, these methods of nonlinear analysis for computing the embedding 
-parameters can break down
-with real-world datasets which have generally different length, 
+where an $n$-dimensional is reconstructed state space using $1-$dimensional 
+time series \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
+frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
+Key to the selection of these parameters is the need to preserve 
+topological properties of an unknown $M$-dimensional state space \cite{takens1981}.
+An common approach to the construction of these state spaces 
+involves Recurrence Plots (RPs), a graphical representation of black and white dots, 
+which shows recurrent patterns of a $n$-dimensional system.
+While RPs provide a human
+interpretable picture of the system, these require further
+analysis to allow the properties of that system to be quantified,
+and so Recurrence Quantification Analysis (RQA) can be applied.
+However, the estimation of the embedding parameters for
+RQA remains an open problem (Bradley et al. 2015 \cite{bradley2015})
+There is no agreed method to estimate embedding parameters
+for RQA and other nonlinear methods \cite{bradley2015} 
+because time series are system-dependent, i.e., these rely
+on the initial conditions, and on the configuration and behaviour of the system.  
+This means that, unless one holds all of the influencing 
+variables constant, embedding parameters computed for one 
+instance may not apply to another instance.
+
+One could apply methods, such as autocorrelation, mutual information, 
+nearest neighbour (and we will apply these in this paper), 
+but these methods assume that the data are well sampled, with 
+little noise and (usually) that the signals are purely deterministic \cite{garland2016, kantz2003}.
+That said, these methods can break-down in the face of real-world datasets 
+which could have 
+different length, 
 different values of accuracy and precision (rounding errors due to finite 
 precision of the measurement apparatus which include frequency 
-acquisition \cite{frank2010}),
-and contaminated data with different sources of noise
-\cite{garland2016}.
-It is then surprising that even with the previous constraints with regard to 
-the quality of data, and the problem with the estimation of embedding parameters,
-the use of nonlinear dynamics have proven to be helpful to understand and 
-to characterise dynamics of time series 
-\cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
-gomezgarcia2014, gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.
-
-%Another point to consider with time series analysis using nonlinear dynamics
-%is the appropriate use of different post-processing techniques such as 
-%interpolation, filtering or normalisation.
-%With that in mind, it can be said that there is little research and 
-%understanding on the effects for post-processing techniques
-%in the interpretation of reconstructed state spaces, RPs and RQA.
-
-For this work, we are interested in exploring and investigating  
-the effects of different features of time series 
-(e.g. levels of smoothness, window data length,
-structures of time series based on types of movements, 
-types of sensors, participants and velocities) to estimate 
-embedded parameters for uniform time-delay embedding, 
-recurrence plots and recurrence quantification analysis.
-Hence, we conducted an experiment with twenty right-handed 
-healthy participants in the context of human-humanoid imitation 
-activities where participants were asked to 
-imitate simple arm movements performed by a humanoid robot.
-The primary aim of this work is to explore the following questions:
+acquisition \cite{frank2010}), or 
+different levels of contamination from exogenous and endogenous 
+sources of ‘noise’ \cite{garland2016}.  What is, perhaps, surprising is that 
+even subject to these problems, the methods are useful and 
+nonlinear dynamics approaches continue to tell us much about 
+human movement \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
+gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.
+
+For this paper, we explore the role of RPs and RQA in the 
+analysis of simple human movement. 
+We compare horizontal and 
+vertical arm movements performed by neurotypical participants 
+who are copying these movements being made by a humanoid robot.  
+This provides an initial point of comparison 
+(between human movement, which we assume to have a degree of variability, 
+and the robot, which we assume to have a degree of consistency).   
+Additionally, our interest lies in the estimation of embedding parameters 
+in light of different features of the data, e.g., levels of smoothing, 
+window length, structure of time-series based on movement, 
+types of sensors, individual differences between participants and 
+the movements that they perform. 
+Specifically, we ask what are 
+the effects of different embedding parameters, recurrence thresholds 
+and characteristics of time-series on nonlinear analysis 
+methods (i.e., reconstructed state space with 
+uniform time-delay embedding, RPs and RQA)?
+
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\section*{Methods}
+\subsection*{State Space Reconstruction}
+The method of state space reconstruction \cite{packard1980} 
+\cite{takens1981} has been applied in many disciplines 
+\cite{aguirre2009, stergiou2011, frank2010, 
+sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
+based on uniform time-delay embedding methodology which is a simple 
+matrix implementation that can reconstruct an unknown $d-$dimensional 
+manifold $M$ from a scalar time series (e.g. one-dimensional 
+time series in $\mathbb{R}$). A manifold, in this context, is a 
+multidimensional curved surface within a space (e.g. a saddle) 
+\cite{guastello-gregson2011}.
+
+The use of a scalar time series is the main advantage of the method of state 
+space reconstruction which in essence preserve dynamic invariants such as 
+correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
+entropy and detrended fluctuation analysis \cite{bradley2015, 
+Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
+However, selecting appropriate embedding parameters which are sued to apply the 
+state space reconstruction is still an open challenge for which we present 
+introductions for the methodologies to compute such embedding parameters 
+With that in mind, in the following subsections, we describe in more detail 
+the state space reconstruction theorem (RSSs), uniform time-delay embedding 
+theorem (UTDE), false nearest neighbours (FNN) and average mutual 
+information (AMI). 
+%and other methodologies for state space reconstruction.
+
+\subsection*{State Space Reconstruction Theorem}
+Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
+uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:ssr}
+  s(t)=f^t [s(0)],
+\end{equation}
+%%********************************[EQUATION]************************************
+where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
+$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
+unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
+function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
+iteration of $f$ that corresponds to an initial position 
+$s(0) \in M $ \cite{takens1981}.
+%$f$ is a smooth dynamical system, where smooth means a 
+%continuously differentiable system
+%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
+Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
+can be obtained with
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:measurement}
+  x(t)=h[s(t)],
+\end{equation}
+%%********************************[EQUATION]************************************
+where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
+Reconstructed state space can then be described as an $n-$dimensional state 
+space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
+$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
+is the uniform time-delay embedding with a dimension embedding $m$
+and delay embedding $\tau$ and
+$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
+of dimensionality (e.g. Principal Component Analysis, 
+Singular Value Decomposition, etc) being $n \leq m$.
+With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
+defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
+$\boldsymbol{X}(t) = \Phi(s(t))$,
+where $\Phi$ is a diffeomorphic map \cite{takens1981}
+whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
+dimension of $M$ \cite{garland2016}.
+% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
+% is called diffeomorphic if it is one-to-one correspondence and its inverse
+% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
+Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
+space then a composition of functions represented with $F^t$ is induced on the
+reconstructed state space determined:
+%by true smooth dynamical system $f^t$ and $\Phi$:
+%********************************[EQUATION]************************************
+\begin{equation}\label{eq:st}
+  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
+\end{equation}
+%%********************************[EQUATION]************************************
+With this in mind, an embedding is defined as "a smooth one-to-one 
+coordinate transformation with a smooth inverse" and the total reconstruction 
+map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
+Fig~\ref{fig:ssr} illustrates the state space reconstruction.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
+    \caption{
+	{\bf State space reconstruction methodology.}
+	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
+	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
+    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
+    	The time-delay embedding represented as the $\Phi$, maps the original
+    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
+	time-delay embedding matrix $\boldsymbol{X}(t)$.
+    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
+	a new state $y(t)$ of dimensions $n < m$.
+    	(A) $M-$dimensional manifold representing the state space 
+	(e.g. Lorenz system);
+    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
+   	(C) $m-$dimensional reconstructed state space with 
+    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
+    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
+	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
+	where $\Phi$ is the delay reconstruction map and 
+	$\Psi$ is the coordinate transformation map.
+	This figure is adapted from 
+   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+    }
+    \label{fig:ssr}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
+Frank et al. and Sama et al. refer to the state space reconstruction
+as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
+However, we consider the term "uniform time-delay embedding"
+as more descriptive and appropriate terminology for our work.
+Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed 
+copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
+the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
+samples of $\{ \boldsymbol{x}_n \}$.
+$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
+The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
+and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
+where $i$ goes from $0,1, \dots, (m-1)$.
+Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
+information of unobserved state variables and encapsulates the information of
+the delayed copies of the available time series in the uniform time-delay 
+embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
+$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:tde}
+\boldsymbol{X}^{m}_{\tau}  =
+\begin{pmatrix}
+\boldsymbol{ \tilde{x} }_n \\
+\boldsymbol{ \tilde{x} }_{n-\tau} \\
+\boldsymbol{ \tilde{x} }_{n-2\tau} \\
+\vdots \\
+\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
+\end{pmatrix}^\intercal, 
+\end{equation}
+%%********************************[EQUATION]************************************
+where $m$ is the embedding dimension, $\tau$ is the embedding delay and
+$ ^\intercal$ denotes the transpose.
+$m$ and $\tau$ are known as embedding parameters.
+The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
+$N-(m-1)\tau$ rows and $m$ columns and 
+$N-(m-1)\tau$ defines the length of each delayed copy 
+of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
+%For further details and explicit examples of uniform time-delay 
+%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
+
+\subsection*{Estimation of Embedding Parameters}
+The estimation of the embedding parameters ($m$ and $\tau$) 
+is a fundamental step for the state space reconstruction with the use
+of uniform time-delay embedding method.
+With this in mind, we review two of the most common algorithms,
+which will be used in our work, to compute the embedding
+parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%*******************************************************************************
+\subsubsection*{False Nearest Neighbours}
+To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
+used the method of false neighbours which can be understood as follows:
+on the one hand, when the embedding dimension is too small to unfold the 
+attractor "not all points that lie close each other will be neighbours and 
+some points appear as neighbours because of the attractor has been projected 
+down into an smaller space", on the other hand, when increasing the embedding 
+dimension "points that are near to each other in the sufficient
+embedding dimension should remain close as the dimension increase from $m$ 
+to $m+1$ \cite{krakovska2015}".
+From a mathematical point of view, the state space reconstruction theorem is 
+done when the attractor is unfolded with either the minimum embedding 
+dimension, $m_0$, or any other embedding dimension value where 
+$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
+leads to excessive computations \cite{bradley2015}. With this in mind, 
+Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
+where only the time-series and one delay embedding value are necessary 
+to select the minimum embedding dimension. Cao's algorithm is based 
+on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:e}
+  \begin{aligned}
+	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
+	 &=
+       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
+       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
+            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
+  \end{aligned}
+\end{equation}
+%%********************************[EQUATION]************************************
+where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
+embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
+From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
+and $\tau$ for which $E_1(m)$ is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:e1}
+E_1(m) = \frac{ E(m+1) } { E(m)}.
+\end{equation}
+%%********************************[EQUATION]************************************
+$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
+in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
+As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
+than some $m_0$, if the time series comes from a multidimensional state space
+then $m_0 + 1$ is the minimum dimension".
+Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
+stochastic signals. $E_2(m)$ is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:e2}
+E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
+\end{equation}
+%%********************************[EQUATION]************************************
+where
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:ee}
+E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
+|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
+\end{equation}
+%%********************************[EQUATION]************************************
+For instance, when the signal comes from random noise (values that are 
+independent from each other), all $E_2(m)$ values are approximately equal 
+to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
+not constant for all $m$ (e.g. $E_2(m) \neq 1$).
+
+As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
+series: the solution for the $x$ variable of the Lorenz system 
+(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
+and a variance of one (Fig~\ref{fig:e1e2}F).
+We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
+The $E_1(m)$ values for the chaotic time series appear to be constant
+after the dimension is equal to six.
+The determination of six is given that any value of $m$ can be used as they 
+are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
+$E_2(m)$ values, for chaotic time series, are different to one 
+(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
+chaotic time series the minimum embedding dimension the time series 
+comes from a deterministic signal. With regard to the noise time series,  
+$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
+which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
+$E_1(m)$ values then indicate the minimum embedding dimension of the 
+noisy time series is thirteen, however all of the $E_2(m)$ values are 
+approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
+concluded that noise time series is a stochastic signal.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
+    \caption{
+	{\bf Minimum dimension embedding values with Cao's method.} 
+	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
+	with variations of $\tau$ values from one to twenty
+	for (E) chaotic and (F) random time series.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:e1e2}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
+computed but also a variation of $\tau$ from 1 to 20 is presented. 
+The purpose of such variation for $\tau$ is to show its independence with
+regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
+(Fig~\ref{fig:e1e2}A,B,C, and D). 
+However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
+a new threshold where $m$ values appear to be constant in $E_1 (m)$.
+In the case of the given examples and reported results, we defined such 
+threshold as 0.05. Further investigation is required for the selection of the 
+threshold in the $E_1(m)$, as the selection of the threshold in this work is 
+base on no particular method but visual inspection.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%*******************************************************************************
+\subsubsection*{Average Mutual Information}
+When selecting the delay dimension parameter, $\tau$, one can consider the 
+following two cases:
+(i) when $\tau$ is too small, the elements of time-delay embedding will be along
+the bisectrix of the phase space and the reconstruction is generally not 
+satisfactory, 
+(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
+time-delay embedding will become spread and uncorrelated which makes 
+recovering the underlying attractor more difficult if not 
+impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
+With regard to the algorithms to compute $\tau$, 
+Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
+function in which the first zero crossing is considered as the minimum delay 
+embedding parameter. However, the autocorrelation function is a linear 
+statistic over which the Average Mutual Information (AMI) algorithm is 
+preferred because the AMI takes into account the nonlinear dynamical 
+correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
+algorithm is described below to estimate the minimum delay embedding 
+parameter, \texorpdfstring{$\tau_0$}{T}.
+
+To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
+and then a probability distribution of data is computed \cite{kantz2003}.
+AMI is therefore denoted by $I(\tau)$ which is the average mutual 
+information between the original time series, $x(n)$, and the delayed 
+time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
+AMI is defined by
+%%********************************[EQUATION]************************************
+\begin{equation}\label{eq:ami}
+I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
+\end{equation}
+%%********************************[EQUATION]************************************
+Probabilities are defined as follows:
+$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
+the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
+the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
+that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
+in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
+For small $\tau$, AMI will be large and it will then decrease more or 
+less rapidly. As $\tau$ increase and goes to a large limit, 
+$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
+factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
+to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
+where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
+where the redundancy is the least" \cite{kantz2003}.
+
+For example, we compute the AMI for two time series:
+A) the $x$ solution of the Lorenz system, and
+B) a noise time series using a normal distribution with mean zero and 
+standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
+concluded that the amount of knowledge for any noise time series is zero 
+for which the first minimum embedding parameter is $\tau_0=1$. 
+On the contrary, the first minimum of the AMI for the chaotic time series 
+is $\tau_0=17$ which is the value that maximize the independence 
+between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
+\cite{bradley2015}.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
+    \caption{
+	{\bf Minimum delay embedding values with AMI's method.} 
+    	(A, B) AMI values where its first minimum value in the curve
+	is the minimum time delay embedding ($\tau_0$), 
+	for (C) a chaotic and (D) noise time series.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:amis}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+Similarly as Cao's algorithm negatives, AMI's algorithm is not an
+exception for negatives, which are worthwhile to mention for further 
+investigations.
+For instance, (i) is not clear why the choose of the first minimum of the AMI 
+is the minimum delay embedding parameter \cite{kantz2003} and 
+(ii) the probability distribution of the AMI function is computed
+with the use of histograms which depends on a heuristic choice of number of bins
+for which AMI depends on partitioning \cite{garcia2005e71}.
+
+%*******************************************************************************
+\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
+\subsection*{Recurrence Plots}
+Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
+in conservative systems, however such discovery was not put into practice 
+until the development of faster computers \cite{marwan2007},
+for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
+where recurrences in the dynamics of a system can be visualised using 
+Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
+propose a tool, called Recurrence Plot (RP), that provides insights into 
+high-dimensional dynamical systems where trajectories are very difficult to 
+visualise. Therefore, "RP helps us to investigate the 
+$m-$dimensional phase space trajectories through a two-dimensional 
+representation of its recurrences" \cite{marwan2015}.
+Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
+to the methodologies of the state space reconstruction and other dynamic 
+invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
+the recurrences of the trajectories in the phase space can provide 
+important clues to characterise natural process that present, for
+instance, periodicities (as Milankovitch cycles) or irregular cycles 
+(as El Ni\~no Southern Oscillation). 
+Such recurrences can not only be presented visually using Recurrence Plots (RP) 
+but also be quantified with Recurrence Quantification metrics, which leads 
+to applications of these in various areas such as economy, physiology, 
+neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
+
+For the creation of a recurrence plot based on time series 
+$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
+reconstruction with uniform time-delay embedding 
+$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
+\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
+where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
+and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
+The recurrence plot is therefore a two-dimensional $N \times N$ square 
+matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
+whenever $X(i)$ is sufficiently close to $X(j)$: 
+%%********************************[EQUATION]************************************
+\begin{equation}
+\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
+\end{equation}
+%%********************************[EQUATION]************************************
+where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
+$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
+function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
+(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
+RP is also characterised with a line of identity (LOI) which is a  
+diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
+%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
+    \caption{
+	{\bf Recurrence Plots.} 
+	(A) State space of the Lorenz system with controlling parameters 
+	($\rho=28, \sigma=10, \beta=8/3$). 
+	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
+	(black circle) of a given point at $i$ is a recurrent point and is 
+	represented as a black dot in the recurrence plot at location 
+	$(i, j)$ or white otherwise.
+	(B) Recurrence plot using the 
+	three components of the Lorenz system and the RP with no embeddings 
+	and threshold $\epsilon=5$.
+	This figure is adapted from \cite{marwan2015}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+	}
+    \label{fig:mrp}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\subsection*{Structures of Recurrence Plots}
+Pattern formations in the RPs can be designated either 
+as topology for large-scale patterns or texture for small-scale patterns.
+In the case of topology, the following pattern formations are commonly presented:
+(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
+uniformly distributed noise (Fig~\ref{fig:rp2}A), 
+(ii) periodic and quasi-periodic systems where diagonal lines and 
+checkerboard structures represent oscillating systems, e.g., sinusoidal 
+signals (Fig~\ref{fig:rp2}B), 
+(iii) drift where paling or darkening recurrence points away from 
+the LOI is caused by drifting systems, 
+e.g., logistic map (Fig~\ref{fig:rp2}C), and
+(iv) disrupted where recurrence points are presented white areas or 
+bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
+(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
+Texture patterns in RPs can be categorised as:
+(i) single or isolated recurrence points that represent rare occurring states, 
+and do not persist for any time or fluctuate heavily,
+(ii) dots forming diagonal lines where the length of the small-scale parallel 
+lines in the diagonal are related to the ratio of determinism or predictability 
+in the dynamics of the system, and
+(iii) dots forming vertical and horizontal lines where the length of the 
+lines represent a time length where a state does not change or change very 
+slowly and these patterns formation represent discontinuities in the signal, 
+and (iv) dots clustering to inscribe rectangular regions which are related 
+to laminar states or singularities \cite{marwan2015}.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
+    \caption{
+	{\bf Patterns in Recurrence Plots.} 
+	Time-series with its respective recurrence plot for:
+	(A) uniformly distributed noise,
+	(B) super-positionet harmonic oscillation 
+	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
+	with a linearly increase term ($0.01*i$), and
+	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	This figure is adapted from \cite{marwan2015}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+	}
+    \label{fig:rp2}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
+
+Although, each of the previous pattern descriptions of the structures in the 
+RP offer an idea of the characteristics of dynamical systems, 
+these might be misinterpreted and conclusions might tend to be subjective 
+as these require the interpretation of a particular researcher(s).
+Because of that, recurrence quantification analyis (RQA) offer objective 
+methodologies to quantify such visual characteristics of previous 
+recurrent pattern structures in the RP \cite{zbilut1992}.
+
+%*******************************************************************************
+\subsection*{Recurrence Quantifications Analysis (RQA)}
+Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
+the density of recurrence points in RPs, then histograms of lengths for 
+diagonal lines in RPs were studied by \cite{trulla1996} which were the 
+introduction to the term recurrence quantification analysis (RQA) 
+\cite{marwan2008}. RQA has been applied in many fields such as life science, 
+engineering, physics, and others \cite{marwan2008}. Particularly in human 
+movement to investigate noise and complexity of postural control 
+\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
+coordination \cite{duran2017}. The success of RQA is not only due to its 
+simple algorithmic implementation but also to its capacity to detect tiny 
+modulations in frequency or phase which are not detectable using standard 
+methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
+RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
+dimension which is verified experimentally by \cite{iwanski1998}.
+RQA metrics comprehend percentage of recurrence, percentage of determinism, 
+ratio, Shannon entropy of the frequency distributions of the line lengths,
+maximal line length and divergence, trend and laminariy 
+\cite{marwan2007, marwan2015}. For this work, we considered only four 
+RQA metrics, due to its consistency with our preliminary experiments, 
+which are described below. Such metrics are computed the nonlinearTseries 
+R package \cite{nonlinearTseries2016}.
+
+\subsubsection*{REC values}
+The percentage of recurrence (REC) is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}
+REC(\epsilon,N) = 
+		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
+		\mathbf{R}^{m}_{i,j}(\epsilon),
+\end{equation}
+%%********************************[EQUATION]************************************
+which enumerates the black dots in the RP excluding the line of identity.
+REC is a measure of the relative density of recurrence points in the sparse 
+matrix \cite{marwan2015}.
+%REC is computed as follow with the nonlinearTseries package 
+% \cite{nonlinearTseries} 
+%  hist = getHistograms(neighs, ntakens, lmin, vmin)
+%  # calculate the number of recurrence points from the recurrence rate. The
+%  # recurrence rate counts the number of points at every distance in a concrete
+%  # side of the main diagonal.
+%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
+%  # add the maindiagonal
+%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
+%  # calculate the recurrence rate dividing the number of recurrent points at a
+%  # given distance by all points that could be at that distance
+%  recurrence_rate_vector = 
+%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
+%  # percentage of recurrent points
+%  REC = (numberRecurrencePoints) / ntakens ^ 2
+
+\subsubsection*{DET values} 
+The percent determinism (DET) is defined as the fraction of recurrence points
+that form diagonal lines and it is determined by
+%%********************************[EQUATION]************************************
+\begin{equation}
+DET=
+	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
+	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
+\end{equation}
+%%********************************[EQUATION]************************************
+where 
+%%********************************[EQUATION]************************************
+\begin{equation}
+H_D(l) = 
+	\sum^{N}_{i,j=1} 
+	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
+	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
+	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
+\end{equation}
+%%********************************[EQUATION]************************************
+is the histogram of the lengths of the diagonal structures in the RP.
+DET can be interpreted as the predictability of the system for periodic signals 
+which, in essence, have longer diagonal lines than the short diagonals lines
+for chaotic signals or absent diagonal lines for stochastic signals 
+\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
+the organisation of points in RPs  \cite{iwanski1998}. 
+%percent determinism (DET) is computed as follow with the nonlinearTseries 
+%package \cite{nonlinearTseries}  
+% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
+%                                       lmin = 2, lDiagonalHistogram,
+%                                       recurrence_rate_vector, maxDistanceMD) {
+%  #begin parameter computations
+%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
+%  DET = num / numberRecurrencePoints
+
+\subsubsection*{RATIO values}
+RATIO is defined as the ratio between DET and REC and it is calculated from 
+the frequency distributions of the lengths of the diagonal lines.
+RATIO is useful to discover dynamic transitions \cite{marwan2015}.
+%  diagP = calculateDiagonalParameters(
+%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
+%    recurrence_rate_vector, maxDistanceMD
+%  )
+% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
+%                                       lmin = 2, lDiagonalHistogram,
+%                                       recurrence_rate_vector, maxDistanceMD) {
+%  #begin parameter computations
+%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
+%  DET = num / numberRecurrencePoints
+% 
+%
+%    RATIO = diagP$DET / REC
+%
+
+\subsubsection*{ENT values}
+ENT is the Shannon entropy of the frequency distribution of the diagonal 
+line lengths and it is defined as
+%%********************************[EQUATION]************************************
+\begin{equation}
+ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with 
+	\quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
+\end{equation}
+%%********************************[EQUATION]************************************
+ENT reflects the complexity of the deterministic structure in the system.
+For instance, for uncorrelated noise or oscillations, 
+the value of ENT is rather small and indicates low complexity of the system,
+therefore "the higher the ENT is the more complex the dynamics are" 
+\cite{marwan2007, marwan2015}.
+%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
+%
+%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
+%                                       lmin = 2, lDiagonalHistogram,
+%                                       recurrence_rate_vector, maxDistanceMD) {
+%
+%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
+%  diff_0 = which(pl > 0)
+%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
+
+\subsection*{Sensitivity and robustness of RPs and RQA.}
+RP and RQA are a very young field in nonlinear dynamics and many questions 
+are still open, for instance, different parameters for window length size 
+of the time series, embedding parameters or recurrence threshold can 
+generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
+
+The selection of recurrence threshold, $\epsilon$, can depend on the system 
+that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
+require to be very small, for trajectory reconstruction $\epsilon$ requires 
+to have a large thresholds or when studying dynamical transition 
+there is little importance about the selection of the threshold 
+\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
+the recurrence threshold  should be five times larger 
+than the standard deviation of the observational noise
+or the use of diagonal structures within the RP is suggested in order
+to find the optimal recurrence threshold for (quasi-)periodic process 
+\cite{marwan2011}.
+Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
+of choosing the right embedding parameters to perform RQA for which 
+many experiments have to be performed using different parameters in order 
+to have a better intuition of the nature of the time series and how 
+this is represented by using RQA.
+
+With that in mind, this work explores the sensitivity and robustness 
+of the window size of time series, embedding parameters for RSS with UTDE 
+and recurrence threshold for RP and RQA in order to gain a better insight 
+into the underlying time series collected from inertial sensors in the 
+context of human-humanoid imitation activities.
+%Such changes of recurrence threshold values 
+%can modify the patterns in RPs and therefore the values of RQA metrics.
+%We therefore computed 3D surfaces to explore the sensibility and robustness of 
+%embedding parameters and recurrence threshold in RQA  metrics. Following the 
+%same methodology of computing 3D surfaces, we also considered variation of 
+%window length size to present RQA metrics dependencies with embedding 
+%parameters, recurrence thresholds and window length size.
+
+%Choosing an appropriate recurrence threshold is crucial so as to get 
+%meaningful representations in RP and RQA, however,
+
+%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
+%metrics for RQA are independent of embedding dimension parameters, 
+%however, that is not the case when using 
+%different recurrence thresholds.
+
+%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
+%determining embedding parameters, their method consisted on creating 3D 
+%surfaces with RQA metrics with an increase of embedding parameters 
+%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
+%fluctuations and gradual changes in the 3D surfaces that provide information 
+%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
+%created 3D surfaces for visual selection of not only embedding parameters 
+%but also recurrence thresholds. Following same methodologies, we explored 
+%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
+%using 3D surfaces by an unitary increase of the pair embedding 
+%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
+%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
+%(Fig.~\ref{fig:topo_rqas}).
+
+%*******************************************************************************
+%*******************************************************************************
+%*******************************************************************************
+\section*{Experiment} \label{sec:experiment}
+We conducted an experiment in the context of human-humanoid imitation (HHI) 
+activities where participants were asked to imitate simple horizontal and 
+vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
+Such simple movements were repeated ten times for the participant 
+who copied NAO's arm movements in a face-to-face imitation activity.
+Also, wearable inertial measurement unit (IMU) sensors were attached 
+to the right hand of the participant and to the left hand of the robot 
+(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
+IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
+accelerometer, gyroscope and magnetometer sensors and quaternions 
+\cite{Comotti2014}.
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+  \centering
+\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
+    \caption{
+	{\bf Human-humanoid imitation activities.} 
+    		Face-to-face human-humanoid imitation (HHI) activities for 
+		(A) HHI of horizontal arm movement, 
+		(B) Humanoid horizontal arm movement,
+		(C) HHI of vertical arm movement, and 
+		(D) Humanoid vertical arm movement.
+        }
+    \label{fig:hri}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+\subsection*{Participants}
+Twenty-three participants,
+from now on defined as $pN$ where $N$ is the number of participant, were 
+invited to do the experiment. However, data for three participants were 
+not used because the instructions for $p01$, who was the only left-handed,
+were mistakenly given in a way that movements were performed different
+from what had been planned, and for participants $p13$ and $p16$ 
+data were corrupted because of problems with Bluetooth communications with sensors. 
+With that in mind, data for twenty participants were analysed in this work.
+Of the twenty participants, all of them are right-handed healthy participants 
+of whom four are females and sixteen are males with a mean and standard 
+deviation (SD) age of mean=19.8 (SD=1.39).
+
+\subsection*{Ethics}
+All participants provided informed consent forms prior to participation 
+in the experiment. Hence, the experiments were conducted in November 2016 and 
+participants confirmed reading and understanding the participant information sheet of the 
+experiments and were able to withdraw from the experiment at any time without giving
+any reason. The design of the experiments adhered to the University of Birmingham
+regulations, data were anonymised and videos were stored only on a 
+personal computer in accordance with the Data Protection Act 1998. 
+%Refer to Appendix D for further information about the ethics, 
+%online participation information sheets and experiment check list.
+
+\subsection*{Human-humanoid imitation activities}
+For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
+two of which were attached to the right hand of the participant and the 
+other two to the left hand of the humanoid robot.
+Then, each participant was asked to imitate repetitions of simple horizontal
+and vertical arm movements performed by the humanoid robot in the following 
+conditions:
+(i) ten repetitions of horizontal arm movement at normal (HN) 
+	and faster (HF) speed (Figure~\ref{fig:hri} A), and
+(ii) ten repetitions of vertical arm movement at normal (VN) 
+	and faster (VF) speed (Figure~\ref{fig:hri} C).
+The normal and faster speed of arm movements is defined by the duration 
+in number of samples of one repetition of NAO's arm movements. 
+We select NAO's arm movements duration to distinguish between normal and 
+faster arm movements as NAO's movements have less variation 
+for such repetitive movements. 
+The duration for one repetition of the horizontal 
+arm movement at normal speed, HN, is 5 seconds considering 
+that each repetition last around 250 samples.
+For horizontal arm movement at faster speed, HF, each repetition were performed 
+in 2 seconds which correspond to 90 samples of data.
+The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
+which is around 300 samples of data. For vertical arm movement at 
+faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
+to 120 samples of data.
+To visualise the distinction between normal and faster speed for horizontal 
+and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
+for axes Z and Y of the gyroscope sensors with four window lengths: 
+2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
+and 15-sec (750-samples).
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht] 
+\centering
+\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
+    \caption{
+	{\bf Time series duration of horizontal and vertical arm movements.} 
+    		Time series of smoothed data from gyroscope sensor 
+		for different speed arm movements performed by NAO: 
+		(A) Horizontal Normal arm movement (HN), 
+		(B) Horizontal Faster arm movement (HF),
+		(C) Vertical Normal arm movement (VN) and 
+		(D) Vertical Faster arm movement (VF).
+		Additionally, 
+		(A) shows window sizes for 2-seconds (100 samples), 
+		5-seconds (250 samples), 10-seconds (500 samples) 
+		and 15-seconds (750 samples)
+		which are also presented in (B), (C) and (D).
+		Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+	\label{fig:sts}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+\subsection*{Data collection from inertial measurement units} 
+\label{sec:experiment:subsec:imu}
+To give insight to the research questions, 
+we considered various conditions of time series collected for this work: 
 \begin{itemize}
-\item What are the effects on three nonlinear methods 
-	(reconstructed state space with uniform time-delay embedding, 
-	Recurrence Plots, and Recurrence Quantification Analysis), 
-	for different embedding parameters, different recurrence thresholds 
-	and different characteristics of time series 
-	(structure of the signal, window length and smoothness of the signal)?, and 
-\item What are the strengths and weaknesses of Recurrence
-	Quantification Analysis for the previous conditions
-	of nonlinear analysis methods?
+\item Three levels of smoothness for normalised data 
+	(e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg 
+	stands for Savitzky-Golay filter with two different filter lengths (29 and 159) 
+	and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal}
+	and zmuv is zero mean unit variance.
+\item four arm movement activities with two velocities: 
+	horizontal normal (HN), horizontal faster (HF), 
+	vertical normal (VN), and vertical faster (VF), and
+\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
+	10-sec (500 samples) and 15-sec (750 samples).
 \end{itemize}
+%After the data collection, raw time series were windowed, normalised and smoothed. 
+Due to space limitations and to have simple visualisation, 
+we only present 10-sec (500 samples) window length time series for 
+three participants (p01, p01 and p03) performing horizontal 
+arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}).
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
+    	\caption{ 
+	{\bf Time series for horizontal and vertical arm movements.}
+		(A/D) raw-normalised (sg0zmuv), 
+		(B/E) normalised-smoothed 1 (sg1zmuv) and
+		(C/F) normalised-smoothed 2 (sg2zmuv).
+		Time series are only for three participants 
+		($p01$, $p02$, and $p03$) 
+		for horizontal and vertical arm movements in normal 
+		and faster velocity (HN, HF, VN, VF) 
+		with the normalised GyroZ or GyroY axis 
+		and with one sensor attached to the participant (HS01) 
+		and other sensor attached to the robot (RS01).	
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:ts}
+\end{figure}
+%%---------------------------------(FIGURE)------------------------------------
+
+\subsubsection*{Raw data}
+%Considering the work of \cite{shoaib2016} which provided evidence of an 
+%improvement in recognition activities when combining data from 
+%accelerometer and gyroscope. 
+We focus our analysis from data of the 
+accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
+leave the data of the magnetometer and quaternions for further investigation 
+because of their possible variations with regard to magnetic disturbances \cite{shoaib2016}.
+Data from the accelerometer is defined by triaxial time series 
+$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
+(Eq.~\ref{eq:AG}), and the same for data from the gyroscope 
+which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
+representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:AG}).
+Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
+its respective axes subscripts $x,y,z$, where $n$ is the sample index 
+and $N$ is the same maximum length of all axes for the time series.
+%Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
+%%---------------------------------(EQUATION)----------------------------------
+\begin{equation}\label{eq:AG}
+\boldsymbol{A} =
+\begin{pmatrix}
+  A_x(n) \\
+  A_y(n) \\
+  A_z(n)
+\end{pmatrix}
+=
+\begin{pmatrix}
+ a_x(1),a_x(2),\dots,a_x(N) \\
+ a_y(1),a_y(2),\dots,a_y(N) \\
+ a_z(1),a_z(2),\dots,a_z(N) 
+\end{pmatrix},
+%\end{equation}
+%%---------------------------------(EQUATION)-----------------------------------
+%and 
+%%---------------------------------(EQUATION)-----------------------------------
+%\begin{equation}\label{eq:G}
+\boldsymbol{G} =
+\begin{pmatrix}
+ G_x(n) \\
+ G_y(n) \\
+ G_z(n)
+\end{pmatrix}
+=
+\begin{pmatrix}
+ g_x(1),g_x(2),\dots,g_x(N) \\
+ g_y(1),g_y(2),\dots,g_y(N) \\
+ g_z(1),g_z(2),\dots,g_z(N) 
+\end{pmatrix}.
+\end{equation}
+%%---------------------------------(EQUATION)-----------------------------------
+
+\subsubsection*{Postprocessing data}
+After the collection of raw data from four NeMEMsi sensors,
+time synchronisation alignment and interpolation were performed 
+in order to create time series with the same length and synchronised time.
+We refer the reader to \cite{Comotti2014} for further
+details about the time synchronisation process.
+
+\subsubsection*{Normalising data}
+Data is normalised to be zero mean and unit variance.
+The sample mean and sample standard deviation using $x(n)$ is given by
+%%---------------------------------(EQUATION)----------------------------------
+\begin{equation}\label{eq:ms}
+\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad 
+	\sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
+\end{equation}
+%%---------------------------------(EQUATION)----------------------------------
+and the normalised data, $\hat{x}(n)$, is computed as follows
+%%---------------------------------(EQUATION)----------------------------------
+\begin{equation}\label{eq:normalization}
+\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
+\end{equation}
+%%---------------------------------(EQUATION)----------------------------------
+
+\subsubsection*{Smoothing data}
+Commonly, a low-pass filter is use either to capture the low 
+frequencies that represent \%99 of the human body energy or to get 
+the gravitational and body motion components of 
+accelerations \cite{anguita2013}. However, the elimination of 
+a range of frequencies is not the main focus of this work but 
+the conservation of the structure of time series in terms of 
+their width and heights where, for instance, Savitzky-Golay filter 
+can be a way to accomplish such task \cite{press1992}. 
+Savitzky-Golay filter is based on the 
+principle of moving window averaging which preserves the area under 
+the curve (the zeroth moment), its mean position in time 
+(the first moment) but the line width (the second moment) is violated 
+and that results, for example, in the case of spectrometric data, into
+a narrow spectral line with reduced height and width. 
+With that in mind, the aim of Savitzky-Golay filtering is to find filter 
+coefficients $c_n$ that preserve higher momentums which are based on local 
+least-square polynomial approximations \cite{savitzkygolay1964, 
+press1992, schafer2011}.
+Therefore, Savitzky-Golay coefficients are computed with \texttt{sgolay(p,n,m)} 
+in R where \texttt{p} is the filter order, 
+\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
+$m$-th derivative of the filter coefficients \cite{Rsignal}. 
+Smoothed signal is represented with a tilde over the original 
+variable of the signal: $\tilde{x}(n)$.
+
+\subsubsection*{Windowing data size}
+With regards to the window size, Shoaib et al. in 2016 \cite{shoaib2016} 
+investigated effects of seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
+and combination of inertial sensors (accelerometer, gyroscope and linear 
+acceleration sensor) to improve the activity recognition performance for 
+repetitive activities (walking, jogging and biking) and less repetitive 
+activities (smoking, eating, giving a talk or drinking a coffee).
+With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
+increase of window length improve the recognition of complex activities 
+because these requires a large window length to learn the repetitive 
+motion patterns. Particularly, one of the recommendations is to use large 
+window size to recognise less repetitive activities which mainly involve 
+random hand gestures. Therefore, for the four activities 
+(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
+we select only four window sizes for analysis: 2-s window (100 samples), 
+5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
+(Fig~\ref{fig:sts}).
 
@@
 \subsection*{Reconstructed State Spaces}
-One of the challenges of the implementation of 
-uniform time-delay embedding to reconstruct the state spaces is the 
-selection of embedding parameters because each time series is unique in 
-terms of its structure (e.g. modulation of amplitude, frequency and phase) 
-\cite{ frank2010, sama2013, bradley2015}.
-With that in mind, the problem for this work is not to compute individual 
-embedding parameters for each of the time series but to deal with a 
-selection of two parameters that can represent all the time series. 
-Our solution for such problem, as explained in below sections, 
-was to compute a sample mean over all 
-values in each of the conditions of the time series 
-%of Figs~\ref{fig:ts}
-for minimum dimension values and 
-%Figs~\ref{fig:caoami} 
-for minimum delay values.
-%, resulting in an average minimum embedding parameters 
-%of ($\overline{m_0}=6$, $\overline{\tau_0}=8$).
+As noted in the Introduction, a challenge in the implementation of uniform time-delay embedding 
+arises from the selection of embedding parameters because of the uniqueness of each time series 
+in terms of its structure (e.g., modulation of amplitude, frequency, phase etc.). 
+With that in mind, the options are to either calculate embedding parameters for each unique 
+instance (which can make comparison challenging) or to find parameters which can apply to all 
+instances in the study.
+We recognise that the latter approach is not without its problems, 
+but our approach is to compute sample mean over all values in each of the conditions of the 
+time series for minimum dimension and minimum delay values.
 
 \subsubsection*{Minimum embedding parameters}
-Minimum embedding parameters were firstly computed 
-with the methods of False Nearest Neighbour and Average Mutual Information.
-%for Uniform Time-Delay Embedding (UTDE),
-%Reconstructed State Space (RSS), 
-%Recurrence Plots (RPs) and 
-%Recurrence Quantification Analysis (RQAs).
-Figures \ref{fig:cao_ami}(A) show box plots for the minimum embedding values 
-for sensors HS01 and RS01. Minimum embedding vaues for HS01 presents more 
-variations than RS01 as interquartile range for HS01 is near to 1 (with the exceptions of HF sg0, VN sg0, and VF sg1)
-whereas interquartile range for RS01 is near to 0.1 (with the exception of two axis HF sg0 and VF sg1).
-Additionally, Figures \ref{fig:cao_ami}(A) show
-a decrease of mean values (rhombus) in the box plots
-as smoothness of time series increase (see sg0, sg1, sg2).
-Figures \ref{fig:cao_ami}(B) show box plots for the first minimum values of AMI values.
-It can be seen that values for HS01 tend to be more spread as the smoothness 
-of the time series is increasing 
-(see the increase of both mean (rhombus) and interquartile range).
-However, AMI values for RS01 do not show such increase in relation with
-the increase of smoothness except for HF and VF
-(Figs \ref{fig:cao_ami}(B)).
-
-We also computed a sample mean for an overall minimum embedding parameters that 
-represent all participants, activities, sensors and levels of smoothness.
-The sample mean for the minimum values of $E_{1}(m)$ from Figs \ref{fig:cao_ami}(A) 
-is $\overline{m}_0=6$ 
-and the sample mean for minimum values of AMIs from Figs~\ref{fig:cao_ami}(B)
-is $\overline{\tau}_0=8$.
-%Hence, we use the average minimum embedding parameters 
-%($\overline{m_0}=6$, $\overline{\tau_0}=8$) to compute 
-%Reconstructed State Spaces (RSSs), Recurrence Plots (RPs) and
-%Recurrence Quantification Analysis (RQA) metrics for human-humanoid activities.
+Minimum embedding parameters were initially computed using False Nearest Neighbour (FNN) 
+and Average Mutual Information (AMI).   We used FNN to calculate a value for $m$
+to unfold the attractors and AMI to calculate a value for $\tau$ and maxime the 
+information in the unfolded attractor.
+To illustrate this, figure \ref{fig:cao_ami} (A) show box plots for minimum embedding 
+values for sensors on the wrist of the human (HS01) and the robot (RS01).   
+As we had assumed, minimum embedding values for HS01 show greater variation than 
+those for RS01 (as indicated by differences in interquartile ranges).  
+Additionally, figure \ref{fig:cao_ami} (A) show a decrease in mean values (rhombus) in the box plots 
+as smoothness of the time-series increases (see sg0, sg1, sg2).  
+Figure \ref{fig:cao_ami} (B) show box plots for Average Mutual Information (AMI).  
+One can see that, in contrast to RS01, minimum values for HS01 tend to spread 
+as the smoothness of the time-series increases.
+The sample mean for minimum value of embedding parameter, 
+$m$ derived from FNN  (figures \ref{fig:cao_ami} A) is $\overline{m}_0=6$ 
+and $\tau$ from AMI (figures \ref{fig:cao_ami} B) is $\overline{\tau}_0=8$.
+
 %---------------------------------(FIGURE)-------------------------------------
@@
 \subsubsection*{Uniform Time-Delay Embedding}
-Uniform time-delay embedding were computed with the overall 
-embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$) and 
-the first three axis of the rotated data of the PCA are shown 
-for the reconstructed state spaces of horizontal arm movements
-(Figs~\ref{fig:rss_aHw10}) and vertical arm movements 
-(Figs~\ref{fig:rss_aVw10}).
+Using the overall embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$), 
+the first three axis of the rotated Principal Components Analysis (PCA) 
+are shown for the reconstructed state spaces of horizontal (figures~\ref{fig:rss_aHw10}) 
+and vertical (figures~\ref{fig:rss_aVw10}) arm movements. 
+While visual inspection of these figures 
+suggests differences in the trajectories of the reconstructed state spaces, 
+we require an objective quantification to determine the extent of the differences.  
+One approach could be use Euclidean distances from the origin for points in these 
+trajectories, but this proved inconclusive and was not able to capture 
+the dynamics of the trajectories. Consequently, we applied 
+Recurrence Plots and Recurrence Quantification Analysis.
+
 %%---------------------------------(FIGURE)-------------------------------------
@@
 
-One can observe by eye the differences in each of the
-trajectories of the reconstructed state spaces 
-(Figs~\ref{fig:rss_aHw10} and \ref{fig:rss_aVw10}), 
-however an objective quantification is required.
-That being said, we tried to quantify those differences 
-using euclidean distances between the origin 
-to each of the points in their trajectories, 
-which results did not capture the dynamics of 
-the trajectories specially 
-for those that looked very messy.
-Such ambiguities lead us to consider the use of 
-Recurrence Plots and Recurrence Quantification Analysis 
-to objectively quantify dynamics the capture activities 
-using time series.
-
 \subsection*{Recurrences Plots}
-Recurrence Plots (RP) were computed for horizontal arm movements (Fig~\ref{fig:rp_aH}) and 
-vertical arm movements (Fig~\ref{fig:rp_aV}) 
-using the average embedding parameters 
+Using the average embedding parameters 
 ($\overline{m}_0=6$, $\overline{\tau}_0=8$) 
 and an recurrence threshold of $\epsilon=1$.
+As our interest is for dynamical transitions, 
+there is little importance on the selection of $\epsilon$ which in this case
+is 1. 
+Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rp_aH}) and 
+vertical (figures~\ref{fig:rp_aV}) arm movements.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
-Four Recurrence Quantification Analysis (RQA) metrics (REC, DET, RATIO and ENTR) 
-were also computed with $\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
-
-\subsubsection*{REC values}
-Figs~\ref{fig:RQABP}(A) show box plots of REC values 
-that represent the \% of black dots in RPs.
-It can be noted that REC values are more spread 
-for HN and VN movements (higher interquartile range) than 
-for HF and VF movements (lower interquartile range) for HS01 sensor. 
-In contrast, REC values for RS01 sensor present little variation 
+Four Recurrence Quantification Analysis metrics (
+percentage recurrence, REC, representing the percentage of black dots in RP; 
+percentage determinism, DET, representing the predictability of the RP; 
+ratio of DET / REC, RATIO; Shannon entropy, ENTR) 
+were computed using the same parameters as for RP.
+%\subsubsection*{REC values}
+Figure~\ref{fig:RQABP}(A) presents box plots of REC values, 
+for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements 
+in Horizontal (HS) or Vertical (VS) than for Fast (i.e., 2 seconds per movement).  
+This suggests greater variation between participants for the Slow movements.  
+For RS01, there is little variation between Slow and Fast movement
 (interquartile range of 0.01). 
-Regarding the increase of smoothness for time series 
-(sg0, sg1 and sg2), REC values present little 
-variation as the smoothness is increasing for time series from HS01 
-(changes of mean values (rhombus)) whereas REC values are 
-more affected with the smoothness for data from RS01 
-(see the incremental changes of mean values (rhombus)).
-%See Figs~\ref{fig:rec_aH} and \ref{fig:rec_aV} in appendix \ref{appendix:e:ep} 
-%for more details about individual REC values for each participant.
-
-\subsubsection*{DET values}
-Figs \ref{fig:RQABP}(B) illustrate DET values 
-that represent the predictability and organisation of RPs.
-Generally, it can be noted little change of DET values 
-(interquartile range is around 0.1) 
-for type of movement, type of sensor 
-but the increase of DET values as the smoothness of 
-the signal increase 
-(see the incremental changes of mean values (rhombus)).
-However, the interquartile range for faster movements
-(HF and VF) with no smoothing (sg0) is lower than the other
-levels of smoothness (sg1 and sg2).
-%See Figs~\ref{fig:det_aH} and \ref{fig:det_aV} in appendix \ref{appendix:e:ep} 
-%for more details about individual DET values for each participant.
-
-\subsubsection*{RATIO values}
-Figs \ref{fig:RQABP}(C) illustrate RATIO values that represent the 
-dynamic transitions in RPs. 
-Generally, it can be seen that RATIO values for HS01 sensor vary less 
-for HN movements (interquartile range around 2)
-than HF movements (interquartile range around 5).
-For faster movements (HF and VF), it can be noted a decrease of 
-the mean values the smoothness of the time series 
-is increasing (rhombus). However, for normal movements (HN and VN), 
-the decrease of mean values is less evident than 
-faster movements (rhombus).
-%See Figs~\ref{fig:ratio_aH} and \ref{fig:ratio_aV} in appendix 
-%\ref{appendix:e:ep} 
-%for more details about individual RATIO values for each participant.
-
-\subsubsection*{ENTR values}
-Figs \ref{fig:RQABP}(D) show ENTR values that represent 
-Shannon entropy values in RPs. 
-Generally, 
-ENTR values for HS01 sensor show a higher variation  
-(interquartile range around 0.5)
-than ENTR values for RS01 sensor 
-(interquartile range 0.1).
-It can also be noted the increase of mean values 
-for ENTR values as the smoothness of time series 
-increase (rhombus) for type of movement, 
-type of sensor and level of smoothness.  
-%See Figs~\ref{fig:entr_aH} and \ref{fig:entr_aV} in appendix
-%\ref{appendix:e:ep}
-%for more details about individual ENTR values for each participant.
-
+In terms of smoothness, there seems little effect of HS01 but RS01 values 
+do show affects of smoothness (see the incremental changes of mean values (rhombus)).
+%\subsubsection*{DET values}
+Figure \ref{fig:RQABP}(B) presents DET values and shows 
+little difference for type of movement or performer.  
+However, DET values are affected by changes in smoothness of the signal, 
+particularly for Fast movement.
+%\subsubsection*{RATIO values}
+Figure \ref{fig:RQABP}(C) presents the ratio of DET / REC. 
+These values, for the human performer, vary less for HN than for HF movement.  
+Additionally, smoothness leads to a decrease in mean values for Fast movements.
+%\subsubsection*{ENTR values}
+Figure \ref{fig:RQABP}(D) shows ENTR values are higher for the human 
+performer than the robot, and vary with the smoothness of the time-series.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \subsection*{3D RQA ENTR}
-To show how that the selection of recurrence threshold affects RQA
-values, 
-we computed ENTR values of RQA metrics for different embedding parameters
+As ENTR appeared to be have higher sensitivity than the other measures, 
+we explored the impact of different embedding parameters 
 $ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
-$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by one,
-with the consideration of three recurrence thresholds $\epsilon=1, 2, 3$ 
-and three levels of smoothness (sg0, sg1, sg2). 
-For instance, 
-Figures \ref{fig:RQA-IND} show the increase of recurrence threshold is 
-associated to the increase of ENTR values in any of the levels 
-of smoothness (see values of the ENTR bar).
-Similarly, it can be noted that the increase of level of 
-smoothness (sg0, sg1 and sg2) is associated with the increase 
-of ENTR values of the 3D surface plots.
+$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ 
+incrementing by 1 each run, with recurrence thresholds $\epsilon=1, 2, 3$ 
+and levels of smoothness (sg0, sg1, sg2).  
+Figure \ref{fig:RQA-IND} shows that increasing recurrence 
+threshold leads to an increase in ENTR regardless of level of smoothness.  
+Similarly, increasing level of smoothness will also increase ENTR (figure \ref{fig:RQA-IND}).
+In terms of movement or performer, RQA ENTR decrease from Slow (HS, VS) to Fast (HF, VF) 
+for both human (HS01) and (RS01) (Figures~\ref{fig:3dRQAENTR_sensoractivities}).
+In terms of individual differences, for human participants, 
+figure \ref{fig:3dRQAENTR_participantsactivities} compares p01, p02 and p03 (for illustrative purposes) --
+both between each other and in comparison with the more consistent performance of the robot.  
+In terms of window length (w100 (2s), w250 (5s), w500(10s), w750 (15s)), 
+figure \ref{fig:3dRQAENTR_windowsactivities} shows that improvement in 
+capture with number of samples, although this has less of an effect on RQA ENTR.
+
 %%---------------------------------(FIGURE)-------------------------------------
@@
 %%---------------------------------(FIGURE)------------------------------------
-Figures~\ref{fig:3dRQAENTR_sensoractivities} show 3D surface plots of 
-ENTR values for different sensors and different activities.
-Figs~\ref{fig:3dRQAENTR_sensoractivities} show that RQA ENTR (lateral bars) 
-decrease from normal (HN, VN) to faster (HF, VF) velocities from both 
-human sensor (HS01) to robot sensor (RS01).
-Also, it can be noted that RQA ENTR values decrease from human sensor (HS01) 
-to robot sensor (RS01).
+
 %%---------------------------------(FIGURE)-------------------------------------
@@
 %%---------------------------------(FIGURE)-------------------------------------
-Figures~\ref{fig:3dRQAENTR_participantsactivities} show the similarity of 3D surface plots in 
-relationship with three participants (p01, p02 and p03)
-performing four activities (HN, HF, VN, and VF) and two sensors (HS01, RS01).
-Figures~\ref{fig:3dRQAENTR_participantsactivities}(A) 
-present subtle differences for each of the participants (see RQA ENTR bar). 
-Figures~\ref{fig:3dRQAENTR_participantsactivities}(B) show 
-less change of RQA ENTR values for each participant as the data is 
-from a sensor attached to the robot. 
-Also, from data of sensors attached to the robot, it can be noted 
-that changes are more notable for faster movements than normal movements. 
+
 %%---------------------------------(FIGURE)-------------------------------------
@@
 %%---------------------------------(FIGURE)-------------------------------------
-Figures \ref{fig:3dRQAENTR_windowsactivities} show the effect of the increase 
-of window length of the time series (w100 (2-sec), w250 (5-sec), w500 (10-sec) and w750 (15-sec))
-on the 3D surface plots. 
-It can be noted that the increase of number of samples improves the quality 
-of the surface plots where 100 samples poorly capture the dynamics of each activity.
-However as the window length increases from 250, 500 to 750 samples, 
-3D surface plots show a similar representation of RQA ENTR values 
-(being the surface plot with 750 samples the best test case).
+
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \section*{Discussion}
-Time series from different sources  
-(e.g., participants, movements, axis type) as well as
-different characteristics (e.g., sample rate, window length or levels of smoothness) 
-result in different embedding parameters and therefore different 
-values for recurrence plots and recurrence quantification analysis metrics.
-That said, while embedding parameters for individual
-time series were successfully computed, the quantification 
-of variability with regard to the shape of the trajectories in 
-reconstructed state spaces require more investigation
-as our initial, approach based on euclidean distances on the trajectories, 
-failed to quantify trajectories which were not well unfolded. 
-To then found out that Recurrence Plots along with 
-Recurrence Quantification Analysis (RQA) metrics (e.g., REC, DET, RATIO and ENTR) 
-and its variation of embedding parameters and recurrence thresholds 
-helped to quantify the variability of different sources of time series.
-However, further investigation is required to be done in order to 
-have better intuition and meaningful interpretation of nonlinear 
-analysis such as to find a right 
-balance among (i) the level of smoothness of the signal, 
+While there are many approaches to estimating embedding parameters for nonlinear analysis, 
+these can be influenced by the structure of the time-series data.   
+We show that, for RSS, RP and RQA, the estimation of embedding parameters can be performed 
+using a sample mean which, together with recurrence threshold, can be shown to be 
+influenced by activity, performer, window length and smoothness of time-series.  
+It is known that time-series from different sources and with different characteristics 
+require different embedding parameters, and this can produce different RSS, RPs and RQAs. 
+Although this work helps to understand the open problem of 
+finding right balance among (i) the level of smoothness of the signal, 
 (ii) the selection of recurrence thresholds and (iii) 
-the range of embedding parameters.
+the range of embedding parameters, 
+we have shown how RQA metrics can help to quantify movement variability.
 
 \section*{Conclusions}
-This work allows to conclude that the choice of 
-nonlinear analysis tool (e.g., RSSs, RPs, RQA metrics) will depend 
-on what one would like to quantify on the time-series 
-data (e.g., predicability, organisation, dynamics transitions, 
-or complexity and determinism).
-Then, time-series data characteristics 
-(e.g., window size length, level of smoothness) plays an important role
-as well on the results that nonlinear analysis tool can provide.
-Similarly, the results of the nonlinear analysis tools
-are associated with the structure of the time-series data 
-(e.g. frequency, amplitude), the position of the sensor 
-and activity performed by either a robot or human being
-(as degrees of freedom from the humanoid are far less than
-human movement). 
-That said, it has been shown that the use of different 
-characteristics of the time-series data 
-(e.g., sensor, activity, level of smoothness and participants)
-has help us to visualise and to quantify 
-with nonlinear tools the variation of 
-movements of, in this work, human-humanoid activities.
-However, some limitation of nonlinear tools are related 
-to the computation of different parameters 
-(e.g., recurrence thresholds, embedding parameters)
-that reflect the dynamics of individual characteristics 
-of activity type, window length and structure of the time series.
-Specifically, the example of DET values which appear 
-to be constants across sensors, activities and levels of smoothness, 
-whereas REC and RATIO, as function of REC, values 
-show variation for certain sensors and movements. 
+In this paper we show how the selection of nonlinear analysis tool  
+(i.e., RSS, RP, RQA metrics) depends on what question one wishes to 
+address with time-series data (e.g., predictability, organisation, dynamics, complexity).  
+Time-series data characteristics (e.g., window length, smoothness), 
+time-series structure (e.g., frequency, amplitude) and 
+data source (e.g., sensor placement, performance, movement) all influence 
+the results that nonlinear analysis methods can provide.
+That said, it has been shown that the use of different characteristics of the data and 
+their collection can help us visualise and quantify variability of movement using 
+methods of nonlinear analysis.   
+There remain limitations of nonlinear methods in 
+relation to the estimation of parameters (e.g., recurrence threshold, embedding parameters) 
+which reflect the dynamics of specific movement and performers, window length 
+and structure of the time-series. 
+We note that RQA DET seems to show low sensitivity 
+to these differences, whereas REC and RATIO (primarily as a result of REC) show 
+variation across performers and movements.   
+RQA ENTR, with different recurrence thresholds, 
+can quantify variation in the time-series data and offers the most appropriate means 
+for analysing variability in movement to allow us to analyse individual differences 
+between human performers.  
 To then found out that RQA ENTR values with different recurrence 
@@
 time-series data.
-Therefore, we can conclude that this work provides a good starting 
-point and reference to the use of Shannon Entropy to quantify 
-human-humanoid imitation activities 
-that can then lead to interesting results on the quantification 
-of movement variability of participants with different ages, 
-state of health and anthropomorphic features.
-
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
-\section*{Methods}
-\subsection*{State Space Reconstruction}
-The method of state space reconstruction \cite{packard1980} 
-\cite{takens1981} has been applied in many disciplines 
-\cite{aguirre2009, stergiou2011, frank2010, 
-sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
-based on uniform time-delay embedding methodology which is a simple 
-matrix implementation that can reconstruct an unknown $d-$dimensional 
-manifold $M$ from a scalar time series (e.g. one-dimensional 
-time series in $\mathbb{R}$). A manifold, in this context, is a 
-multidimensional curved surface within a space (e.g. a saddle) 
-\cite{guastello-gregson2011}.
-
-The use of a scalar time series is the main advantage of the method of state 
-space reconstruction which in essence preserve dynamic invariants such as 
-correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
-entropy and detrended fluctuation analysis \cite{bradley2015, 
-Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
-However, selecting appropriate embedding parameters which are sued to apply the 
-state space reconstruction is still an open challenge for which we present 
-introductions for the methodologies to compute such embedding parameters 
-With that in mind, in the following subsections, we describe in more detail 
-the state space reconstruction theorem (RSSs), uniform time-delay embedding 
-theorem (UTDE), false nearest neighbours (FNN) and average mutual 
-information (AMI). 
-%and other methodologies for state space reconstruction.
-
-\subsection*{State Space Reconstruction Theorem}
-Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
-uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:ssr}
-  s(t)=f^t [s(0)],
-\end{equation}
-%%********************************[EQUATION]************************************
-where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
-$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
-unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
-function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
-iteration of $f$ that corresponds to an initial position 
-$s(0) \in M $ \cite{takens1981}.
-%$f$ is a smooth dynamical system, where smooth means a 
-%continuously differentiable system
-%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
-Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
-can be obtained with
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:measurement}
-  x(t)=h[s(t)],
-\end{equation}
-%%********************************[EQUATION]************************************
-where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
-Reconstructed state space can then be described as an $n-$dimensional state 
-space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
-$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
-is the uniform time-delay embedding with a dimension embedding $m$
-and delay embedding $\tau$ and
-$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
-of dimensionality (e.g. Principal Component Analysis, 
-Singular Value Decomposition, etc) being $n \leq m$.
-With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
-defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
-$\boldsymbol{X}(t) = \Phi(s(t))$,
-where $\Phi$ is a diffeomorphic map \cite{takens1981}
-whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
-dimension of $M$ \cite{garland2016}.
-% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
-% is called diffeomorphic if it is one-to-one correspondence and its inverse
-% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
-Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
-space then a composition of functions represented with $F^t$ is induced on the
-reconstructed state space determined:
-%by true smooth dynamical system $f^t$ and $\Phi$:
-%********************************[EQUATION]************************************
-\begin{equation}\label{eq:st}
-  \boldsymbol{X}(t)=F^t [\boldsymbol{X}(0)] = \Phi \circ f^t \circ \Phi ^{-1}[\boldsymbol{X}(0)].
-\end{equation}
-%%********************************[EQUATION]************************************
-With this in mind, an embedding is defined as "a smooth one-to-one 
-coordinate transformation with a smooth inverse" and the total reconstruction 
-map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
-Fig~\ref{fig:ssr} illustrates the state space reconstruction.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
-    \caption{
-	{\bf State space reconstruction methodology.}
-	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
-	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
-    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
-    	The time-delay embedding represented as the $\Phi$, maps the original
-    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
-	time-delay embedding matrix $\boldsymbol{X}(t)$.
-    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
-	a new state $y(t)$ of dimensions $n < m$.
-    	(A) $M-$dimensional manifold representing the state space 
-	(e.g. Lorenz system);
-    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
-   	(C) $m-$dimensional reconstructed state space with 
-    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
-    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
-	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
-	where $\Phi$ is the delay reconstruction map and 
-	$\Psi$ is the coordinate transformation map.
-	This figure is adapted from 
-   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-    }
-    \label{fig:ssr}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-\subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
-Frank et al. and Sama et al. refer to the state space reconstruction
-as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
-However, we consider the term "uniform time-delay embedding"
-as more descriptive and appropriate terminology for our work.
-Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed 
-copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
-the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
-samples of $\{ \boldsymbol{x}_n \}$.
-$\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
-The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
-and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
-where $i$ goes from $0,1, \dots, (m-1)$.
-Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
-information of unobserved state variables and encapsulates the information of
-the delayed copies of the available time series in the uniform time-delay 
-embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
-$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:tde}
-\boldsymbol{X}^{m}_{\tau}  =
-\begin{pmatrix}
-\boldsymbol{ \tilde{x} }_n \\
-\boldsymbol{ \tilde{x} }_{n-\tau} \\
-\boldsymbol{ \tilde{x} }_{n-2\tau} \\
-\vdots \\
-\boldsymbol{ \tilde{x} }_{n- (m-1) \tau} \\
-\end{pmatrix}^\intercal, 
-\end{equation}
-%%********************************[EQUATION]************************************
-where $m$ is the embedding dimension, $\tau$ is the embedding delay and
-$ ^\intercal$ denotes the transpose.
-$m$ and $\tau$ are known as embedding parameters.
-The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
-$N-(m-1)\tau$ rows and $m$ columns and 
-$N-(m-1)\tau$ defines the length of each delayed copy 
-of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
-%For further details and explicit examples of uniform time-delay 
-%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
-
-\subsection*{Estimation of Embedding Parameters}
-The estimation of the embedding parameters ($m$ and $\tau$) 
-is a fundamental step for the state space reconstruction with the use
-of uniform time-delay embedding method.
-With this in mind, we review two of the most common algorithms,
-which will be used in our work, to compute the embedding
-parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%*******************************************************************************
-\subsubsection*{False Nearest Neighbours}
-To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
-used the method of false neighbours which can be understood as follows:
-on the one hand, when the embedding dimension is too small to unfold the 
-attractor "not all points that lie close each other will be neighbours and 
-some points appear as neighbours because of the attractor has been projected 
-down into an smaller space", on the other hand, when increasing the embedding 
-dimension "points that are near to each other in the sufficient
-embedding dimension should remain close as the dimension increase from $m$ 
-to $m+1$ \cite{krakovska2015}".
-From a mathematical point of view, the state space reconstruction theorem is 
-done when the attractor is unfolded with either the minimum embedding 
-dimension, $m_0$, or any other embedding dimension value where 
-$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
-leads to excessive computations \cite{bradley2015}. With this in mind, 
-Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
-where only the time-series and one delay embedding value are necessary 
-to select the minimum embedding dimension. Cao's algorithm is based 
-on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:e}
-  \begin{aligned}
-	E(m) &= \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau} a(i,m) \\
-	 &=
-       \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
-       \frac{ || \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) || }
-            { || \boldsymbol{X}_i(m) - \boldsymbol{X}_{n(i,m)}(m) ||  }
-  \end{aligned}
-\end{equation}
-%%********************************[EQUATION]************************************
-where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
-embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
-From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
-and $\tau$ for which $E_1(m)$ is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:e1}
-E_1(m) = \frac{ E(m+1) } { E(m)}.
-\end{equation}
-%%********************************[EQUATION]************************************
-$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
-in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
-As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
-than some $m_0$, if the time series comes from a multidimensional state space
-then $m_0 + 1$ is the minimum dimension".
-Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
-stochastic signals. $E_2(m)$ is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:e2}
-E_2(m) = \frac{ E^* (m+1) } { E^*(m)},
-\end{equation}
-%%********************************[EQUATION]************************************
-where
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:ee}
-E^*(m) = \frac{1}{N-m\tau} \sum_{i=1}^{N-m\tau}
-|  \boldsymbol{X}_i(m+1) - \boldsymbol{X}_{n(i,m)}(m+1) |.
-\end{equation}
-%%********************************[EQUATION]************************************
-For instance, when the signal comes from random noise (values that are 
-independent from each other), all $E_2(m)$ values are approximately equal 
-to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
-not constant for all $m$ (e.g. $E_2(m) \neq 1$).
-
-As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
-series: the solution for the $x$ variable of the Lorenz system 
-(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
-and a variance of one (Fig~\ref{fig:e1e2}F).
-We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
-The $E_1(m)$ values for the chaotic time series appear to be constant
-after the dimension is equal to six.
-The determination of six is given that any value of $m$ can be used as they 
-are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
-$E_2(m)$ values, for chaotic time series, are different to one 
-(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
-chaotic time series the minimum embedding dimension the time series 
-comes from a deterministic signal. With regard to the noise time series,  
-$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
-$E_1(m)$ values then indicate the minimum embedding dimension of the 
-noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
-concluded that noise time series is a stochastic signal.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
-    \caption{
-	{\bf Minimum dimension embedding values with Cao's method.} 
-	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
-	with variations of $\tau$ values from one to twenty
-	for (E) chaotic and (F) random time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:e1e2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
-computed but also a variation of $\tau$ from 1 to 20 is presented. 
-The purpose of such variation for $\tau$ is to show its independence with
-regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-(Fig~\ref{fig:e1e2}A,B,C, and D). 
-However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
-a new threshold where $m$ values appear to be constant in $E_1 (m)$.
-In the case of the given examples and reported results, we defined such 
-threshold as 0.05. Further investigation is required for the selection of the 
-threshold in the $E_1(m)$, as the selection of the threshold in this work is 
-base on no particular method but visual inspection.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%*******************************************************************************
-\subsubsection*{Average Mutual Information}
-When selecting the delay dimension parameter, $\tau$, one can consider the 
-following two cases:
-(i) when $\tau$ is too small, the elements of time-delay embedding will be along
-the bisectrix of the phase space and the reconstruction is generally not 
-satisfactory, 
-(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
-time-delay embedding will become spread and uncorrelated which makes 
-recovering the underlying attractor more difficult if not 
-impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
-With regard to the algorithms to compute $\tau$, 
-Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
-function in which the first zero crossing is considered as the minimum delay 
-embedding parameter. However, the autocorrelation function is a linear 
-statistic over which the Average Mutual Information (AMI) algorithm is 
-preferred because the AMI takes into account the nonlinear dynamical 
-correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
-algorithm is described below to estimate the minimum delay embedding 
-parameter, \texorpdfstring{$\tau_0$}{T}.
-
-To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
-and then a probability distribution of data is computed \cite{kantz2003}.
-AMI is therefore denoted by $I(\tau)$ which is the average mutual 
-information between the original time series, $x(n)$, and the delayed 
-time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
-AMI is defined by
-%%********************************[EQUATION]************************************
-\begin{equation}\label{eq:ami}
-I(\tau) = \sum_{i,j}^N p_{ij} log_2 \frac{ p_{ij} }{ p_i p_j }.
-\end{equation}
-%%********************************[EQUATION]************************************
-Probabilities are defined as follows:
-$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
-the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
-the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
-that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
-in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
-For small $\tau$, AMI will be large and it will then decrease more or 
-less rapidly. As $\tau$ increase and goes to a large limit, 
-$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
-factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
-to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
-where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
-where the redundancy is the least" \cite{kantz2003}.
-
-For example, we compute the AMI for two time series:
-A) the $x$ solution of the Lorenz system, and
-B) a noise time series using a normal distribution with mean zero and 
-standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
-concluded that the amount of knowledge for any noise time series is zero 
-for which the first minimum embedding parameter is $\tau_0=1$. 
-On the contrary, the first minimum of the AMI for the chaotic time series 
-is $\tau_0=17$ which is the value that maximize the independence 
-between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
-\cite{bradley2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
-    \caption{
-	{\bf Minimum delay embedding values with AMI's method.} 
-    	(A, B) AMI values where its first minimum value in the curve
-	is the minimum time delay embedding ($\tau_0$), 
-	for (C) a chaotic and (D) noise time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:amis}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-Similarly as Cao's algorithm negatives, AMI's algorithm is not an
-exception for negatives, which are worthwhile to mention for further 
-investigations.
-For instance, (i) is not clear why the choose of the first minimum of the AMI 
-is the minimum delay embedding parameter \cite{kantz2003} and 
-(ii) the probability distribution of the AMI function is computed
-with the use of histograms which depends on a heuristic choice of number of bins
-for which AMI depends on partitioning \cite{garcia2005e71}.
-
-%*******************************************************************************
-\section*{Recurrence Quantification}\label{sec:recurrence-quantification}
-\subsection*{Recurrence Plots}
-Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
-in conservative systems, however such discovery was not put into practice 
-until the development of faster computers \cite{marwan2007},
-for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
-where recurrences in the dynamics of a system can be visualised using 
-Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
-propose a tool, called Recurrence Plot (RP), that provides insights into 
-high-dimensional dynamical systems where trajectories are very difficult to 
-visualise. Therefore, "RP helps us to investigate the 
-$m-$dimensional phase space trajectories through a two-dimensional 
-representation of its recurrences" \cite{marwan2015}.
-Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
-to the methodologies of the state space reconstruction and other dynamic 
-invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
-the recurrences of the trajectories in the phase space can provide 
-important clues to characterise natural process that present, for
-instance, periodicities (as Milankovitch cycles) or irregular cycles 
-(as El Ni\~no Southern Oscillation). 
-Such recurrences can not only be presented visually using Recurrence Plots (RP) 
-but also be quantified with Recurrence Quantification metrics, which leads 
-to applications of these in various areas such as economy, physiology, 
-neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
-
-For the creation of a recurrence plot based on time series 
-$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
-reconstruction with uniform time-delay embedding 
-$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
-\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
-where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
-and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
-The recurrence plot is therefore a two-dimensional $N \times N$ square 
-matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
-whenever $X(i)$ is sufficiently close to $X(j)$: 
-%%********************************[EQUATION]************************************
-\begin{equation}
-\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
-\end{equation}
-%%********************************[EQUATION]************************************
-where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
-$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
-function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
-RP is also characterised with a line of identity (LOI) which is a  
-diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
-%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
-    \caption{
-	{\bf Recurrence Plots.} 
-	(A) State space of the Lorenz system with controlling parameters 
-	($\rho=28, \sigma=10, \beta=8/3$). 
-	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
-	(black circle) of a given point at $i$ is a recurrent point and is 
-	represented as a black dot in the recurrence plot at location 
-	$(i, j)$ or white otherwise.
-	(B) Recurrence plot using the 
-	three components of the Lorenz system and the RP with no embeddings 
-	and threshold $\epsilon=5$.
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:mrp}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
-\subsection*{Structures of Recurrence Plots}
-Pattern formations in the RPs can be designated either 
-as topology for large-scale patterns or texture for small-scale patterns.
-In the case of topology, the following pattern formations are commonly presented:
-(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-uniformly distributed noise (Fig~\ref{fig:rp2}A), 
-(ii) periodic and quasi-periodic systems where diagonal lines and 
-checkerboard structures represent oscillating systems, e.g., sinusoidal 
-signals (Fig~\ref{fig:rp2}B), 
-(iii) drift where paling or darkening recurrence points away from 
-the LOI is caused by drifting systems, 
-e.g., logistic map (Fig~\ref{fig:rp2}C), and
-(iv) disrupted where recurrence points are presented white areas or 
-bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
-Texture patterns in RPs can be categorised as:
-(i) single or isolated recurrence points that represent rare occurring states, 
-and do not persist for any time or fluctuate heavily,
-(ii) dots forming diagonal lines where the length of the small-scale parallel 
-lines in the diagonal are related to the ratio of determinism or predictability 
-in the dynamics of the system, and
-(iii) dots forming vertical and horizontal lines where the length of the 
-lines represent a time length where a state does not change or change very 
-slowly and these patterns formation represent discontinuities in the signal, 
-and (iv) dots clustering to inscribe rectangular regions which are related 
-to laminar states or singularities \cite{marwan2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
-    \caption{
-	{\bf Patterns in Recurrence Plots.} 
-	Time-series with its respective recurrence plot for:
-	(A) uniformly distributed noise,
-	(B) super-positionet harmonic oscillation 
-	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-	with a linearly increase term ($0.01*i$), and
-	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:rp2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
-Although, each of the previous pattern descriptions of the structures in the 
-RP offer an idea of the characteristics of dynamical systems, 
-these might be misinterpreted and conclusions might tend to be subjective 
-as these require the interpretation of a particular researcher(s).
-Because of that, recurrence quantification analyis (RQA) offer objective 
-methodologies to quantify such visual characteristics of previous 
-recurrent pattern structures in the RP \cite{zbilut1992}.
-
-%*******************************************************************************
-\subsection*{Recurrence Quantifications Analysis (RQA)}
-Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
-the density of recurrence points in RPs, then histograms of lengths for 
-diagonal lines in RPs were studied by \cite{trulla1996} which were the 
-introduction to the term recurrence quantification analysis (RQA) 
-\cite{marwan2008}. RQA has been applied in many fields such as life science, 
-engineering, physics, and others \cite{marwan2008}. Particularly in human 
-movement to investigate noise and complexity of postural control 
-\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
-coordination \cite{duran2017}. The success of RQA is not only due to its 
-simple algorithmic implementation but also to its capacity to detect tiny 
-modulations in frequency or phase which are not detectable using standard 
-methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
-RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
-dimension which is verified experimentally by \cite{iwanski1998}.
-RQA metrics comprehend percentage of recurrence, percentage of determinism, 
-ratio, Shannon entropy of the frequency distributions of the line lengths,
-maximal line length and divergence, trend and laminariy 
-\cite{marwan2007, marwan2015}. For this work, we considered only four 
-RQA metrics, due to its consistency with our preliminary experiments, 
-which are described below. Such metrics are computed the nonlinearTseries 
-R package \cite{nonlinearTseries2016}.
-
-\subsubsection*{REC values}
-The percentage of recurrence (REC) is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}
-REC(\epsilon,N) = 
-		\frac{1}{N^2 - N} \sum^{N}_{i \neq j = 1} 
-		\mathbf{R}^{m}_{i,j}(\epsilon),
-\end{equation}
-%%********************************[EQUATION]************************************
-which enumerates the black dots in the RP excluding the line of identity.
-REC is a measure of the relative density of recurrence points in the sparse 
-matrix \cite{marwan2015}.
-%REC is computed as follow with the nonlinearTseries package 
-% \cite{nonlinearTseries} 
-%  hist = getHistograms(neighs, ntakens, lmin, vmin)
-%  # calculate the number of recurrence points from the recurrence rate. The
-%  # recurrence rate counts the number of points at every distance in a concrete
-%  # side of the main diagonal.
-%  # Thus, sum all points for all distances, multiply by 2 (count both sides) and
-%  # add the maindiagonal
-%  numberRecurrencePoints = sum(hist$recurrenceHist) + ntakens
-%  # calculate the recurrence rate dividing the number of recurrent points at a
-%  # given distance by all points that could be at that distance
-%  recurrence_rate_vector = 
-%  hist$recurrenceHist[1:(ntakens - 1)] / ((ntakens - 1):1)
-%  # percentage of recurrent points
-%  REC = (numberRecurrencePoints) / ntakens ^ 2
-
-\subsubsection*{DET values} 
-The percent determinism (DET) is defined as the fraction of recurrence points
-that form diagonal lines and it is determined by
-%%********************************[EQUATION]************************************
-\begin{equation}
-DET=
-	\frac{\sum^{N}_{l=d_{min}} l H_D{l} }
-	     {\sum^{N}_{i,j=1} \mathbf{R}_{i,j}(\epsilon) },
-\end{equation}
-%%********************************[EQUATION]************************************
-where 
-%%********************************[EQUATION]************************************
-\begin{equation}
-H_D(l) = 
-	\sum^{N}_{i,j=1} 
-	(1- \mathbf{R}_{i-1,j-1}(\epsilon) ) 
-	(1- \mathbf{R}_{i+l,j+l}(\epsilon) ) 
-	\prod^{l-1}_{k=0}  \mathbf{R}_{i+k,j+k}(\epsilon)
-\end{equation}
-%%********************************[EQUATION]************************************
-is the histogram of the lengths of the diagonal structures in the RP.
-DET can be interpreted as the predictability of the system for periodic signals 
-which, in essence, have longer diagonal lines than the short diagonals lines
-for chaotic signals or absent diagonal lines for stochastic signals 
-\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
-the organisation of points in RPs  \cite{iwanski1998}. 
-%percent determinism (DET) is computed as follow with the nonlinearTseries 
-%package \cite{nonlinearTseries}  
-% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%                                       lmin = 2, lDiagonalHistogram,
-%                                       recurrence_rate_vector, maxDistanceMD) {
-%  #begin parameter computations
-%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
-%  DET = num / numberRecurrencePoints
-
-\subsubsection*{RATIO values}
-RATIO is defined as the ratio between DET and REC and it is calculated from 
-the frequency distributions of the lengths of the diagonal lines.
-RATIO is useful to discover dynamic transitions \cite{marwan2015}.
-%  diagP = calculateDiagonalParameters(
-%    ntakens, numberRecurrencePoints, lmin, hist$diagonalHist,
-%    recurrence_rate_vector, maxDistanceMD
-%  )
-% calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%                                       lmin = 2, lDiagonalHistogram,
-%                                       recurrence_rate_vector, maxDistanceMD) {
-%  #begin parameter computations
-%  num = sum((lmin:ntakens) * lDiagonalHistogram[lmin:ntakens])
-%  DET = num / numberRecurrencePoints
-% 
-%
-%    RATIO = diagP$DET / REC
-%
-
-\subsubsection*{ENT values}
-ENT is the Shannon entropy of the frequency distribution of the diagonal 
-line lengths and it is defined as
-%%********************************[EQUATION]************************************
-\begin{equation}
-ENT= - \sum^{N}_{l=d_{min}} p(l) ln p(l) \quad with 
-	\quad p(l)=\frac{ H_D(l) }{ \sum^{N}_{ l=d_{min} } H_D(l) }.
-\end{equation}
-%%********************************[EQUATION]************************************
-ENT reflects the complexity of the deterministic structure in the system.
-For instance, for uncorrelated noise or oscillations, 
-the value of ENT is rather small and indicates low complexity of the system,
-therefore "the higher the ENT is the more complex the dynamics are" 
-\cite{marwan2007, marwan2015}.
-%#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
-%
-%calculateDiagonalParameters = function(ntakens, numberRecurrencePoints,
-%                                       lmin = 2, lDiagonalHistogram,
-%                                       recurrence_rate_vector, maxDistanceMD) {
-%
-%  pl = lDiagonalHistogram / sum(lDiagonalHistogram)
-%  diff_0 = which(pl > 0)
-%  ENTR = -sum(pl[diff_0] * log(pl[diff_0]))
-
-\subsection*{Sensitivity and robustness of RPs and RQA.}
-RP and RQA are a very young field in nonlinear dynamics and many questions 
-are still open, for instance, different parameters for window length size 
-of the time series, embedding parameters or recurrence threshold can 
-generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
-
-The selection of recurrence threshold, $\epsilon$, can depend on the system 
-that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
-require to be very small, for trajectory reconstruction $\epsilon$ requires 
-to have a large thresholds or when studying dynamical transition 
-there is little importance about the selection of the threshold 
-\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
-the recurrence threshold  should be five times larger 
-than the standard deviation of the observational noise
-or the use of diagonal structures within the RP is suggested in order
-to find the optimal recurrence threshold for (quasi-)periodic process 
-\cite{marwan2011}.
-Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
-of choosing the right embedding parameters to perform RQA for which 
-many experiments have to be performed using different parameters in order 
-to have a better intuition of the nature of the time series and how 
-this is represented by using RQA.
-
-With that in mind, this work explores the sensitivity and robustness 
-of the window size of time series, embedding parameters for RSS with UTDE 
-and recurrence threshold for RP and RQA in order to gain a better insight 
-into the underlying time series collected from inertial sensors in the 
-context of human-humanoid imitation activities.
-%Such changes of recurrence threshold values 
-%can modify the patterns in RPs and therefore the values of RQA metrics.
-%We therefore computed 3D surfaces to explore the sensibility and robustness of 
-%embedding parameters and recurrence threshold in RQA  metrics. Following the 
-%same methodology of computing 3D surfaces, we also considered variation of 
-%window length size to present RQA metrics dependencies with embedding 
-%parameters, recurrence thresholds and window length size.
-
-%Choosing an appropriate recurrence threshold is crucial so as to get 
-%meaningful representations in RP and RQA, however,
-
-%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
-%metrics for RQA are independent of embedding dimension parameters, 
-%however, that is not the case when using 
-%different recurrence thresholds.
-
-%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
-%determining embedding parameters, their method consisted on creating 3D 
-%surfaces with RQA metrics with an increase of embedding parameters 
-%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
-%fluctuations and gradual changes in the 3D surfaces that provide information 
-%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
-%created 3D surfaces for visual selection of not only embedding parameters 
-%but also recurrence thresholds. Following same methodologies, we explored 
-%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
-%using 3D surfaces by an unitary increase of the pair embedding 
-%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
-%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
-%(Fig.~\ref{fig:topo_rqas}).
-
-%*******************************************************************************
-%*******************************************************************************
-%*******************************************************************************
-\section*{Experiment} \label{sec:experiment}
-We conducted an experiment in the context of human-humanoid imitation (HHI) 
-activities where participants were asked to imitate simple horizontal and 
-vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
-Such simple movements were repeated ten times for the participant 
-who copied NAO's arm movements in a face-to-face imitation activity.
-Also, wearable inertial measurement unit (IMU) sensors were attached 
-to the right hand of the participant and to the left hand of the robot 
-(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
-IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
-accelerometer, gyroscope and magnetometer sensors and quaternions 
-\cite{Comotti2014}.
-
-\subsection*{Ethics}
-The experiments of this work were conducted in November 2016 and participants
-confirmed reading and understanding the participant information sheet of the 
-experiments and were able to withdraw from the experiment at any time without giving
-any reason. The design of the experiments adhered to the University of X
-regulations, data were anonymised and videos were stored only on a 
-personal computer in accordance with the Data Protection Act 1998. 
-%Refer to Appendix D for further information about the ethics, 
-%online participation information sheets and experiment check list.
-
-%A further description of the NeMEMSi IMU sensors 
-%is given in \nameref{S1_AppendixB}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-  \centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
-    \caption{
-	{\bf Human-humanoid imitation activities.} 
-    		Face-to-face human-humanoid imitation (HHI) activities for 
-		(A) HHI of horizontal arm movement, 
-		(B) Humanoid horizontal arm movement,
-		(C) HHI of vertical arm movement, and 
-		(D) Humanoid vertical arm movement.
-        }
-    \label{fig:hri}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-\subsection*{Participants}
-Twenty-three participants,
-from now on defined as $pN$ where $N$ is the number of participant, were 
-invited to do the experiment. However, data for three participants were 
-not used because the instructions for $p01$, who was the only left-handed,
-were mistakenly given in a way that movements were performed different
-from what had been planned, and for participants $p13$ and $p16$ 
-data were corrupted because bluetooth communications problems with the 
-sensors. With that in mind, data for twenty participants were analysed in 
-this work.
-
-Of the twenty participants, all of them are right-handed healthy participants 
-of whom four are females and sixteen are males with a mean and standard 
-deviation (SD) age of mean=19.8 (SD=1.39). All participants provided 
-informed consent forms prior to participation in the experiment.
-
-\subsection*{Human-humanoid imitation activities}
-For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
-two of which were attached to the right hand of the participant and the 
-other two to the left hand of the humanoid robot.
-Then, each participant was asked to imitate repetitions of simple horizontal
-and vertical arm movements performed by the humanoid robot in the following 
-conditions:
-(i) ten repetitions of horizontal arm movement at normal (HN) 
-	and faster (HF) speed (Figure~\ref{fig:hri} A), and
-(ii) ten repetitions of vertical arm movement at normal (VN) 
-	and faster (VF) speed (Figure~\ref{fig:hri} C).
-The normal and faster speed of arm movements is defined by the duration 
-in number of samples of one repetition of NAO's arm movements. 
-We select NAO's arm movements duration to distinguish between normal and 
-faster arm movements as NAO's movements have less variation between 
-repetition to repetition. The duration for one repetition of the horizontal 
-arm movement at normal speed, HN, is about 5 seconds considering 
-that each repetition last around 250 samples.
-For horizontal arm movement at faster speed, HF, each repetition were performed 
-in around 2 seconds which correspond to 90 samples of data.
-The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
-which is around 300 samples of data. For vertical arm movement at 
-faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
-to 120 samples of data.
-To visualise the distinction between normal and faster speed for horizontal 
-and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
-for axes Z and Y of the gyroscope sensors with four window lengths: 
-2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
-and 15-sec (750-samples).
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht] 
-\centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
-    \caption{
-	{\bf Time series duration of horizontal and vertical arm movements.} 
-    		Time series of smoothed data from gyroscope sensor 
-		for different speed arm movements performed by NAO: 
-		(A) Horizontal Normal arm movement (HN), 
-		(B) Horizontal Faster arm movement (HF),
-		(C) Vertical Normal arm movement (VN) and 
-		(D) Vertical Faster arm movement (VF).
-		Additionally, 
-		(A) shows window sizes for 2-seconds (100 samples), 
-		5-seconds (250 samples), 10-seconds (500 samples) 
-		and 15-seconds (750 samples)
-		which are also presented in (B), (C) and (D).
-		Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-	\label{fig:sts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-\subsection*{Data from Inertial Measurement Units} 
-	\label{sec:experiment:subsec:imu}
-
-To give insight to the research questions, 
-we considered various conditions of time series collected for this work 
-(see Experiment section for more details) which are described as follows
-\begin{itemize}
-\item Three levels of smoothness for the normalised data 
-	(sg0zmuv, sg1zmuv and sg2zmuv), computed from two different filter 
-	lengths (29 and 159) with the same polynomial degree 
-	of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal},
-\item four velocities arm movement activities: 
-	horizontal normal (HN), horizontal faster (HF), 
-	vertical normal (VN), and vertical faster (VF), and
-\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
-	10-sec (500 samples) and 15-sec (750 samples).
-\end{itemize}
-%\subsection*{Time series}
-After the data collection, raw time series were windowed, normalised and 
-smoothed. Then, due to space limitations and to have simple visualisation, 
-we only present 10-sec (500 samples) window length time series for 
-three participants (p01, p01 and p03) performing horizontal 
-arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}). 
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
-    	\caption{ 
-	{\bf Time series for horizontal and vertical arm movements.}
-		(A/D) raw-normalised (sg0zmuv), 
-		(B/E) normalised-smoothed 1 (sg1zmuv) and
-		(C/F) normalised-smoothed 2 (sg2zmuv).
-		Time series are only for three participants 
-		($p01$, $p02$, and $p03$) 
-		for horizontal and vertical arm movements in normal 
-		and faster velocity (HN, HF, VN, VF) 
-		with the normalised GyroZ or GyroY axis 
-		and with one sensor attached to the participant (HS01) 
-		and other sensor attached to the robot (RS01).	
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:ts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-
-\subsubsection*{Raw data}
-Considering the work of \cite{shoaib2016} which provided evidence of an 
-improvement in recognition activities when combining data from 
-accelerometer and gyroscope. We focus our analysis from data of the 
-accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
-leave the data of the magnetometer and quaternions for further investigation 
-because of their possible variations with regard to magnetic disturbances.
-
-Data from the accelerometer is defined by triaxial time series 
-$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
-(Eq.~\ref{eq:A}), and the same for data from the gyroscope 
-which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
-representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:G}).
-Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
-its respective axes subscripts $x,y,z$, where $n$ is the sample index 
-and $N$ is the same maximum length of all axes for the time series.
-Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
-%%---------------------------------(EQUATION)----------------------------------
-\begin{equation}\label{eq:A}
-\boldsymbol{A} =
-\begin{pmatrix}
-  A_x(n) \\
-  A_y(n) \\
-  A_z(n)
-\end{pmatrix}
-=
-\begin{pmatrix}
- a_x(1),a_x(2),\dots,a_x(N) \\
- a_y(1),a_y(2),\dots,a_y(N) \\
- a_z(1),a_z(2),\dots,a_z(N) 
-\end{pmatrix},
-\end{equation}
-%%---------------------------------(EQUATION)-----------------------------------
-and 
-%%---------------------------------(EQUATION)-----------------------------------
-\begin{equation}\label{eq:G}
-\boldsymbol{G} =
-\begin{pmatrix}
- G_x(n) \\
- G_y(n) \\
- G_z(n)
-\end{pmatrix}
-=
-\begin{pmatrix}
- g_x(1),g_x(2),\dots,g_x(N) \\
- g_y(1),g_y(2),\dots,g_y(N) \\
- g_z(1),g_z(2),\dots,g_z(N) 
-\end{pmatrix}.
-\end{equation}
-%%---------------------------------(EQUATION)-----------------------------------
-
-\subsubsection*{Postprocessing data}
-After the collection of raw data from four NeMEMsi sensors,
-time synchronisation alignment and interpolation were performed 
-in order to create time series with the same length and synchronised time.
-We refer the reader to \cite{Comotti2014} for further
-details about the time synchronisation process.
-
-\subsubsection*{Data normalization}
-Data is normalised to have zero mean and unit variance 
-using sample mean and sample standard deviation.
-The sample mean and sample standard deviation using $x(n)$ is given by
-%%---------------------------------(EQUATION)----------------------------------
-\begin{equation}\label{eq:ms}
-\mu_{x(n)}= \frac{1}{N} ( \sum_{i=1}^N x(i) ), \quad 
-	\sigma_{x(n)} =  \sqrt{ \frac{  \sum_{1=1}^N ( x(i) - \mu_{x(n)} )^2 }{ N-1 }  },      
-\end{equation}
-%%---------------------------------(EQUATION)----------------------------------
-and the normalised data, $\hat{x}(n)$, is computed as follows
-%%---------------------------------(EQUATION)----------------------------------
-\begin{equation}\label{eq:normalization}
-\hat{x} (n) = \frac{   x(n) -  \mu_{x(n)}  }{   \sigma_{x(n)} }.   
-\end{equation}
-%%---------------------------------(EQUATION)----------------------------------
-
-\subsubsection*{Smoothing data}
-Commonly, a low-pass filter is the method either to capture the low 
-frequencies that represent \%99 of the human body energy or to get 
-the gravitational and body motion components of 
-accelerations \cite{anguita2013}. However, for this work the elimination of 
-certain range of frequencies is not the main focus but the conservation of 
-the structure in the time series in terms of the width and heights where, 
-for instance, Savitzky-Golay filter can help to accomplish 
-such task \cite{press1992}. Savitzky-Golay filter is based on the 
-principle of moving window averaging which preserves the area under 
-the curve (the zeroth moment) and its mean position in time 
-(the first moment) but the line width (the second moment) is violated 
-and that results, for example, in the case of spectrometric data
-where a narrow spectral line is presented with reduced height and width. 
-With that in mind, the aim of Savitzky-Golay filtering is to find filter 
-coefficients $c_n$ that preserve higher momentums which are based on local 
-least-square polynomial approximations \cite{savitzkygolay1964, 
-press1992, schafer2011}.
-Therefore, Savitzky-Golay coefficients are therefore computed using 
-an R function \texttt{sgolay(p,n,m)} where \texttt{p} is the filter order, 
-\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
-$m$-th derivative of the filter coefficients  \cite{Rsignal}. 
-Smoothed signal is represented with a tilde over the original 
-signal: $\tilde{x}(n)$.
-
-\subsubsection*{Window size data}
-With regard to the window size, \cite{shoaib2016} investigated its effects 
-using seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
-and combination of inertial sensors (accelerometer, gyroscope and linear 
-acceleration sensor) to improve the activity recognition performance for 
-repetitive activities (walking, jogging and biking) and less repetitive 
-activities (smoking, eating, giving a talk or drinking a coffee).
-With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
-increase of window length improve the recognition of complex activities 
-because these requires a large window length to learn the repetitive 
-motion patterns. Particularly, one of the recommendations is to use large 
-window size to recognise less repetitive activities which mainly involve 
-random hand gestures. Therefore, for the four activities 
-(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
-we select only four window sizes for analysis: 2-s window (100 samples), 
-5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
-(Fig~\ref{fig:sts}).
+Therefore, the use of Shannon Entropy could be applied 
+to analyse human participants who might vary in age, 
+state of health, anthropometric features and capability to perform movement.
 

### 2020-10-28 docs/submission/arxiv/v02/main-arxiv.tex
axiv submission v02

--- 2020-10-18 docs/manuscript/main.tex
@@
 \usepackage[T1]{fontenc}
-\graphicspath{{../}} %goes to path: docs/
+%\graphicspath{{../}} %goes to path: docs/
 
@@
 \author[1,*]{Miguel Xochicale}
-\author[1]{Chris Baber}
-\affil[1]{University of Birmingham, 
-	School of Engineering, 
+\author[2]{Chris Baber}
+\affil[1]{King's College London, 
+	School of Biomedical Engineering amd Imaging Sciences,
+	London, SE1 7EU, UK
+	} 
+\affil[2]{University of Birmingham,
+	School of Computer Science,
 	Birmingham, 
@@
 	UK}
-\affil[*]{email@user.com}
+\affil[*]{miguel.xochicale@kcl.ac.uk}
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
+\includegraphics[width=1.0\textwidth]{rss.pdf}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
+\includegraphics[width=1.0\textwidth]{cao.pdf}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
+\includegraphics[width=1.0\textwidth]{ami.pdf}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
+\includegraphics[width=1.0\textwidth]{rp}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
+\includegraphics[width=1.0\textwidth]{rpsp}
     \caption{
@@
   \centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
+\includegraphics[width=1.0\textwidth]{hri}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
+\includegraphics[width=1.0\textwidth]{sts}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
+\includegraphics[width=1.0\textwidth]{fig1}
     	\caption{ 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/caoami/pdf/fig2}
+\includegraphics[width=1.0\textwidth]{fig2}
 	\caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig3}
+\includegraphics[width=1.0\textwidth]{fig3}
 \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig4}
+\includegraphics[width=1.0\textwidth]{fig4}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aH}
+\includegraphics[width=1.0\textwidth]{rp_aH}
 \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aV}
+\includegraphics[width=1.0\textwidth]{rp_aV}
 \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-bp}
+\includegraphics[width=1.0\textwidth]{rqa-bp}
     \caption
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-epsilons}
+\includegraphics[width=1.0\textwidth]{rqa-epsilons}
     \caption
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
+\includegraphics[width=1.0\textwidth]{rqa-sensors-activities}
     \caption{
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/rqa/output/rqa-participants}
+\includegraphics[width=0.9\textwidth]{rqa-participants}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-windows}
+\includegraphics[width=1.0\textwidth]{rqa-windows}
     \caption{
@@
 %\bibliography{sample}
-%\bibliography{references}
-\bibliography{../references/references}
+%\bibliography{references.bib}
+\input{main-arxiv.bbl}
 

### 2020-10-28 docs/submission/arxiv/reducepdfsize.sh

@@ -1,0 +1,11 @@
+
+cp ../../manuscript/main.pdf . 
+PDFQUALITY=printer
+#-dPDFSETTINGS=/screen lower quality, smaller size. (72 dpi)
+#-dPDFSETTINGS=/ebook for better quality, but slightly larger pdfs. (150 dpi)
+#-dPDFSETTINGS=/prepress output similar to Acrobat Distiller "Prepress Optimized" setting (300 dpi)
+#-dPDFSETTINGS=/printer selects output similar to the Acrobat Distiller "Print Optimized" setting (300 dpi)
+#-dPDFSETTINGS=/default selects output intended to be useful across a wide variety of uses, possibly at the expense of a larger output file
+
+gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.5 -dPDFSETTINGS=/$PDFQUALITY -dNOPAUSE -dQUIET -dBATCH -sOutputFile=manuscript-v02.pdf main.pdf
+rm -rf main.pdf

### 2020-12-17 docs/submission/srep/coverletter/letter-tex/covlet.tex
adds docs/submission/srep/coverletter/

\documentclass[10pt]{letter}
\usepackage{fontawesome}
\usepackage{hyperref}

% Margins
\topmargin=-1in % Moves the top of the document 1 inch above the default
\textheight=8.5in % Total height of the text on the page before text goes on to the next page, this can be increased in a longer letter
\oddsidemargin=-10pt % Position of the left margin, can be negative or positive if you want more or less room
\textwidth=6.5in % Total width of the text, increase this if the left margin was decreased and vice-versa
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\begin{document}
%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------
\begin{letter}{Dr. Richard White \\
Chief Editor of Scientific Reports \\
Springer Nature Limited \\
The Campus, 4 Crinan Street\\
London, N1 9XW \\
UK
}

%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------
\begin{flushright}
\large\bf Miguel Xochicale \\ % Your name
%\vspace{20pt} \hrule height 1pt % If you would like a horizontal line separating the name from the address, uncomment the line to the left of this text
4th Floor \\
Lambeth Wing  St. Thomas' Hospital \\
Westminster Bridge Road \\
London SE1 7EH \\
\faEnvelopeO  miguel.xochicale@kcl.ac.uk \\
\faMobile  (+44) 0744 281 7616 \\
\faHome \href{http://mxochicale.github.io}{http://mxochicale.github.io}
%\vspace{20pt}
%\large\bf Chris Baber, Professor \\ % Your name
%%\vspace{20pt} \hrule height 1pt % If you would like a horizontal line separating the name from the address, uncomment the line to the left of this text
%123 Broadway \\ City, State 12345 \\ (000) 111-1111 % Your address and phone number
\end{flushright}

%\vfill
\vspace{20pt}
%\author{Miguel Xochicale, King's College London}
\signature{Miguel Xochicale} % Your name for the signature at the bottom

%----------------------------------------------------------------------------------------
%	LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Dear Editor Richard White,}

%* The affiliation and contact information of your corresponding author
I have the pleasure of sending you the manuscript entitled 
"Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities"
authored by Miguel Xochicale and Chris Baber to be considered for publication as a research article 
in your prestigious journal Scientific Reports.

%* A brief explanation of why the work is appropriate for Scientific Reports
We selected Scientific Reports for 
its reputation on publishing works that are algigment to the principles of reproducibility and open science. 
as well as the various topics and readers in areas of natural and clinical science that our work can be reached. 
Paper is containing original research and has not been submitted or published earlier 
in any journal and is not being considered for publication elsewhere. 
All authors have seen and approved the manuscript and have contributed significantly for the paper.

%* The names and contact information of any reviewers you consider suitable
%* The names of any referees you would like excluded from reviewing
Two potential reviewers that we consider suitable for our work can be:
Bedartha Goswami, bedartha.goswami@uni-tuebingen.de,
at University of Tübingen investigating nonlinear time series analysis; recurrence plots, and
Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin investigating 
dynamics time delay networks of dynamical systems; pattern formation; coupled systems.

%BLURS
%The manuscript is original (i.e., my co-author and myself wrote it, not copied it).
%https://www.aje.com/en/arc/writing-cover-letter/

\closing{Sincerely yours,}

\encl{manuscript and supplementary document} 
% List your enclosed documents here, comment this out to get rid of the "encl:"
%----------------------------------------------------------------------------------------

\end{letter}
\end{document}

### 2020-12-17 docs/manuscript/main.tex
axiv submission v02
merge fig2fig3 into fig2 for caoami

@@
 series: the solution for the $x$ variable of the Lorenz system 
-(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
-and a variance of one (Fig~\ref{fig:e1e2}F).
+(Fig~\ref{fig:caoami}A), and a Gaussian noise time series with zero mean 
+and a variance of one (Fig~\ref{fig:caoami}B).
 We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
@@
 The determination of six is given that any value of $m$ can be used as they 
-are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
+are within the threshold of $1\pm0.05$ (Fig~\ref{fig:caoami}C).
 $E_2(m)$ values, for chaotic time series, are different to one 
-(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
+(Fig~\ref{fig:caoami}E), for which, it can be concluded that for the 
 chaotic time series the minimum embedding dimension the time series 
@@
 $E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
+which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:caoami}D).
 $E_1(m)$ values then indicate the minimum embedding dimension of the 
 noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
+approximately equal to one (Figure~\ref{fig:caoami}F) for which it can be 
 concluded that noise time series is a stochastic signal.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
-    \caption{
-	{\bf Minimum dimension embedding values with Cao's method.} 
-	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
-	with variations of $\tau$ values from one to twenty
-	for (E) chaotic and (F) random time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:e1e2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
 It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
@@
 regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-(Fig~\ref{fig:e1e2}A,B,C, and D). 
+(Fig~\ref{fig:caoami}C,D,E and F). 
 However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
@@
 B) a noise time series using a normal distribution with mean zero and 
-standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
+standard deviation equal to one. From Fig~\ref{fig:caoami}(G, H), it can then be 
 concluded that the amount of knowledge for any noise time series is zero 
@@
 \cite{bradley2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
-    \caption{
-	{\bf Minimum delay embedding values with AMI's method.} 
-    	(A, B) AMI values where its first minimum value in the curve
-	is the minimum time delay embedding ($\tau_0$), 
-	for (C) a chaotic and (D) noise time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:amis}
-\end{figure}
 %%---------------------------------(FIGURE)-------------------------------------
@@
 for which AMI depends on partitioning \cite{garcia2005e71}.
+
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/caoami/versions/drawing-v00.pdf}
+    \caption{
+	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
+	(A) chaotic and (B) random time series.
+	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
+	with variations of $\tau$ values from one to twenty
+	for chaotic and random time series.
+	(G, H) AMI values where its first minimum value in the curve
+	is the minimum time delay embedding ($\tau_0$).
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:caoami}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
 

### 2020-12-17 docs/manuscript/main.tex
merge fig4andfig5 into fig3 for rps

@@
 function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
+(Fig~\ref{fig:rps}(A,B)) \cite{eckmann1987, marwan2007,marwan2015}.
 RP is also characterised with a line of identity (LOI) which is a  
 diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$.
-%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
-    \caption{
-	{\bf Recurrence Plots.} 
-	(A) State space of the Lorenz system with controlling parameters 
-	($\rho=28, \sigma=10, \beta=8/3$). 
-	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
-	(black circle) of a given point at $i$ is a recurrent point and is 
-	represented as a black dot in the recurrence plot at location 
-	$(i, j)$ or white otherwise.
-	(B) Recurrence plot using the 
-	three components of the Lorenz system and the RP with no embeddings 
-	and threshold $\epsilon=5$.
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:mrp}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
 
@@
 (i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-uniformly distributed noise (Fig~\ref{fig:rp2}A), 
+uniformly distributed noise (Fig~\ref{fig:rps}C), 
 (ii) periodic and quasi-periodic systems where diagonal lines and 
 checkerboard structures represent oscillating systems, e.g., sinusoidal 
-signals (Fig~\ref{fig:rp2}B), 
+signals (Fig~\ref{fig:rps}D), 
 (iii) drift where paling or darkening recurrence points away from 
 the LOI is caused by drifting systems, 
-e.g., logistic map (Fig~\ref{fig:rp2}C), and
+e.g., logistic map (Fig~\ref{fig:rps}E), and
 (iv) disrupted where recurrence points are presented white areas or 
 bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
+(Fig~\ref{fig:rps}F) \cite{eckmann1987, marwan2015}.
 Texture patterns in RPs can be categorised as:
@@
 to laminar states or singularities \cite{marwan2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
-    \caption{
-	{\bf Patterns in Recurrence Plots.} 
-	Time-series with its respective recurrence plot for:
-	(A) uniformly distributed noise,
-	(B) super-positionet harmonic oscillation 
-	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-	with a linearly increase term ($0.01*i$), and
-	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:rp2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
 
@@
 recurrent pattern structures in the RP \cite{zbilut1992}.
+
+%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{figures/methods/rps/versions/drawing-v00.pdf}
+    \caption{
+	{\bf Recurrence Plots and its patterns.} 
+	(A) State space of the Lorenz system with controlling parameters 
+	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
+	(black circle) of a given point at $i$ is a recurrent point and is 
+	represented as a black dot in the recurrence plot at location 
+	$(i, j)$ or white otherwise.
+	(B) Recurrence plot using the 
+	three components of the Lorenz system and the RP with no embeddings 
+	and threshold $\epsilon=5$.
+	Time-series with its respective recurrence plot for:
+	(C) uniformly distributed noise,
+	(D) super-positionet harmonic oscillation 
+	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
+	with a linearly increase term ($0.01*i$), and
+	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	This figure is adapted from \cite{marwan2015}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+	}
+    \label{fig:rps}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
 

### 2020-12-17 docs/manuscript/main.tex
adds merge rss figures into one

@@
 the first three axis of the rotated Principal Components Analysis (PCA) 
-are shown for the reconstructed state spaces of horizontal (figures~\ref{fig:rss_aHw10}) 
-and vertical (figures~\ref{fig:rss_aVw10}) arm movements. 
+are shown for the reconstructed state spaces of horizontal (figures~\ref{fig:rsss}(A)) 
+and vertical (figures~\ref{fig:rsss}(B)) arm movements. 
 While visual inspection of these figures 
@@
 Recurrence Plots and Recurrence Quantification Analysis.
-
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig3}
+\includegraphics[width=1.0\textwidth]{figures/rss/versions/drawing-v00.pdf}
 \caption{
-	{\bf RSSs for horizontal arm movements.}
+	{\bf RSSs for horizontal and vertical arm movements.}
 	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
@@
 	embedding parameters 
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$
+	for (A) horizontal and (B) vertical arm movements.
 	Code and data to reproduce the figure is available in \cite{srep2020}.	
         }
-    \label{fig:rss_aHw10}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig4}
-    \caption{
-	{\bf RSSs for vertical arm movements.}
-	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
-	Reconstructed state spaces were computed with 
-	embedding parameters
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:rss_aVw10}
+    \label{fig:rsss}
 \end{figure}

### 2020-12-17 docs/manuscript/main.tex
merge RPs figures into one

@@
 is 1. 
-Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rp_aH}) and 
-vertical (figures~\ref{fig:rp_aV}) arm movements.
+Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rps}(A)) and 
+vertical (figures~\ref{fig:rps}(B)) arm movements.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aH}
+\includegraphics[width=1.0\textwidth]{figures/rps/versions/drawing-v00.pdf}
 \caption{
-	{\bf RPs for horizontal arm movements.}	
+	{\bf RPs for horizontal (A) and vertical (B) arm movements.}	
 	Recurrence plots were computed with 
@@
         }
-    \label{fig:rp_aH}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/rps/pdf/rp_aV}
-\caption{
-	{\bf RPs for vertical arm movements.}	
-	Recurrence plots were computed with 
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:rp_aV}
+    \label{fig:rps}
 \end{figure}

### 2020-12-21 docs/submission/srep/coverletter/letter-tex/covlet.tex
polishes docs/submission/srep/coverletter/letter-tex/covlet.tex

@@
 %* A brief explanation of why the work is appropriate for Scientific Reports
-We selected Scientific Reports for 
-its reputation on publishing works that are algigment to the principles of reproducibility and open science. 
-as well as the various topics and readers in areas of natural and clinical science that our work can be reached. 
-Paper is containing original research and has not been submitted or published earlier 
+Our work is appropriate to be considered to be published in Scientific Reports due to
+its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility).
+Similarly, we chosen Scientific Reports because of the outcomes of our work might lead to 
+potential applications in areas of natural and clinical science.
+Our manuscript is containing original research and has not been submitted or published earlier 
 in any journal and is not being considered for publication elsewhere. 
-All authors have seen and approved the manuscript and have contributed significantly for the paper.
+Authors have seen and approved the manuscript and have contributed significantly to it.
 
@@
 %* The names of any referees you would like excluded from reviewing
-Two potential reviewers that we consider suitable for our work can be:
+Finally, two potential reviewers for our work can be:
 Bedartha Goswami, bedartha.goswami@uni-tuebingen.de,

### 2020-12-23 docs/manuscript/main.tex
abstract is 200words docs/manuscript/main.tex

@@
 Human movement variability arises from the process of 
-mastering redundant (bio)mechanical degrees of freedom (DOF)
+mastering redundant (bio)mechanical degrees of freedom
 to successfully accomplish any given motor task
-(from the most complex to even the simplest one) 
 where flexibility and stability of many possible joint 
@@
 popular as a diagnostic tool or skill performance evaluation, 
-there are remain challenges in terms of defining 
-the most appropriate methods and parameters to apply. 
-For this work, we therefore investigate nonlinear dynamics methods 
+there are remain challenges on applying the most appropriate methods. 
+We therefore investigate nonlinear methods 
 such as reconstructed state space (RSSs), uniform time-delay embedding, 
-recurrence plots (RPs) and recurrence quantification analysis metrics (RQAs)
+recurrence plots (RPs) and recurrence quantification analysis (RQAs)
 with real-world time-series data of wearable inertial sensors. 
-That said, twenty right-handed healthy participants 
-imitated simple vertical and horizontal arm movements in normal 
+That said, twenty healthy participants 
+imitated vertical and horizontal arm movements in normal 
 and faster velocity from an humanoid robot.
-We applied nonlinear methods to four activities, 
-four window lengths and three levels of smoothed time-series data,
-to then found visual differences in the patterns of RSSs and RPs
-and statistical differences with RQA metrics.
-We therefore conclude that Shannon Entropy for RQA is a robust method 
-that help us to quantify activities, types of sensors, windows lengths 
+We applied nonlinear methods to the collected data 
+to found visual differences in the patterns of RSSs and RPs
+and statistical differences with RQAs.
+We conclude that Shannon Entropy with RQA is a robust method 
+that helps to quantify activities, types of sensors, windows lengths 
 and level of smoothness.

### 2020-12-27 docs/supplementary-information/report/report.tex
polishes docs/supplementary-information/

@@
 \usepackage{graphicx}
+\usepackage{url}
 \graphicspath{{../../}} %%%goes to the path: docs/
@@
 \begin{abstract}
-Section 1 shows surface plots of 3D RQA ENTR values for 3 participants (p01, p02 and p03).
-Datasets are for 
-(a) window size of 100, 250, 500 and 750 samples (w100, w250, w500, w750);
-(b) smoothnes: sg0, sg1 and sg2;
-(c) movement: Horizontal Normal (HN), Horizontal Faster (HF), Vertical Normal (VN) and Vertical Faster (VF); and 
-(d) sensors: Sensor attached to Human (HS01), sensor attached to Robot (RS01).
+Section 1 describes datasets and section 2 shows surface plots of 3D RQA ENTR values for 3 participants (p01, p02 and p03).
 \end{abstract}
@@
 %\newpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Datasets}
+Datasets are for 
+(a) window size of 100, 250, 500 and 750 samples (w100, w250, w500, w750);
+(b) smoothness: sg0, sg1 and sg2;
+(c) movement: Horizontal Normal (HN), Horizontal Faster (HF), Vertical Normal (VN) and Vertical Faster (VF); and 
+(d) sensors: Sensor attached to Human (HS01), sensor attached to Robot (RS01).
+
+The location of the dataset is shown below with the first two lines of xdatav00.dt denoting labels.
+\begin{verbatim}
+~/srep2020/data/dataset$ tree --si
+[ 46M]  xdata_v00.dt
+.
+.
+.
+"Participant" "Activity" "Sensor" "Sample" "Time" 
+"sg0zmuvGyroY" "sg1zmuvGyroY" "sg2zmuvGyroY" 
+"sg0zmuvGyroZ" "sg1zmuvGyroZ" "sg2zmuvGyroZ"
+...
+"p01" "HN" "HS01" 1 0 
+0.0110396263359954 0.00606430191548277 0.0385586376765087 
+0.00467559072085554 0.00509907428630649 0.168927517477539
+...
+\end{verbatim}
+
+%├── [4.1k]  dataset
+%│   └── [ 46M]  xdata_v00.dt
+%.
+%.
+%.
 
@@
 ## Data
-$HOME/srep2019/data/rqa$ tree -s
+$HOME/srep2020/data/rqa$ tree -s
 
@@
 
-Datasets for NN participants where NN is 01, 02 and 03 are:
+For the following plots, datasets for NN participants where NN is 01, 02 and 03 are:
 \begin{verbatim}

### 2020-12-30 docs/submission/srep/coverletter/letter-tex/covlet.tex
polishes README.md; polishes coverletter; updates abstracts and re-submit to arxiv with v04

@@
 %* The affiliation and contact information of your corresponding author
-I have the pleasure of sending you the manuscript entitled 
-"Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities"
-authored by Miguel Xochicale and Chris Baber to be considered for publication as a research article 
-in your prestigious journal Scientific Reports.
+I have the pleasure of sending you the manuscript entitled "Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities" authored by Miguel Xochicale and Chris Baber to be considered for publication as a research article in your prestigious journal.
 
 %* A brief explanation of why the work is appropriate for Scientific Reports
-Our work is appropriate to be considered to be published in Scientific Reports due to
-its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility).
-Similarly, we chosen Scientific Reports because of the outcomes of our work might lead to 
-potential applications in areas of natural and clinical science.
-Our manuscript is containing original research and has not been submitted or published earlier 
-in any journal and is not being considered for publication elsewhere. 
+Our work is appropriate for Scientific Reports due to its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility).   
+Similarly, we choose Scientific Reports because of the potential reachability to readers in areas of natural and clinical sciences along with potential application of our work and the spark for collaborations.
+
+Our manuscript is containing original research and has not been submitted or published earlier in any journal and is not being considered for publication elsewhere.   
 Authors have seen and approved the manuscript and have contributed significantly to it.
@@
 %* The names of any referees you would like excluded from reviewing
-Finally, two potential reviewers for our work can be:
-Bedartha Goswami, bedartha.goswami@uni-tuebingen.de,
-at University of Tübingen investigating nonlinear time series analysis; recurrence plots, and
-Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin investigating 
-dynamics time delay networks of dynamical systems; pattern formation; coupled systems.
+Finally, two potential reviewers for our work can be: Bedartha Goswami, bedartha.goswami@uni-tuebingen.de, at University of Tübingen investigating nonlinear time series analysis; recurrence plots, and Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin investigating dynamics time delay networks of dynamical systems; pattern formation; coupled systems.
 

### 2021-03-15 docs/manuscript/main.tex
polishes README.md; polishes coverletter; updates abstracts and re-submit to arxiv with v04
fused fig1,2,3 into fig01

@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}
+\includegraphics[width=1.0\textwidth]{figures/fig01/versions/drawing-v00}
     \caption{
@@
    	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.	
+	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
+	(A) chaotic and (B) random time series.
+	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
+	with variations of $\tau$ values from one to twenty
+	for chaotic and random time series.
+	(G, H) AMI values where its first minimum value in the curve
+	is the minimum time delay embedding ($\tau_0$).
+	{\bf Recurrence Plots and its patterns.} 
+	(A) State space of the Lorenz system with controlling parameters 
+	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
+	(black circle) of a given point at $i$ is a recurrent point and is 
+	represented as a black dot in the recurrence plot at location 
+	$(i, j)$ or white otherwise.
+	(B) Recurrence plot using the 
+	three components of the Lorenz system and the RP with no embeddings 
+	and threshold $\epsilon=5$.
+	Time-series with its respective recurrence plot for:
+	(C) uniformly distributed noise,
+	(D) super-positionet harmonic oscillation 
+	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
+	with a linearly increase term ($0.01*i$), and
+	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	This figure is adapted from \cite{marwan2015}.
 	Code and data to reproduce the figure is available in \cite{srep2020}.
@@
 
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/caoami/versions/drawing-v00.pdf}
-    \caption{
-	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
-	(A) chaotic and (B) random time series.
-	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
-	with variations of $\tau$ values from one to twenty
-	for chaotic and random time series.
-	(G, H) AMI values where its first minimum value in the curve
-	is the minimum time delay embedding ($\tau_0$).
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:caoami}
-\end{figure}
+%%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=1.0\textwidth]{figures/methods/caoami/versions/drawing-v00.pdf}
+%    \caption{
+%	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
+%	(A) chaotic and (B) random time series.
+%	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
+%	with variations of $\tau$ values from one to twenty
+%	for chaotic and random time series.
+%	(G, H) AMI values where its first minimum value in the curve
+%	is the minimum time delay embedding ($\tau_0$).
+%	Code and data to reproduce the figure is available in \cite{srep2020}.
+%        }
+%    \label{fig:caoami}
+%\end{figure}
 %%---------------------------------(FIGURE)-------------------------------------
@@
 
-%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/methods/rps/versions/drawing-v00.pdf}
-    \caption{
-	{\bf Recurrence Plots and its patterns.} 
-	(A) State space of the Lorenz system with controlling parameters 
-	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
-	(black circle) of a given point at $i$ is a recurrent point and is 
-	represented as a black dot in the recurrence plot at location 
-	$(i, j)$ or white otherwise.
-	(B) Recurrence plot using the 
-	three components of the Lorenz system and the RP with no embeddings 
-	and threshold $\epsilon=5$.
-	Time-series with its respective recurrence plot for:
-	(C) uniformly distributed noise,
-	(D) super-positionet harmonic oscillation 
-	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-	with a linearly increase term ($0.01*i$), and
-	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:rps}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
+%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=1.0\textwidth]{figures/methods/rps/versions/drawing-v00.pdf}
+%    \caption{
+%	{\bf Recurrence Plots and its patterns.} 
+%	(A) State space of the Lorenz system with controlling parameters 
+%	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
+%	(black circle) of a given point at $i$ is a recurrent point and is 
+%	represented as a black dot in the recurrence plot at location 
+%	$(i, j)$ or white otherwise.
+%	(B) Recurrence plot using the 
+%	three components of the Lorenz system and the RP with no embeddings 
+%	and threshold $\epsilon=5$.
+%	Time-series with its respective recurrence plot for:
+%	(C) uniformly distributed noise,
+%	(D) super-positionet harmonic oscillation 
+%	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+%	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
+%	with a linearly increase term ($0.01*i$), and
+%	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+%	This figure is adapted from \cite{marwan2015}.
+%	Code and data to reproduce the figure is available in \cite{srep2020}.
+%	}
+%    \label{fig:rps}
+%\end{figure}
+%%%---------------------------------(FIGURE)-------------------------------------
 
@@
 is 1. 
-Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rps}(A)) and 
-vertical (figures~\ref{fig:rps}(B)) arm movements.
+Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rpsr}(A)) and 
+vertical (figures~\ref{fig:rpsr}(B)) arm movements.
 %%---------------------------------(FIGURE)-------------------------------------
@@
         }
-    \label{fig:rps}
+    \label{fig:rpsr}
 \end{figure}

### 2021-03-18 docs/manuscript/main.tex
polishes fig01 and updates its caption

@@
 	{\bf State space reconstruction methodology.}
-	State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$
-	where $f^t$ is the true dynamical system, $s(t)$ indicates the state,
-    	$s$, at time, $t$,  and $h[ ]$ the measurement function. 
-    	The time-delay embedding represented as the $\Phi$, maps the original
-    	$d-$dimensional state $s(t)$ into the $m-$dimensional uniform 
-	time-delay embedding matrix $\boldsymbol{X}(t)$.
-    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into 
-	a new state $y(t)$ of dimensions $n < m$.
-    	(A) $M-$dimensional manifold representing the state space 
-	(e.g. Lorenz system);
-    	(B) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
-   	(C) $m-$dimensional reconstructed state space with 
-    	\texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and 
-    	(D) $y(t)$ is the $n-$dimensional reconstructed state space.
-	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $
-	where $\Phi$ is the delay reconstruction map and 
-	$\Psi$ is the coordinate transformation map.
-	This figure is adapted from 
-   	\cite{Quintana-Duque2012, casdagli1991, uzal2011}.	
+		State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$ where $f^t$ is the true dynamical system, $s(t)$ indicates the state, $s$, at time, $t$,  and $h[ ]$ the measurement function.
+		The time-delay embedding represented as the $\Phi$, maps the original $d-$dimensional state $s(t)$ into the $m-$dimensional uniform time-delay embedding matrix $\boldsymbol{X}(t)$.
+    	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into a new state $y(t)$ of dimensions $n < m$.
+    	(A.1) $M-$dimensional manifold representing the state space (e.g. Lorenz system);
+    	(B.1) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
+		(C.1) $m-$dimensional reconstructed state space with \texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and
+    	(D.1) $y(t)$ is the $n-$dimensional reconstructed state space.
+	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $where $\Phi$ is the delay reconstruction map and $\Psi$ is the coordinate transformation map.
+	A.1 to D.1 figures are adapted from \cite{Quintana-Duque2012, casdagli1991, uzal2011}.
 	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
-	(A) chaotic and (B) random time series.
-	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
-	with variations of $\tau$ values from one to twenty
-	for chaotic and random time series.
-	(G, H) AMI values where its first minimum value in the curve
-	is the minimum time delay embedding ($\tau_0$).
+	(A.2) chaotic and (B.2) random time series.
+	(C.2, D.2) $E_1 (m)$ values and (E.2, F.2) $E_2(m)$ values with variations of $\tau$ values from one to twenty for chaotic and random time series.
+	(G.2, H.2) AMI values where its first minimum value in the curve is the minimum time delay embedding ($\tau_0$).
 	{\bf Recurrence Plots and its patterns.} 
-	(A) State space of the Lorenz system with controlling parameters 
-	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
-	(black circle) of a given point at $i$ is a recurrent point and is 
-	represented as a black dot in the recurrence plot at location 
-	$(i, j)$ or white otherwise.
-	(B) Recurrence plot using the 
-	three components of the Lorenz system and the RP with no embeddings 
-	and threshold $\epsilon=5$.
+	(A.3) State space of the Lorenz system with controlling parameters ($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood (black circle) of a given point at $i$ is a recurrent point and is represented as a black dot in the recurrence plot at location $(i, j)$ or white otherwise.
+	(B.3) Recurrence plot using the three components of the Lorenz system and the RP with no embeddings and threshold $\epsilon=5$.
 	Time-series with its respective recurrence plot for:
-	(C) uniformly distributed noise,
-	(D) super-positionet harmonic oscillation 
-	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-	with a linearly increase term ($0.01*i$), and
-	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	This figure is adapted from \cite{marwan2015}.
+	(C.3) uniformly distributed noise,
+	(D.3) super-positionet harmonic oscillation ($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(E.3) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted with a linearly increase term ($0.01*i$), and
+	(F.3) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	Figures A.3 to F.3 are adapted from \cite{marwan2015}.
 	Code and data to reproduce the figure is available in \cite{srep2020}.

### 2021-03-18 docs/manuscript/main.tex
updates references for github to srep2021
fixes references for figure01

@@
 map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
-Fig~\ref{fig:ssr} illustrates the state space reconstruction.
+Figures~\ref{fig01}(A.1 to D.1) illustrates the state space reconstruction.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 	Figures A.3 to F.3 are adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
     }
-    \label{fig:ssr}
+    \label{fig01}
 \end{figure}
@@
 series: the solution for the $x$ variable of the Lorenz system 
-(Fig~\ref{fig:caoami}A), and a Gaussian noise time series with zero mean 
-and a variance of one (Fig~\ref{fig:caoami}B).
+(Fig~\ref{fig01}A.2), and a Gaussian noise time series with zero mean
+and a variance of one (Fig~\ref{fig01}B.2).
 We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
@@
 The determination of six is given that any value of $m$ can be used as they 
-are within the threshold of $1\pm0.05$ (Fig~\ref{fig:caoami}C).
+are within the threshold of $1\pm0.05$ (Fig~\ref{fig01}C.2).
 $E_2(m)$ values, for chaotic time series, are different to one 
-(Fig~\ref{fig:caoami}E), for which, it can be concluded that for the 
+(Fig~\ref{fig01}E.2), for which, it can be concluded that for the
 chaotic time series the minimum embedding dimension the time series 
@@
 $E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:caoami}D).
+which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig01}D.2).
 $E_1(m)$ values then indicate the minimum embedding dimension of the 
 noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig:caoami}F) for which it can be 
+approximately equal to one (Figure~\ref{fig01}F.2) for which it can be
 concluded that noise time series is a stochastic signal.
@@
 regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-(Fig~\ref{fig:caoami}C,D,E and F). 
+(Fig~\ref{fig01}C.2,D.2,E.2 and F.2).
 However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
@@
 B) a noise time series using a normal distribution with mean zero and 
-standard deviation equal to one. From Fig~\ref{fig:caoami}(G, H), it can then be 
+standard deviation equal to one. From Fig~\ref{fig01}(G.2, H.2), it can then be
 concluded that the amount of knowledge for any noise time series is zero 
@@
 %	is the minimum time delay embedding ($\tau_0$).
-%	Code and data to reproduce the figure is available in \cite{srep2020}.
+%	Code and data to reproduce the figure is available in \cite{srep2021}.
 %        }
@@
 function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig:rps}(A,B)) \cite{eckmann1987, marwan2007,marwan2015}.
+(Fig~\ref{fig01}(A.3,B.3)) \cite{eckmann1987, marwan2007,marwan2015}.
 RP is also characterised with a line of identity (LOI) which is a  
@@
 (i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-uniformly distributed noise (Fig~\ref{fig:rps}C), 
+uniformly distributed noise (Fig~\ref{fig01}C.3),
 (ii) periodic and quasi-periodic systems where diagonal lines and 
 checkerboard structures represent oscillating systems, e.g., sinusoidal 
-signals (Fig~\ref{fig:rps}D), 
+signals (Fig~\ref{fig01}D.3),
 (iii) drift where paling or darkening recurrence points away from 
 the LOI is caused by drifting systems, 
-e.g., logistic map (Fig~\ref{fig:rps}E), and
+e.g., logistic map (Fig~\ref{fig01}E.3), and
 (iv) disrupted where recurrence points are presented white areas or 
 bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-(Fig~\ref{fig:rps}F) \cite{eckmann1987, marwan2015}.
+(Fig~\ref{fig01}F.3) \cite{eckmann1987, marwan2015}.
 Texture patterns in RPs can be categorised as:
@@
 %	This figure is adapted from \cite{marwan2015}.
-%	Code and data to reproduce the figure is available in \cite{srep2020}.
+%	Code and data to reproduce the figure is available in \cite{srep2021}.
 %	}
@@
 		which are also presented in (B), (C) and (D).
-		Code and data to reproduce the figure is available in \cite{srep2020}.
+		Code and data to reproduce the figure is available in \cite{srep2021}.
         }
@@
 		and other sensor attached to the robot (RS01).	
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
@@
 	and window length of 10-sec (500 samples).
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
@@
 	for (A) horizontal and (B) vertical arm movements.
-	Code and data to reproduce the figure is available in \cite{srep2020}.	
+	Code and data to reproduce the figure is available in \cite{srep2021}.	
         }
@@
 	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
@@
 	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$. 
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
@@
 	a window size of 10-secs (500 samples).
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
@@
 	window size of 10-secs (500 samples).
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
        }
@@
 	RQA ENTR values were computed from data of sg0 and window size of 10-secs (500 samples).
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
        }
@@
 	RQA ENTR values were computed from data of $p01$ and sg0.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
        }

### 2021-03-18 docs/manuscript/main.tex
fused fig2,3,4 into fig02 and polishes its caption

@@
   \centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
+\includegraphics[width=1.0\textwidth]{figures/fig02/versions/drawing-v00.png}
     \caption{
 	{\bf Human-humanoid imitation activities.} 
-    		Face-to-face human-humanoid imitation (HHI) activities for 
-		(A) HHI of horizontal arm movement, 
-		(B) Humanoid horizontal arm movement,
-		(C) HHI of vertical arm movement, and 
-		(D) Humanoid vertical arm movement.
+		Face-to-face human-humanoid imitation (HHI) activities for (A.1) HHI of horizontal arm movement, (B.1) Humanoid horizontal arm movement, (C.1) HHI of vertical arm movement, and (D.1) Humanoid vertical arm movement.
+	{\bf Time series duration for arm movements.}
+    	Time series of smoothed data from gyroscope sensor for different speed arm movements performed by NAO: (A.2) Horizontal Normal arm movement (HN), (B.2) Horizontal Faster arm movement (HF), (C.2) Vertical Normal arm movement (VN) and (D.2) Vertical Faster arm movement (VF).
+		(A.2) shows window sizes for 2-seconds (100 samples), 5-seconds (250 samples), 10-seconds (500 samples) and 15-seconds (750 samples) which are also presented in (B.2), (C.2) and (D.2).
+	{\bf Time series for horizontal and vertical arm movements.}
+		(A.3/D.3) raw-normalised (sg0zmuv), (B.3/E.3) normalised-smoothed 1 (sg1zmuv) and (C.3/F.3) normalised-smoothed 2 (sg2zmuv).
+		Time series are only for three participants	($p01$, $p02$, and $p03$) for horizontal and vertical arm movements in normal and faster velocity (HN, HF, VN, VF) with the normalised GyroZ or GyroY axis and with one sensor attached to the participant (HS01) and other sensor attached to the robot (RS01).
+	Code and data to reproduce the figure is available in \cite{srep2021}.
 }
@@
 and 15-sec (750-samples).
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht] 
-\centering
-\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
-    \caption{
-	{\bf Time series duration of horizontal and vertical arm movements.} 
-    		Time series of smoothed data from gyroscope sensor 
-		for different speed arm movements performed by NAO: 
-		(A) Horizontal Normal arm movement (HN), 
-		(B) Horizontal Faster arm movement (HF),
-		(C) Vertical Normal arm movement (VN) and 
-		(D) Vertical Faster arm movement (VF).
-		Additionally, 
-		(A) shows window sizes for 2-seconds (100 samples), 
-		5-seconds (250 samples), 10-seconds (500 samples) 
-		and 15-seconds (750 samples)
-		which are also presented in (B), (C) and (D).
-		Code and data to reproduce the figure is available in \cite{srep2021}.
-        }
-	\label{fig:sts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
+%%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
+%    \caption{
+%	{\bf Time series duration of horizontal and vertical arm movements.}
+%    		Time series of smoothed data from gyroscope sensor
+%		for different speed arm movements performed by NAO:
+%		(A) Horizontal Normal arm movement (HN),
+%		(B) Horizontal Faster arm movement (HF),
+%		(C) Vertical Normal arm movement (VN) and
+%		(D) Vertical Faster arm movement (VF).
+%		Additionally,
+%		(A) shows window sizes for 2-seconds (100 samples),
+%		5-seconds (250 samples), 10-seconds (500 samples)
+%		and 15-seconds (750 samples)
+%		which are also presented in (B), (C) and (D).
+%		Code and data to reproduce the figure is available in \cite{srep2021}.
+%        }
+%	\label{fig:sts}
+%\end{figure}
+%%%---------------------------------(FIGURE)------------------------------------
 
@@
 arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}).
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
-    	\caption{ 
-	{\bf Time series for horizontal and vertical arm movements.}
-		(A/D) raw-normalised (sg0zmuv), 
-		(B/E) normalised-smoothed 1 (sg1zmuv) and
-		(C/F) normalised-smoothed 2 (sg2zmuv).
-		Time series are only for three participants 
-		($p01$, $p02$, and $p03$) 
-		for horizontal and vertical arm movements in normal 
-		and faster velocity (HN, HF, VN, VF) 
-		with the normalised GyroZ or GyroY axis 
-		and with one sensor attached to the participant (HS01) 
-		and other sensor attached to the robot (RS01).	
-	Code and data to reproduce the figure is available in \cite{srep2021}.
-        }
-    \label{fig:ts}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
+%%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
+%    	\caption{
+%	{\bf Time series for horizontal and vertical arm movements.}
+%		(A/D) raw-normalised (sg0zmuv),
+%		(B/E) normalised-smoothed 1 (sg1zmuv) and
+%		(C/F) normalised-smoothed 2 (sg2zmuv).
+%		Time series are only for three participants
+%		($p01$, $p02$, and $p03$)
+%		for horizontal and vertical arm movements in normal
+%		and faster velocity (HN, HF, VN, VF)
+%		with the normalised GyroZ or GyroY axis
+%		and with one sensor attached to the participant (HS01)
+%		and other sensor attached to the robot (RS01).
+%	Code and data to reproduce the figure is available in \cite{srep2021}.
+%       }
+%    \label{fig:ts}
+%\end{figure}
+%%%---------------------------------(FIGURE)------------------------------------
 

### 2021-03-18 docs/manuscript/main.tex
fixes references for figure02

@@
 to the right hand of the participant and to the left hand of the robot 
-(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
+(Figure~\ref{fig02}A.1,C.1). Data were then collected with four NeMEMSi
 IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
@@
 }
-    \label{fig:hri}
+    \label{fig02}
 \end{figure}
@@
 (i) ten repetitions of horizontal arm movement at normal (HN) 
-	and faster (HF) speed (Figure~\ref{fig:hri} A), and
+	and faster (HF) speed (Figure~\ref{fig02}A.1), and
 (ii) ten repetitions of vertical arm movement at normal (VN) 
-	and faster (VF) speed (Figure~\ref{fig:hri} C).
+	and faster (VF) speed (Figure~\ref{fig02}C.1).
 The normal and faster speed of arm movements is defined by the duration 
@@
 To visualise the distinction between normal and faster speed for horizontal 
-and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
+and vertical arm movements, Fig~\ref{fig02}A.2-D.2 shows smoothed time series
 for axes Z and Y of the gyroscope sensors with four window lengths: 
@@
 three participants (p01, p01 and p03) performing horizontal 
-arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}).
+arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}A.3-F.3).
 %%%---------------------------------(FIGURE)-------------------------------------
@@
 5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
-(Fig~\ref{fig:sts}).
+(Figures~\ref{fig02}A.2-D.2).
 
@@
 	{\bf RSSs for horizontal and vertical arm movements.}
-	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
+	Reconstructed state spaces for time series for p01 of Figure \ref{fig02}A.3-F.3.
 	Reconstructed state spaces were computed with 

### 2021-03-18 docs/manuscript/main.tex
fused fig5and6 into fig05 and ammends captions and fixes references

@@
 is 1. 
-Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rpsr}(A)) and 
-vertical (figures~\ref{fig:rpsr}(B)) arm movements.
+Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig05}(A)) and
+vertical (figures~\ref{fig05}(B)) arm movements.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rps/versions/drawing-v00.pdf}
+\includegraphics[width=1.0\textwidth]{figures/fig05/versions/drawing-v00}
 \caption{
@@
 	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
+	{\bf Box plots for RQA metrics (C).}
+	RQA metrics for (C.1) REC, (C.2) DET, (C.3) RATIO, and (C.4) ENTR of
+	20 participants performing HN, HF, VN and VF movements
+	with sensors HS01, RS01 and three smoothed-normalised
+	time series (sg0, sg1 and sg2).
+	RQA values were computed with
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
 	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
-    \label{fig:rpsr}
+    \label{fig05}
 \end{figure}
@@
 %\subsubsection*{REC values}
-Figure~\ref{fig:RQABP}(A) presents box plots of REC values, 
+Figure~\ref{fig05}(C.1) presents box plots of REC values,
 for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements 
@@
 %\subsubsection*{DET values}
-Figure \ref{fig:RQABP}(B) presents DET values and shows 
+Figure \ref{fig05}(C.2) presents DET values and shows
 little difference for type of movement or performer.  
@@
 %\subsubsection*{RATIO values}
-Figure \ref{fig:RQABP}(C) presents the ratio of DET / REC. 
+Figure \ref{fig05}(C.3) presents the ratio of DET / REC.
 These values, for the human performer, vary less for HN than for HF movement.  
@@
 %\subsubsection*{ENTR values}
-Figure \ref{fig:RQABP}(D) shows ENTR values are higher for the human 
+Figure \ref{fig05}(C.4) shows ENTR values are higher for the human
 performer than the robot, and vary with the smoothness of the time-series.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}
-\centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-bp}
-    \caption
-	[Box plots for RQA values]{
-	{\bf Box plots for RQA metrics.}
-	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of 
-	20 participants performing HN, HF, VN and VF movements
-	with sensors HS01, RS01 and three smoothed-normalised  
-	time series (sg0, sg1 and sg2).
-	RQA values were computed with 
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$. 
-	Code and data to reproduce the figure is available in \cite{srep2021}.
-        }
-    \label{fig:RQABP}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
+%%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}
+%\centering
+%\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-bp}
+%    \caption
+%	[Box plots for RQA values]{
+%	{\bf Box plots for RQA metrics.}
+%	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of
+%	20 participants performing HN, HF, VN and VF movements
+%	with sensors HS01, RS01 and three smoothed-normalised
+%	time series (sg0, sg1 and sg2).
+%	RQA values were computed with
+%	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
+%	Code and data to reproduce the figure is available in \cite{srep2021}.
+%       }
+%    \label{fig:RQABP}
+%\end{figure}
+%%%---------------------------------(FIGURE)------------------------------------
 

### 2021-03-18 docs/manuscript/main.tex
fused fig6and7 into fig06 and ammends captions and fixes references

@@
 and levels of smoothness (sg0, sg1, sg2).  
-Figure \ref{fig:RQA-IND} shows that increasing recurrence 
+Figure \ref{fig06}(A) shows that increasing recurrence
 threshold leads to an increase in ENTR regardless of level of smoothness.  
-Similarly, increasing level of smoothness will also increase ENTR (figure \ref{fig:RQA-IND}).
+Similarly, increasing level of smoothness will also increase ENTR (figure \ref{fig06}(A)).
 In terms of movement or performer, RQA ENTR decrease from Slow (HS, VS) to Fast (HF, VF) 
-for both human (HS01) and (RS01) (Figures~\ref{fig:3dRQAENTR_sensoractivities}).
+for both human (HS01) and (RS01) (Figures~\ref{fig06}(B)).
 In terms of individual differences, for human participants, 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-epsilons}
+\includegraphics[width=1.0\textwidth]{figures/fig06/versions/drawing-v00}
     \caption
@@
 	{\bf 
-	3D surface plots of RQA ENTR values for different recurrence threshold and smoothness levels.}
-	RQA ENTR values are
-	for embedding parameters
-	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $,
-	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \} $
-	incrementing by one and three recurrence thresholds $\epsilon=1, 2, 3$.
-	RQA ENTR values were computed with data from $p03$, sensor HS01, with 
-	a window size of 10-secs (500 samples).
+	3D surface plots of RQA ENTR values for different recurrence threshold and smoothness levels (A) and for different sensors and activities (B).}
+	RQA ENTR values are for embedding parameters $ \{ m \in \mathbb{R} | 0 \le m \le 10  \} $, $ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$ incrementing by one and three recurrence thresholds $\epsilon=1, 2, 3$.
+	RQA ENTR values were for A were computed with data from $p03$, sensor HS01, with a window size of 10-secs (500 samples).
+	RQA ENTR values are for embedding parameters $ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$, $ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$ with $\epsilon = 1 $ considering four activities: Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), Vertical Faster (VF), and sensors: Human Sensor 01 (HS01) and Robot Sensor (RS01).
+	RQA ENTR values for B were computed from data of $p03$, sg0 and window size of 10-secs (500 samples).
 	Code and data to reproduce the figure is available in \cite{srep2021}.
         }
-    \label{fig:RQA-IND}
+    \label{fig06}
 \end{figure}
@@
 
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
-    \caption{
-	{\bf 3D surface plots of RQA ENTR values for different sensors and activities.}
-	RQA ENTR values are for embedding parameters
-	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
-	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
-	with $\epsilon = 1 $ considering four activities 
-	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and 
-	Vertical Faster (VF) and sensors Human Sensor 01 (HS01) and 
-	Robot Sensor (RS01).
-	RQA ENTR values were computed from data of $p03$, sg0 and 
-	window size of 10-secs (500 samples).
-	Code and data to reproduce the figure is available in \cite{srep2021}.
-       }
-\label{fig:3dRQAENTR_sensoractivities}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
+%%%---------------------------------(FIGURE)-------------------------------------
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=1.0\textwidth]{figures/rqa/output/rqa-sensors-activities}
+%    \caption{
+%	{\bf 3D surface plots of RQA ENTR values for different sensors and activities.}
+%	RQA ENTR values are for embedding parameters
+%	$ \{ m \in \mathbb{R} | 0 \le m \le 10  \}$,
+%	$ \{ \tau \in \mathbb{R} | 0 \le \tau \le 10  \}$
+%	with $\epsilon = 1 $ considering four activities
+%	Horizontal Normal (HN), Horizontal Faster(HF), Vertical Normal(VN), and
+%	Vertical Faster (VF) and sensors Human Sensor 01 (HS01) and
+%	Robot Sensor (RS01).
+%	RQA ENTR values were computed from data of $p03$, sg0 and
+%	window size of 10-secs (500 samples).
+%	Code and data to reproduce the figure is available in \cite{srep2021}.
+%       }
+%\label{fig:3dRQAENTR_sensoractivities}
+%\end{figure}
+%%%---------------------------------(FIGURE)-------------------------------------
 

### 2021-03-18 docs/manuscript/main.tex
ammends affiliations and correspondance email
figure01 harmonised and updated

@@
 \author[1,*]{Miguel Xochicale}
-\author[2]{Chris Baber}
-\affil[1]{King's College London, 
-	School of Biomedical Engineering amd Imaging Sciences,
-	London, SE1 7EU, UK
-	} 
-\affil[2]{University of Birmingham,
+\author{Chris Baber}
+\affil[1]{University of Birmingham,
 	School of Computer Science,
@@
 	UK}
-\affil[*]{miguel.xochicale@kcl.ac.uk}
+\affil[*]{perez.xochicale@gmail.com}
 
@@
 map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
-Figures~\ref{fig01}(A.1 to D.1) illustrates the state space reconstruction.
+Figures~\ref{fig01}A illustrates the state space reconstruction.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/fig01/versions/drawing-v00}
+\includegraphics[width=1.0\textwidth]{figures/fig01/versions/drawing-v01}
     \caption{
-	{\bf State space reconstruction methodology.}
+	{\bf State space reconstruction methodology (A),
+		Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method (B), and
+		Recurrence Plots and its patterns (C).
+	}
 		State space reconstruction is based on $x(t)=h[s(t)]= h[f^t [s(0)]]$ where $f^t$ is the true dynamical system, $s(t)$ indicates the state, $s$, at time, $t$,  and $h[ ]$ the measurement function.
@@
     	The transformation map $\Psi$ then maps $\boldsymbol{X}(t)$ into a new state $y(t)$ of dimensions $n < m$.
-    	(A.1) $M-$dimensional manifold representing the state space (e.g. Lorenz system);
-    	(B.1) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
-		(C.1) $m-$dimensional reconstructed state space with \texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and
-    	(D.1) $y(t)$ is the $n-$dimensional reconstructed state space.
+    	(a1) $M-$dimensional manifold representing the state space (e.g. Lorenz system);
+    	(a2) Delayed copies of $1-$dimensional $x(t)$ from the Lorenz system;
+		(a3) $m-$dimensional reconstructed state space with \texorpdfstring{$m$}{m} and    \texorpdfstring{$\tau$}{T}, and
+    	(a4) $y(t)$ is the $n-$dimensional reconstructed state space.
 	The total reconstruction map is represented as $\Xi = \Psi \circ \Phi $where $\Phi$ is the delay reconstruction map and $\Psi$ is the coordinate transformation map.
-	A.1 to D.1 figures are adapted from \cite{Quintana-Duque2012, casdagli1991, uzal2011}.
-	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
-	(A.2) chaotic and (B.2) random time series.
-	(C.2, D.2) $E_1 (m)$ values and (E.2, F.2) $E_2(m)$ values with variations of $\tau$ values from one to twenty for chaotic and random time series.
-	(G.2, H.2) AMI values where its first minimum value in the curve is the minimum time delay embedding ($\tau_0$).
-	{\bf Recurrence Plots and its patterns.} 
-	(A.3) State space of the Lorenz system with controlling parameters ($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood (black circle) of a given point at $i$ is a recurrent point and is represented as a black dot in the recurrence plot at location $(i, j)$ or white otherwise.
-	(B.3) Recurrence plot using the three components of the Lorenz system and the RP with no embeddings and threshold $\epsilon=5$.
+	a1 to a4 figures are adapted from \cite{Quintana-Duque2012, casdagli1991, uzal2011}.
+	(b1) chaotic and (b5) random time series.
+	(b2, b6) $E_1 (m)$ values and (b3, b7) $E_2(m)$ values with variations of $\tau$ values from one to twenty for chaotic and random time series.
+	(b4, b8) AMI values where its first minimum value in the curve is the minimum time delay embedding ($\tau_0$).
+	(c1) State space of the Lorenz system with controlling parameters ($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood (black circle) of a given point at $i$ is a recurrent point and is represented as a black dot in the recurrence plot at location $(i, j)$ or white otherwise.
+	(c2) Recurrence plot using the three components of the Lorenz system and the RP with no embeddings and threshold $\epsilon=5$.
 	Time-series with its respective recurrence plot for:
-	(C.3) uniformly distributed noise,
-	(D.3) super-positionet harmonic oscillation ($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(E.3) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted with a linearly increase term ($0.01*i$), and
-	(F.3) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	Figures A.3 to F.3 are adapted from \cite{marwan2015}.
+	(c3) uniformly distributed noise,
+	(c4) super-positionet harmonic oscillation ($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(c5) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted with a linearly increase term ($0.01*i$), and
+	(c6) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	Figures in C are adapted from \cite{marwan2015}.
 	Code and data to reproduce the figure is available in \cite{srep2021}.
@@
 series: the solution for the $x$ variable of the Lorenz system 
-(Fig~\ref{fig01}A.2), and a Gaussian noise time series with zero mean
-and a variance of one (Fig~\ref{fig01}B.2).
+(Fig~\ref{fig01}b1), and a Gaussian noise time series with zero mean
+and a variance of one (Fig~\ref{fig01}b5).
 We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
@@
 The determination of six is given that any value of $m$ can be used as they 
-are within the threshold of $1\pm0.05$ (Fig~\ref{fig01}C.2).
+are within the threshold of $1\pm0.05$ (Fig~\ref{fig01}b2).
 $E_2(m)$ values, for chaotic time series, are different to one 
-(Fig~\ref{fig01}E.2), for which, it can be concluded that for the
+(Fig~\ref{fig01}b3), for which, it can be concluded that for the
 chaotic time series the minimum embedding dimension the time series 
@@
 $E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig01}D.2).
+which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig01}b6).
 $E_1(m)$ values then indicate the minimum embedding dimension of the 
 noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig01}F.2) for which it can be
+approximately equal to one (Figure~\ref{fig01}b7) for which it can be
 concluded that noise time series is a stochastic signal.
@@
 regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-(Fig~\ref{fig01}C.2,D.2,E.2 and F.2).
+(Fig~\ref{fig01}b2,b3,b6, and b7).
 However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
@@
 B) a noise time series using a normal distribution with mean zero and 
-standard deviation equal to one. From Fig~\ref{fig01}(G.2, H.2), it can then be
+standard deviation equal to one. From Fig~\ref{fig01}(b4, b8), it can then be
 concluded that the amount of knowledge for any noise time series is zero 
@@
 function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig01}(A.3,B.3)) \cite{eckmann1987, marwan2007,marwan2015}.
+(Fig~\ref{fig01}(c1,c2)) \cite{eckmann1987, marwan2007,marwan2015}.
 RP is also characterised with a line of identity (LOI) which is a  
@@
 (i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-uniformly distributed noise (Fig~\ref{fig01}C.3),
+uniformly distributed noise (Fig~\ref{fig01}c3),
 (ii) periodic and quasi-periodic systems where diagonal lines and 
 checkerboard structures represent oscillating systems, e.g., sinusoidal 
-signals (Fig~\ref{fig01}D.3),
+signals (Fig~\ref{fig01}c4),
 (iii) drift where paling or darkening recurrence points away from 
 the LOI is caused by drifting systems, 
-e.g., logistic map (Fig~\ref{fig01}E.3), and
+e.g., logistic map (Fig~\ref{fig01}c5), and
 (iv) disrupted where recurrence points are presented white areas or 
 bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-(Fig~\ref{fig01}F.3) \cite{eckmann1987, marwan2015}.
+(Fig~\ref{fig01}c7) \cite{eckmann1987, marwan2015}.
 Texture patterns in RPs can be categorised as:

### 2021-03-19 docs/manuscript/main.tex
figure02 harmonised and updated

@@
 to the right hand of the participant and to the left hand of the robot 
-(Figure~\ref{fig02}A.1,C.1). Data were then collected with four NeMEMSi
+(Figure~\ref{fig02}a1,a3). Data were then collected with four NeMEMSi
 IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
@@
   \centering
-\includegraphics[width=1.0\textwidth]{figures/fig02/versions/drawing-v00.png}
+\includegraphics[width=1.0\textwidth]{figures/fig02/versions/drawing-v01.png}
     \caption{
-	{\bf Human-humanoid imitation activities.} 
-		Face-to-face human-humanoid imitation (HHI) activities for (A.1) HHI of horizontal arm movement, (B.1) Humanoid horizontal arm movement, (C.1) HHI of vertical arm movement, and (D.1) Humanoid vertical arm movement.
-	{\bf Time series duration for arm movements.}
-    	Time series of smoothed data from gyroscope sensor for different speed arm movements performed by NAO: (A.2) Horizontal Normal arm movement (HN), (B.2) Horizontal Faster arm movement (HF), (C.2) Vertical Normal arm movement (VN) and (D.2) Vertical Faster arm movement (VF).
-		(A.2) shows window sizes for 2-seconds (100 samples), 5-seconds (250 samples), 10-seconds (500 samples) and 15-seconds (750 samples) which are also presented in (B.2), (C.2) and (D.2).
-	{\bf Time series for horizontal and vertical arm movements.}
-		(A.3/D.3) raw-normalised (sg0zmuv), (B.3/E.3) normalised-smoothed 1 (sg1zmuv) and (C.3/F.3) normalised-smoothed 2 (sg2zmuv).
+	{\bf Human-humanoid imitation activities (A). Time series duration for arm movements (B).
+		Time series for horizontal and vertical arm movements (C).
+	}
+	Face-to-face human-humanoid imitation (HHI) activities for (a1) HHI of horizontal arm movement, (a2) Humanoid horizontal arm movement, (a3) HHI of vertical arm movement, and (a4) Humanoid vertical arm movement.
+    Time series of smoothed data from gyroscope sensor for different speed arm movements performed by NAO: (b1) Horizontal Normal arm movement (HN), (b2) Horizontal Faster arm movement (HF), (b3) Vertical Normal arm movement (VN) and (b4) Vertical Faster arm movement (VF).
+	(B) shows window sizes for 2-seconds (100 samples), 5-seconds (250 samples), 10-seconds (500 samples) and 15-seconds (750 samples).
+	(c1,c4) raw-normalised (sg0zmuv), (c2,c5) normalised-smoothed 1 (sg1zmuv) and (c3,c7) normalised-smoothed 2 (sg2zmuv).
 	Time series are only for three participants	($p01$, $p02$, and $p03$) for horizontal and vertical arm movements in normal and faster velocity (HN, HF, VN, VF) with the normalised GyroZ or GyroY axis and with one sensor attached to the participant (HS01) and other sensor attached to the robot (RS01).
@@
 (i) ten repetitions of horizontal arm movement at normal (HN) 
-	and faster (HF) speed (Figure~\ref{fig02}A.1), and
+	and faster (HF) speed (Figure~\ref{fig02}a1), and
 (ii) ten repetitions of vertical arm movement at normal (VN) 
-	and faster (VF) speed (Figure~\ref{fig02}C.1).
+	and faster (VF) speed (Figure~\ref{fig02}a3).
 The normal and faster speed of arm movements is defined by the duration 
@@
 To visualise the distinction between normal and faster speed for horizontal 
-and vertical arm movements, Fig~\ref{fig02}A.2-D.2 shows smoothed time series
+and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series
 for axes Z and Y of the gyroscope sensors with four window lengths: 
@@
 three participants (p01, p01 and p03) performing horizontal 
-arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}A.3-F.3).
+arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}C).
 %%%---------------------------------(FIGURE)-------------------------------------
@@
 5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
-(Figures~\ref{fig02}A.2-D.2).
+(Figures~\ref{fig02}B).
 
@@
 	{\bf RSSs for horizontal and vertical arm movements.}
-	Reconstructed state spaces for time series for p01 of Figure \ref{fig02}A.3-F.3.
+	Reconstructed state spaces for time series for p01 of Figure \ref{fig02}C.
 	Reconstructed state spaces were computed with 

### 2021-03-19 docs/manuscript/main.tex
figure04 harmonised and updated

@@
 the first three axis of the rotated Principal Components Analysis (PCA) 
-are shown for the reconstructed state spaces of horizontal (figures~\ref{fig:rsss}(A)) 
-and vertical (figures~\ref{fig:rsss}(B)) arm movements. 
+are shown for the reconstructed state spaces of horizontal (figures~\ref{fig04}(A))
+and vertical (figures~\ref{fig04}(B)) arm movements.
 While visual inspection of these figures 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/rss/versions/drawing-v00.pdf}
+\includegraphics[width=1.0\textwidth]{figures/rss/versions/drawing-v01}
 \caption{
@@
         }
-    \label{fig:rsss}
+    \label{fig04}
 \end{figure}

### 2021-03-19 docs/manuscript/main.tex
figure05 harmonised and updated

@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/fig05/versions/drawing-v00}
+\includegraphics[width=1.0\textwidth]{figures/fig05/versions/drawing-v01}
 \caption{
@@
 	{\bf Box plots for RQA metrics (C).}
-	RQA metrics for (C.1) REC, (C.2) DET, (C.3) RATIO, and (C.4) ENTR of
+	RQA metrics for (c1) REC, (c2) DET, (c3) RATIO, and (c4) ENTR of
 	20 participants performing HN, HF, VN and VF movements
@@
 %\subsubsection*{REC values}
-Figure~\ref{fig05}(C.1) presents box plots of REC values,
+Figure~\ref{fig05}(c1) presents box plots of REC values,
 for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements 
@@
 %\subsubsection*{DET values}
-Figure \ref{fig05}(C.2) presents DET values and shows
+Figure \ref{fig05}(c2) presents DET values and shows
 little difference for type of movement or performer.  
@@
 %\subsubsection*{RATIO values}
-Figure \ref{fig05}(C.3) presents the ratio of DET / REC.
+Figure \ref{fig05}(c3) presents the ratio of DET / REC.
 These values, for the human performer, vary less for HN than for HF movement.  
@@
 %\subsubsection*{ENTR values}
-Figure \ref{fig05}(C.4) shows ENTR values are higher for the human
+Figure \ref{fig05}(c4) shows ENTR values are higher for the human
 performer than the robot, and vary with the smoothness of the time-series.

### 2021-03-19 docs/manuscript/main.tex
#4 cleans main.tex

@@
 \begin{abstract}
-Human movement variability arises from the process of 
-mastering redundant (bio)mechanical degrees of freedom
-to successfully accomplish any given motor task
-where flexibility and stability of many possible joint 
-combinations helps to adapt to environment conditions. 
-While the analysis of movement of variability is becoming increasingly 
-popular as a diagnostic tool or skill performance evaluation, 
-there are remain challenges on applying the most appropriate methods. 
-We therefore investigate nonlinear methods 
-such as reconstructed state space (RSSs), uniform time-delay embedding, 
-recurrence plots (RPs) and recurrence quantification analysis (RQAs)
-with real-world time-series data of wearable inertial sensors. 
-That said, twenty healthy participants 
-imitated vertical and horizontal arm movements in normal 
-and faster velocity from an humanoid robot.
-We applied nonlinear methods to the collected data 
-to found visual differences in the patterns of RSSs and RPs
-and statistical differences with RQAs.
-We conclude that Shannon Entropy with RQA is a robust method 
-that helps to quantify activities, types of sensors, windows lengths 
-and level of smoothness.
-Hence this work might enhance the development of 
-better diagnostic tools for applications in 
-rehabilitation and sport science for skill performance
-or new forms of human-humanoid interaction for
-quantification of movement adaptations and motor pathologies.
+Human movement variability arises from the process of mastering redundant (bio)mechanical degrees of freedom to successfully accomplish any given motor task where flexibility and stability of many possible joint combinations helps to adapt to environment conditions.
+While the analysis of movement of variability is becoming increasingly popular as a diagnostic tool or skill performance evaluation, there are remain challenges on applying the most appropriate methods.
+We therefore investigate nonlinear methods such as reconstructed state space (RSSs), uniform time-delay embedding, recurrence plots (RPs) and recurrence quantification analysis (RQAs) with real-world time-series data of wearable inertial sensors.
+That said, twenty healthy participants imitated vertical and horizontal arm movements in normal and faster velocity from an humanoid robot.
+We applied nonlinear methods to the collected data to found visual differences in the patterns of RSSs and RPs and statistical differences with RQAs.
+We conclude that Shannon Entropy with RQA is a robust method that helps to quantify activities, types of sensors, windows lengths and level of smoothness.
+Hence this work might enhance the development of better diagnostic tools for applications in rehabilitation and sport science for skill performance or new forms of human-humanoid interaction for quantification of movement adaptations and motor pathologies.
 \end{abstract}
@@
 \section*{Introduction}
-The complexity of human movement arises from the process of mastering 
-redundant (bio)mechanical degrees of freedom (DOF) to successfully 
-accomplish any given motor task (Bernstein, 1967).
-Such DOF are independent coordinates to uniquely 
-describe the configurations of the system which provides flexibility 
-and stability to adapt to a change environment conditions but 
-that leads many possible combinations (Newell and Vaillancourt, 2001).
-Hence, human movement complexity can be seen as a balance 
-between the required DOF "to generate a stable (persistent) 
-and flexible (variable) behavioural output in response to 
-changing intentions and dynamic environmental
-conditions" \cite{davids2003}. 
-Consequently, one can see much variability in human movement even in the 
-simplest of movements.  
-While the analysis of movement of variability is becoming increasingly 
-popular as a diagnostic tool or skill performance evaluation, 
-there remain challenges in terms of defining 
-the most appropriate methods and parameters to apply. 
-In part, 
-these challenges stem from the fact that the identification movement 
-variability requires analysis of signals which are time-series of 
-of $1-$dimension in $\mathbb{R}$ which are 
-noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
-Further problems arise from assigning a plausible locus of control to 
-the movements, e.g., in order to determine whether variability is the 
-result of deliberative control by the performer or whether it arises 
-from exogenous or endogenous disturbances.  For this paper, our focus 
-is on the analysis of signals; specifically, in terms of objectively 
-quantifying variability of lower dimension signals using time-series analysis.
-
-Methods for nonlinear time series analysis generally involve the estimation of the 
-embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
-reconstruct the state space,
-where an $n$-dimensional is reconstructed state space using $1-$dimensional 
-time series \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
-frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
-Key to the selection of these parameters is the need to preserve 
-topological properties of an unknown $M$-dimensional state space \cite{takens1981}.
-An common approach to the construction of these state spaces 
-involves Recurrence Plots (RPs), a graphical representation of black and white dots, 
-which shows recurrent patterns of a $n$-dimensional system.
-While RPs provide a human
-interpretable picture of the system, these require further
-analysis to allow the properties of that system to be quantified,
-and so Recurrence Quantification Analysis (RQA) can be applied.
-However, the estimation of the embedding parameters for
-RQA remains an open problem (Bradley et al. 2015 \cite{bradley2015})
-There is no agreed method to estimate embedding parameters
-for RQA and other nonlinear methods \cite{bradley2015} 
-because time series are system-dependent, i.e., these rely
-on the initial conditions, and on the configuration and behaviour of the system.  
-This means that, unless one holds all of the influencing 
-variables constant, embedding parameters computed for one 
-instance may not apply to another instance.
-
-One could apply methods, such as autocorrelation, mutual information, 
-nearest neighbour (and we will apply these in this paper), 
-but these methods assume that the data are well sampled, with 
-little noise and (usually) that the signals are purely deterministic \cite{garland2016, kantz2003}.
-That said, these methods can break-down in the face of real-world datasets 
-which could have 
-different length, 
-different values of accuracy and precision (rounding errors due to finite 
-precision of the measurement apparatus which include frequency 
-acquisition \cite{frank2010}), or 
-different levels of contamination from exogenous and endogenous 
-sources of ‘noise’ \cite{garland2016}.  What is, perhaps, surprising is that 
-even subject to these problems, the methods are useful and 
-nonlinear dynamics approaches continue to tell us much about 
-human movement \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010,
-gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.
-
-For this paper, we explore the role of RPs and RQA in the 
-analysis of simple human movement. 
-We compare horizontal and 
-vertical arm movements performed by neurotypical participants 
-who are copying these movements being made by a humanoid robot.  
-This provides an initial point of comparison 
-(between human movement, which we assume to have a degree of variability, 
-and the robot, which we assume to have a degree of consistency).   
-Additionally, our interest lies in the estimation of embedding parameters 
-in light of different features of the data, e.g., levels of smoothing, 
-window length, structure of time-series based on movement, 
-types of sensors, individual differences between participants and 
-the movements that they perform. 
-Specifically, we ask what are 
-the effects of different embedding parameters, recurrence thresholds 
-and characteristics of time-series on nonlinear analysis 
-methods (i.e., reconstructed state space with 
-uniform time-delay embedding, RPs and RQA)?
+The complexity of human movement arises from the process of mastering redundant (bio)mechanical degrees of freedom (DOF) to successfully accomplish any given motor task (Bernstein, 1967).
+Such DOF are independent coordinates to uniquely describe the configurations of the system which provides flexibility and stability to adapt to a change environment conditions but that leads many possible combinations (Newell and Vaillancourt, 2001).
+Hence, human movement complexity can be seen as a balance between the required DOF "to generate a stable (persistent) and flexible (variable) behavioural output in response to changing intentions and dynamic environmental conditions" \cite{davids2003}.
+Consequently, one can see much variability in human movement even in the simplest of movements.
+While the analysis of movement of variability is becoming increasingly popular as a diagnostic tool or skill performance evaluation, there remain challenges in terms of defining the most appropriate methods and parameters to apply.
+In part, these challenges stem from the fact that the identification movement variability requires analysis of signals which are time-series of $1-$dimension in $\mathbb{R}$ which are noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
+Further problems arise from assigning a plausible locus of control to the movements, e.g., in order to determine whether variability is the result of deliberative control by the performer or whether it arises from exogenous or endogenous disturbances.
+For this paper, our focus is on the analysis of signals; specifically, in terms of objectively quantifying variability of lower dimension signals using time-series analysis.
+
+Methods for nonlinear time series analysis generally involve the estimation of the embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to reconstruct the state space, where an $n$-dimensional is reconstructed state space using $1-$dimensional time series \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
+Key to the selection of these parameters is the need to preserve topological properties of an unknown $M$-dimensional state space \cite{takens1981}.
+An common approach to the construction of these state spaces involves Recurrence Plots (RPs), a graphical representation of black and white dots, which shows recurrent patterns of a $n$-dimensional system.
+While RPs provide a human interpretable picture of the system, these require further analysis to allow the properties of that system to be quantified, and so Recurrence Quantification Analysis (RQA) can be applied.
+However, the estimation of the embedding parameters for RQA remains an open problem (Bradley et al. 2015 \cite{bradley2015}).
+There is no agreed method to estimate embedding parameters for RQA and other nonlinear methods \cite{bradley2015} because time series are system-dependent, i.e., these rely on the initial conditions, and on the configuration and behaviour of the system.
+This means that, unless one holds all of the influencing variables constant, embedding parameters computed for one instance may not apply to another instance.
+
+One could apply methods, such as autocorrelation, mutual information, nearest neighbour (and we will apply these in this paper), but these methods assume that the data are well sampled, with little noise and (usually) that the signals are purely deterministic \cite{garland2016, kantz2003}.
+That said, these methods can break-down in the face of real-world datasets which could have different length, different values of accuracy and precision (rounding errors due to finite precision of the measurement apparatus which include frequency acquisition \cite{frank2010}), or different levels of contamination from exogenous and endogenous sources of ‘noise’ \cite{garland2016}.
+What is, perhaps, surprising is that even subject to these problems, the methods are useful and nonlinear dynamics approaches continue to tell us much about human movement \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010, gomezgarcia2014, marwan2011, stergiou2011, bradley2015}.
+
+For this paper, we explore the role of RPs and RQA in the analysis of simple human movement.
+We compare horizontal and vertical arm movements performed by neurotypical participants who are copying these movements being made by a humanoid robot.
+This provides an initial point of comparison (between human movement, which we assume to have a degree of variability, and the robot, which we assume to have a degree of consistency).
+Additionally, our interest lies in the estimation of embedding parameters in light of different features of the data, e.g., levels of smoothing, window length, structure of time-series based on movement, types of sensors, individual differences between participants and the movements that they perform.
+Specifically, we ask what are the effects of different embedding parameters, recurrence thresholds and characteristics of time-series on nonlinear analysis methods (i.e., reconstructed state space with uniform time-delay embedding, RPs and RQA)?
 
@@
 \subsection*{State Space Reconstruction}
-The method of state space reconstruction \cite{packard1980} 
-\cite{takens1981} has been applied in many disciplines 
-\cite{aguirre2009, stergiou2011, frank2010, 
-sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
-based on uniform time-delay embedding methodology which is a simple 
-matrix implementation that can reconstruct an unknown $d-$dimensional 
-manifold $M$ from a scalar time series (e.g. one-dimensional 
-time series in $\mathbb{R}$). A manifold, in this context, is a 
-multidimensional curved surface within a space (e.g. a saddle) 
-\cite{guastello-gregson2011}.
-
-The use of a scalar time series is the main advantage of the method of state 
-space reconstruction which in essence preserve dynamic invariants such as 
-correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
-entropy and detrended fluctuation analysis \cite{bradley2015, 
-Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
-However, selecting appropriate embedding parameters which are sued to apply the 
-state space reconstruction is still an open challenge for which we present 
-introductions for the methodologies to compute such embedding parameters 
-With that in mind, in the following subsections, we describe in more detail 
-the state space reconstruction theorem (RSSs), uniform time-delay embedding 
-theorem (UTDE), false nearest neighbours (FNN) and average mutual 
-information (AMI). 
-%and other methodologies for state space reconstruction.
+The method of state space reconstruction \cite{packard1980, takens1981} has been applied in many disciplines \cite{aguirre2009, stergiou2011, frank2010, sama2013, Quintana-Duque2016}.
+The method of state space reconstruction is based on uniform time-delay embedding methodology which is a simple matrix implementation that can reconstruct an unknown $d-$dimensional manifold $M$ from a scalar time series (e.g.
+one-dimensional time series in $\mathbb{R}$).
+A manifold, in this context, is a multidimensional curved surface within a space (e.g. a saddle) \cite{guastello-gregson2011}.
+
+The use of a scalar time series is the main advantage of the method of state space reconstruction which in essence preserve dynamic invariants such as correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai entropy and detrended fluctuation analysis \cite{bradley2015, Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
+However, selecting appropriate embedding parameters which are sued to apply the state space reconstruction is still an open challenge for which we present introductions for the methodologies to compute such embedding parameters.
+With that in mind, in the following subsections, we describe in more detail the state space reconstruction theorem (RSSs), uniform time-delay embedding theorem (UTDE), false nearest neighbours (FNN) and average mutual information (AMI).
 
 \subsection*{State Space Reconstruction Theorem}
-Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
-uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
+Following the notation employed in \cite{casdagli1991, garland2016, gibson1992, uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:
 %%********************************[EQUATION]************************************
@@
 %%********************************[EQUATION]************************************
-where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
-$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
-unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
-function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
-iteration of $f$ that corresponds to an initial position 
-$s(0) \in M $ \cite{takens1981}.
-%$f$ is a smooth dynamical system, where smooth means a 
-%continuously differentiable system
-%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
-Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
-can be obtained with
+where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and $M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th iteration of $f$ that corresponds to an initial position $s(0) \in M $ \cite{takens1981}.
+%$f$ is a smooth dynamical system, where smooth means a continuously differentiable system (e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
+Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, can be obtained with
 %%********************************[EQUATION]************************************
@@
 where $h$ is a function, $h: M \rightarrow \mathbb{R}$, defined on the trajectory $s(t)$.
-Reconstructed state space can then be described as an $n-$dimensional state 
-space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
-$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
-is the uniform time-delay embedding with a dimension embedding $m$
-and delay embedding $\tau$ and
-$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
-of dimensionality (e.g. Principal Component Analysis, 
+Reconstructed state space can then be described as an $n-$dimensional state space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where $\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ is the uniform time-delay embedding with a dimension embedding $m$ and delay embedding $\tau$ and $ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation of dimensionality (e.g.
+Principal Component Analysis,
 Singular Value Decomposition, etc) being $n \leq m$.
-With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
-defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
-$\boldsymbol{X}(t) = \Phi(s(t))$,
-where $\Phi$ is a diffeomorphic map \cite{takens1981}
-whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
-dimension of $M$ \cite{garland2016}.
-% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
-% is called diffeomorphic if it is one-to-one correspondence and its inverse
-% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
-Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
-space then a composition of functions represented with $F^t$ is induced on the
-reconstructed state space determined:
+With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that $\boldsymbol{X}(t) = \Phi(s(t))$, where $\Phi$ is a diffeomorphic map \cite{takens1981} whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting dimension of $M$ \cite{garland2016}.
+%"Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$ is called diffeomorphic if it is one-to-one correspondence and its inverse $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
+Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed space then a composition of functions represented with $F^t$ is induced on the reconstructed state space determined:
 %by true smooth dynamical system $f^t$ and $\Phi$:
@@
 %%********************************[EQUATION]************************************
-With this in mind, an embedding is defined as "a smooth one-to-one 
-coordinate transformation with a smooth inverse" and the total reconstruction 
-map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
+With this in mind, an embedding is defined as "a smooth one-to-one coordinate transformation with a smooth inverse" and the total reconstruction map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.
 Figures~\ref{fig01}A illustrates the state space reconstruction.
@@
 	{\bf State space reconstruction methodology (A),
-		Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method (B), and
-		Recurrence Plots and its patterns (C).
+		Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method (B), and Recurrence Plots and its patterns (C).
 	}
@@
 \subsection*{Uniform Time-Delay Embedding (UTDE)}\label{sec:utimedelayembedding}
-Frank et al. and Sama et al. refer to the state space reconstruction
-as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}. 
-However, we consider the term "uniform time-delay embedding"
-as more descriptive and appropriate terminology for our work.
-Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed 
-copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is 
-the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the 
-samples of $\{ \boldsymbol{x}_n \}$.
+Frank et al. and Sama et al. refer to the state space reconstruction as "time-delay embeddings" or "delay coordinates" \cite{frank2010, sama2013}.
+However, we consider the term "uniform time-delay embedding" as more descriptive and appropriate terminology for our work.
+Hence, the uniform time-delay embedding is represented as a matrix of uniform delayed copies of the time series $\{ \boldsymbol{x}_n \}_{n=1}^N$ where $N$ is the sample length of $\{ \boldsymbol{x}_n \}$ and $n$ is index for the samples of $\{ \boldsymbol{x}_n \}$.
 $\{ \boldsymbol{x}_n \}_{n=1}^N$ has a sample rate of $T$.
-The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$
-and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$
-where $i$ goes from $0,1, \dots, (m-1)$.
-Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains
-information of unobserved state variables and encapsulates the information of
-the delayed copies of the available time series in the uniform time-delay 
-embedding matrix $\boldsymbol{X}^{m}_{\tau}$, 
-$\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
+The delayed copies of $\{ \boldsymbol{x}_n \}$ are uniformly separated by $\tau$ and represented as $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ where $i$ goes from $0,1, \dots, (m-1)$.
+Generally speaking, $\{\boldsymbol{ \tilde{x} }_{n- i\tau} \}$ contains information of unobserved state variables and encapsulates the information of the delayed copies of the available time series in the uniform time-delay embedding matrix $\boldsymbol{X}^{m}_{\tau}$, $\boldsymbol{X}^{m}_{\tau} \in \mathbb{R}^m$, defined as
 %%********************************[EQUATION]************************************
@@
 %%********************************[EQUATION]************************************
-where $m$ is the embedding dimension, $\tau$ is the embedding delay and
-$ ^\intercal$ denotes the transpose.
+where $m$ is the embedding dimension, $\tau$ is the embedding delay and $ ^\intercal$ denotes the transpose.
 $m$ and $\tau$ are known as embedding parameters.
-The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by
-$N-(m-1)\tau$ rows and $m$ columns and 
-$N-(m-1)\tau$ defines the length of each delayed copy 
-of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
-%For further details and explicit examples of uniform time-delay 
-%embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
+The matrix dimension of $ \boldsymbol{X}_{\tau}^{m} $ is defined by $N-(m-1)\tau$ rows and $m$ columns and $N-(m-1)\tau$ defines the length of each delayed copy of $\{ \boldsymbol{ \tilde{x} }_n \}$ in $\boldsymbol{X}^{m}_{\tau}$.
+%For further details and explicit examples of uniform time-delay %embedding methodology, we refer the reader to the \nameref{S1_AppendixA}.
 
 \subsection*{Estimation of Embedding Parameters}
-The estimation of the embedding parameters ($m$ and $\tau$) 
-is a fundamental step for the state space reconstruction with the use
-of uniform time-delay embedding method.
-With this in mind, we review two of the most common algorithms,
-which will be used in our work, to compute the embedding
-parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
+The estimation of the embedding parameters ($m$ and $\tau$) is a fundamental step for the state space reconstruction with the use of uniform time-delay embedding method.
+With this in mind, we review two of the most common algorithms, which will be used in our work, to compute the embedding parameters: the false nearest neighbour (FNN) and the average mutual information (AMI).
 
@@
 \subsubsection*{False Nearest Neighbours}
-To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992}
-used the method of false neighbours which can be understood as follows:
-on the one hand, when the embedding dimension is too small to unfold the 
-attractor "not all points that lie close each other will be neighbours and 
-some points appear as neighbours because of the attractor has been projected 
-down into an smaller space", on the other hand, when increasing the embedding 
-dimension "points that are near to each other in the sufficient
-embedding dimension should remain close as the dimension increase from $m$ 
-to $m+1$ \cite{krakovska2015}".
-From a mathematical point of view, the state space reconstruction theorem is 
-done when the attractor is unfolded with either the minimum embedding 
-dimension, $m_0$, or any other embedding dimension value where 
-$m \ge m_0$ \cite{kennel1992}. On the contrary, any large value of $m_0$ 
-leads to excessive computations \cite{bradley2015}. With this in mind, 
-Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method 
-where only the time-series and one delay embedding value are necessary 
-to select the minimum embedding dimension. Cao's algorithm is based 
-on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
+To select the minimum embedding dimension $m_0$, Kennel et al. \cite{kennel1992} used the method of false neighbours which can be understood as follows:
+on the one hand, when the embedding dimension is too small to unfold the attractor "not all points that lie close each other will be neighbours and some points appear as neighbours because of the attractor has been projected down into an smaller space", on the other hand, when increasing the embedding dimension "points that are near to each other in the sufficient embedding dimension should remain close as the dimension increase from $m$ to $m+1$ \cite{krakovska2015}".
+From a mathematical point of view, the state space reconstruction theorem is done when the attractor is unfolded with either the minimum embedding dimension, $m_0$, or any other embedding dimension value where $m \ge m_0$ \cite{kennel1992}.
+On the contrary, any large value of $m_0$ leads to excessive computations \cite{bradley2015}.
+With this in mind, Cao \cite{Cao1997} proposed an algorithm based on the false neighbour method where only the time-series and one delay embedding value are necessary to select the minimum embedding dimension.
+Cao's algorithm is based on $E(m)$  which is the mean value of all $a(i,m)$, both defined as follows
 %%********************************[EQUATION]************************************
@@
 %%********************************[EQUATION]************************************
-where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay
-embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
-From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ 
-and $\tau$ for which $E_1(m)$ is defined as
+where $\boldsymbol{X}_i(m)$ and $\boldsymbol{X}_{n(i,m)}(m)$ are the time-delay embeddings with $i=1,2,\dots,N-(m-1)\tau$ and $ n(i,m)= 1 \le n(i,m) \le N-m\tau$.
+From Eq.~\ref{eq:e}, it can be seen that $E(m)$ is only dependent on $m$ and $\tau$ for which $E_1(m)$ is defined as
 %%********************************[EQUATION]************************************
@@
 %%********************************[EQUATION]************************************
-$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$
-in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
-As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater
-than some $m_0$, if the time series comes from a multidimensional state space
-then $m_0 + 1$ is the minimum dimension".
-Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from
-stochastic signals. $E_2(m)$ is defined as
+$E_1(m)$ is therefore considered to investigate the variation from $m$ to $m+1$ in order to find the minimum embedding dimension $m_0$ (Eq.~\ref{eq:e1}).
+As described in \cite{Cao1997}: "$E_1(m)$ stops changing when $m$ is greater than some $m_0$, if the time series comes from a multidimensional state space then $m_0 + 1$ is the minimum dimension".
+Additionally, Cao proposed $E_2(m)$ to distinguish deterministic signals from stochastic signals. $E_2(m)$ is defined as
 %%********************************[EQUATION]************************************
@@
 %%********************************[EQUATION]************************************
-For instance, when the signal comes from random noise (values that are 
-independent from each other), all $E_2(m)$ values are approximately equal 
-to 1 (e.g. $E_2(m) \approx 1$). However, for deterministic data $E_2(m)$ is 
-not constant for all $m$ (e.g. $E_2(m) \neq 1$).
-
-As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
-series: the solution for the $x$ variable of the Lorenz system 
-(Fig~\ref{fig01}b1), and a Gaussian noise time series with zero mean
-and a variance of one (Fig~\ref{fig01}b5).
+For instance, when the signal comes from random noise (values that are independent from each other), all $E_2(m)$ values are approximately equal to 1 (e.g.
+$E_2(m) \approx 1$).
+However, for deterministic data $E_2(m)$ is not constant for all $m$ (e.g.
+$E_2(m) \neq 1$).
+
+As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time series: the solution for the $x$ variable of the Lorenz system (Fig~\ref{fig01}b1), and a Gaussian noise time series with zero mean and a variance of one (Fig~\ref{fig01}b5).
 We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
-The $E_1(m)$ values for the chaotic time series appear to be constant
-after the dimension is equal to six.
-The determination of six is given that any value of $m$ can be used as they 
-are within the threshold of $1\pm0.05$ (Fig~\ref{fig01}b2).
-$E_2(m)$ values, for chaotic time series, are different to one 
-(Fig~\ref{fig01}b3), for which, it can be concluded that for the
-chaotic time series the minimum embedding dimension the time series 
-comes from a deterministic signal. With regard to the noise time series,  
-$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig01}b6).
-$E_1(m)$ values then indicate the minimum embedding dimension of the 
-noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig01}b7) for which it can be
-concluded that noise time series is a stochastic signal.
-It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
-computed but also a variation of $\tau$ from 1 to 20 is presented. 
-The purpose of such variation for $\tau$ is to show its independence with
-regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing 
-(Fig~\ref{fig01}b2,b3,b6, and b7).
-However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of 
-a new threshold where $m$ values appear to be constant in $E_1 (m)$.
-In the case of the given examples and reported results, we defined such 
-threshold as 0.05. Further investigation is required for the selection of the 
-threshold in the $E_1(m)$, as the selection of the threshold in this work is 
-base on no particular method but visual inspection.
+The $E_1(m)$ values for the chaotic time series appear to be constant after the dimension is equal to six.
+The determination of six is given that any value of $m$ can be used as they are within the threshold of $1\pm0.05$ (Fig~\ref{fig01}b2).
+$E_2(m)$ values, for chaotic time series, are different to one (Fig~\ref{fig01}b3), for which, it can be concluded that for the chaotic time series the minimum embedding dimension the time series comes from a deterministic signal.
+With regard to the noise time series, $E_1(m)$ values appeared to be constant when $m$ is close to thirteen, which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig01}b6).
+$E_1(m)$ values then indicate the minimum embedding dimension of the noisy time series is thirteen, however all of the $E_2(m)$ values are approximately equal to one (Figure~\ref{fig01}b7) for which it can be concluded that noise time series is a stochastic signal.
+It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are computed but also a variation of $\tau$ from 1 to 20 is presented.
+The purpose of such variation for $\tau$ is to show its independence with regard to $E_1(m)$ and $E_2(m)$ values as $\tau$ is increasing (Fig~\ref{fig01}b2,b3,b6, and b7).
+However, one negative of the Cao's algorithm \cite{Cao1997} is the definition of a new threshold where $m$ values appear to be constant in $E_1 (m)$.
+In the case of the given examples and reported results, we defined such threshold as 0.05.
+Further investigation is required for the selection of the threshold in the $E_1(m)$, as the selection of the threshold in this work is base on no particular method but visual inspection.
 
@@
 \subsubsection*{Average Mutual Information}
-When selecting the delay dimension parameter, $\tau$, one can consider the 
-following two cases:
-(i) when $\tau$ is too small, the elements of time-delay embedding will be along
-the bisectrix of the phase space and the reconstruction is generally not 
-satisfactory, 
-(ii) on the contraty, when $\tau$ is too large the elements of the uniform 
-time-delay embedding will become spread and uncorrelated which makes 
-recovering the underlying attractor more difficult if not 
-impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
-With regard to the algorithms to compute $\tau$, 
-Emrani et al. \cite{emrani2014a}, for instance, used the autocorrelation 
-function in which the first zero crossing is considered as the minimum delay 
-embedding parameter. However, the autocorrelation function is a linear 
-statistic over which the Average Mutual Information (AMI) algorithm is 
-preferred because the AMI takes into account the nonlinear dynamical 
-correlations \cite{afraser1986,krakovska2015}. With this in mind, the AMI 
-algorithm is described below to estimate the minimum delay embedding 
-parameter, \texorpdfstring{$\tau_0$}{T}.
-
-To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated
-and then a probability distribution of data is computed \cite{kantz2003}.
-AMI is therefore denoted by $I(\tau)$ which is the average mutual 
-information between the original time series, $x(n)$, and the delayed 
-time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}. 
+When selecting the delay dimension parameter, $\tau$, one can consider the following two cases:
+(i) when $\tau$ is too small, the elements of time-delay embedding will be along the bisectrix of the phase space and the reconstruction is generally not satisfactory,
+(ii) on the contraty, when $\tau$ is too large the elements of the uniform time-delay embedding will become spread and uncorrelated which makes recovering the underlying attractor more difficult if not impossible \cite{casdagli1991, emrani2014a, garcia2005e71}.
+With regard to the algorithms to compute $\tau$, Emrani et al.
+\cite{emrani2014a}, for instance, used the autocorrelation function in which the first zero crossing is considered as the minimum delay embedding parameter.
+However, the autocorrelation function is a linear statistic over which the Average Mutual Information (AMI) algorithm is preferred because the AMI takes into account the nonlinear dynamical correlations \cite{afraser1986,krakovska2015}.
+With this in mind, the AMI algorithm is described below to estimate the minimum delay embedding parameter, \texorpdfstring{$\tau_0$}{T}.
+
+To compute the AMI, an histogram of $x(n)$ using $n$ bins is calculated and then a probability distribution of data is computed \cite{kantz2003}.
+AMI is therefore denoted by $I(\tau)$ which is the average mutual information between the original time series, $x(n)$, and the delayed time series, $x(n-\tau)$, delayed by $\tau$ \cite{kabiraj2012}.
 AMI is defined by
@@
 Probabilities are defined as follows:
-$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of 
-the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside 
-the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability 
-that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$. The AMI is measured 
-in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
-For small $\tau$, AMI will be large and it will then decrease more or 
-less rapidly. As $\tau$ increase and goes to a large limit, 
-$x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is 
-factorised as $p_ip_j$ for which AMI is close to zero. Then, in order 
-to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ 
-where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or,
-where the redundancy is the least" \cite{kantz2003}.
-
-For example, we compute the AMI for two time series:
-A) the $x$ solution of the Lorenz system, and
-B) a noise time series using a normal distribution with mean zero and 
-standard deviation equal to one. From Fig~\ref{fig01}(b4, b8), it can then be
-concluded that the amount of knowledge for any noise time series is zero 
-for which the first minimum embedding parameter is $\tau_0=1$. 
-On the contrary, the first minimum of the AMI for the chaotic time series 
-is $\tau_0=17$ which is the value that maximize the independence 
-between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
-\cite{bradley2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-Similarly as Cao's algorithm negatives, AMI's algorithm is not an
-exception for negatives, which are worthwhile to mention for further 
-investigations.
-For instance, (i) is not clear why the choose of the first minimum of the AMI 
-is the minimum delay embedding parameter \cite{kantz2003} and 
-(ii) the probability distribution of the AMI function is computed
-with the use of histograms which depends on a heuristic choice of number of bins
-for which AMI depends on partitioning \cite{garcia2005e71}.
-
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/methods/caoami/versions/drawing-v00.pdf}
-%    \caption{
-%	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
-%	(A) chaotic and (B) random time series.
-%	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
-%	with variations of $\tau$ values from one to twenty
-%	for chaotic and random time series.
-%	(G, H) AMI values where its first minimum value in the curve
-%	is the minimum time delay embedding ($\tau_0$).
-%	Code and data to reproduce the figure is available in \cite{srep2021}.
-%        }
-%    \label{fig:caoami}
-%\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
+$p_i$ is the probability that $x(n)$ has a value inside the $i$-th bin of the histogram, $p_j$ is the probability that $x(n+\tau)$ has a value inside the $j$-th bin of the histogram and $p_{ij}(\tau)$ the probability that $x(n)$ is in bin $i$ and $x(n+\tau)$ is in bin $j$.
+The AMI is measured in bits (base 2, also called shannons) \cite{kantz2003, nonlinearTseries2016}.
+For small $\tau$, AMI will be large and it will then decrease more or less rapidly.
+As $\tau$ increase and goes to a large limit, $x(n)$ and $x(n+\tau)$ have nothing to do with each other and $p_(ij)$ is factorised as $p_ip_j$ for which AMI is close to zero.
+Then, in order to obtain $\tau_0$, "it has to be found the first minimum of $I(\tau)$ where $x(n+\tau)$ adds maximal information to the knowledge from $x(n)$, or, where the redundancy is the least" \cite{kantz2003}.
+
+For example, we compute the AMI for two time series: A) the $x$ solution of the Lorenz system, and B) a noise time series using a normal distribution with mean zero and standard deviation equal to one.
+From Fig~\ref{fig01}(b4, b8), it can then be concluded that the amount of knowledge for any noise time series is zero for which the first minimum embedding parameter is $\tau_0=1$.
+On the contrary, the first minimum of the AMI for the chaotic time series is $\tau_0=17$ which is the value that maximize the independence between $x(n)$ and $x(n+\tau)$ in the reconstructed state space \cite{bradley2015}.
+Similarly as Cao's algorithm negatives, AMI's algorithm is not an exception for negatives, which are worthwhile to mention for further investigations.
+For instance, (i) is not clear why the choose of the first minimum of the AMI is the minimum delay embedding parameter \cite{kantz2003} and (ii) the probability distribution of the AMI function is computed with the use of histograms which depends on a heuristic choice of number of bins for which AMI depends on partitioning \cite{garcia2005e71}.
 
@@
 \subsection*{Recurrence Plots}
-Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences 
-in conservative systems, however such discovery was not put into practice 
-until the development of faster computers \cite{marwan2007},
-for which Eckmann et al. \cite{eckmann1987} in 1987 introduced a method
-where recurrences in the dynamics of a system can be visualised using 
-Recurrence Plots. The intention of Eckmann et al. \cite{eckmann1987} was to 
-propose a tool, called Recurrence Plot (RP), that provides insights into 
-high-dimensional dynamical systems where trajectories are very difficult to 
-visualise. Therefore, "RP helps us to investigate the 
-$m-$dimensional phase space trajectories through a two-dimensional 
-representation of its recurrences" \cite{marwan2015}.
-Similarly, Marwan et al. \cite{marwan2015} pointed out that additionally 
-to the methodologies of the state space reconstruction and other dynamic 
-invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, 
-the recurrences of the trajectories in the phase space can provide 
-important clues to characterise natural process that present, for
-instance, periodicities (as Milankovitch cycles) or irregular cycles 
-(as El Ni\~no Southern Oscillation). 
-Such recurrences can not only be presented visually using Recurrence Plots (RP) 
-but also be quantified with Recurrence Quantification metrics, which leads 
-to applications of these in various areas such as economy, physiology, 
-neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
-
-For the creation of a recurrence plot based on time series 
-$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
-reconstruction with uniform time-delay embedding 
-$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
-\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
-where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
-and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
-The recurrence plot is therefore a two-dimensional $N \times N$ square 
-matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
-whenever $X(i)$ is sufficiently close to $X(j)$: 
+Originally, Henri Poincar\'e in 1890 introduced the concept of recurrences in conservative systems, however such discovery was not put into practice until the development of faster computers \cite{marwan2007}, for which Eckmann et al.
+\cite{eckmann1987} in 1987 introduced a method where recurrences in the dynamics of a system can be visualised using
+Recurrence Plots.
+The intention of Eckmann et al.
+\cite{eckmann1987} was to propose a tool, called Recurrence Plot (RP), that provides insights into high-dimensional dynamical systems where trajectories are very difficult to visualise.
+Therefore, "RP helps us to investigate the $m-$dimensional phase space trajectories through a two-dimensional representation of its recurrences" \cite{marwan2015}.
+Similarly, Marwan et al.
+\cite{marwan2015} pointed out that additionally to the methodologies of the state space reconstruction and other dynamic invariants such as Lyapunov exponent, Kolmogorov-Sinai entropy, the recurrences of the trajectories in the phase space can provide important clues to characterise natural process that present, for instance, periodicities (as Milankovitch cycles) or irregular cycles (as El Ni\~no Southern Oscillation).
+Such recurrences can not only be presented visually using Recurrence Plots (RP) but also be quantified with Recurrence Quantification metrics, which leads to applications of these in various areas such as economy, physiology, neuroscience, earth science, astrophysics and engineering \cite{marwan2007}.
+
+For the creation of a recurrence plot based on time series $\{ \boldsymbol{x}_n \}$, it is first computed the state space reconstruction with uniform time-delay embedding $X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots, \boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$ where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
+The recurrence plot is therefore a two-dimensional $N \times N$ square matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ whenever $X(i)$ is sufficiently close to $X(j)$:
 %%********************************[EQUATION]************************************
@@
 %%********************************[EQUATION]************************************
-where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
-$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
-function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig01}(c1,c2)) \cite{eckmann1987, marwan2007,marwan2015}.
-RP is also characterised with a line of identity (LOI) which is a  
-diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$.
+where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, $|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside function (i.e.
+$\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) (Fig~\ref{fig01}(c1,c2)) \cite{eckmann1987, marwan2007,marwan2015}.
+RP is also characterised with a line of identity (LOI) which is a diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$.
 
@@
 \subsection*{Structures of Recurrence Plots}
-Pattern formations in the RPs can be designated either 
-as topology for large-scale patterns or texture for small-scale patterns.
-In the case of topology, the following pattern formations are commonly presented:
-(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
-uniformly distributed noise (Fig~\ref{fig01}c3),
-(ii) periodic and quasi-periodic systems where diagonal lines and 
-checkerboard structures represent oscillating systems, e.g., sinusoidal 
-signals (Fig~\ref{fig01}c4),
-(iii) drift where paling or darkening recurrence points away from 
-the LOI is caused by drifting systems, 
-e.g., logistic map (Fig~\ref{fig01}c5), and
-(iv) disrupted where recurrence points are presented white areas or 
-bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
-(Fig~\ref{fig01}c7) \cite{eckmann1987, marwan2015}.
-Texture patterns in RPs can be categorised as:
-(i) single or isolated recurrence points that represent rare occurring states, 
-and do not persist for any time or fluctuate heavily,
-(ii) dots forming diagonal lines where the length of the small-scale parallel 
-lines in the diagonal are related to the ratio of determinism or predictability 
-in the dynamics of the system, and
-(iii) dots forming vertical and horizontal lines where the length of the 
-lines represent a time length where a state does not change or change very 
-slowly and these patterns formation represent discontinuities in the signal, 
-and (iv) dots clustering to inscribe rectangular regions which are related 
-to laminar states or singularities \cite{marwan2015}.
-
-Although, each of the previous pattern descriptions of the structures in the 
-RP offer an idea of the characteristics of dynamical systems, 
-these might be misinterpreted and conclusions might tend to be subjective 
-as these require the interpretation of a particular researcher(s).
-Because of that, recurrence quantification analyis (RQA) offer objective 
-methodologies to quantify such visual characteristics of previous 
-recurrent pattern structures in the RP \cite{zbilut1992}.
-
-%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/methods/rps/versions/drawing-v00.pdf}
-%    \caption{
-%	{\bf Recurrence Plots and its patterns.} 
-%	(A) State space of the Lorenz system with controlling parameters 
-%	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
-%	(black circle) of a given point at $i$ is a recurrent point and is 
-%	represented as a black dot in the recurrence plot at location 
-%	$(i, j)$ or white otherwise.
-%	(B) Recurrence plot using the 
-%	three components of the Lorenz system and the RP with no embeddings 
-%	and threshold $\epsilon=5$.
-%	Time-series with its respective recurrence plot for:
-%	(C) uniformly distributed noise,
-%	(D) super-positionet harmonic oscillation 
-%	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-%	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-%	with a linearly increase term ($0.01*i$), and
-%	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-%	This figure is adapted from \cite{marwan2015}.
-%	Code and data to reproduce the figure is available in \cite{srep2021}.
-%	}
-%    \label{fig:rps}
-%\end{figure}
-%%%---------------------------------(FIGURE)-------------------------------------
+Pattern formations in the RPs can be designated either as topology for large-scale patterns or texture for small-scale patterns.
+In the case of topology, the following pattern formations are commonly presented: (i) homogeneous where uniform recurrence points are spread in the RP e.g., uniformly distributed noise (Fig~\ref{fig01}c3), (ii) periodic and quasi-periodic systems where diagonal lines and checkerboard structures represent oscillating systems, e.g., sinusoidal signals (Fig~\ref{fig01}c4), (iii) drift where paling or darkening recurrence points away from the LOI is caused by drifting systems, e.g., logistic map (Fig~\ref{fig01}c5), and (iv) disrupted where recurrence points are presented white areas or bands that indicate abrupt changes in the dynamics, e.g.
+Brownian motion (Fig~\ref{fig01}c7) \cite{eckmann1987, marwan2015}.
+Texture patterns in RPs can be categorised as: (i) single or isolated recurrence points that represent rare occurring states, and do not persist for any time or fluctuate heavily, (ii) dots forming diagonal lines where the length of the small-scale parallel lines in the diagonal are related to the ratio of determinism or predictability in the dynamics of the system, and (iii) dots forming vertical and horizontal lines where the length of the lines represent a time length where a state does not change or change very slowly and these patterns formation represent discontinuities in the signal, and (iv) dots clustering to inscribe rectangular regions which are related to laminar states or singularities \cite{marwan2015}.
+
+Although, each of the previous pattern descriptions of the structures in the RP offer an idea of the characteristics of dynamical systems, these might be misinterpreted and conclusions might tend to be subjective as these require the interpretation of a particular researcher(s).
+Because of that, recurrence quantification analyis (RQA) offer objective methodologies to quantify such visual characteristics of previous recurrent pattern structures in the RP \cite{zbilut1992}.
 
@@
 \subsection*{Recurrence Quantifications Analysis (RQA)}
-Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
-the density of recurrence points in RPs, then histograms of lengths for 
-diagonal lines in RPs were studied by \cite{trulla1996} which were the 
-introduction to the term recurrence quantification analysis (RQA) 
-\cite{marwan2008}. RQA has been applied in many fields such as life science, 
-engineering, physics, and others \cite{marwan2008}. Particularly in human 
-movement to investigate noise and complexity of postural control 
-\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
-coordination \cite{duran2017}. The success of RQA is not only due to its 
-simple algorithmic implementation but also to its capacity to detect tiny 
-modulations in frequency or phase which are not detectable using standard 
-methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
-RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
-dimension which is verified experimentally by \cite{iwanski1998}.
-RQA metrics comprehend percentage of recurrence, percentage of determinism, 
-ratio, Shannon entropy of the frequency distributions of the line lengths,
-maximal line length and divergence, trend and laminariy 
-\cite{marwan2007, marwan2015}. For this work, we considered only four 
-RQA metrics, due to its consistency with our preliminary experiments, 
-which are described below. Such metrics are computed the nonlinearTseries 
-R package \cite{nonlinearTseries2016}.
+Originally, Zbilut et al.
+\cite{zbilut1992} proposed metrics to investigate the density of recurrence points in RPs, then histograms of lengths for diagonal lines in RPs were studied by \cite{trulla1996} which were the introduction to the term recurrence quantification analysis (RQA) \cite{marwan2008}.
+RQA has been applied in many fields such as life science, engineering, physics, and others \cite{marwan2008}.
+Particularly in human movement to investigate noise and complexity of postural control \cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal coordination \cite{duran2017}.
+The success of RQA is not only due to its simple algorithmic implementation but also to its capacity to detect tiny modulations in frequency or phase which are not detectable using standard methods e.g.
+spectral or wavelet analysis \cite{marwan2011}, and that RQA's metrics are quantitatively and qualitatively independent of embedding dimension which is verified experimentally by \cite{iwanski1998}.
+RQA metrics comprehend percentage of recurrence, percentage of determinism, ratio, Shannon entropy of the frequency distributions of the line lengths, maximal line length and divergence, trend and laminariy \cite{marwan2007, marwan2015}.
+For this work, we considered only four RQA metrics, due to its consistency with our preliminary experiments, which are described below.
+Such metrics are computed the nonlinearTseries R package \cite{nonlinearTseries2016}.
 
@@
 which enumerates the black dots in the RP excluding the line of identity.
-REC is a measure of the relative density of recurrence points in the sparse 
-matrix \cite{marwan2015}.
+REC is a measure of the relative density of recurrence points in the sparse matrix \cite{marwan2015}.
 %REC is computed as follow with the nonlinearTseries package 
@@
 \subsubsection*{DET values} 
-The percent determinism (DET) is defined as the fraction of recurrence points
-that form diagonal lines and it is determined by
+The percent determinism (DET) is defined as the fraction of recurrence points that form diagonal lines and it is determined by
 %%********************************[EQUATION]************************************
@@
 is the histogram of the lengths of the diagonal structures in the RP.
-DET can be interpreted as the predictability of the system for periodic signals 
-which, in essence, have longer diagonal lines than the short diagonals lines
-for chaotic signals or absent diagonal lines for stochastic signals 
-\cite{marwan2007, marwan2015}. Similarly, DET is considered as a measurement for 
-the organisation of points in RPs  \cite{iwanski1998}. 
+DET can be interpreted as the predictability of the system for periodic signals which, in essence, have longer diagonal lines than the short diagonals lines for chaotic signals or absent diagonal lines for stochastic signals \cite{marwan2007, marwan2015}.
+Similarly, DET is considered as a measurement for the organisation of points in RPs  \cite{iwanski1998}.
 %percent determinism (DET) is computed as follow with the nonlinearTseries 
@@
 \subsubsection*{RATIO values}
-RATIO is defined as the ratio between DET and REC and it is calculated from 
-the frequency distributions of the lengths of the diagonal lines.
+RATIO is defined as the ratio between DET and REC and it is calculated from the frequency distributions of the lengths of the diagonal lines.
 RATIO is useful to discover dynamic transitions \cite{marwan2015}.
@@
 \subsubsection*{ENT values}
-ENT is the Shannon entropy of the frequency distribution of the diagonal 
-line lengths and it is defined as
+ENT is the Shannon entropy of the frequency distribution of the diagonal line lengths and it is defined as
 %%********************************[EQUATION]************************************
@@
 ENT reflects the complexity of the deterministic structure in the system.
-For instance, for uncorrelated noise or oscillations, 
-the value of ENT is rather small and indicates low complexity of the system,
-therefore "the higher the ENT is the more complex the dynamics are" 
-\cite{marwan2007, marwan2015}.
+For instance, for uncorrelated noise or oscillations, the value of ENT is rather small and indicates low complexity of the system, therefore "the higher the ENT is the more complex the dynamics are" \cite{marwan2007, marwan2015}.
 %#'  \item \emph{ENTR}: Shannon entropy of the diagonal line lengths distribution
@@
 \subsection*{Sensitivity and robustness of RPs and RQA.}
-RP and RQA are a very young field in nonlinear dynamics and many questions 
-are still open, for instance, different parameters for window length size 
-of the time series, embedding parameters or recurrence threshold can 
-generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
-
-The selection of recurrence threshold, $\epsilon$, can depend on the system 
-that is analysed. For instance, when studying dynamical invariants $\epsilon$ 
-require to be very small, for trajectory reconstruction $\epsilon$ requires 
-to have a large thresholds or when studying dynamical transition 
-there is little importance about the selection of the threshold 
-\cite{marwan2011}. Other criteria for the selection of $\epsilon$ is that 
-the recurrence threshold  should be five times larger 
-than the standard deviation of the observational noise
-or the use of diagonal structures within the RP is suggested in order
-to find the optimal recurrence threshold for (quasi-)periodic process 
-\cite{marwan2011}.
-Similarly, Iwanski et al. \cite{iwanski1998} highlighted the importance 
-of choosing the right embedding parameters to perform RQA for which 
-many experiments have to be performed using different parameters in order 
-to have a better intuition of the nature of the time series and how 
-this is represented by using RQA.
-
-With that in mind, this work explores the sensitivity and robustness 
-of the window size of time series, embedding parameters for RSS with UTDE 
-and recurrence threshold for RP and RQA in order to gain a better insight 
-into the underlying time series collected from inertial sensors in the 
-context of human-humanoid imitation activities.
-%Such changes of recurrence threshold values 
-%can modify the patterns in RPs and therefore the values of RQA metrics.
-%We therefore computed 3D surfaces to explore the sensibility and robustness of 
-%embedding parameters and recurrence threshold in RQA  metrics. Following the 
-%same methodology of computing 3D surfaces, we also considered variation of 
-%window length size to present RQA metrics dependencies with embedding 
-%parameters, recurrence thresholds and window length size.
-
-%Choosing an appropriate recurrence threshold is crucial so as to get 
-%meaningful representations in RP and RQA, however,
-
-%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
-%metrics for RQA are independent of embedding dimension parameters, 
-%however, that is not the case when using 
-%different recurrence thresholds.
-
-%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
-%determining embedding parameters, their method consisted on creating 3D 
-%surfaces with RQA metrics with an increase of embedding parameters 
-%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
-%fluctuations and gradual changes in the 3D surfaces that provide information 
-%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
-%created 3D surfaces for visual selection of not only embedding parameters 
-%but also recurrence thresholds. Following same methodologies, we explored 
-%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
-%using 3D surfaces by an unitary increase of the pair embedding 
-%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
-%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
-%(Fig.~\ref{fig:topo_rqas}).
+RP and RQA are a very young field in nonlinear dynamics and many questions are still open, for instance, different parameters for window length size of the time series, embedding parameters or recurrence threshold can generate different results in RQA's metrics \cite{marwan2011, eckmann1987}.
+
+The selection of recurrence threshold, $\epsilon$, can depend on the system that is analysed.
+For instance, when studying dynamical invariants $\epsilon$ require to be very small, for trajectory reconstruction $\epsilon$ requires to have a large thresholds or when studying dynamical transition there is little importance about the selection of the threshold \cite{marwan2011}.
+Other criteria for the selection of $\epsilon$ is that the recurrence threshold  should be five times larger than the standard deviation of the observational noise or the use of diagonal structures within the RP is suggested in order to find the optimal recurrence threshold for (quasi-)periodic process \cite{marwan2011}.
+Similarly, Iwanski et al.
+\cite{iwanski1998} highlighted the importance of choosing the right embedding parameters to perform RQA for which many experiments have to be performed using different parameters in order to have a better intuition of the nature of the time series and how this is represented by using RQA.
+
+With that in mind, this work explores the sensitivity and robustness of the window size of time series, embedding parameters for RSS with UTDE and recurrence threshold for RP and RQA in order to gain a better insight into the underlying time series collected from inertial sensors in the context of human-humanoid imitation activities.
+%Such changes of recurrence threshold values can modify the patterns in RPs and therefore the values of RQA metrics.
+%We therefore computed 3D surfaces to explore the sensibility and robustness of embedding parameters and recurrence threshold in RQA  metrics.
+%Following the same methodology of computing 3D surfaces, we also considered variation of window length size to present RQA metrics dependencies with embedding parameters, recurrence thresholds and window length size.
+%Zbilut et al.
+\cite{zbilut1992} established RQA metrics with the aim of determining embedding parameters, their method consisted on creating 3D surfaces with RQA metrics with an increase of embedding parameters ($m$ and $\tau$), then Zbilut et al.
+\cite{zbilut1992} explored fluctuations and gradual changes in the 3D surfaces that provide information about the embeddings.
+Much recently, Marwan et al.
+\cite{marwan2015} created 3D surfaces for visual selection of not only embedding parameters but also recurrence thresholds.
+%Following same methodologies, we explored the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR) using 3D surfaces by an unitary increase of the pair embedding parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) (Fig.~\ref{fig:topo_rqas}).
 
@@
 \section*{Experiment} \label{sec:experiment}
-We conducted an experiment in the context of human-humanoid imitation (HHI) 
-activities where participants were asked to imitate simple horizontal and 
-vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
-Such simple movements were repeated ten times for the participant 
-who copied NAO's arm movements in a face-to-face imitation activity.
-Also, wearable inertial measurement unit (IMU) sensors were attached 
-to the right hand of the participant and to the left hand of the robot 
-(Figure~\ref{fig02}a1,a3). Data were then collected with four NeMEMSi
-IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
-accelerometer, gyroscope and magnetometer sensors and quaternions 
-\cite{Comotti2014}.
+We conducted an experiment in the context of human-humanoid imitation (HHI) activities where participants were asked to imitate simple horizontal and vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
+Such simple movements were repeated ten times for the participant who copied NAO's arm movements in a face-to-face imitation activity.
+Also, wearable inertial measurement unit (IMU) sensors were attached to the right hand of the participant and to the left hand of the robot (Figure~\ref{fig02}a1,a3).
+Data were then collected with four NeMEMSi IMU sensors with sampling rate of 50Hz provinding tri-axial data of the accelerometer, gyroscope and magnetometer sensors and quaternions \cite{Comotti2014}.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 \subsection*{Participants}
-Twenty-three participants,
-from now on defined as $pN$ where $N$ is the number of participant, were 
-invited to do the experiment. However, data for three participants were 
-not used because the instructions for $p01$, who was the only left-handed,
-were mistakenly given in a way that movements were performed different
-from what had been planned, and for participants $p13$ and $p16$ 
-data were corrupted because of problems with Bluetooth communications with sensors. 
+Twenty-three participants, from now on defined as $pN$ where $N$ is the number of participant, were invited to do the experiment.
+However, data for three participants were not used because the instructions for $p01$, who was the only left-handed, were mistakenly given in a way that movements were performed different from what had been planned, and for participants $p13$ and $p16$ data were corrupted because of problems with Bluetooth communications with sensors.
 With that in mind, data for twenty participants were analysed in this work.
-Of the twenty participants, all of them are right-handed healthy participants 
-of whom four are females and sixteen are males with a mean and standard 
-deviation (SD) age of mean=19.8 (SD=1.39).
+Of the twenty participants, all of them are right-handed healthy participants of whom four are females and sixteen are males with a mean and standard deviation (SD) age of mean=19.8 (SD=1.39).
 
 \subsection*{Ethics}
-All participants provided informed consent forms prior to participation 
-in the experiment. Hence, the experiments were conducted in November 2016 and 
-participants confirmed reading and understanding the participant information sheet of the 
-experiments and were able to withdraw from the experiment at any time without giving
-any reason. The design of the experiments adhered to the University of Birmingham
-regulations, data were anonymised and videos were stored only on a 
-personal computer in accordance with the Data Protection Act 1998. 
-%Refer to Appendix D for further information about the ethics, 
-%online participation information sheets and experiment check list.
+All participants provided informed consent forms prior to participation in the experiment.
+Hence, the experiments were conducted in November 2016 and participants confirmed reading and understanding the participant information sheet of the experiments and were able to withdraw from the experiment at any time without giving any reason.
+The design of the experiments adhered to the University of Birmingham regulations, data were anonymised and videos were stored only on a personal computer in accordance with the Data Protection Act 1998.
+%Refer to Appendix D for further information about the ethics, online participation information sheets and experiment check list.
 
 \subsection*{Human-humanoid imitation activities}
-For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
-two of which were attached to the right hand of the participant and the 
-other two to the left hand of the humanoid robot.
-Then, each participant was asked to imitate repetitions of simple horizontal
-and vertical arm movements performed by the humanoid robot in the following 
-conditions:
-(i) ten repetitions of horizontal arm movement at normal (HN) 
-	and faster (HF) speed (Figure~\ref{fig02}a1), and
-(ii) ten repetitions of vertical arm movement at normal (VN) 
-	and faster (VF) speed (Figure~\ref{fig02}a3).
-The normal and faster speed of arm movements is defined by the duration 
-in number of samples of one repetition of NAO's arm movements. 
-We select NAO's arm movements duration to distinguish between normal and 
-faster arm movements as NAO's movements have less variation 
-for such repetitive movements. 
-The duration for one repetition of the horizontal 
-arm movement at normal speed, HN, is 5 seconds considering 
-that each repetition last around 250 samples.
-For horizontal arm movement at faster speed, HF, each repetition were performed 
-in 2 seconds which correspond to 90 samples of data.
-The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
-which is around 300 samples of data. For vertical arm movement at 
-faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
-to 120 samples of data.
-To visualise the distinction between normal and faster speed for horizontal 
-and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series
-for axes Z and Y of the gyroscope sensors with four window lengths: 
-2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
-and 15-sec (750-samples).
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
-%    \caption{
-%	{\bf Time series duration of horizontal and vertical arm movements.}
-%    		Time series of smoothed data from gyroscope sensor
-%		for different speed arm movements performed by NAO:
-%		(A) Horizontal Normal arm movement (HN),
-%		(B) Horizontal Faster arm movement (HF),
-%		(C) Vertical Normal arm movement (VN) and
-%		(D) Vertical Faster arm movement (VF).
-%		Additionally,
-%		(A) shows window sizes for 2-seconds (100 samples),
-%		5-seconds (250 samples), 10-seconds (500 samples)
-%		and 15-seconds (750 samples)
-%		which are also presented in (B), (C) and (D).
-%		Code and data to reproduce the figure is available in \cite{srep2021}.
-%        }
-%	\label{fig:sts}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
+For human-humanoid imitation (HHI) activities four neMEMSi sensors were used, two of which were attached to the right hand of the participant and the other two to the left hand of the humanoid robot.
+Then, each participant was asked to imitate repetitions of simple horizontal and vertical arm movements performed by the humanoid robot in the following conditions:
+(i) ten repetitions of horizontal arm movement at normal (HN) and faster (HF) speed (Figure~\ref{fig02}a1), and
+(ii) ten repetitions of vertical arm movement at normal (VN) and faster (VF) speed (Figure~\ref{fig02}a3).
+The normal and faster speed of arm movements is defined by the duration in number of samples of one repetition of NAO's arm movements.
+We select NAO's arm movements duration to distinguish between normal and faster arm movements as NAO's movements have less variation for such repetitive movements.
+The duration for one repetition of the horizontal arm movement at normal speed, HN, is 5 seconds considering that each repetition last around 250 samples.
+For horizontal arm movement at faster speed, HF, each repetition were performed in 2 seconds which correspond to 90 samples of data.
+The vertical arm movement at normal speed, VN, were performed  in 6 seconds which is around 300 samples of data.
+For vertical arm movement at faster speed, VF, each repetition lasts about 2.4 seconds which correspond to 120 samples of data.
+To visualise the distinction between normal and faster speed for horizontal and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series for axes Z and Y of the gyroscope sensors with four window lengths: 2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) and 15-sec (750-samples).
 
@@
 \label{sec:experiment:subsec:imu}
-To give insight to the research questions, 
-we considered various conditions of time series collected for this work: 
+To give insight to the research questions, we considered various conditions of time series collected for this work:
 \begin{itemize}
-\item Three levels of smoothness for normalised data 
-	(e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg 
-	stands for Savitzky-Golay filter with two different filter lengths (29 and 159) 
-	and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal}
-	and zmuv is zero mean unit variance.
-\item four arm movement activities with two velocities: 
-	horizontal normal (HN), horizontal faster (HF), 
-	vertical normal (VN), and vertical faster (VF), and
-\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
-	10-sec (500 samples) and 15-sec (750 samples).
+\item Three levels of smoothness for normalised data (e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg stands for Savitzky-Golay filter with two different filter lengths (29 and 159) and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal} and zmuv is zero mean unit variance.
+\item four arm movement activities with two velocities: horizontal normal (HN), horizontal faster (HF), vertical normal (VN), and vertical faster (VF), and
+\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 10-sec (500 samples) and 15-sec (750 samples).
 \end{itemize}
 %After the data collection, raw time series were windowed, normalised and smoothed. 
-Due to space limitations and to have simple visualisation, 
-we only present 10-sec (500 samples) window length time series for 
-three participants (p01, p01 and p03) performing horizontal 
-arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}C).
-%%%---------------------------------(FIGURE)-------------------------------------
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
-%    	\caption{
-%	{\bf Time series for horizontal and vertical arm movements.}
-%		(A/D) raw-normalised (sg0zmuv),
-%		(B/E) normalised-smoothed 1 (sg1zmuv) and
-%		(C/F) normalised-smoothed 2 (sg2zmuv).
-%		Time series are only for three participants
-%		($p01$, $p02$, and $p03$)
-%		for horizontal and vertical arm movements in normal
-%		and faster velocity (HN, HF, VN, VF)
-%		with the normalised GyroZ or GyroY axis
-%		and with one sensor attached to the participant (HS01)
-%		and other sensor attached to the robot (RS01).
-%	Code and data to reproduce the figure is available in \cite{srep2021}.
-%       }
-%    \label{fig:ts}
-%\end{figure}
-%%%---------------------------------(FIGURE)------------------------------------
+Due to space limitations and to have simple visualisation, we only present 10-sec (500 samples) window length time series for three participants (p01, p01 and p03) performing horizontal arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}C).
 
 \subsubsection*{Raw data}
-%Considering the work of \cite{shoaib2016} which provided evidence of an 
-%improvement in recognition activities when combining data from 
-%accelerometer and gyroscope. 
-We focus our analysis from data of the 
-accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and 
-leave the data of the magnetometer and quaternions for further investigation 
-because of their possible variations with regard to magnetic disturbances \cite{shoaib2016}.
-Data from the accelerometer is defined by triaxial time series 
-$A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ 
-(Eq.~\ref{eq:AG}), and the same for data from the gyroscope 
-which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ 
-representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:AG}).
-Both triaxial time series of each sensor, $a$ and $g$, are denoted with 
-its respective axes subscripts $x,y,z$, where $n$ is the sample index 
-and $N$ is the same maximum length of all axes for the time series.
+%Considering the work of \cite{shoaib2016} which provided evidence of an improvement in recognition activities when combining data from accelerometer and gyroscope.
+We focus our analysis from data of the accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and leave the data of the magnetometer and quaternions for further investigation because of their possible variations with regard to magnetic disturbances \cite{shoaib2016}.
+Data from the accelerometer is defined by triaxial time series $A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ (Eq.~\ref{eq:AG}), and the same for data from the gyroscope which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:AG}).
+Both triaxial time series of each sensor, $a$ and $g$, are denoted with its respective axes subscripts $x,y,z$, where $n$ is the sample index and $N$ is the same maximum length of all axes for the time series.
 %Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
@@
 \subsubsection*{Postprocessing data}
-After the collection of raw data from four NeMEMsi sensors,
-time synchronisation alignment and interpolation were performed 
-in order to create time series with the same length and synchronised time.
-We refer the reader to \cite{Comotti2014} for further
-details about the time synchronisation process.
+After the collection of raw data from four NeMEMsi sensors, time synchronisation alignment and interpolation were performed in order to create time series with the same length and synchronised time.
+We refer the reader to \cite{Comotti2014} for further details about the time synchronisation process.
 
@@
 \subsubsection*{Smoothing data}
-Commonly, a low-pass filter is use either to capture the low 
-frequencies that represent \%99 of the human body energy or to get 
-the gravitational and body motion components of 
-accelerations \cite{anguita2013}. However, the elimination of 
-a range of frequencies is not the main focus of this work but 
-the conservation of the structure of time series in terms of 
-their width and heights where, for instance, Savitzky-Golay filter 
-can be a way to accomplish such task \cite{press1992}. 
-Savitzky-Golay filter is based on the 
-principle of moving window averaging which preserves the area under 
-the curve (the zeroth moment), its mean position in time 
-(the first moment) but the line width (the second moment) is violated 
-and that results, for example, in the case of spectrometric data, into
-a narrow spectral line with reduced height and width. 
-With that in mind, the aim of Savitzky-Golay filtering is to find filter 
-coefficients $c_n$ that preserve higher momentums which are based on local 
-least-square polynomial approximations \cite{savitzkygolay1964, 
-press1992, schafer2011}.
-Therefore, Savitzky-Golay coefficients are computed with \texttt{sgolay(p,n,m)} 
-in R where \texttt{p} is the filter order, 
-\texttt{n} is the filter length (must be odd) and \texttt{m} is the 
-$m$-th derivative of the filter coefficients \cite{Rsignal}. 
-Smoothed signal is represented with a tilde over the original 
-variable of the signal: $\tilde{x}(n)$.
+Commonly, a low-pass filter is use either to capture the low frequencies that represent \%99 of the human body energy or to get the gravitational and body motion components of accelerations \cite{anguita2013}.
+However, the elimination of a range of frequencies is not the main focus of this work but the conservation of the structure of time series in terms of their width and heights where, for instance, Savitzky-Golay filter can be a way to accomplish such task \cite{press1992}.
+Savitzky-Golay filter is based on the principle of moving window averaging which preserves the area under the curve (the zeroth moment), its mean position in time (the first moment) but the line width (the second moment) is violated and that results, for example, in the case of spectrometric data, into a narrow spectral line with reduced height and width.
+With that in mind, the aim of Savitzky-Golay filtering is to find filter coefficients $c_n$ that preserve higher momentums which are based on local least-square polynomial approximations \cite{savitzkygolay1964, press1992, schafer2011}.
+Therefore, Savitzky-Golay coefficients are computed with \texttt{sgolay(p,n,m)} in R where \texttt{p} is the filter order, \texttt{n} is the filter length (must be odd) and \texttt{m} is the $m$-th derivative of the filter coefficients \cite{Rsignal}.
+Smoothed signal is represented with a tilde over the original variable of the signal: $\tilde{x}(n)$.
 
 \subsubsection*{Windowing data size}
-With regards to the window size, Shoaib et al. in 2016 \cite{shoaib2016} 
-investigated effects of seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds)
-and combination of inertial sensors (accelerometer, gyroscope and linear 
-acceleration sensor) to improve the activity recognition performance for 
-repetitive activities (walking, jogging and biking) and less repetitive 
-activities (smoking, eating, giving a talk or drinking a coffee).
-With that in mind, Shoaib et al. \cite{shoaib2016} concluded that the 
-increase of window length improve the recognition of complex activities 
-because these requires a large window length to learn the repetitive 
-motion patterns. Particularly, one of the recommendations is to use large 
-window size to recognise less repetitive activities which mainly involve 
-random hand gestures. Therefore, for the four activities 
-(HN, HF, VN, and VF) in this work, which are mainly repetitive, 
-we select only four window sizes for analysis: 2-s window (100 samples), 
-5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) 
-(Figures~\ref{fig02}B).
+With regards to the window size, Shoaib et al.
+in 2016 \cite{shoaib2016} investigated effects of seven window lengths (2, 5, 10, 15, 20, 25, 30 seconds) and combination of inertial sensors (accelerometer, gyroscope and linear
+acceleration sensor) to improve the activity recognition performance for repetitive activities (walking, jogging and biking) and less repetitive activities (smoking, eating, giving a talk or drinking a coffee).
+With that in mind, Shoaib et al.
+\cite{shoaib2016} concluded that the increase of window length improve the recognition of complex activities because these requires a large window length to learn the repetitive motion patterns.
+Particularly, one of the recommendations is to use large window size to recognise less repetitive activities which mainly involve random hand gestures.
+Therefore, for the four activities (HN, HF, VN, and VF) in this work, which are mainly repetitive, we select only four window sizes for analysis: 2-s window (100 samples), 5-s window (250 samples), 10-s (500 samples) and 15-s window (750 samples) (Figures~\ref{fig02}B).
 
@@
 \subsection*{Reconstructed State Spaces}
-As noted in the Introduction, a challenge in the implementation of uniform time-delay embedding 
-arises from the selection of embedding parameters because of the uniqueness of each time series 
-in terms of its structure (e.g., modulation of amplitude, frequency, phase etc.). 
-With that in mind, the options are to either calculate embedding parameters for each unique 
-instance (which can make comparison challenging) or to find parameters which can apply to all 
-instances in the study.
-We recognise that the latter approach is not without its problems, 
-but our approach is to compute sample mean over all values in each of the conditions of the 
-time series for minimum dimension and minimum delay values.
+As noted in the Introduction, a challenge in the implementation of uniform time-delay embedding arises from the selection of embedding parameters because of the uniqueness of each time series in terms of its structure (e.g., modulation of amplitude, frequency, phase etc.).
+With that in mind, the options are to either calculate embedding parameters for each unique instance (which can make comparison challenging) or to find parameters which can apply to all instances in the study.
+We recognise that the latter approach is not without its problems, but our approach is to compute sample mean over all values in each of the conditions of the time series for minimum dimension and minimum delay values.
 
 \subsubsection*{Minimum embedding parameters}
-Minimum embedding parameters were initially computed using False Nearest Neighbour (FNN) 
-and Average Mutual Information (AMI).   We used FNN to calculate a value for $m$
-to unfold the attractors and AMI to calculate a value for $\tau$ and maxime the 
-information in the unfolded attractor.
-To illustrate this, figure \ref{fig:cao_ami} (A) show box plots for minimum embedding 
-values for sensors on the wrist of the human (HS01) and the robot (RS01).   
-As we had assumed, minimum embedding values for HS01 show greater variation than 
-those for RS01 (as indicated by differences in interquartile ranges).  
-Additionally, figure \ref{fig:cao_ami} (A) show a decrease in mean values (rhombus) in the box plots 
-as smoothness of the time-series increases (see sg0, sg1, sg2).  
+Minimum embedding parameters were initially computed using False Nearest Neighbour (FNN) and Average Mutual Information (AMI).
+We used FNN to calculate a value for $m$ to unfold the attractors and AMI to calculate a value for $\tau$ and maxime the information in the unfolded attractor.
+To illustrate this, figure \ref{fig:cao_ami} (A) show box plots for minimum embedding values for sensors on the wrist of the human (HS01) and the robot (RS01).
+As we had assumed, minimum embedding values for HS01 show greater variation than those for RS01 (as indicated by differences in interquartile ranges).
+Additionally, figure \ref{fig:cao_ami} (A) show a decrease in mean values (rhombus) in the box plots as smoothness of the time-series increases (see sg0, sg1, sg2).
 Figure \ref{fig:cao_ami} (B) show box plots for Average Mutual Information (AMI).
-One can see that, in contrast to RS01, minimum values for HS01 tend to spread 
-as the smoothness of the time-series increases.
-The sample mean for minimum value of embedding parameter, 
-$m$ derived from FNN  (figures \ref{fig:cao_ami} A) is $\overline{m}_0=6$ 
-and $\tau$ from AMI (figures \ref{fig:cao_ami} B) is $\overline{\tau}_0=8$.
+One can see that, in contrast to RS01, minimum values for HS01 tend to spread as the smoothness of the time-series increases.
+The sample mean for minimum value of embedding parameter, $m$ derived from FNN  (figures \ref{fig:cao_ami} A) is $\overline{m}_0=6$ and $\tau$ from AMI (figures \ref{fig:cao_ami} B) is $\overline{\tau}_0=8$.
 
@@
 	{\bf Box plots of minimum embedding parameters.} 
-	Box plots of (A) minimum embedding dimensions 
-	and (B) first minimum AMI values for 
-	Horizontal Normal (HN), Horizontal Faster (HF),
-	Vertical Normal (VN) and Vertical Faster (VF)
-	with sensors attached to participants (HS01) and
-	sensor attached to robot (RS01).
-	Minimum embedding parameters are for twenty participants 
-	($p01$ to $p20$) with three smoothed signals 
-	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
-	and window length of 10-sec (500 samples).
+	Box plots of (A) minimum embedding dimensions and (B) first minimum AMI values for Horizontal Normal (HN), Horizontal Faster (HF), Vertical Normal (VN) and Vertical Faster (VF) with sensors attached to participants (HS01) and sensor attached to robot (RS01).
+	Minimum embedding parameters are for twenty participants ($p01$ to $p20$) with three smoothed signals (sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ) and window length of 10-sec (500 samples).
 	Code and data to reproduce the figure is available in \cite{srep2021}.
@@
 \subsubsection*{Uniform Time-Delay Embedding}
-Using the overall embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$), 
-the first three axis of the rotated Principal Components Analysis (PCA) 
-are shown for the reconstructed state spaces of horizontal (figures~\ref{fig04}(A))
-and vertical (figures~\ref{fig04}(B)) arm movements.
-While visual inspection of these figures 
-suggests differences in the trajectories of the reconstructed state spaces, 
-we require an objective quantification to determine the extent of the differences.  
-One approach could be use Euclidean distances from the origin for points in these 
-trajectories, but this proved inconclusive and was not able to capture 
-the dynamics of the trajectories. Consequently, we applied 
-Recurrence Plots and Recurrence Quantification Analysis.
+Using the overall embedding parameters ($\overline{m_0}=6$, $\overline{\tau_0}=8$), the first three axis of the rotated Principal Components Analysis (PCA) are shown for the reconstructed state spaces of horizontal (figures~\ref{fig04}(A)) and vertical (figures~\ref{fig04}(B)) arm movements.
+While visual inspection of these figures suggests differences in the trajectories of the reconstructed state spaces, we require an objective quantification to determine the extent of the differences.
+One approach could be use Euclidean distances from the origin for points in these trajectories, but this proved inconclusive and was not able to capture the dynamics of the trajectories.
+Consequently, we applied Recurrence Plots and Recurrence Quantification Analysis.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 	Reconstructed state spaces for time series for p01 of Figure \ref{fig02}C.
-	Reconstructed state spaces were computed with 
-	embedding parameters 
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$
+	Reconstructed state spaces were computed with embedding parameters $\overline{m}_0=6$, $\overline{\tau}_0=8$
 	for (A) horizontal and (B) vertical arm movements.
@@
 \subsection*{Recurrences Plots}
-Using the average embedding parameters 
-($\overline{m}_0=6$, $\overline{\tau}_0=8$) 
-and an recurrence threshold of $\epsilon=1$.
-As our interest is for dynamical transitions, 
-there is little importance on the selection of $\epsilon$ which in this case
-is 1. 
-Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig05}(A)) and
-vertical (figures~\ref{fig05}(B)) arm movements.
+Using the average embedding parameters ($\overline{m}_0=6$, $\overline{\tau}_0=8$) and an recurrence threshold of $\epsilon=1$.
+As our interest is for dynamical transitions, there is little importance on the selection of $\epsilon$ which in this case is 1.
+Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig05}(A)) and vertical (figures~\ref{fig05}(B)) arm movements.
 %%---------------------------------(FIGURE)-------------------------------------
@@
 	{\bf RPs for horizontal (A) and vertical (B) arm movements.}	
-	Recurrence plots were computed with 
-	embedding parameters 
+	Recurrence plots were computed with embedding parameters
 	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
 	{\bf Box plots for RQA metrics (C).}
-	RQA metrics for (c1) REC, (c2) DET, (c3) RATIO, and (c4) ENTR of
-	20 participants performing HN, HF, VN and VF movements
-	with sensors HS01, RS01 and three smoothed-normalised
-	time series (sg0, sg1 and sg2).
-	RQA values were computed with
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
+	RQA metrics for (c1) REC, (c2) DET, (c3) RATIO, and (c4) ENTR of 20 participants performing HN, HF, VN and VF movements with sensors HS01, RS01 and three smoothed-normalised time series (sg0, sg1 and sg2).
+	RQA values were computed with $\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
 	Code and data to reproduce the figure is available in \cite{srep2021}.
@@
 \subsection*{Recurrence Quantification Analysis} \label{ch6:rqas}
-Four Recurrence Quantification Analysis metrics (
-percentage recurrence, REC, representing the percentage of black dots in RP; 
-percentage determinism, DET, representing the predictability of the RP; 
-ratio of DET / REC, RATIO; Shannon entropy, ENTR) 
-were computed using the same parameters as for RP.
+Four Recurrence Quantification Analysis metrics (percentage recurrence, REC, representing the percentage of black dots in RP; percentage determinism, DET, representing the predictability of the RP; ratio of DET / REC, RATIO; Shannon entropy, ENTR) were computed using the same parameters as for RP.
 %\subsubsection*{REC values}
-Figure~\ref{fig05}(c1) presents box plots of REC values,
-for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements 
-in Horizontal (HS) or Vertical (VS) than for Fast (i.e., 2 seconds per movement).  
+Figure~\ref{fig05}(c1) presents box plots of REC values, for HS01, are more spread for Slow (i.e., 5 seconds per movement) movements in Horizontal (HS) or Vertical (VS) than for Fast (i.e., 2 seconds per movement).
 This suggests greater variation between participants for the Slow movements.  
-For RS01, there is little variation between Slow and Fast movement
-(interquartile range of 0.01). 
-In terms of smoothness, there seems little effect of HS01 but RS01 values 
-do show affects of smoothness (see the incremental changes of mean values (rhombus)).
+For RS01, there is little variation between Slow and Fast movement (interquartile range of 0.01).
+In terms of smoothness, there seems little effect of HS01 but RS01 values do show affects of smoothness (see the incremental changes of mean values (rhombus)).
 %\subsubsection*{DET values}
-Figure \ref{fig05}(c2) presents DET values and shows
-little difference for type of movement or performer.  
-However, DET values are affected by changes in smoothness of the signal, 
-particularly for Fast movement.
+Figure \ref{fig05}(c2) presents DET values and shows little difference for type of movement or performer.
+However, DET values are affected by changes in smoothness of the signal, particularly for Fast movement.
 %\subsubsection*{RATIO values}
@@
 %	{\bf Box plots for RQA metrics.}
-%	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of
-%	20 participants performing HN, HF, VN and VF movements
-%	with sensors HS01, RS01 and three smoothed-normalised
-%	time series (sg0, sg1 and sg2).
-%	RQA values were computed with
-%	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
+%	RQA metrics for (A) REC, (B) DET, (C) RATIO, and (D) ENTR of 20 participants performing HN, HF, VN and VF movements with sensors HS01, RS01 and three smoothed-normalised time series (sg0, sg1 and sg2).
+%	RQA values were computed with $\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
 %	Code and data to reproduce the figure is available in \cite{srep2021}.
@@
 %\subsection*{RQA metrics with fixed parameters}
-%Considering that RQA metrics were computed with fixed embedding parameters 
-%($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found 
-%the following. REC values, which represents the \% of black points in the RPs, 
-%were more affected with and increase in normal speed movements (HN and VN) 
-%than faster movements (HF and VF) for the sensor attached to the participants 
-%(HS01). Such decrease of REC values from normal speed to faster speed 
-%movements is also presented in data from sensor attached to the robot (RS01), 
-%and little can be said with regard to the dynamics of the time series coming 
-%from RS01 (Fig \ref{fig:RQABP}A).
-%Similarly, DET values, representing predictability and 
-%organisation in the RPs, present little variation in the any of the time 
-%series where little can be said (Fig \ref{fig:RQABP}B).
-%In contrast, RATIO values, which represent 
-%dynamic transitions, were more variable for faster movements (HF and VF) 
-%than normal speed movements (HN and VN) with sensors attached to the 
-%participants (HS01). For data coming from sensors attached to the robot 
-%(RS01), RATIO values from horizontal movements (HN, HF) appear to vary 
-%more than values coming from vertical movmentes (VN, VF) 
-%(Fig \ref{fig:RQABP}C).
-%With that, it can be said that RATIO values can represent a bit better
-%than REC or DET metrics for the variability of imitation activities in 
-%each of the conditions for time series.
-%Similarly, ENTR values for HN were higher than values for HF
-%and ENTR values varied more for sensor attached to participants 
-%than ENTR values for sensors of the robot. It is evidently that 
-%the higher the entropy the more complex the dynamics are, 
-%however, ENTR values for HN appear a bit higher than HF values, 
-%for which we believe this happens because of the structure the time series
-%which appear more complex for HN than  HF movements which presented a 
-%more consistence repetition (Fig \ref{fig:RQABP}D).
+%Considering that RQA metrics were computed with fixed embedding parameters ($m=6$ and $\tau=8$) and recurrence thresholds ($\epsilon=1$), we found the following.
+%REC values, which represents the \% of black points in the RPs, were more affected with and increase in normal speed movements (HN and VN) than faster movements (HF and VF) for the sensor attached to the participants (HS01).
+%Such decrease of REC values from normal speed to faster speed movements is also presented in data from sensor attached to the robot (RS01), and little can be said with regard to the dynamics of the time series coming from RS01 (Fig \ref{fig:RQABP}A).
+%Similarly, DET values, representing predictability and organisation in the RPs, present little variation in the any of the time series where little can be said (Fig \ref{fig:RQABP}B).
+%In contrast, RATIO values, which represent dynamic transitions, were more variable for faster movements (HF and VF) than normal speed movements (HN and VN) with sensors attached to the
+%participants (HS01).
+For data coming from sensors attached to the robot (RS01), RATIO values from horizontal movements (HN, HF) appear to vary more than values coming from vertical movmentes (VN, VF) (Fig \ref{fig:RQABP}C).
+%With that, it can be said that RATIO values can represent a bit better than REC or DET metrics for the variability of imitation activities in each of the conditions for time series.
+%Similarly, ENTR values for HN were higher than values for HF and ENTR values varied more for sensor attached to participants than ENTR values for sensors of the robot.
+It is evidently that the higher the entropy the more complex the dynamics are,
+%however, ENTR values for HN appear a bit higher than HF values, for which we believe this happens because of the structure the time series which appear more complex for HN than  HF movements which presented a more consistence repetition (Fig \ref{fig:RQABP}D).
 
 \subsection*{3D RQA ENTR}
-As ENTR appeared to be have higher sensitivity than the other measures, 
-we explored the impact of different embedding parameters 
-$ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $,
-$  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ 
-incrementing by 1 each run, with recurrence thresholds $\epsilon=1, 2, 3$ 
-and levels of smoothness (sg0, sg1, sg2).  
-Figure \ref{fig06}(A) shows that increasing recurrence
-threshold leads to an increase in ENTR regardless of level of smoothness.  
+As ENTR appeared to be have higher sensitivity than the other measures, we explored the impact of different embedding parameters $ \{ m \in \mathbb{R} | 1 \le m \le 10  \} $, $  \{ \tau \in \mathbb{R} | 1 \le \tau \le 10  \} $ incrementing by 1 each run, with recurrence thresholds $\epsilon=1, 2, 3$ and levels of smoothness (sg0, sg1, sg2).
+Figure \ref{fig06}(A) shows that increasing recurrence threshold leads to an increase in ENTR regardless of level of smoothness.
 Similarly, increasing level of smoothness will also increase ENTR (figure \ref{fig06}(A)).
-In terms of movement or performer, RQA ENTR decrease from Slow (HS, VS) to Fast (HF, VF) 
-for both human (HS01) and (RS01) (Figures~\ref{fig06}(B)).
-In terms of individual differences, for human participants, 
-figure \ref{fig:3dRQAENTR_participantsactivities} compares p01, p02 and p03 (for illustrative purposes) --
-both between each other and in comparison with the more consistent performance of the robot.  
-In terms of window length (w100 (2s), w250 (5s), w500(10s), w750 (15s)), 
-figure \ref{fig:3dRQAENTR_windowsactivities} shows that improvement in 
-capture with number of samples, although this has less of an effect on RQA ENTR.
+In terms of movement or performer, RQA ENTR decrease from Slow (HS, VS) to Fast (HF, VF) for both human (HS01) and (RS01) (Figures~\ref{fig06}(B)).
+In terms of individual differences, for human participants, figure \ref{fig:3dRQAENTR_participantsactivities} compares p01, p02 and p03 (for illustrative purposes) -- both between each other and in comparison with the more consistent performance of the robot.
+In terms of window length (w100 (2s), w250 (5s), w500(10s), w750 (15s)), figure \ref{fig:3dRQAENTR_windowsactivities} shows that improvement in capture with number of samples, although this has less of an effect on RQA ENTR.
 
@@
 
-To summarise this section of results, it can be said that computing
-embedding parameters for individual structure of time-series 
-data is already a solved problem \cite{frank2010, sama2013, bradley2015}. 
-However, it has been shown the challenge of finding embedding parameters
-for nonlinear dynamic tools that represent a set of different time-series data.
-That said, we proposed the use of sample mean of the set of embedding parameters
-for RSSs, RP and RQA to then noticed that the selection of recurrence 
-threshold, $\epsilon$, is also an open problem.
-For which, this work proposed the variation of recurrence thresholds 
-and embedding parameters to show the relationships of these to different datasets 
-(participants, activities, windows lengths and sensors).
+To summarise this section of results, it can be said that computing embedding parameters for individual structure of time-series data is already a solved problem \cite{frank2010, sama2013, bradley2015}.
+However, it has been shown the challenge of finding embedding parameters for nonlinear dynamic tools that represent a set of different time-series data.
+That said, we proposed the use of sample mean of the set of embedding parameters for RSSs, RP and RQA to then noticed that the selection of recurrence threshold, $\epsilon$, is also an open problem.
+For which, this work proposed the variation of recurrence thresholds and embedding parameters to show the relationships of these to different datasets (participants, activities, windows lengths and sensors).
 
@@
 \section*{Discussion}
-While there are many approaches to estimating embedding parameters for nonlinear analysis, 
-these can be influenced by the structure of the time-series data.   
-We show that, for RSS, RP and RQA, the estimation of embedding parameters can be performed 
-using a sample mean which, together with recurrence threshold, can be shown to be 
-influenced by activity, performer, window length and smoothness of time-series.  
-It is known that time-series from different sources and with different characteristics 
-require different embedding parameters, and this can produce different RSS, RPs and RQAs. 
-Although this work helps to understand the open problem of 
-finding right balance among (i) the level of smoothness of the signal, 
-(ii) the selection of recurrence thresholds and (iii) 
-the range of embedding parameters, 
-we have shown how RQA metrics can help to quantify movement variability.
+While there are many approaches to estimating embedding parameters for nonlinear analysis, these can be influenced by the structure of the time-series data.
+We show that, for RSS, RP and RQA, the estimation of embedding parameters can be performed using a sample mean which, together with recurrence threshold, can be shown to be influenced by activity, performer, window length and smoothness of time-series.
+It is known that time-series from different sources and with different characteristics require different embedding parameters, and this can produce different RSS, RPs and RQAs.
+Although this work helps to understand the open problem of finding right balance among (i) the level of smoothness of the signal, (ii) the selection of recurrence thresholds and (iii) the range of embedding parameters, we have shown how RQA metrics can help to quantify movement variability.
 
 \section*{Conclusions}
-In this paper we show how the selection of nonlinear analysis tool  
-(i.e., RSS, RP, RQA metrics) depends on what question one wishes to 
-address with time-series data (e.g., predictability, organisation, dynamics, complexity).  
-Time-series data characteristics (e.g., window length, smoothness), 
-time-series structure (e.g., frequency, amplitude) and 
-data source (e.g., sensor placement, performance, movement) all influence 
-the results that nonlinear analysis methods can provide.
-That said, it has been shown that the use of different characteristics of the data and 
-their collection can help us visualise and quantify variability of movement using 
-methods of nonlinear analysis.   
-There remain limitations of nonlinear methods in 
-relation to the estimation of parameters (e.g., recurrence threshold, embedding parameters) 
-which reflect the dynamics of specific movement and performers, window length 
-and structure of the time-series. 
-We note that RQA DET seems to show low sensitivity 
-to these differences, whereas REC and RATIO (primarily as a result of REC) show 
-variation across performers and movements.   
-RQA ENTR, with different recurrence thresholds, 
-can quantify variation in the time-series data and offers the most appropriate means 
-for analysing variability in movement to allow us to analyse individual differences 
-between human performers.  
-To then found out that RQA ENTR values with different recurrence 
-thresholds were appropriate to quantify 
-the different changes and variations of the characteristics of 
-time-series data.
-Therefore, the use of Shannon Entropy could be applied 
-to analyse human participants who might vary in age, 
-state of health, anthropometric features and capability to perform movement.
-
-%% As a guideline references should be limited to 60 (this is not 
-%% strictly enforced).
-%\bibliography{sample}
-%\bibliography{references}
+In this paper we show how the selection of nonlinear analysis tool (i.e., RSS, RP, RQA metrics) depends on what question one wishes to address with time-series data (e.g., predictability, organisation, dynamics, complexity).
+Time-series data characteristics (e.g., window length, smoothness), time-series structure (e.g., frequency, amplitude) and data source (e.g., sensor placement, performance, movement) all influence the results that nonlinear analysis methods can provide.
+That said, it has been shown that the use of different characteristics of the data and their collection can help us visualise and quantify variability of movement using methods of nonlinear analysis.
+There remain limitations of nonlinear methods in relation to the estimation of parameters (e.g., recurrence threshold, embedding parameters) which reflect the dynamics of specific movement and performers, window length  and structure of the time-series.
+We note that RQA DET seems to show low sensitivity to these differences, whereas REC and RATIO (primarily as a result of REC) show variation across performers and movements.
+RQA ENTR, with different recurrence thresholds, can quantify variation in the time-series data and offers the most appropriate means for analysing variability in movement to allow us to analyse individual differences between human performers.
+To then found out that RQA ENTR values with different recurrence thresholds were appropriate to quantify the different changes and variations of the characteristics of time-series data.
+Therefore, the use of Shannon Entropy could be applied to analyse human participants who might vary in age, state of health, anthropometric features and capability to perform movement.
+
+%% As a guideline references should be limited to 60 (this is not strictly enforced).
 \bibliography{../references/references}
@@
 \section*{Acknowledgements}
-
-This work was funded by the University of Birmingham and the 
-Mexican National Council of Science and Technology as part of 
-Miguel Xochicale's PhD degree. I would like to thank to Dr. Dolores 
-Columba Perez Flores and Prof. Martin J Russell for their helpful 
-comments to polish the use of the language of mathematics and to 
-Constantino Antonio Garcia Martinez et al. for developing the R 
-package nonlinearTseries that help to accelerate the 
-analysis of the nonlinear time series in this work.
+This work was funded by the University of Birmingham and the Mexican National Council of Science and Technology as part of Miguel Xochicale's PhD degree.
+I would like to thank to Dr.
+Dolores Columba Perez Flores and Prof.
+Martin J Russell for their helpful comments to polish the use of the language of mathematics and to Constantino Antonio Garcia Martinez et al.
+for developing the R package nonlinearTseries that help to accelerate the analysis of the nonlinear time series in this work.
 
 \section*{Author contributions statement}
-Contributions for this work of 
-Miguel Xochicale (MX) and Chris Baber (CB) are as follows:
+Contributions for this work of Miguel Xochicale (MX) and Chris Baber (CB) are as follows:
 \begin{description}
@@
 
-%
 %\section*{Additional information}
-%
 %To include, in this order: \textbf{Accession codes} 
 %(where applicable); \textbf{Competing interests} (mandatory statement).
-%
 %The corresponding author is responsible for submitting a 
@@
 %in the submitted article file.
-%
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=\linewidth]{stream}
-%\caption{Legend (350 words max). Example legend text.}
-%\label{fig:stream}
-%\end{figure}
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{|l|l|l|}
-%\hline
-%Condition & n & p \\
-%\hline
-%A & 5 & 0.1 \\
-%\hline
-%B & 10 & 0.01 \\
-%\hline
-%\end{tabular}
-%\caption{\label{tab:example}Legend (350 words max). Example legend text.}
-%\end{table}
-%
-%Figures and tables can be referenced in LaTeX using the ref command, 
-%e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.
 

### 2021-03-21 docs/submission/srep/coverletter/letter-tex/covlet.tex
#6 polishes cover letter

@@
 %\vspace{20pt} \hrule height 1pt % If you would like a horizontal line separating the name from the address, uncomment the line to the left of this text
-4th Floor \\
-Lambeth Wing  St. Thomas' Hospital \\
-Westminster Bridge Road \\
-London SE1 7EH \\
-\faEnvelopeO  miguel.xochicale@kcl.ac.uk \\
+%School of Electronic, Electrical and Systems Engineering\\
+Edgbaston\\
+Birmingham\\
+B15 2TT\\
+UK\\
+\faEnvelopeO  perez.xochicale@gmail.com \\
 \faMobile  (+44) 0744 281 7616 \\
@@
 Our work is appropriate for Scientific Reports due to its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility).   
-Similarly, we choose Scientific Reports because of the potential reachability to readers in areas of natural and clinical sciences along with potential application of our work and the spark for collaborations.
+Similarly, we choose Scientific Reports as venue where potential interest of readers in areas of natural and clinical sciences along with potential application of our work and the spark for collaborations.
 
@@
 %* The names of any referees you would like excluded from reviewing
-Finally, two potential reviewers for our work can be: Bedartha Goswami, bedartha.goswami@uni-tuebingen.de, at University of Tübingen investigating nonlinear time series analysis; recurrence plots, and Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin investigating dynamics time delay networks of dynamical systems; pattern formation; coupled systems.
+Finally, two suitable reviewers for our work would be: Bedartha Goswami, bedartha.goswami@uni-tuebingen.de, at University of Tübingen, for his investigation in nonlinear time series analysis and recurrence plots; and Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin, for his research in dynamics time delay networks of dynamical systems, pattern formation, and coupled systems.
 
@@
 
-\encl{manuscript and supplementary document} 
+\encl{Manuscript and supplementary document} 
 % List your enclosed documents here, comment this out to get rid of the "encl:"

### 2021-03-21 docs/supplementary-information/report/supplementary-information.tex
#6 polishes and prepare for submission suplementary-information

--- 2020-12-27 docs/supplementary-information/report/report.tex
@@
 \begin{abstract}
-Section 1 describes datasets and section 2 shows surface plots of 3D RQA ENTR values for 3 participants (p01, p02 and p03).
+Report for supplementary material where section 1 describes datasets and section 2 shows surface plots of 3D RQA ENTR values for 3 participants (p01, p02 and p03).
 \end{abstract}
@@
 \begin{verbatim}
-~/srep2020/data/dataset$ tree --si
+~/srep2021/data/dataset$ tree --si
 [ 46M]  xdata_v00.dt
@@
 ## Code
-$HOME/srep2020/code/rscripts/G_3Drqa
+$HOME/srep2021/code/rscripts/G_3Drqa
 `> source(  paste( getwd(), '/C_3Drqa_plots_epsilons.R', sep=''), echo=TRUE )`
@@
 ## Data
-$HOME/srep2020/data/rqa$ tree -s
+$HOME/srep2021/data/rqa$ tree -s
 
 ## Figures
-$HOME/srep2020/docs/figures/rqa/src/3drqa_epsilons$ tree -s
+$HOME/srep2021/docs/figures/rqa/src/3drqa_epsilons$ tree -s
 \end{verbatim}
@@
 		%and other sensor attached to the robot (RS01).
-	Code and data to reproduce the figure is available in \cite{srep2020}.
+	Code and data to reproduce the figure is available in \cite{srep2021}.
         }

### 2021-03-21 docs/supplementary-information/report/reduce-pdf-size.sh

@@ -1,0 +1,10 @@
+gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dPDFSETTINGS=/ebook -dNOPAUSE -dQUIET -dBATCH -sOutputFile=supplementary-information-r.pdf supplementary-information.pdf
+#    -dPDFSETTINGS=/screen lower quality, smaller size. (72 dpi)
+#    -dPDFSETTINGS=/ebook for better quality, but slightly larger pdfs. (150 dpi)
+#    -dPDFSETTINGS=/prepress output similar to Acrobat Distiller "Prepress Optimized" setting (300 dpi)
+#    -dPDFSETTINGS=/printer selects output similar to the Acrobat Distiller "Print Optimized" setting (300 dpi)
+#    -dPDFSETTINGS=/default selects output intended to be useful across a wide variety of uses, possibly at the expense of a larger output file
+mv supplementary-information-r.pdf supplementary-information.pdf
+
+## Reference
+# https://askubuntu.com/questions/113544/how-can-i-reduce-the-file-size-of-a-scanned-pdf-file

### 2021-03-21 docs/manuscript/reduce-pdf-size.sh
#6 polishes docs/manuscript/README.md and adds docs/manuscript/reduce-pdf-size.sh

@@ -1,0 +1,10 @@
+gs -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dPDFSETTINGS=/ebook -dNOPAUSE -dQUIET -dBATCH -sOutputFile=main-r.pdf main.pdf
+#    -dPDFSETTINGS=/screen lower quality, smaller size. (72 dpi)
+#    -dPDFSETTINGS=/ebook for better quality, but slightly larger pdfs. (150 dpi)
+#    -dPDFSETTINGS=/prepress output similar to Acrobat Distiller "Prepress Optimized" setting (300 dpi)
+#    -dPDFSETTINGS=/printer selects output similar to the Acrobat Distiller "Print Optimized" setting (300 dpi)
+#    -dPDFSETTINGS=/default selects output intended to be useful across a wide variety of uses, possibly at the expense of a larger output file
+mv main-r.pdf main.pdf
+
+## Reference
+# https://askubuntu.com/questions/113544/how-can-i-reduce-the-file-size-of-a-scanned-pdf-file

### 2021-03-21 code/anthropometrics/rscript/computing_statistics.R
clean pdfs branch

@@ -1,66 +1,0 @@
-###############################################################################	
-#
-#  Computing age statistics of twenty participants.
-#
-#
-#
-#
-# Written by Miguel P Xochicale [http://mxochicale.github.io]
-#
-# If you see any errors or have any questions
-# please create an issue at https://github.com/mxochicale/phd-thesis-code-data/issues
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Statistics for 20 participants 
-	# (2) Filtering particpants (NOT USED)
-
-
-library('data.table') #
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-setwd('../')
-main_path <- getwd()
-data_path <- paste(main_path,'/data', sep='')
-
-
-
-################################################################################
-# (1) Statistics for 20 participants 
-
-datafilename <- 'data_20p.csv'
-#datafilename <- 'data_06p.csv'
-
-full_anthropometric_path_and_datafilename <- paste( data_path, '/', datafilename, sep='')
-pNNanthropometric <- fread(full_anthropometric_path_and_datafilename)
-
-all_pNNanthropometric <- pNNanthropometric[, head(.SD, .N), ] # replace .N for the number of participants
-summary(all_pNNanthropometric)
-
-m <- mean(all_pNNanthropometric$Age)
-sd <- sd(all_pNNanthropometric$Age)
-message('mean age: ', m, '   SD age: ' , sd)
-
-
-#################################################################################
-## (2) Filtering particpants (NOT USED)
-#males_pNNanthropometric <- all_pNNanthropometric[Gender == 'Male' &  Handeness == 'Right', !c("MusicTraining") ]
-#summary(males_pNNanthropometric)
-#
-#females_pNNanthropometric <- all_pNNanthropometric[Gender == 'Female' &  Handeness == 'Right', !c("MusicTraining") ]
-#summary(females_pNNanthropometric)
-#
-#mf_pNNanthropometric <- all_pNNanthropometric[Gender == c('Male','Female') &  Handeness == 'Right', !c("MusicTraining") ]
-#summary(mf_pNNanthropometric)
-#
-
-
-
-
-
-
-#############################################
-setwd(r_scripts_path) ## go back to the r-script source path

### 2021-03-21 code/creation_of_curated_timeseries/code_raw2aligned/A_mscripts/A_hri_rawData_TO_TimeAlignedDataForSeparateActivities_p01_to_p22_octave_linux.m

@@ -1,2889 +1,0 @@
-%
-% octave --no-gui
-% A_hri_rawData_TO_TimeAlignedDataForSeparateActivities_p01_to_p22_octave_linux
-%
-% sensors_timealignment_different_length_p16() function is used to correct the
-% time idex for participant 16
-%
-%
-% Written by Miguel P Xochicale [http://mxochicale.github.io]
-%
-% If you see any errors or have any questions
-% please create an issue at https://github.com/mxochicale/phd-thesis-code-data/issues
-%
-
-
-
-format long;
-clear; % Clear all variables and functions from memory
-
-m_scripts_path = pwd;
-
-source("MyFunctions_octave_linux.m");
-% two cd.. to go the main path
-
-cd ..;
-cd ..;
-main_path = pwd;
-cd ..;
-cd ..;
-cd ..;
-cd ..;
-user_path = pwd;
-
-
-data_path='/data_raw_p01-p22'
-datapath=strcat(main_path, data_path);
-
-tmpdatapath='/tmp/phdtmpdata';
-outcomes_path = strcat(user_path, tmpdatapath, '/time_aligned_different_length_data_for_separate_activities_p01_to_p22_octave_linux','/');
-
-outcomes_path_char= char(outcomes_path);
-if ~exist(outcomes_path_char,'dir');
-  {
-    myk =  cstrcat('system(',' ''mkdir', ' -p ',outcomes_path_char,''')');
-    eval( sprintf(myk) );
-  }
-end
-
-
-p01_data_path = strcat(datapath,'/p01/p01-HRI/');
-[p01_s01m, p01_s02m, p01_s03m, p01_s04m, p01_max_length_sensors] = cvs2arrays(p01_data_path);
-
-p02_data_path = strcat(datapath,'/p02/p02-HRI/');
-[p02_s01m, p02_s02m, p02_s03m, p02_s04m, p02_max_length_sensors] = cvs2arrays(p02_data_path);
-
[…]

### 2021-03-21 code/creation_of_curated_timeseries/code_raw2aligned/A_mscripts/MyFunctions_octave_linux.m

@@ -1,696 +1,0 @@
-
-
-% To use this collection of functions, you have to source the m-file, e.g.
-% source("MyFunctions_octave_linux.m");
-% In MATLAB, you can try to group functions in one main file
-% then calling subfunctions  [stackoverflow.com/questions/3569933]
-
-
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%  FUNCTIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-function [s01, s02, s03, s04, max_length_sensors] = cvs2arrays(sensor_data_path)
-  % This function reads the values from the sensor_data_path
-  % and convert them into arrays s01m, s02m, s03m, s04m
-  %%% Define files variables which save an structure of the current files of
-  %%% the path
-  files=dir([sensor_data_path,'*']);
-
-  clocks  = dlmread(  [sensor_data_path,files(2).name]  );
-  s01  = dlmread(  [sensor_data_path,files(3).name]  );
-  s02  = dlmread(  [sensor_data_path,files(4).name]  );
-  s03  = dlmread(  [sensor_data_path,files(5).name]  );
-  s04  = dlmread(  [sensor_data_path,files(6).name]  );
-
-  #Delete first row
-  clocks(1,:)=[];
-  s01(1,:)=[];
-  s02(1,:)=[];
-  s03(1,:)=[];
-  s04(1,:)=[];
-
-  %%%% Update the time variable that sinchronise the data
-  tnew = (s01(:,2) /clocks(1,2)) - clocks(1,3);
-  s01(:,2) = tnew;
-
-  tnew = (s02(:,2) /clocks(2,2)) - clocks(2,3);
-  s02(:,2) = tnew;
-
[…]

### 2021-03-21 code/creation_of_curated_timeseries/code_raw2aligned/A_mscripts/test_hri_readrawdata.m

@@ -1,175 +1,0 @@
-%
-
-% USAGE
-% octave --no-gui
-% readrawdataHRI
-
-%  M-Script to read data files for SIX participant from \data_P01-P22.
-%  Data is separated by using labels00.txt in four activitities (a1,...,a4).
-%
-
-% Written by Miguel P Xochicale [http://mxochicale.github.io]
-%
-% If you see any errors or have any questions
-% please create an issue at https://github.com/mxochicale/phd-thesis-code-data/issues
-%
-
-
-format long;
-clear; % Clear all variables and functions from memory
-
-
-m_scripts_path = pwd;
-source("MyFunctions_octave_linux.m");
-
-cd ..;
-cd ..;
-main_path = pwd;
-
-data_path='/data_raw_p01-p22';
-datapath=strcat(main_path, data_path);
-
-
-p01_data_path = strcat(datapath,'/p01/p01-HRI/');
-[p01_s01m, p01_s02m, p01_s03m, p01_s04m, p01_max_length_sensors] = cvs2arrays(p01_data_path);
-
-%p02_data_path = strcat(datapath,'/p02/p02-HRI/');
-%[p02_s01m, p02_s02m, p02_s03m, p02_s04m, p02_max_length_sensors] = cvs2arrays(p02_data_path);
-%
-%p03_data_path = strcat(datapath,'/p03/p03-HRI/');
-%[p03_s01m, p03_s02m, p03_s03m, p03_s04m, p03_max_length_sensors] = cvs2arrays(p03_data_path);
-%
-%p04_data_path = strcat(datapath,'/p04/p04-HRI/');
-%[p04_s01m, p04_s02m, p04_s03m, p04_s04m, p04_max_length_sensors] = cvs2arrays(p04_data_path);
-%
-%p05_data_path = strcat(datapath,'/p05/p05-HRI/');
-%[p05_s01m, p05_s02m, p05_s03m, p05_s04m, p05_max_length_sensors] = cvs2arrays(p05_data_path);
-%
[…]

### 2021-03-21 code/creation_of_curated_timeseries/code_raw2aligned/B_rscripts/BA_creating_TidiedInterpolatedDATA_p01p22.R

@@ -1,384 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:           BA_creating_TidiedInterpolatedDATA_p01p22.R
-# FileDescription:
-#
-# Variables:
-#     participantsNN number
-#     aNN: activities
-#     sNN: sensors
-#
-# NOTE: Before running the current script, make sure that you have run:
-#           octave --no-gui
-#           /mscripts/rawData_TO_TimeAlignedDataForSeparateActivities16_p01_to_p22_octave_linux
-#
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Defining participants, activities and sensors to read
-participantsNN <- c(1:22)
-#participantsNN <- c(1:12,14:15,17:22) # Discarding p13 and p16
-
-
-aNN <- c(1,2,3,4) # available activities 1,2,3,4
-sNN <- c(1,2,3,4)# available sensors 1,2,3,4
-
-participant_number <- c("p01_","p02_", "p03_", "p04_", "p05_","p06_","p07_","p08_","p09_","p10_","p11_","p12_","p13_","p14_","p15_","p16_","p17_","p18_","p19_","p20_","p21_","p22_")
-#participant_index <-  c(4,      6,      8,      10,     12,     14,   16,     18,   20,    22,    24,    26,    28,    30,    32,    34,    36,    38,    40,    42,    44,    46)
-#participant_index <-  c(2,      3,      4,      5,     6,     7,    8,     9,   10,    11,    12,    13,    14,    15,    16,    17,    18,    19,    20,    21,    22,    23)
-participant_index <-  c(3,      5,      7,      9,     11,     13,   15,     17,   19,    21,    23,    25,    27,    29,    31,    33,    35,    37,    39,    41,    43,    45)
-
-
-activity_number <- c("a01_","a02_","a03_","a04_")
-sensor_number <- c("s01_","s02_","s03_","s04_")
-
-
-
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(akima) # for interpolation function aspline()
-
-
-options(digits=15)
-
-
-
[…]

### 2021-03-21 code/dependencies/A_install-R.sh

@@ -1,48 +1,0 @@
-#!/bin/bash  
-
-##USAGE:
-#chmod +x file
-#./file
-
-
-
-sudo apt-key adv –keyserver keyserver.ubuntu.com –recv-keys E084DAB9
-##or
-#gpg --keyserver keyserver.ubuntu.com --recv-key E084DAB9
-#gpg -a --export E084DAB9 | sudo apt-key add -
-
-
-
-#SELECT UBUNTU VERSION
-
-# Basic format of next line deb https://<my.favorite.cran.mirror>/bin/linux/ubuntu <enter your ubuntu version>/
-#sudo add-apt-repository 'deb https://cran.ma.imperial.ac.uk/bin/linux/ubuntu precise/'
-
-## Ubuntu 12.04: precise
-#sudo add-apt-repository 'deb https://cloud.r-project.org/bin/linux/ubuntu precise/'
-
-## Ubuntu 14.04: trusty
-#sudo add-apt-repository 'deb https://cloud.r-project.org/bin/linux/ubuntu precise/'
-
-## Ubuntu 16.04: xenial
-sudo add-apt-repository 'deb-src https://cloud.r-project.org/bin/linux/ubuntu xenial'
-
-
-
-sudo apt-get update
-
-
-## DEPENDENCIES
-sudo apt-get update
-sudo apt-get --yes install gfortran
-sudo apt-get --yes install libblas-dev libatlas-base-dev liblapack-dev libatlas-base-dev libncurses5-dev
-sudo apt-get --yes install cdbs
-sudo apt-get --yes install r-base r-base-dev
-
-
-## devtools dependencies
-sudo apt-get --yes install libssl-dev libcurl4-openssl-dev
-sudo apt-get --yes install libxml2-dev # for xml2 calling load_all()
-sudo apt-get --yes install r-cran-rgl # rgl for nonlinearTseries
-
-

### 2021-03-21 code/dependencies/B_install-basic-packages.R

@@ -1,63 +1,0 @@
-
-mirror_repo <- 'https://www.stats.bris.ac.uk/R/'
-
-## OTHER UK MIRRORS
-#  https://www.stats.bris.ac.uk/R/
-#  http://www.stats.bris.ac.uk/R/
-#  https://mirrors.ebi.ac.uk/CRAN/
-#  http://mirrors.ebi.ac.uk/CRAN/
-#  https://cran.ma.imperial.ac.uk/
-#  http://cran.ma.imperial.ac.uk/
-#  http://mirror.mdx.ac.uk/R/
-
-
-
-### devtools
-install.packages("devtools", repos=mirror_repo, dependencies = TRUE)
-
-#### [data.table]([https://github.com/Rdatatable/data.table/wiki/Installation)
-#[MORE](https://github.com/Rdatatable/data.table/wiki/Installation#why-we-recommend-installpackages-above-rather-than-devtoolsinstall_github)
-library(devtools)
-install_github("Rdatatable/data.table", build_vignettes=FALSE)
-
-
-# for timeseries
-#### Interpolation of Irregular and Regularly Spaced Data
-install.packages("akima", repos=mirror_repo, dependencies = TRUE)
-
-####  for savitzly-golay filter
-install.packages("signal", repos=mirror_repo, dependencies = TRUE)
-
-### ggplot2
-install.packages("ggplot2", repos=mirror_repo, dependencies = TRUE)
-
-### RColorBrewer) #for brewer.pal
-install.packages("RColorBrewer", repos=mirror_repo, dependencies = TRUE)
-
-
-
-
-##for rfunctions
-### plot3D
-.packages <- c('car', 'scatterplot3d', 'plot3D')
-lapply(.packages, install.packages, dependencies = TRUE, repos=mirror_repo)
-
-
-####  for deSolve
-install.packages("deSolve", repos=mirror_repo, dependencies = TRUE)
-
-
-### tseriesChaos
-install.packages("tseriesChaos", repos=mirror_repo, dependencies = TRUE)
-
-
-
-### pals for colors functions in scatter3D plots
-install.packages("pals", repos=mirror_repo, dependencies = TRUE)
-
-
-## Latex 
-install.packages("latex2exp", repos=mirror_repo, dependencies = TRUE)
-
-
-

### 2021-03-21 code/dependencies/C_install-nonlinearTseries.R

@@ -1,34 +1,0 @@
-
-mirror_repo <- 'https://www.stats.bris.ac.uk/R/'
-
-## OTHER UK MIRRORS
-#  https://www.stats.bris.ac.uk/R/
-#  http://www.stats.bris.ac.uk/R/
-#  https://mirrors.ebi.ac.uk/CRAN/
-#  http://mirrors.ebi.ac.uk/CRAN/
-#  https://cran.ma.imperial.ac.uk/
-#  http://cran.ma.imperial.ac.uk/
-#  http://mirror.mdx.ac.uk/R/
-
-
-
-#xml2 for roxygen2
-install.packages("xml2", repos=mirror_repo, dependencies = TRUE)
-
-#### roxygen2 for load_all()
-install.packages("roxygen2", repos=mirror_repo, dependencies = TRUE)
-
-
-
-# dependencies for nonlinearTseries
-
-install.packages("tseries", repos=mirror_repo, dependencies = TRUE)
-install.packages("TSA", repos=mirror_repo, dependencies = TRUE)
-install.packages("RcppArmadillo", repos=mirror_repo, dependencies = TRUE)
-
-
-install.packages("nonlinearTseries", repos=mirror_repo, dependencies = TRUE)
-
-
-
-

### 2021-03-21 code/libraries_functions/nonlinearTseries/rfunctions/embedding_parameters/mututalInformation/comparison-3methods.R

@@ -1,299 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-# The following times series are generated using /timeseries/periodicTimeSeries/periodic03.R
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-#################
-### Load libraries
-require(tseriesChaos)
-library(deSolve)
-library(devtools)
-load_all('~/mxochicale/github/nonlinearTseries')
-
-library(data.table)
-library(ggplot2)
-
-
-
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-plots_path <- paste(r_scripts_path,"/plots_path",sep="")
-
-
-###################
-## Main R Script
-
-
-
-
-######################
-### Lorenz Function
-###
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-########################
-### define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-########################
-### define initial state
-# state <- c(X=1, Y=1, Z=1)
-state <- c(X=20, Y=41, Z=20)
-
-########################
-### define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,100, by=0.01)
-
-########################
-### perform the integration and assign it to variable 'out'
[…]

### 2021-03-21 code/libraries_functions/nonlinearTseries/rfunctions/embedding_parameters/withCao1997/cao97sub.f

@@ -1,46 +1,0 @@
-c
-c e1 and e2 values in fortran application differ
-c slighly from the values in the wrapped function in R,
-c since the the later use 16 decimal places
-c
-
-      subroutine cao97sub(x,ndp,maxd,tau,e1,e2)
-
-      integer ndp, d, maxd, tau, maxtau
-      double precision x(1:ndp),e1(1:maxd),e2(1:maxd),v(50),a(50),ya(50),v0,vij
-
-      do d=1,maxd
-        a(d)=0.
-        ya(d)=0.
-        ng=ndp-d*tau
-        do i=1,ng
-          v0=1.e+30
-        do j=1,ng
-          if(j.eq.i) go to 20
-            do k=1,d
-              v(k)=abs(x(i+(k-1)*tau)-x(j+(k-1)*tau))
-            enddo
-          if (d.eq.1) go to 50
-            do k=1,d-1
-               v(k+1)=max(v(k),v(k+1))
-            enddo
- 50         vij=v(d)
-              if (vij.lt.v0.and.vij.ne.0.) then
-                n0=j
-                v0=vij
-             endif
- 20         continue
-          enddo
-          a(d)=a(d)+max(v0,abs(x(i+d*tau)-x(n0+d*tau)))/v0
-          ya(d)=ya(d)+abs(x(i+d*tau)-x(n0+d*tau))
-        enddo
-        a(d)=a(d)/float(ng)
-        ya(d)=ya(d)/float(ng)
-        !write (*,*) a(d), ya(d)
-        if (d.ge.2) then
-         !write(*,*) d-1,' E1=', a(d)/a(d-1), ' E2=', ya(d)/ya(d-1)
-         e1(d-1) = a(d) / a(d-1)
-         e2(d-1) = ya(d) / ya(d-1)
-      endif
-      enddo
-      end

### 2021-03-21 code/libraries_functions/nonlinearTseries/rfunctions/embedding_parameters/withCao1997/examples/exampleCAO97.R

@@ -1,122 +1,0 @@
-r_scripts_path <- getwd()
-setwd("../../../../../")
-github_repo_path <- getwd()
-
-
-library(data.table)
-library(deSolve)
-library(ggplot2)
-
-
-
-################################################################################
-# Lorenz Function
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-
-################################################################################
-# Lorenz time-series
-
-#define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-#define initial state
-state <- c(X=1, Y=1, Z=1)
-# yini <- c(X = 0.01, Y = 0.001, Z = 0.001)
-# state <- c(X=20, Y=41, Z=20)
-
-
-# define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,150, by=0.1)
-# timeserie <- seq(0, 150, by = 0.1)
-
-#perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-maxLength <- dim(out)[1]
-
-
-# Lorenz Time series as data.table object
-lts <- as.data.table(out)
-func <-function(x) {list("Lorenz")}
-lts[,c("type"):=func(), ]
-lts[,n:=seq(.N)]
-setcolorder(lts, c(5,6,1:4))
-
-
-
-################################################################################
-### (4.1) Windowing Data
-###
-windowframe = 1000:maxLength;
-lts <- lts[,.SD[windowframe],by=.(type)];
-
-
-################################################################################
-### Plot time series
-p <- ggplot(lts) +
-   geom_line(aes(x=n,y=X,col='x'),lwd = 1,alpha=0.8)+
-   geom_line(aes(x=n,y=Y,col='y'),lwd = 1,alpha=0.8)+
-   geom_line(aes(x=n,y=Z,col='z'),lwd = 1,alpha=0.8)+
[…]

### 2021-03-21 code/libraries_functions/nonlinearTseries/rfunctions/extra_rqa.R

@@ -1,133 +1,0 @@
-#------------------------------------------------------------
-#   Extra Functions For Recurrent Quantification Analysis using 
-#   `nonlinearTseries` package
-#   
-#
-#
-# Written by Miguel P Xochicale [http://mxochicale.github.io]
-# email:perez.xochicale@gmail.com
-#
-# If you see any errors or have any questions
-# please create an issue at https://github.com/mxochicale/hwum-dataset/issues
-#
-#
-#
-#------------------------------------------------------------
-
-
-## LOADING FUNCTIONS
-#github_path <- getwd()
-#source( paste(github_path,'/hwum-dataset/r-scripts/functions/extra_rqa.R',sep='') )
-
-## LOADING LIBRARIES
-#################################################################################
-#mirror_repo <- 'https://www.stats.bris.ac.uk/R/'
-#if (!require("ggplot2")) {
-#install.packages('ggplot2', repos=mirror_repo, dependencies = TRUE) 
-#library(ggplot2)
-#}
-
-library(ggplot2) 
-
-
-## USAGE EXAMPLE
-
-#
-#rqa.analysis=rqa(time.series = xn, embedding.dim=1, time.lag=1,
-#                radius=5,lmin=2,vmin=2,do.plot=FALSE,distanceToBorder=2)
-#
-#rm <- as.matrix(rqa.analysis$recurrence.matrix)
-#maxsamplerp <- dim(rm)[1]
-#RM <- as.data.table( melt(rm, varnames=c('a','b'),value.name='Recurrence') )
-#rplot <- plotRecurrencePlot(RM,maxsamplerp)
-#
-#width = 1000
-#height = 1000
-#savePlot(filename_extension,width,height,rplot)
-#
-#
-
-
-
-## output of Ymt using RSS()
-#  output<-  list(  [1] ,      [2]  ,            [3]  ,    ...
-#  output   list(  PC , singular_values  ,  rotateddata  , ...
-
-
-
-################################################################################
-plotRecurrencePlot <-function (RM, maxsamplerp)
-{
-
-
-axis_start<- as.numeric(RM[1,1])
-axis_end<- as.numeric(RM[.N,1])
-
-ggplot(RM)+
-	geom_raster( aes(x=a,y=b,fill=Recurrence) ) +  
-	scale_fill_manual( values = c("#ffffff", "#000000") ) + # black and white colours
-	labs(x="Time (secs)", y="Time (secs)") +
-	#labs(x="Sample", y="Sample") +
[…]

### 2021-03-21 code/libraries_functions/rfunctions/embedding_parameters/mututalInformation/lorenzts.R

@@ -1,150 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-#################
-### Load libraries
-library(deSolve)
-library(data.table)
-library(ggplot2)
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-plots_path <- paste(r_scripts_path,"/plots_path",sep="")
-
-
-###################
-## Main R Script
-
-
-
-######################
-### Lorenz Function
-###
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-########################
-### define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-
-
-########################
-### define initial state
-state <- c(X=1, Y=1, Z=1)
-# state <- c(X=20, Y=41, Z=20)
-# X=0.01; Y=0.01; Z=0.01
-
-
-########################
-### define integrations times
-# dt=0.01
-# times <- seq(0,100, by=0.001)
-times <- seq(0,100, by=0.01)
-
-########################
-### perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-
[…]

### 2021-03-21 code/libraries_functions/rfunctions/embedding_parameters/mututalInformation/minfo5.c

@@ -1,290 +1,0 @@
-/*******************************************************************
-
-To compile, use
-cc -o minfo minfo5.c -lm
-
-To run, use
-./minfo datafile -b 100 -t 100 > r.mi
-
-# Reference
-# http://www.physics.emory.edu/faculty/weeks//software/minfo.html
-
-
-minfo.c -- Eric R. Weeks -- started 2/28/97
-
-does the mutual information algorithm discussed by Fraser & Swinney
-(Phys Rev A 33 (1986) p1134-1140)
-
-v01:  2-28-97: taken from shell.c (7/1/96)
-			quicksort routine taken from sane.c (original version)
-v02:  2-28-97: revised sorting for s[] (different than q[])
-			sped up math
-v03:  2-28-97: add in tau loop
-	 3-01-97: fix for variable number of input; add -b option
-v04:  3-01-97: take out chi2 tests for substructure
-v05:  3-01-97: realize that with chi2 tests taken out, number()
-			function doesn't need to be called very often.  remove
-			a[] and b[][] arrays!  Much faster now.
-
-This program is public domain, although please leave my name and
-email address attached.
-
-email: weeks@physics.emory.edu
-web: http://www.physics.emory.edu/~weeks/
-
-explanation of how to use this program:
-    http://www.physics.emory.edu/~weeks/software/minfo.html
-
- *******************************************************************/
-#include <stdio.h>
-#include <math.h>
-#include <stdlib.h>
-#define PI 3.14159265358979323846264338328
-#define EE 2.71828182845904523536
-#define MAXNUM 100000
-#define KMAX 25
-
-int s[MAXNUM],q[MAXNUM];
-int numpts,n0,mmax;
-float pop[MAXNUM];
-int pindex[MAXNUM];
-int pow2[KMAX];
-int bin;
-int globalivalue[4];
-float logtwo;
-
-int number(int karray2[],int m2);
-float findisq();
-float ffunct(int kmarray[],int m);
-void saneqsort(int p,int r);
-int qpartition_neurons(int p,int r);
-
-main(argc,argv)
[…]

### 2021-03-21 code/libraries_functions/rfunctions/embedding_parameters/withCao1997/examples/exampleCAO97.R

@@ -1,125 +1,0 @@
-
-#### Setting up paths
-r_scripts_path <- getwd()
-setwd("../../../../../../")
-github_path <- getwd()
-
-
-homepath <- Sys.getenv("HOME")
-path_cao97_functions_R <- '/code/libraries_functions/rfunctions/embedding_parameters/withCao1997/cao97_functions.R'
-
-# Libraries
-library(data.table)
-library(deSolve)
-library(ggplot2)
-
-################################################################################
-# Lorenz Function
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-
-################################################################################
-# Lorenz time-series
-
-#define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-#define initial state
-state <- c(X=1, Y=1, Z=1)
-# yini <- c(X = 0.01, Y = 0.001, Z = 0.001)
-# state <- c(X=20, Y=41, Z=20)
-
-
-# define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,150, by=0.1)
-# timeserie <- seq(0, 150, by = 0.1)
-
-#perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-maxLength <- dim(out)[1]
-
-
-# Lorenz Time series as data.table object
-lts <- as.data.table(out)
-func <-function(x) {list("Lorenz")}
-lts[,c("type"):=func(), ]
-lts[,n:=seq(.N)]
-setcolorder(lts, c(5,6,1:4))
-
-
-
-################################################################################
-### (4.1) Windowing Data
-###
-windowframe = 1000:maxLength;
-lts <- lts[,.SD[windowframe],by=.(type)];
-
-
-################################################################################
-### Plot time series
-p <- ggplot(lts) +
[…]

### 2021-03-21 code/libraries_functions/rfunctions/extra_rqa.R

@@ -1,133 +1,0 @@
-#------------------------------------------------------------
-#   Extra Functions For Recurrent Quantification Analysis using 
-#   `nonlinearTseries` package
-#   
-#
-#
-# Written by Miguel Xochicale
-# 
-#------------------------------------------------------------
-
-
-## LOADING FUNCTIONS
-#github_path <- getwd()
-#libfun_path <- '/code/libraries_functions'
-#rfunctions_extra_rqa_R <- '/rfunctions/extra_rqa.R'
-
-#### Calling `functions_extra_nonlinearTseries` 
-#source( paste(github_repo_path, libfun_path, rfunctions_extra_rqa_R , sep='') )
-
-
-## LOADING LIBRARIES
-#################################################################################
-#mirror_repo <- 'https://www.stats.bris.ac.uk/R/'
-#if (!require("ggplot2")) {
-#install.packages('ggplot2', repos=mirror_repo, dependencies = TRUE) 
-#library(ggplot2)
-#}
-
-library(ggplot2) 
-
-
-## USAGE EXAMPLE
-
-#
-#rqa.analysis=rqa(time.series = xn, embedding.dim=1, time.lag=1,
-#                radius=5,lmin=2,vmin=2,do.plot=FALSE,distanceToBorder=2)
-#
-#rm <- as.matrix(rqa.analysis$recurrence.matrix)
-#maxsamplerp <- dim(rm)[1]
-#RM <- as.data.table( melt(rm, varnames=c('a','b'),value.name='Recurrence') )
-#rplot <- plotRecurrencePlot(RM,maxsamplerp)
-#
-#width = 1000
-#height = 1000
-#savePlot(filename_extension,width,height,rplot)
-#
-#
-
-
-
-## output of Ymt using RSS()
-#  output<-  list(  [1] ,      [2]  ,            [3]  ,    ...
-#  output   list(  PC , singular_values  ,  rotateddata  , ...
-
-
-
-################################################################################
-plotRecurrencePlot <-function (RM, maxsamplerp)
-{
-
-
-axis_start<- as.numeric(RM[1,1])
-axis_end<- as.numeric(RM[.N,1])
-
-ggplot(RM)+
-	geom_raster( aes(x=a,y=b,fill=Recurrence) ) +  
-	scale_fill_manual( values = c("#ffffff", "#000000") ) + # black and white colours
-	labs(x="Time (secs)", y="Time (secs)") +
-	#labs(x="Sample", y="Sample") +
[…]

### 2021-03-21 code/libraries_functions/rfunctions/ollin_cencah.R

@@ -1,304 +1,0 @@
-#------------------------------------------------------------
-#   Functions
-#   Ollin Cencah <-> Movimiento eterno <-> Eternal Movement
-#
-#
-#
-# Written by Miguel Xochicale [http://mxochicale.github.io]
-#
-#
-#------------------------------------------------------------
-
-
-## LOADING FUNCTIONS
-#github_path <- getwd()
-#source( paste(github_path,'/hwum-dataset/r-scripts/functions/ollin_cencah.R',sep='') )
-
-## LOADING LIBRARIES
-library(ggplot2) ## percentage of variance bar plot
-require(tseriesChaos) # for embed()
-library(plot3D) # for plotRSS2D_rotateddata <-function (PCAMatrix,maxlimit)
-
-
-
-################################################################################
-################################################################################
-################################################################################
-#-------------------- zero mean unit variance function  --------------------------
-
-#Sphering the data (whitening)
-zeromean_unitvariance <- function ( x )
-{
-  # convert data as a matrix object
-  x <- as.matrix(x)
-
-  #column zero mean
-  mx <- mean( x )
-
-  zmx <-  x - matrix(rep( mx,dim(x)[1] ),ncol=dim(x)[2])
-
-  sd_zmx <- sd(zmx)
-  zmuv <- zmx/sd_zmx
-  return( zmuv )
-}
-
-
-
-
-
-
-################################################################################
-################################################################################
-################################################################################
-#--------------------    Uniform Time Delay Embedding  --------------------------
-#--------------------    Takens Theorem  --------------------------
-#
-
-UTDE <- function(timeserie,dim,tau,print_Xdims=FALSE)
-{
-  timedelayembedded <- embedd(timeserie,  m=dim, d=tau)
-
-  if (print_Xdims == TRUE ){
-    message("  --------------")
-    message("  Embedded Parameters:   " ,"m=",dim," tau=",tau,sep="" )
-    message("  Embedded Matrix dimension:  ",dim(timedelayembedded)[1], 'x' ,dim(timedelayembedded)[2] )
[…]

### 2021-03-21 code/rscripts/A_dataset/A_dataset.R

@@ -1,211 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading raw tidied interpolated data
-	# (3) Creating paths
-	# (4) Selecting Variables in data.table
-		# (4.1)	Selecting Participants
-	# (5) Adding vectors
-		# (5.1) Deleting some Magnetomer and quaternion data
-		# (5.2) zero mean and unit variance
-		# (5.3) Savitzky-Golay filter
-	# (6) Selecting Axis after postprocessing
-	# (7) Creating preprocessed data path
-	# (8) Writing data.table object to a file
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-libfun_path <- '/code/libraries_functions'
-rfunctions_ollin_cencah_R <- '/rfunctions/ollin_cencah.R'
-
-##VERSION 
-version <- '00'
-feature_path <- '/data/dataset'
-
-### Outcomes Data Path
-outcomes_data_path <- paste(github_repo_path, feature_path, sep="")
-### Raw Data Path
-rawdata_path <- paste(github_repo_path,'/data/timeseries',sep='')
-setwd(file.path(rawdata_path))
-#
-#
-#################################################################################
-## (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-#source( paste(github_repo_path,'/code/rfunctions/ollin_cencah.R',sep='') )
-source( paste(github_repo_path, libfun_path, rfunctions_ollin_cencah_R, sep='') )
-
-
-
-
-################################################################################
-# (2) Reading raw tidied  interpolated data
[…]

### 2021-03-21 code/rscripts/C_cao/B_cao.R

@@ -1,597 +1,0 @@
-###############################################################################	
-#
-# 
-# 
-#
-#
-#
-#
-###############################################################################	
-	# TABLE OF CONTENT:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading data
-	
-	# (4) Filtering data.table variables
-		# (4.1) Windowing data
-		# (4.2) Filtering variables
-			# (4.2.1) Activities Selection
-			# (4.2.2) Sensor Selection
-			# (4.2.3) Participant Selection
-			# (4.2.4) Axis Selection
-
-	# (5) Average Mutual Information 
-		# (5.1) Parameter selections
-		# (5.2) MAIN LOOP for CAOS 
-
-	# (6) Creating Outcome Data Path
-	# (7)  Writing Data
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-home_path <- Sys.getenv("HOME") 
-
-setwd("../../../")
-github_repo_path <- getwd()
-
-##VERSION 
-version <- '00'
-feature_path <- '/data/utde'
-dataset_path <- '/data/dataset'
-
-## Outcomes Data Path
-outcomes_data_path <- paste(github_repo_path, feature_path, sep="")
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-
-
-################################################################################
-# (2) Reading data
-
-file_ext <- paste('xdata_v', version, '.dt',sep='')
-data <- fread( file_ext, header=TRUE)
-
-
-################################################################################
-# (4) Filtering out variables in xdata data.table
-
-
-data <- data[,.(
-	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY,
[…]

### 2021-03-21 code/rscripts/C_cao/CB_cao_mm_V.R

@@ -1,378 +1,0 @@
-###############################################################################	
-#
-# 
-# 
-#
-#
-#
-#
-###############################################################################	
-	# TABLE OF CONTENT:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading data
-
-	# (6) Plot E values
-		# (6.1) main if for:
-			# (6.1.0) Creating paths for plots
-			# (6.1.1) E1 and E2 values
-			# (6.1.2) Minimum Embedding Dimensions Plots
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-setwd("../")
-github_path <- getwd()
-
-##VERSION 
-version <- '00'
-dataset_path <- '/data/utde'
-
-## Outcomes Data Path
-outcomes_data_path <- paste(github_repo_path, dataset_path, sep="")
-
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-
-################################################################################
-# (2) Reading data
-
-
-#acts <- 'V'
-#windowl<-'w2'
-#wdata <- fread('MED-w2.dt', header=TRUE)
-
-
-#acts <- 'V'
-#windowl<-'w5'
-#wdata <- fread('MED-w5.dt', header=TRUE)
-
-acts <- 'V'
-windowl<-'w10'
-wdata <- fread('MED-w10.dt', header=TRUE)
-
-#acts <- 'V'
-#windowl<-'w15'
-#wdata <- fread('MED-w15.dt', header=TRUE)
-
-
-
-
-
-################################################################################
-################################################################################
-################################################################################
-################################################################################
-	## (4.2.1) Activities Selection
-
-	ACPSAmdim <- NULL
-
-
-	activities <- c('VN','VF')
-
-	#########################################################
-	for (activity_k in 1:length(activities) ) {
-
-	activityk <- activities[activity_k]
-	message(activityk)
-	awdata <- wdata
-
-	if (activityk == 'HN' ) {
[…]

### 2021-03-21 code/rscripts/C_cao/EA_plotcao_avMED_H.R

@@ -1,315 +1,0 @@
-###############################################################################	
-#
-# 
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# TABLE OF CONTENT:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading data
-
-	# (6) Plot E values
-		# (6.1) main if for:
-			# (6.1.0) Creating paths for plots
-			# (6.1.1) E1 and E2 values
-			# (6.1.2) Minimum Embedding Dimensions Plots
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-
-##VERSION 
-version <- '00'
-plots_path <- '/docs/figures/caoami/src'
-dataset_path <- '/data/utde'
-
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path, plots_path, sep="")
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(ggplot2)
-library(data.table) # for manipulating data
-library(RColorBrewer) #for brewer.pal
-
-################################################################################
-# (2) Reading data
-
-
-
-
-#
-#windowl<-'w2'
-#aMED <- fread('aMED-H-w2.dt', header=TRUE)
-
-
-#windowl<-'w5'
-#aMED <- fread('aMED-H-w5.dt', header=TRUE)
-
-
-windowl<-'w10'
-aMED <- fread('aMED-H-w10.dt', header=TRUE)
-
-#windowl<-'w15'
-#aMED <- fread('aMED-H-w15.dt', header=TRUE)
-
-
-
-
-
-
-
-
-#################################################################################
-#### (6) Plot E values
-
-
-# (6.1.1)  Creating paths for plots
-if (file.exists(outcomes_plot_path)){
[…]

### 2021-03-21 code/rscripts/C_cao/EB_plotcao_avMED_V.R

@@ -1,328 +1,0 @@
-###############################################################################	
-#
-# 
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# TABLE OF CONTENT:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading data
-
-	# (6) Plot E values
-		# (6.1) main if for:
-			# (6.1.0) Creating paths for plots
-			# (6.1.1) E1 and E2 values
-			# (6.1.2) Minimum Embedding Dimensions Plots
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-
-##VERSION 
-version <- '00'
-plots_path <- '/docs/figures/caoami/src'
-dataset_path <- '/data/utde'
-
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path, plots_path, sep="")
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(ggplot2)
-library(data.table) # for manipulating data
-library(RColorBrewer) #for brewer.pal
-
-################################################################################
-# (2) Reading data
-
-
-
-
-
-#windowl<-'w2'
-#aMED <- fread('aMED-V-w2.dt', header=TRUE)
-#
-#
-#windowl<-'w5'
-#aMED <- fread('aMED-V-w5.dt', header=TRUE)
-
-
-windowl<-'w10'
-aMED <- fread('aMED-V-w10.dt', header=TRUE)
-
-
-#windowl<-'w15'
-#aMED <- fread('aMED-V-w15.dt', header=TRUE)
-
-
-
-
-
-
-
-
-#################################################################################
-#### (6) Plot E values
-
-
-# (6.1.1)  Creating paths for plots
-if (file.exists(outcomes_plot_path)){
[…]

### 2021-03-21 code/rscripts/D_ami/CA_plotamiH.R

@@ -1,318 +1,0 @@
-###############################################################################	
-#
-# 
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# TABLE OF CONTENT:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading data
-	
-	# (4) Filtering data.table variables
-		# (4.1) Windowing data
-		# (4.2) Filtering variables
-			# (4.2.1) Sensor Selection
-			# (4.2.2) Activities Selection
-			# (4.2.3) Participant Selection
-			# (4.2.4) Axis Selection
-
-	# (5) Average Mutual Information 
-		# (5.1) Plot Avarage Mutual Information 
-		# (5.2) Creating and Changing to Preprosseced Data Path
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/docs/figures/caoami/src'
-dataset_path <- '/data/utde'
-
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
-
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-
-library(ggplot2)
-library(RColorBrewer) #for brewer.pal
-
-
-###############################################################################
-## (2) Reading data
-
-
-noparticipants <- 20
-
-
-#AMI <- fread('AMI-w2.dt', header=TRUE)
-#MTD <- fread('MTD-w2.dt', header=TRUE)
-#windowl<-'w2'
-#
-
-#
-#AMI <- fread('AMI-w5.dt', header=TRUE)
-#MTD <- fread('MTD-w5.dt', header=TRUE)
-#windowl<-'w5'
-#
-#
-AMI <- fread('AMI-w10.dt', header=TRUE)
[…]

### 2021-03-21 code/rscripts/E_rss/BC_mean_amiH.R

@@ -1,154 +1,0 @@
-###############################################################################	
-#
-# 
-# 
-#
-#
-#
-#
-###############################################################################	
-	# TABLE OF CONTENT:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading data
-	
-	# (4) Filtering data.table variables
-		# (4.1) Windowing data
-		# (4.2) Filtering variables
-			# (4.2.1) Sensor Selection
-			# (4.2.2) Activities Selection
-			# (4.2.3) Participant Selection
-			# (4.2.4) Axis Selection
-
-	# (5) Average Mutual Information 
-		# (5.1) Plot Avarage Mutual Information 
-		# (5.2) Creating and Changing to Preprosseced Data Path
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-
-##VERSION 
-version <- '00'
-dataset_path <- '/data/utde'
-
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-
-library(ggplot2)
-library(RColorBrewer) #for brewer.pal
-
-
-################################################################################
-# (2) Reading data
-
-
-noparticipants <- 20
-
-
-#AMI <- fread('AMI-w2.dt', header=TRUE)
-#MTD <- fread('MTD-w2.dt', header=TRUE)
-#windowl<-'w2'
-#
-
-#
-#AMI <- fread('AMI-w5.dt', header=TRUE)
-#MTD <- fread('MTD-w5.dt', header=TRUE)
-#windowl<-'w5'
-#
-
-AMI <- fread('AMI-w10.dt', header=TRUE)
-MTD <- fread('MTD-w10.dt', header=TRUE)
-windowl<-'w10'
-
-
-#AMI <- fread('AMI-w15.dt', header=TRUE)
[…]

### 2021-03-21 code/rscripts/F_rps/Ba_timeseries_aH.R

@@ -1,323 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-	# (3) Creating paths
-	# (4) Selecting Variables in data.table
-		# (4.1)	Selecting Participants
-	# (5) Adding vectors
-		# (5.1) Deleting some Magnetomer and quaternion data
-		# (5.2) zero mean and unit variance
-		# (5.3) Savitzky-Golay filter
-	# (6) Selecting Axis after postprocessing
-	# (7) Creating preprocessed data path
-	# (8) Writing data.table object to a file
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-libfun_path <- '/code/libraries_functions'
-rfunctions_extra_rqa_R <- '/rfunctions/extra_rqa.R'
-
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/docs/figures/rps/src'
-dataset_path <- '/data/dataset'
-
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path, feature_path, sep="")
-
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-source( paste(github_repo_path, libfun_path, rfunctions_extra_rqa_R , sep='') )
-
-################################################################################
-# (2) Reading data
-file_ext <- paste('xdata_v', version, '.dt',sep='')
-data <- fread( file_ext, header=TRUE)
-
-
-
-# axis for horizontal movments
-data <- data[,.(
-	sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
-	), by=. (Participant,Activity,Sensor,Sample,Time)]
-
-
-setkey(data, Participant)
-pNN <- c('p01')
-#pNN <- 		c(	'p01', 'p02', 'p03', 'p04', 'p05', 'p06', 'p07', 'p08', 'p09', 'p10',
[…]

### 2021-03-21 code/rscripts/F_rps/Eb_rqa_plots_aV.R

@@ -1,381 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-	# (3) Creating paths
-	# (4) Selecting Variables in data.table
-		# (4.1)	Selecting Participants
-	# (5) Adding vectors
-		# (5.1) Deleting some Magnetomer and quaternion data
-		# (5.2) zero mean and unit variance
-		# (5.3) Savitzky-Golay filter
-	# (6) Selecting Axis after postprocessing
-	# (7) Creating preprocessed data path
-	# (8) Writing data.table object to a file
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-
-homepath <- Sys.getenv("HOME")
-main_home_path <- '/quetzalcoalt'
-r_scripts_path <- getwd()
-
-setwd("../../../")
-github_repo_path <- getwd()
-libfun_path <- '/code/libraries_functions'
-rfunctions_extra_rqa_R <- '/rfunctions/extra_rqa.R'
-
-
-
-
-##VERSION 
-version <- '00'
-feature_plot_path <- '/docs/figures/rqa/src/V'
-#feature_path <- '/docs/figures/rps/src'
-dataset_path <- '/data/rqa'
-
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path, feature_plot_path, sep="")
-
-## RQA Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-#rqa_path <- paste(data_path, '/rqa', sep="")
-setwd(file.path(data_path))
-
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(ggplot2)
-library(RColorBrewer)
-
-library(devtools)
-load_all(paste(github_repo_path, libfun_path, '/nonlinearTseries', sep=""))
-source( paste(github_repo_path, libfun_path, rfunctions_extra_rqa_R , sep='') )
-
-
-
-
-
-
-
[…]

### 2021-03-21 code/rscripts/G_3Drqa/A_3Drqa_datatable.R

@@ -1,547 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-libfun_path <- '/code/libraries_functions'
-rfunctions_extra_rqa_R <- '/rfunctions/extra_rqa.R'
-
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/data/rqa'
-dataset_path <- '/data/dataset'
-
-## Outcomes Data Path
-outcomes_data_path <- paste(github_repo_path, feature_path,  sep="")
-## Data Path
-data_path <- paste(github_repo_path, dataset_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(devtools)
-load_all(paste(github_repo_path, libfun_path, '/nonlinearTseries', sep=""))
-source( paste(github_repo_path, libfun_path, rfunctions_extra_rqa_R , sep='') )
-
-library(ggplot2)
-library(RColorBrewer)
-
-
-################################################################################
-# (2) Reading data
-file_ext <- paste('xdata_v', version, '.dt',sep='')
-data <- fread( file_ext, header=TRUE)
-
-
-
-data <- data[,.(
-	#sg0zmuvGyroZ
-	#sg0zmuvGyroZ, sg1zmuvGyroZ
-	#sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
-	sg0zmuvGyroY, sg1zmuvGyroY, sg2zmuvGyroY, sg0zmuvGyroZ, sg1zmuvGyroZ, sg2zmuvGyroZ
-	), by=. (Participant,Activity,Sensor,Sample,Time)]
-
-
-################################################################################
-################################################################################
-################################################################################
[…]

### 2021-03-21 code/rscripts/G_3Drqa/D_3Drqa_plots_sensors_activities.R

@@ -1,386 +1,0 @@
-###############################################################################	
-#
-# 
-#
-#
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
- 	# (0) Definifing paths 
-	# (1) Loading libraries and functions
-	# (2) Reading 
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-
-################################################################################
-# (0) Defining paths for main_path, r_scripts_path, ..., etc.
-
-r_scripts_path <- getwd()
-setwd("../../../")
-github_repo_path <- getwd()
-
-
-
-
-
-##VERSION 
-version <- '00'
-feature_path <- '/docs/figures/rqa/src/3drqa_sensors_activities'
-data_path <- '/data/rqa'
-
-## Outcomes Plot Path
-outcomes_plot_path <- paste(github_repo_path,  feature_path, sep="")
-## Data Path
-data_path <- paste(github_repo_path, data_path, sep="")
-setwd(file.path(data_path))
-
-
-
-################################################################################
-# (1) Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(plot3D)
-
-
-
-################################################################################
-# (2) Reading data
-#
-#file_ext <- paste('RQAs_p01w100.dt',sep='')
-#W100 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w250.dt',sep='')
-#W250 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w500.dt',sep='')
-#W500 <- fread( file_ext, header=TRUE)
-#
-#file_ext <- paste('RQAs_p01w750.dt',sep='')
-#W750 <- fread( file_ext, header=TRUE)
-#
-#W <- rbind(W100, W250, W500, W750)
-#
-
-#file_ext <- paste('RQAs_p02w500.dt',sep='')
-#W <- fread( file_ext, header=TRUE)
-#
-
-file_ext <- paste('RQAs_p03w500.dt',sep='')
-W <- fread( file_ext, header=TRUE)
-
-
-
-################################################################################
-################################################################################
-################################################################################
-################################################################################
-## (x.x.x) RQA Metric Selection
-rqas <- c('ENTR')
-#rqas <- c('REC','DET', 'RATIO', 'ENTR')
-
-
-#########################################################
-for (rqas_k in 1:length(rqas) ) {
-
[…]

### 2021-03-21 docs/figures/figures-originals/appendices/appendixA/r-scripts/figure_2E_e1ab_v00.R

@@ -1,218 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:           DataTidiedInterpolatedSmoothed_p01_to_p22.R
-# FileDescription:
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(ggplot2)
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Defining paths for main_path, r_scripts_path, ..., etc.
-setwd("../")
-main_path <- getwd()
-setwd("../")
-main_outcomesdata_path <- getwd()
-
-r_scripts_path <- paste(main_path,"/r-scripts/",sep="")
-data_path <-  paste(main_path,"/data/second_experiment/muse/",sep="")
-outcomes_path <- paste(main_path,"/outcomes",sep="")
-
-
-############################################################################
-###   Reading RAW Data from Participants
-###
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Setting DataSets paths
-setwd(data_path)
-
-
-details = file.info(list.files(pattern=""))
-details = details[with(details, order(as.POSIXct(mtime))), ]
-files = rownames(details)
-
-e1a <- fread(files[ 1 ] )
-e1a <- cbind(experiment="e1a", e1a)
-e1a[,sample:=seq(.N)]
-setcolorder(e1a, c(20,1:19))
-
-
-e1b <- fread(files[ 2 ] )
-e1b <- cbind(experiment="e1b", e1b)
-e1b[,sample:=seq(.N)]
[…]

### 2021-03-21 docs/figures/figures-originals/appendices/appendixA/r-scripts/time_delay_embedding/ami.R

@@ -1,236 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:          ami.R
-# FileDescription:
-#
-#
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(ggplot2)
-library(plot3D) #scatter3D
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Defining paths for main_path, r_scripts_path, ..., etc.
-
-r_scripts_path <- getwd()
-setwd("../")
-setwd("../")
-main_path <- getwd()
-setwd("../")
-setwd("../")
-setwd("../")
-setwd("../")
-github_path <- getwd()
-
-
-# r_scripts_path <- paste(main_path,"/r-scripts/",sep="")
-data_path <-  paste(main_path,"/data/second_experiment/muse/",sep="")
-outcomes_path <- paste(main_path,"/outcomes",sep="")
-
-
-
-
-############################################################################
-###   Reading RAW Data from Participants
-###
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Setting DataSets paths
-setwd(data_path)
-
-
-details = file.info(list.files(pattern=""))
-details = details[with(details, order(as.POSIXct(mtime))), ]
-files = rownames(details)
-
-image_width =  1654 #3508 #595 #877
-image_height = 1240#2480 #842 #620
-resolution = 300
-
-image_width_p3d =  2000 #3508 #595 #877
-image_height_p3d = 700#2480 #842 #620
-
-
-
-
-
-# y <- c(3:5) # AccX AccY AccZ
[…]

### 2021-03-21 docs/figures/figures-originals/appendices/appendixA/r-scripts/time_delay_embedding/cao.R

@@ -1,262 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:          cao.R
-# FileDescription:
-#
-#
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-library(signal)# for butterworth filter and sgolay
-library(ggplot2)
-library(plot3D) #scatter3D
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Defining paths for main_path, r_scripts_path, ..., etc.
-
-r_scripts_path <- getwd()
-setwd("../")
-setwd("../")
-main_path <- getwd()
-setwd("../")
-setwd("../")
-setwd("../")
-setwd("../")
-github_path <- getwd()
-
-
-# r_scripts_path <- paste(main_path,"/r-scripts/",sep="")
-data_path <-  paste(main_path,"/data/second_experiment/muse/",sep="")
-outcomes_path <- paste(main_path,"/outcomes",sep="")
-
-
-
-
-############################################################################
-###   Reading RAW Data from Participants
-###
-
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Setting DataSets paths
-setwd(data_path)
-
-
-details = file.info(list.files(pattern=""))
-details = details[with(details, order(as.POSIXct(mtime))), ]
-files = rownames(details)
-
-image_width =  1654 #3508 #595 #877
-image_height = 1240#2480 #842 #620
-resolution = 300
-
-image_width_p3d =  2000 #3508 #595 #877
-image_height_p3d = 700#2480 #842 #620
-
-
-
-
-
[…]

### 2021-03-21 docs/figures/figures-originals/averageMutualInformation/example/ami_chaoticTimeSeries_plot.R

@@ -1,264 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-#################
-### Load libraries
-require(tseriesChaos)
-library(deSolve)
-library(devtools)
-load_all('~/mxochicale/github/nonlinearTseries')
-
-library(data.table)
-library(ggplot2)
-library(latex2exp)
-
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-plots_path <- paste(r_scripts_path,"/ami_plots",sep="")
-
-
-###################
-## Main R Script
-
-
-
-######################
-### Lorenz Function
-###
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-########################
-### define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-########################
-### define initial state
-state <- c(X=1, Y=1, Z=1)
-# state <- c(X=20, Y=41, Z=20)
-
-########################
-### define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,100, by=0.01)
-
-########################
-### perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-
-########################
[…]

### 2021-03-21 docs/figures/figures-originals/averageMutualInformation/example/ami_nonlinearTseries.R

@@ -1,147 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-### Load libraries
-library(deSolve)
-library(devtools)
-load_all('~/mxochicale/github/nonlinearTseries')
-
-
-library(data.table)
-library(ggplot2)
-
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-plots_path <- paste(r_scripts_path,"/plots_path",sep="")
-
-
-
-
-### Orginal source code for Plotting
-# old.par = par(mfrow = c(1, 2))
-# # tau-delay estimation based on the autocorrelation function
-# tau.acf = timeLag(inputtimseries, technique = "acf",
-#                  lag.max = 100, do.plot = T)
-# # tau-delay estimation based on the mutual information function
-# tau.ami = timeLag(inputtimseries, technique = "ami",
-#                  lag.max = 100, do.plot = T)
-
-
-
-####TESTING
-
-# lagmax <- 50
-
-# tau-delay estimation based on the autocorrelation function
-tau.acf <- timeLag(time.series = inputtimseries, technique = "acf", selection.method = "first.minimum", lag.max = 100, do.plot = F)
-# tau.acf = 56 with lag.max = NULL
-# tau.acf = 56 with lag.max = 100
-
-# tau-delay estimation based on the mutual information function
-tau.ami <- timeLag(time.series = inputtimseries, technique = "ami", selection.method = "first.minimum", lag.max = 50, do.plot = F, n.partitions = 1000, units = "Bits")
-# with lag.max = NULL
-# tau.ami = 17
-# with lag.max = 100
[…]

### 2021-03-21 docs/figures/figures-originals/averageMutualInformation/example/comparison00.R

@@ -1,40 +1,0 @@
-require(tseriesChaos)
-library(lattice) ##xyplot
-
-# rawdata <- NULL;
-# rawdata<-read.csv("pitch_wL1001_imu1_29012015-201358_h.dat", header=TRUE, sep=',');
-# X <-rawdata[,]
-# rawdata <- NULL;
-
-temp <- NULL;
-temp<-read.csv("datafile", header=FALSE, sep=' ');
-lts <-temp[,]
-temp <- NULL;
-
-maxtau <- 100
-b  <- 100
-
-tsmitemp <- mutual(lts, partitions= b , lag.max = maxtau, plot=FALSE)
-tsmi <- unname(tsmitemp[])
-
-m1 <- xyplot( tsmi[1:maxtau+1] ~ (1:maxtau+1),
-        pch=16, col.line = c('royalblue'), type = c("l","g"), lwd=5,
-        xlab=list(label="samples", cex=4),
-        ylab=list(label="", cex=4),
-        scales=list(font=1, cex=2.5)
-)
-
-
-temp <- NULL;
-temp <-read.csv("r.mi", header=FALSE, sep=' ');
-X <-temp[,]
-temp <- NULL;
-
-m2 <- xyplot( X[,2] +  tsmi[1:maxtau+1] ~ X[,1],
-  pch=16, col.line = c('royalblue'), type = c("l","g"), lwd=5,
-  xlab=list(label="samples", cex=4),
-  ylab=list(label="", cex=4),
-  scales=list(font=1, cex=2.5)
-)
-
-plot(m2)

### 2021-03-21 docs/figures/figures-originals/falseNearestNeighbours/cao97/forRandomTimeSeries/cao.R

@@ -1,199 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-# The following times series are generated using /timeseries/periodicTimeSeries/periodic03.R
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-### Load libraries
-library(data.table)
-library(ggplot2)
-require(tseriesChaos)
-library(plot3D)
-library(RColorBrewer)
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-plots_path <- paste(r_scripts_path,"/plots_path",sep="")
-
-setwd("../../../../../")
-main_repository_path <- getwd()
-setwd("../")
-github_path <- getwd()
-
-
-
-
-
-
-##################################################################
-# Time Domain setup
-
-windowLenght <- 2000
-
-
-
-# acq.freq <- 50  # 50 Hertz
-# dt <- 1/acq.freq # 0.02 seconds or 20 miliseconds
-#
-# T <- 40   ## Maximum Time in seconds
-# df <- 1/T
-# n <- T/dt
-# t <- seq(0,T,by=dt)
-
-
-
-################################################################################
-# Create Waveforms in a data.table object
-
-allSDnoise <- 0.00
-
-rts <- data.table(
-  n=0:windowLenght,
-  xn= rnorm(windowLenght+1)
-  # xn= runif(windowLenght+1)
-  )
-
-
-
-################################################################################
-## Plotting time series
-pts <- ggplot(rts) +
-  geom_line(  aes(x=n,y=xn), lwd = 2, alpha=0.7)+      # line
-  geom_point( aes(x=n,y=xn), shape=21, size=2.5, stroke =0.5 ) + # stem type end
[…]

### 2021-03-21 docs/figures/figures-originals/falseNearestNeighbours/timeseries/chaoticTimeSeries/chaotic03.R

@@ -1,129 +1,0 @@
-##################################################
-#
-#
-#
-# Miguel P Xochicale <http://mxochicale.github.io/>
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmigham, U.K. (2014-2018)
-#
-
-library(data.table)
-library(ggplot2)
-
-library(fNonlinear)
-library(plot3D)
-
-# times <- seq(0, 100, by = 0.001)
-tn <- seq(0, 50, by = 0.03)
-NN <- length(tn)
-
-
-## 3D
-## Roessler Attractor
-r <- roesslerSim(
-  times = tn,
-  parms = c(a = 0.2, b = 0.2, c = 8.0),
-  start = c(-1.894, -9.920, 0.0250),
-  doplot = FALSE)
-r <- as.data.table(r)
-
-fsNNtmp <-function(x) {list("Rossler")}
-r[,c("type"):=fsNNtmp(), ]
-r[,n:=seq(.N)]
-setcolorder(r, c(5,6,1:4))
-
-
-
-
-## 3D
-## Lorenz Attractor
-l <- lorentzSim(
-times = tn,
-parms = c(sigma = 16, r = 45.92, b = 4),
-start = c(-14, -13, 47),
-doplot = FALSE)
-l <- as.data.table(l)
-
-fsNNtmp <-function(x) {list("Lorenz")}
-l[,c("type"):=fsNNtmp(), ]
-l[,n:=seq(.N)]
-setcolorder(l, c(5,6,1:4))
-
-
-
-cts3d <- rbind(r,l)
-
-
-# # ################################################################################
-# # ## Plotting State Spaces
-# par(mfrow = c(2, 2) )
-#
-# sslinewidth <- 2
-# scatter3D(
-#   l$x, l$y, l$z, colvar = l$t, bty = "u", type = "l", lwd=sslinewidth,
-#  	axis.scales = TRUE,
-#    main = "XYZ", xlab = 'x', ylab ='y', zlab = 'z'
-#    )
-#    #  colkey = list(length = 0.3, width = 0.8, cex.clab = 0.75)
-#
-# scatter2D(
-#     l$x, l$y, colvar = l$t, bty = "n", pch = ".", type='l', lwd=sslinewidth,
-#     cex = 5, colkey = TRUE, col= jet.col(n=50, alpha=0.9),
-#     main = "x~y", xlab = 'x', ylab ='y'
-#     )
-#
-# scatter2D(
[…]

### 2021-03-21 docs/figures/figures-originals/falseNearestNeighbours/timeseries/periodicTimeSeries/periodic02.R

@@ -1,117 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-#
-#
-#
-#
-#@article{England2007,
-#         author = {England, Scott A. and Granata, Kevin P.},
-#         title = {{The influence of gait speed on local dynamic stability of walking}}
-#}
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-### Load libraries
-library(data.table)
-library(ggplot2)
-
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-
-freq1 <- 1;
-period1 <- 1 / freq1;
-w1 <- 2 * pi * freq1;
-
-freq2 <- 1/2;
-period2 <- 1 / freq2;
-w2 <- 2 * pi * freq2;
-
-phase <- 0
-
-sampling_frequency <- 50
-timeinterval     <- 31         # measuring time interval (seconds)
-nT       <- seq(0,timeinterval-1/sampling_frequency, 1/sampling_frequency ) # vector of sampling time-points (s)
-
-noise <- rnorm( length(nT), 0, .1)
-
-
-
-sin_ts <- data.table(
-    n=0:(length(nT)-1),
-    nT=nT,
-    xn=sin(w1*nT + phase)
-    )
-sin_ts <- cbind(type="sine",sin_ts) # Addind type column
-
-
-sinnoise_ts <- data.table(
-    n=0:(length(nT)-1),
-    nT=nT,
-    xn=sin(w1*nT + phase) + noise
-    )
-sinnoise_ts <- cbind(type="sinenoise",sinnoise_ts) # Addind type column
-
-
-sincos_ts <- data.table(
-    n=0:(length(nT)-1),
-    nT=nT,
-    xn=sin(w1*nT + phase) +  cos(w2*nT + phase)
-    )
-sincos_ts <- cbind(type="sinecosine",sincos_ts) # Addind type column
-
-
-
[…]

### 2021-03-21 docs/figures/figures-originals/falseNearestNeighbours/timeseries/periodicTimeSeries/periodic04.R

@@ -1,174 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-### Load libraries
-library(data.table)
-library(ggplot2)
-
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-
-
-################################################################################
-# Nyquist Theorem
-# said that the frequency of the signal
-# should be the double as the sampling frequency to recover the signal
-# and avoid problems with aliasing. For instance, for the
-# maximum hearning frequency of  20Hz, the
-# sampling rate for those signals should be 40 Hz.
-# FSampling > 2*F max
-
-# Fmax=5
-# 50 > 2*5
-# 50 > 10
-
-# For our case, the Maximum Frequencies shohould be
-# Fmapling/2 which is a maximum freqneices of 25
-#
-
-
-
-################################################################################
-## SineWaves Function
-# dc.component <- 0
-# amplitudes <- c(1,1,1)  # strength of signal components
-# frequencies <- c(4,2,0) # frequency of signal components (Hz)
-# delays <- c(0,0,0)      # delay of signal components (radians)
-sinewaves.func <- function(t,dc.component,amplitudes,frequencies,delays, mean.additivenoise, sd.additivenoise ) {
-    dc.component +
-    sum( amplitudes * sin(2*pi*frequencies*t + delays) +
[…]

### 2021-03-21 docs/figures/figures-originals/rqa/figure12-rp/D.R

@@ -1,129 +1,0 @@
-###############################################################################	
-#
-# Recurrence plot for disrupted (Brownian motion)
-# 
-#
-#
-#
-#
-# * If you have a question or comment that might be relevant 
-#	to other people, please open an issue.
-# * If you have any specific questions about the information 
-#	of this repository, you can me by vising: https://mxochicale.github.io/
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
-	# (0) Loading libraries and functions
- 	# (1) Definifing paths
-	# (2) Lorez Time Series
-	# (3) Plotting State Spaces
-	# (4) Computing Recurrence Quantification Parameters
-	# (5) Plotting Recurrence Plot
-
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-setwd('../')
-functions_path <- getwd()
-
-setwd("../../../../../")
-github_path <- getwd()
-figures_path <- r_scripts_path
-
-
-################################################################################
-# (0) Loading Functions and Libraries
-library(deSolve) # call deSolve first to avoid (Error in .C("unlock_solver") : )
-library(data.table) # for manipulating data
-library(plot3D)
-library("RColorBrewer")
-library(ggplot2)
-library(reshape2)#for melt
-library(nonlinearTseries)
-
-
-################################################################################
-# (1) Disrupted brownian motion
-## `emailing-authors/exrps.m` LINE 55-59:
-##x=0; 
-##for i=1:2000; 
-##   x(i+1)=x(i)+(2*rand-1);
-##end
-##X=crp(x,1,1,.2,'sil');
-
-N <- 400
-M <- 2000
-x<- 0
-for (i in (1:M) ){
-	#x[i+1] <- x[i] + ( (2*rnorm(1))-1  )
-	#x[i+1] <- x[i] +  ( 2*rnorm(1)-1   )
-	x[i+1] <- x[i] +  ( 2*rnorm(1)  )
-}
-dis <- x[1:N]
-
-#N = 2000
-#dis = rnorm(N, 0, 1)
-#dis = cumsum(dis)
-### REF https://www.stat.berkeley.edu/~aldous/Research/Ugrad/ZY1.pdf
-
-
-#################################################################################
[…]

### 2021-03-21 docs/figures/figures-originals/rqa/figure12-timeseries/C.R

@@ -1,148 +1,0 @@
-###############################################################################	
-#
-# Recurrence plot for drift (logistic map xi+1 = 4xi(1-xi) ) 
-# corrupted with a linearly increasing term ( 0.01*(1:150) )
-# 
-#
-#
-#
-#
-# * If you have a question or comment that might be relevant 
-#	to other people, please open an issue.
-# * If you have any specific questions about the information 
-#	of this repository, you can me by vising: https://mxochicale.github.io/
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
-	# (0) Loading libraries and functions
- 	# (1) Definifing paths
-	# (2) Lorez Time Series
-	# (3) Plotting State Spaces
-	# (4) Computing Recurrence Quantification Parameters
-	# (5) Plotting Recurrence Plot
-
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-version <- '01'
-plots_path <- paste(r_scripts_path,'/v', version,sep="")
-
-
-
-
-
-################################################################################
-# (0) Loading Functions and Libraries
-
-library(data.table) # for manipulating data
-library(plot3D)
-library("RColorBrewer")
-library(ggplot2)
-library(reshape2)#for melt
-library(nonlinearTseries)
-
-
-
-################################################################################
-# (1) Logistic Map
-## `emailing-authors/exrps.m` LINE 46-49:
-##x=.617; 
-##for i=2:500; 
-##   x(i)=4*x(i-1)*(1-x(i-1)); 
-##end
-##X=crp(x(201:200+150)+.01*[1:150],1,1,.2,'sil');
-
-#N=150
-#epsilon 0.2 sigma
-
-
-logistic.map <- function(r, x, N, M){
-  ## r: bifurcation parameter
-  ## x: initial value
-  ## N: Number of iteration
-  ## M: Number of iteration points to be returned
-  z <- 1:N
-  z[1] <- x
-  for(i in c(1:(N-1))){
-    z[i+1] <- r *z[i]  * (1 - z[i])
-  }
-  ## Return the last M iterations 
-  z[c((N-M):N)]
-}
-#REF: https://www.r-bloggers.com/logistic-map-feigenbaum-diagram/
-
[…]

### 2021-03-21 docs/figures/figures-originals/rqa/figure12-timeseries/D.R

@@ -1,135 +1,0 @@
-###############################################################################	
-#
-# Recurrence plot for disrupted (Brownian motion)
-# 
-#
-#
-#
-#
-# * If you have a question or comment that might be relevant 
-#	to other people, please open an issue.
-# * If you have any specific questions about the information 
-#	of this repository, you can me by vising: https://mxochicale.github.io/
-#
-#
-#
-###############################################################################	
-	# OUTLINE:
-	# (0) Loading libraries and functions
- 	# (1) Definifing paths
-	# (2) Lorez Time Series
-	# (3) Plotting State Spaces
-	# (4) Computing Recurrence Quantification Parameters
-	# (5) Plotting Recurrence Plot
-
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-version <- '01'
-plots_path <- paste(r_scripts_path,'/v', version,sep="")
-
-
-
-
-################################################################################
-# (0) Loading Functions and Libraries
-
-
-library(data.table) # for manipulating data
-library(plot3D)
-library("RColorBrewer")
-library(ggplot2)
-library(reshape2)#for melt
-library(nonlinearTseries)
-
-
-
-
-
-################################################################################
-# (1) Disrupted brownian motion
-## `emailing-authors/exrps.m` LINE 55-59:
-##x=0; 
-##for i=1:2000; 
-##   x(i+1)=x(i)+(2*rand-1);
-##end
-##X=crp(x,1,1,.2,'sil');
-
-N <- 400
-M <- 2000
-x<- 0
-for (i in (1:M) ){
-	#x[i+1] <- x[i] + ( (2*rnorm(1))-1  )
-	#x[i+1] <- x[i] +  ( 2*rnorm(1)-1   )
-	x[i+1] <- x[i] +  ( 2*rnorm(1)  )
-}
-dis <- x[1:N]
-
-#N = 2000
-#dis = rnorm(N, 0, 1)
-#dis = cumsum(dis)
-### REF https://www.stat.berkeley.edu/~aldous/Research/Ugrad/ZY1.pdf
-
-
-
-tsn <- as.data.table(dis)
-
-tsn [,n:= 0:(.N-1),]
-setcolorder(tsn, c(2,1))
-
-
-#################################################################################
-## (3) Create 
-##
[…]

### 2021-03-21 docs/figures/figures-originals/rqa/functions/functions_extra_nonlinearTseries.R

@@ -1,96 +1,0 @@
-#------------------------------------------------------------
-#   Extra Functions For nonlinearTseries
-#   v0.0.0 - 28april2018.13h29m 
-#
-#
-#                        miguel [http://mxochicale.github.io]
-#
-#
-#------------------------------------------------------------
-
-
-############
-## To use `functions_extra_nonlinearTseries.R` add the following line to your main script:
-## github_path <- getwd()
-## source( paste(github_path,'/tavand/functions/functions_extra_nonlinearTseries.R',sep='') )
-
-
-#################################################################################
-#mirror_repo <- 'https://www.stats.bris.ac.uk/R/'
-#if (!require("ggplot2")) {
-#install.packages('ggplot2', repos=mirror_repo, dependencies = TRUE) 
-#library(ggplot2)
-#}
-
-
-
-################################################################################
-plotRecurrencePlot <-function (RM,maxsamplerp)
-{
-
-ggplot(RM)+
-	geom_raster( aes(x=a,y=b,fill=Recurrence) ) +  
-	scale_fill_manual( values = c("#ffffff", "#000000") ) + # black and white colours
-	labs(x="Sample", y="Sample") +
-	scale_x_continuous(limits = c(0,maxsamplerp), expand = c(0, 0)) + # remove space between axis and area plot	
-	scale_y_continuous(limits = c(0,maxsamplerp), expand = c(0, 0)) + # remove space between axis and area plot
-	theme_bw(base_size=20) + 
-	theme(legend.position="none",
-	plot.margin=grid::unit(c(10,10,10,10), "mm"), # increase the border between axis and the image
-	axis.line = element_line(colour = "black")
-	)
-}
-
-# Usage of plotRecurrencePlot <-function (RM,maxsamplerp)
-# Example: plotRecurrencePlot(RM,maxsamplerp)
-
-
-
-################################################################################
-plotOnlyRecurrencePlot <-function (RM,maxsamplerp)
-{
-
-ggplot(RM)+
-	geom_raster( aes(x=a,y=b,fill=Recurrence) ) +  
-	scale_fill_manual( values = c("#ffffff", "#000000") ) + # black and white colours
-	theme_void()+	
-	theme(legend.position="none")+
-	scale_x_continuous(limits = c(0,maxsamplerp), expand = c(0, 0)) + # remove space between axis and area plot	
-	scale_y_continuous(limits = c(0,maxsamplerp), expand = c(0, 0)) # remove space between axis and area plot
-
-}
-
[…]

### 2021-03-21 docs/figures/figures-originals/rss/allowed_embedding_parameters/01embedding_parameters.R

@@ -1,139 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# source(paste(getwd(),"/01embedding_parameters.R", sep=""), echo=TRUE)
-# Execution Time
-#
-# Create a representation for a 1 to 190 window lenght for the embedding parameters
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-
-#################
-# Start the clock!
-start.time <- Sys.time()
-
-
-
-### Load libraries
-
-require(ggplot2)
-
-
-################################################################################
-# Loading Functions and Libraries and Setting up digits
-library(data.table) # for manipulating data
-options(digits=15)
-source('~/mxochicale/github/r-code_repository/functions/ollin_cencah.R')
-
-
-
-
-################################################################################
-# (1) Defining paths for main_path, r_scripts_path, ..., etc.
-r_scripts_path <- getwd()
-
-
-
-# Time series parameters
-MinTimeSeriesLenght <- 1
-MaxTimeSeriesLenght <- 190
-
-
-TEM <- NULL #TimeSeries Embedding MAtrix
-
-################################################################################
-################################################################################
-################################################################################
-###[timeseriesK]################################################################
-for (k in MinTimeSeriesLenght:MaxTimeSeriesLenght){
-    timeseriesK <- 1:k
-    L <- length(timeseriesK)
-
-
-
-### Embedding Parameters
-
-MaxDimension <- 50
-MaxDelay <- 50
-dimension_k <- c(1:MaxDimension)
-embedding_k <- c(1:MaxDelay)
-
-
-C <- data.table()
-################################################################################
-######State Space Reconstruction [for(dim_i... and for(tau_j...]################
-        for (dim_i in (1:500)[dimension_k]){
-
-            N <- data.table()
-
-                for (tau_j in (1:500)[embedding_k]){
-                # message('        State Space Reconstruction.  ', 'dimension: ', dim_i, ' delay: ', tau_j)
-
-
[…]

### 2021-03-21 docs/figures/figures-originals/rss/quintana-duque2012/figure1/A_lorenz_system.R

@@ -1,98 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-library(deSolve)
-library(data.table)
-library(plot3D)
-library("RColorBrewer")
-
-
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-#define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-#define initial state
-state <- c(X=1, Y=1, Z=1)
-# state <- c(X=20, Y=41, Z=20)
-
-
-# define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,100, by=0.01)
-
-
-#perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-
-
-LorenzStateSpace <- as.data.table(out)
-
-
-fsNNtmp <-function(x) {list("Lorenz")}
-LorenzStateSpace[,c("type"):=fsNNtmp(), ]
-LorenzStateSpace[,Sample:=seq(.N)]
-setcolorder(LorenzStateSpace, c(5,6,1:4))
-
-
-
-
-
-# ################################################################################
-# # Plotting State Spaces
-#
-
-png(filename="A.png",
-  bg = "transparent",
-  type="cairo",
-  width = 800, height = 800, units = "px",
[…]

### 2021-03-21 docs/figures/figures-originals/rss/quintana-duque2012/figure1/B_lorenz_timeseries.R

@@ -1,108 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-library(data.table)
-library(ggplot2)
-library(deSolve)
-
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-#define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-#define initial state
-state <- c(X=1, Y=1, Z=1)
-# state <- c(X=20, Y=41, Z=20)
-
-
-# define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,100, by=0.01)
-
-
-#perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-
-
-LorenzStateSpace <- as.data.table(out)
-
-
-fsNNtmp <-function(x) {list("Lorenz")}
-LorenzStateSpace[,c("type"):=fsNNtmp(), ]
-LorenzStateSpace[,Sample:=seq(.N)]
-setcolorder(LorenzStateSpace, c(5,6,1:4))
-
-
-
-
-################################
-### (4.2) Windowing Data
-windowframe = 5000:5700
-LorenzStateSpace <- LorenzStateSpace[,.SD[windowframe]]
-
-
-
-# ################################################################################
-# # Plotting time series
-
-
-
-
-plotlinewidth <- 7
-image_width <- 800
-image_height <- 500
-image_dpi <- 300
[…]

### 2021-03-21 docs/figures/figures-originals/rss/quintana-duque2012/figure2/A_LorenzTimeSeries.R

@@ -1,132 +1,0 @@
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-#
-#
-# FileName:
-# FileDescription:
-#
-#
-#
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-# Miguel P. Xochicale [http://mxochicale.github.io]
-# Doctoral Researcher in Human-Robot Interaction
-# University of Birmingham, U.K. (2014-2018)
-#+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
-
-
-library(data.table)
-library(ggplot2)
-library(deSolve)
-library(plot3D)
-
-Lorenz <- function(t, state, parameters){
-          with(as.list( c(state,parameters)),
-              {
-              #rate of change
-              dX <- sigma*(Y-X)
-              dY <- rho*X - X*Z - Y
-              dZ <- X*Y - beta*Z
-
-              # return the rate of change
-              list( c(dX, dY, dZ) )
-              }
-          )# end with(as.list...
-}
-
-#define controlling parameters
-# rho     - Scaled Rayleigh number.
-# sigma   - Prandtl number.
-# beta   - Geometry ascpet ratio.
-parameters <- c(rho=28, sigma= 10, beta=8/3)
-
-#define initial state
-state <- c(X=1, Y=1, Z=1)
-# state <- c(X=20, Y=41, Z=20)
-
-
-# define integrations times
-# times <- seq(0,100, by=0.001)
-times <- seq(0,100, by=0.01)
-
-
-#perform the integration and assign it to variable 'out'
-out <- ode(y=state, times= times, func=Lorenz, parms=parameters)
-
-
-LorenzStateSpace <- as.data.table(out)
-
-
-fsNNtmp <-function(x) {list("Lorenz")}
-LorenzStateSpace[,c("type"):=fsNNtmp(), ]
-LorenzStateSpace[,Sample:=seq(.N)]
-setcolorder(LorenzStateSpace, c(5,6,1:4))
-
-
-
-
-# ################################
-# ### (4.2) Windowing Data
-# windowframe = 5000:5700
-# LorenzStateSpace <- LorenzStateSpace[,.SD[windowframe]]
-
-
-
-# # ################################################################################
-# # # Plotting time series
-#
-# plotlinewidth <- 2
-# image_width <- 800
[…]

### 2021-03-21 docs/submission/arxiv/versions/v03/main-arxiv.tex
#8 adds arxiv v05 ready to be submitted

--- 2020-10-28 docs/submission/arxiv/v02/main-arxiv.tex
@@
 \centering
-\includegraphics[width=1.0\textwidth]{rss.pdf}
+\includegraphics[width=1.0\textwidth]{fig-rss.pdf}
     \caption{
@@
 $E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
-which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
+which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:caoami}D).
 $E_1(m)$ values then indicate the minimum embedding dimension of the 
 noisy time series is thirteen, however all of the $E_2(m)$ values are 
-approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
+approximately equal to one (Figure~\ref{fig:caoami}F) for which it can be 
 concluded that noise time series is a stochastic signal.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{cao.pdf}
-    \caption{
-	{\bf Minimum dimension embedding values with Cao's method.} 
-	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
-	with variations of $\tau$ values from one to twenty
-	for (E) chaotic and (F) random time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:e1e2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
-
 It is important to note that for this work not only $E_1(m)$ and $E_2(m)$ are 
@@
 \cite{bradley2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{ami.pdf}
-    \caption{
-	{\bf Minimum delay embedding values with AMI's method.} 
-    	(A, B) AMI values where its first minimum value in the curve
-	is the minimum time delay embedding ($\tau_0$), 
-	for (C) a chaotic and (D) noise time series.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:amis}
-\end{figure}
 %%---------------------------------(FIGURE)-------------------------------------
@@
 for which AMI depends on partitioning \cite{garcia2005e71}.
+
+%%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{fig-caoami.pdf}
+    \caption{
+	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
+	(A) chaotic and (B) random time series.
+	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
+	with variations of $\tau$ values from one to twenty
+	for chaotic and random time series.
+	(G, H) AMI values where its first minimum value in the curve
+	is the minimum time delay embedding ($\tau_0$).
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+        }
+    \label{fig:caoami}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
 
@@
 function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
-(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
+(Fig~\ref{fig:rps}(A,B)) \cite{eckmann1987, marwan2007,marwan2015}.
 RP is also characterised with a line of identity (LOI) which is a  
 diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$.
-%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{rp}
-    \caption{
-	{\bf Recurrence Plots.} 
-	(A) State space of the Lorenz system with controlling parameters 
-	($\rho=28, \sigma=10, \beta=8/3$). 
-	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
-	(black circle) of a given point at $i$ is a recurrent point and is 
-	represented as a black dot in the recurrence plot at location 
-	$(i, j)$ or white otherwise.
-	(B) Recurrence plot using the 
-	three components of the Lorenz system and the RP with no embeddings 
-	and threshold $\epsilon=5$.
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:mrp}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
 
@@
 to laminar states or singularities \cite{marwan2015}.
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{rpsp}
-    \caption{
-	{\bf Patterns in Recurrence Plots.} 
-	Time-series with its respective recurrence plot for:
-	(A) uniformly distributed noise,
-	(B) super-positionet harmonic oscillation 
-	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
-	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
-	with a linearly increase term ($0.01*i$), and
-	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
-	This figure is adapted from \cite{marwan2015}.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-	}
-    \label{fig:rp2}
-\end{figure}
-%%---------------------------------(FIGURE)-------------------------------------
 
@@
 recurrent pattern structures in the RP \cite{zbilut1992}.
+
+%---------------------------------(FIGURE)-------------------------------------
+\begin{figure}[ht]
+\centering
+\includegraphics[width=1.0\textwidth]{fig-rps.pdf}
+    \caption{
+	{\bf Recurrence Plots and its patterns.} 
+	(A) State space of the Lorenz system with controlling parameters 
+	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
+	(black circle) of a given point at $i$ is a recurrent point and is 
+	represented as a black dot in the recurrence plot at location 
+	$(i, j)$ or white otherwise.
+	(B) Recurrence plot using the 
+	three components of the Lorenz system and the RP with no embeddings 
+	and threshold $\epsilon=5$.
+	Time-series with its respective recurrence plot for:
+	(C) uniformly distributed noise,
+	(D) super-positionet harmonic oscillation 
+	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
+	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
+	with a linearly increase term ($0.01*i$), and
+	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
+	This figure is adapted from \cite{marwan2015}.
+	Code and data to reproduce the figure is available in \cite{srep2020}.
+	}
+    \label{fig:rps}
+\end{figure}
+%%---------------------------------(FIGURE)-------------------------------------
 
@@
   \centering
-\includegraphics[width=1.0\textwidth]{hri}
+\includegraphics[width=1.0\textwidth]{fig-hri}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{sts}
+\includegraphics[width=1.0\textwidth]{fig-sts.pdf}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{fig1}
+\includegraphics[width=1.0\textwidth]{fig-fig1.pdf}
     	\caption{ 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{fig2}
+\includegraphics[width=1.0\textwidth]{fig-fig2.pdf}
 	\caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{fig3}
+\includegraphics[width=1.0\textwidth]{fig-fig3.pdf}
 \caption{
-	{\bf RSSs for horizontal arm movements.}
+	{\bf RSSs for horizontal and vertical arm movements.}
 	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
@@
 	embedding parameters 
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
+	$\overline{m}_0=6$, $\overline{\tau}_0=8$
+	for (A) horizontal and (B) vertical arm movements.
 	Code and data to reproduce the figure is available in \cite{srep2020}.	
         }
-    \label{fig:rss_aHw10}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{fig4}
-    \caption{
-	{\bf RSSs for vertical arm movements.}
-	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
-	Reconstructed state spaces were computed with 
-	embedding parameters
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:rss_aVw10}
+    \label{fig:rsss}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{rp_aH}
+\includegraphics[width=1.0\textwidth]{fig-fig4.pdf}
 \caption{
-	{\bf RPs for horizontal arm movements.}	
+	{\bf RPs for horizontal (A) and vertical (B) arm movements.}	
 	Recurrence plots were computed with 
@@
         }
-    \label{fig:rp_aH}
-\end{figure}
-%%---------------------------------(FIGURE)------------------------------------
-%%---------------------------------(FIGURE)-------------------------------------
-\begin{figure}[ht]
-\centering
-\includegraphics[width=1.0\textwidth]{rp_aV}
-\caption{
-	{\bf RPs for vertical arm movements.}	
-	Recurrence plots were computed with 
-	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
-	Code and data to reproduce the figure is available in \cite{srep2020}.
-        }
-    \label{fig:rp_aV}
+    \label{fig:rps}
 \end{figure}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{rqa-bp}
+\includegraphics[width=1.0\textwidth]{fig-fig5.pdf}
     \caption
@@
 \centering
-\includegraphics[width=1.0\textwidth]{rqa-epsilons}
+\includegraphics[width=1.0\textwidth]{fig-rqa-epsilons.pdf}
     \caption
@@
 \centering
-\includegraphics[width=1.0\textwidth]{rqa-sensors-activities}
+\includegraphics[width=1.0\textwidth]{fig-rqa-sensors-activities.pdf}
     \caption{
@@
 \centering
-\includegraphics[width=0.9\textwidth]{rqa-participants}
+\includegraphics[width=0.9\textwidth]{fig-rqa-participants.pdf}
     \caption{
@@
 \centering
-\includegraphics[width=1.0\textwidth]{rqa-windows}
+\includegraphics[width=1.0\textwidth]{fig-rqa-windows}
     \caption{
@@
 %\bibliography{sample}
-%\bibliography{references.bib}
+%\bibliography{../references/references}
+%\bibliography{references}
 \input{main-arxiv.bbl}

### 2021-03-21 docs/submission/arxiv/A_copy-tex-figures.sh

@@ -1,0 +1,33 @@
+## copy tex file
+mkdir -p versions/v05
+cp ../../manuscript/main.tex versions/v05
+cp ../../manuscript/jabbrv* versions/v05
+cp ../../manuscript/naturemag-doi.bst versions/v05
+cp ../../manuscript/wlscirep.cls versions/v05
+
+
+### cp image files to arxiv path
+cp ../../figures/fig01/versions/drawing-v01.png versions/v05
+mv versions/v05/drawing-v01.png versions/v05/fig01.png
+
+cp ../../figures/fig02/versions/drawing-v01.png versions/v05
+mv versions/v05/drawing-v01.png versions/v05/fig02.png
+
+cp ../../figures/caoami/pdf/fig2.pdf versions/v05
+mv versions/v05/fig2.pdf versions/v05/fig03.pdf
+
+cp ../../figures/rss/versions/drawing-v01.png versions/v05
+mv versions/v05/drawing-v01.png versions/v05/fig04.png
+
+cp ../../figures/fig05/versions/drawing-v01.png versions/v05
+mv versions/v05/drawing-v01.png versions/v05/fig05.png
+
+cp ../../figures/fig06/versions/drawing-v00.png versions/v05
+mv versions/v05/drawing-v00.png versions/v05/fig06.png
+
+cp ../../figures/rqa/output/rqa-participants.pdf versions/v05
+mv versions/v05/rqa-participants.pdf versions/v05/fig07.pdf
+
+cp ../../figures/rqa/output/rqa-windows.pdf versions/v05
+mv versions/v05/rqa-windows.pdf versions/v05/fig08.pdf
+

### 2021-03-21 docs/submission/arxiv/B_pdflatex-bibtex.sh

@@ -1,0 +1,2 @@
+pdflatex main.tex
+bibtex main

### 2021-03-21 docs/submission/arxiv/E_zip_files.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+VERSION=$1
+zip -r versions/$VERSION.zip versions/$VERSION

### 2021-03-22 docs/submission/arxiv/A_copy-tex-figures.sh
#8 fixes mixed figure file format

@@ -1,33 +1,0 @@
-## copy tex file
-mkdir -p versions/v05
-cp ../../manuscript/main.tex versions/v05
-cp ../../manuscript/jabbrv* versions/v05
-cp ../../manuscript/naturemag-doi.bst versions/v05
-cp ../../manuscript/wlscirep.cls versions/v05
-
-
-### cp image files to arxiv path
-cp ../../figures/fig01/versions/drawing-v01.png versions/v05
-mv versions/v05/drawing-v01.png versions/v05/fig01.png
-
-cp ../../figures/fig02/versions/drawing-v01.png versions/v05
-mv versions/v05/drawing-v01.png versions/v05/fig02.png
-
-cp ../../figures/caoami/pdf/fig2.pdf versions/v05
-mv versions/v05/fig2.pdf versions/v05/fig03.pdf
-
-cp ../../figures/rss/versions/drawing-v01.png versions/v05
-mv versions/v05/drawing-v01.png versions/v05/fig04.png
-
-cp ../../figures/fig05/versions/drawing-v01.png versions/v05
-mv versions/v05/drawing-v01.png versions/v05/fig05.png
-
-cp ../../figures/fig06/versions/drawing-v00.png versions/v05
-mv versions/v05/drawing-v00.png versions/v05/fig06.png
-
-cp ../../figures/rqa/output/rqa-participants.pdf versions/v05
-mv versions/v05/rqa-participants.pdf versions/v05/fig07.pdf
-
-cp ../../figures/rqa/output/rqa-windows.pdf versions/v05
-mv versions/v05/rqa-windows.pdf versions/v05/fig08.pdf
-

### 2021-03-22 docs/submission/arxiv/E_zip_files.sh

@@ -1,4 +1,5 @@
 #!/bin/bash
 
 VERSION=$1
+rm versions/$VERSION.zip 
 zip -r versions/$VERSION.zip versions/$VERSION

### 2021-03-25 docs/submission/arxiv/C_pdflatex-pdflatex.sh
adds docs/submission/arxiv/versions/v06/

@@ -1,2 +1,4 @@
 pdflatex main.tex
 pdflatex main.tex
+pdflatex supplementary-material.tex
+pdflatex supplementary-material.tex

### 2021-03-25 docs/submission/arxiv/D_clean-tex-project.sh

@@ -1,1 +1,1 @@
-rm -f *.aux *.blg *.log *.out main.pdf comment.cut
+rm -f *.aux *.blg *.log *.out *.toc main.pdf supplementary-material.pdf comment.cut

### 2021-04-05 docs/manuscript/main.tex
#13 addresses reference of figures; and adds workflows; README details and sh for local build

@@
 In the case of topology, the following pattern formations are commonly presented: (i) homogeneous where uniform recurrence points are spread in the RP e.g., uniformly distributed noise (Fig~\ref{fig01}c3), (ii) periodic and quasi-periodic systems where diagonal lines and checkerboard structures represent oscillating systems, e.g., sinusoidal signals (Fig~\ref{fig01}c4), (iii) drift where paling or darkening recurrence points away from the LOI is caused by drifting systems, e.g., logistic map (Fig~\ref{fig01}c5), and (iv) disrupted where recurrence points are presented white areas or bands that indicate abrupt changes in the dynamics, e.g.
-Brownian motion (Fig~\ref{fig01}c7) \cite{eckmann1987, marwan2015}.
+Brownian motion (Fig~\ref{fig01}c6) \cite{eckmann1987, marwan2015}.
 Texture patterns in RPs can be categorised as: (i) single or isolated recurrence points that represent rare occurring states, and do not persist for any time or fluctuate heavily, (ii) dots forming diagonal lines where the length of the small-scale parallel lines in the diagonal are related to the ratio of determinism or predictability in the dynamics of the system, and (iii) dots forming vertical and horizontal lines where the length of the lines represent a time length where a state does not change or change very slowly and these patterns formation represent discontinuities in the signal, and (iv) dots clustering to inscribe rectangular regions which are related to laminar states or singularities \cite{marwan2015}.
@@
 \section*{Experiment} \label{sec:experiment}
-We conducted an experiment in the context of human-humanoid imitation (HHI) activities where participants were asked to imitate simple horizontal and vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
+We conducted an experiment in the context of human-humanoid imitation (HHI) activities where participants were asked to imitate simple horizontal (Figure~\ref{fig02}a1) and vertical arm movements (Figure~\ref{fig02}a3) performed by NAO, a humanoid robot (Figure~\ref{fig02}a2,~\ref{fig02}a4) \cite{gouaillier2009}.
 Such simple movements were repeated ten times for the participant who copied NAO's arm movements in a face-to-face imitation activity.
-Also, wearable inertial measurement unit (IMU) sensors were attached to the right hand of the participant and to the left hand of the robot (Figure~\ref{fig02}a1,a3).
+Also, wearable inertial measurement unit (IMU) sensors were attached to the right hand of the participant and to the left hand of the robot (Figures~\ref{fig02}a1,~\ref{fig02}a3).
 Data were then collected with four NeMEMSi IMU sensors with sampling rate of 50Hz provinding tri-axial data of the accelerometer, gyroscope and magnetometer sensors and quaternions \cite{Comotti2014}.
@@
 \subsection*{Human-humanoid imitation activities}
-For human-humanoid imitation (HHI) activities four neMEMSi sensors were used, two of which were attached to the right hand of the participant and the other two to the left hand of the humanoid robot.
+For human-humanoid imitation (HHI) activities four neMEMSi sensors were used, two of which were attached to the right hand of the participant and the other two to the left hand of the humanoid robot (Figure~\ref{fig02}A).
 Then, each participant was asked to imitate repetitions of simple horizontal and vertical arm movements performed by the humanoid robot in the following conditions:

### 2021-04-05 docs/manuscript/local-tex-build.sh

@@ -1,0 +1,3 @@
+make clean 
+make 
+make view

### 2021-04-05 docs/manuscript/main.tex
#18 adds details on the guidelines for human experiments at UoB

@@
 Hence, the experiments were conducted in November 2016 and participants confirmed reading and understanding the participant information sheet of the experiments and were able to withdraw from the experiment at any time without giving any reason.
-The design of the experiments adhered to the University of Birmingham regulations, data were anonymised and videos were stored only on a personal computer in accordance with the Data Protection Act 1998.
+The design of the experiments adhered to the University of Birmingham guidelines and regulations for experiments with humans that includes: participant information sheet, checklist of the experiment and anonymised data collection in accordance with the Data Protection Act 1998.
+All data in this study were anonymised and videos were stored only on a personal computer.
 %Refer to Appendix D for further information about the ethics, online participation information sheets and experiment check list.

### 2021-06-03 docs/manuscript/main.tex
#13 adds ethics position statement. Ready for new submission. Build CITEX

@@
 All data in this study were anonymised and videos were stored only on a personal computer.
+
+A review of the way in which the project was conducted was undertaken by the Head of Research Governance at the University of Birmingham. 
+An ethics position statement was prepared to confirm that ethical issues for this project were considered sufficiently to allow data from this project to be used for the research (ERN 21-0774).
+This was approved in line with University processes.
 %Refer to Appendix D for further information about the ethics, online participation information sheets and experiment check list.

### 2021-06-04 docs/submission/srep/coverletter/letter-tex/covlet.tex
#15 address requirements for new manuscript submission

@@
 %\vspace{20pt} \hrule height 1pt % If you would like a horizontal line separating the name from the address, uncomment the line to the left of this text
-%School of Electronic, Electrical and Systems Engineering\\
 Edgbaston\\
@@
 %%\vspace{20pt} \hrule height 1pt % If you would like a horizontal line separating the name from the address, uncomment the line to the left of this text
-%123 Broadway \\ City, State 12345 \\ (000) 111-1111 % Your address and phone number
 \end{flushright}
@@
 %* The affiliation and contact information of your corresponding author
-I have the pleasure of sending you the manuscript entitled "Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities" authored by Miguel Xochicale and Chris Baber to be considered for publication as a research article in your prestigious journal.
+I have the pleasure of sending you the manuscript entitled "Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities" authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
 
 %* A brief explanation of why the work is appropriate for Scientific Reports
-Our work is appropriate for Scientific Reports due to its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility).   
-Similarly, we choose Scientific Reports as venue where potential interest of readers in areas of natural and clinical sciences along with potential application of our work and the spark for collaborations.
+Our work is appropriate for Scientific Reports due to its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility) in the fields of physical, biological, health and engineering sciences.
+Similarly, we choose Scientific Reports as venue where potential interest of readers in above mentioned areas along with raising interested and spark collaborations.
 
-Our manuscript is containing original research and has not been submitted or published earlier in any journal and is not being considered for publication elsewhere.   
+Our manuscript contains original research and has not been submitted or published earlier in any journal and is not being considered for publication elsewhere.   
 Authors have seen and approved the manuscript and have contributed significantly to it.
@@
 %* The names of any referees you would like excluded from reviewing
-Finally, two suitable reviewers for our work would be: Bedartha Goswami, bedartha.goswami@uni-tuebingen.de, at University of Tübingen, for his investigation in nonlinear time series analysis and recurrence plots; and Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin, for his research in dynamics time delay networks of dynamical systems, pattern formation, and coupled systems.
-
-%BLURS
-%The manuscript is original (i.e., my co-author and myself wrote it, not copied it).
-%https://www.aje.com/en/arc/writing-cover-letter/
+Finally, we think that two suitable reviewers for our work would be: 
+Bedartha Goswami, bedartha.goswami@uni-tuebingen.de, at University of Tübingen, for his investigation in nonlinear time series analysis and recurrence plots; 
+and Serhiy Yanchuk, yanchuk@math.tu-berlin.de, at Technical University of Berlin, for his research in dynamics time delay networks of dynamical systems, pattern formation, and coupled systems.
 
@@
 
+% List your enclosed documents here, comment this out to get rid of the "encl:"
 \encl{Manuscript and supplementary document}
-% List your enclosed documents here, comment this out to get rid of the "encl:"
-%----------------------------------------------------------------------------------------
 

### 2021-06-27 docs/manuscript/main.tex
rephrase title and abstract

@@
 
-\title{Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities}
+\title{Recurrence Quantification Analysis to quantify Movement Variability in Human-Humanoid Interaction Activities}
 \author[1,*]{Miguel Xochicale}
@@
 Human movement variability arises from the process of mastering redundant (bio)mechanical degrees of freedom to successfully accomplish any given motor task where flexibility and stability of many possible joint combinations helps to adapt to environment conditions.
-While the analysis of movement of variability is becoming increasingly popular as a diagnostic tool or skill performance evaluation, there are remain challenges on applying the most appropriate methods.
-We therefore investigate nonlinear methods such as reconstructed state space (RSSs), uniform time-delay embedding, recurrence plots (RPs) and recurrence quantification analysis (RQAs) with real-world time-series data of wearable inertial sensors.
-That said, twenty healthy participants imitated vertical and horizontal arm movements in normal and faster velocity from an humanoid robot.
-We applied nonlinear methods to the collected data to found visual differences in the patterns of RSSs and RPs and statistical differences with RQAs.
-We conclude that Shannon Entropy with RQA is a robust method that helps to quantify activities, types of sensors, windows lengths and level of smoothness.
-Hence this work might enhance the development of better diagnostic tools for applications in rehabilitation and sport science for skill performance or new forms of human-humanoid interaction for quantification of movement adaptations and motor pathologies.
+While the analysis of movement of variability is becoming increasingly useful as a skill performance evaluation or diagnostic tool, there are remain challenges on applying the most appropriate methods for their quantification.
+We therefore investigate reconstructed state space (RSSs), uniform time-delay embedding, recurrence plots (RPs) and recurrence quantification analysis (RQAs) with real-world time-series data of wearable inertial sensors.
+To which twenty healthy participants imitated vertical and horizontal arm movements in normal and faster velocity from an humanoid robot.  
+We applied RSSs, RPs and RQAs to the collected data to found visual differences in the patterns of RSSs and RPs and statistical differences with RQAs. 
+We conclude that RQA with Shannon Entropy is suitable to quantify activities, types of sensors, windows lengths and level of smoothness.       
+Hence this work might enhance the development of better diagnostic tools for applications in (a) new forms of human-humanoid interaction for quantification of movement adaptations and motor pathologies or (b) rehabilitation and sport science for skill performance.
 \end{abstract}

### 2021-06-27 docs/manuscript/main.tex
provides statement for experimental protocol

@@
 A review of the way in which the project was conducted was undertaken by the Head of Research Governance at the University of Birmingham. 
-An ethics position statement was prepared to confirm that ethical issues for this project were considered sufficiently to allow data from this project to be used for the research (ERN 21-0774).
-This was approved in line with University processes.
+An ethics position statement was prepared to confirm that ethical issues for this project were considered sufficiently to allow data from this project to be used for the research.    
+Hence the ethics position statement and experimental protocol was approved in line with the University of Birmingham processes (ERN 21-0774).
 %Refer to Appendix D for further information about the ethics, online participation information sheets and experiment check list.

### 2021-06-27 docs/submission/srep/coverletter/letter-tex/covlet.tex
updates cover letter matching title

@@
 %* The affiliation and contact information of your corresponding author
-I have the pleasure of sending you the manuscript entitled "Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities" authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
+I have the pleasure of sending you the manuscript entitled 
+"Recurrence Quantification Analysis to quantify Movement Variability in Human-Humanoid Interaction Activities"
+authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
 

### 2021-06-28 docs/manuscript/main.tex
polishes title and abstract and amends ci-workflow

@@
 
-\title{Recurrence Quantification Analysis to quantify Movement Variability in Human-Humanoid Interaction Activities}
+\title{Recurrence Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
 \author[1,*]{Miguel Xochicale}
@@
 \begin{abstract}
-Human movement variability arises from the process of mastering redundant (bio)mechanical degrees of freedom to successfully accomplish any given motor task where flexibility and stability of many possible joint combinations helps to adapt to environment conditions.
-While the analysis of movement of variability is becoming increasingly useful as a skill performance evaluation or diagnostic tool, there are remain challenges on applying the most appropriate methods for their quantification.
+Human movement variability arises from the process of mastering redundant (bio)mechanical degrees of freedom to successfully accomplish any given motor task where stability and flexibility of many possible joint combinations helps to adapt to environment conditions.
+While the analysis of movement of variability is becoming increasingly useful as a diagnostic tool or skill performance evaluation, there are remain challenges on applying the most appropriate methods for their quantification.
 We therefore investigate reconstructed state space (RSSs), uniform time-delay embedding, recurrence plots (RPs) and recurrence quantification analysis (RQAs) with real-world time-series data of wearable inertial sensors.
 To which twenty healthy participants imitated vertical and horizontal arm movements in normal and faster velocity from an humanoid robot.  
-We applied RSSs, RPs and RQAs to the collected data to found visual differences in the patterns of RSSs and RPs and statistical differences with RQAs. 
-We conclude that RQA with Shannon Entropy is suitable to quantify activities, types of sensors, windows lengths and level of smoothness.       
+We applied RSSs, RPs and RQAs to the time-series data to found visual differences in the patterns of RSSs and RPs and statistical differences with RQAs. 
+We conclude that RQA with Shannon Entropy is a suitable method to quantify activities, types of sensors, windows lengths and level of smoothness.         
 Hence this work might enhance the development of better diagnostic tools for applications in (a) new forms of human-humanoid interaction for quantification of movement adaptations and motor pathologies or (b) rehabilitation and sport science for skill performance.

### 2021-06-28 docs/submission/srep/coverletter/letter-tex/covlet.tex
#15 polishes cover letter and hopefully build CITEX

--- 2021-06-04 docs/submission/srep/coverletter/letter-tex/covlet.tex
@@
 %* The affiliation and contact information of your corresponding author
-I have the pleasure of sending you the manuscript entitled "Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities" authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
+I have the pleasure of sending you the manuscript entitled 
+"Recurrence Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities"
+authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
 
@@
 Our work is appropriate for Scientific Reports due to its alignment to the principles of open science (e.g., reproducibility, transparency, reusability and open accessibility) in the fields of physical, biological, health and engineering sciences.
-Similarly, we choose Scientific Reports as venue where potential interest of readers in above mentioned areas along with raising interested and spark collaborations.
+Similarly, we choose Scientific Reports as venue for publication of multidisciplinary areas where potential interest of readers may be interested and spark collaborations of our work.
 

### 2021-06-29 docs/manuscript/main.tex
#15 reverts title and polishes abstract CITEX

@@
 
-\title{Recurrence Quantification Analysis of Movement Variability in Human-Humanoid Interaction Activities}
+\title{Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities}
 \author[1,*]{Miguel Xochicale}
@@
 We conclude that RQA with Shannon Entropy is a suitable method to quantify activities, types of sensors, windows lengths and level of smoothness.         
-Hence this work might enhance the development of better diagnostic tools for applications in (a) new forms of human-humanoid interaction for quantification of movement adaptations and motor pathologies or (b) rehabilitation and sport science for skill performance.
+Hence this work might enhance the development of better diagnostic tools for applications in new forms of human-humanoid interaction for quantification of movement adaptations and motor pathologies or rehabilitation and sport science for skill performance.
 \end{abstract}

### 2021-06-29 docs/submission/srep/coverletter/letter-tex/covlet.tex
#15 updates title in coverletter

--- 2021-06-04 docs/submission/srep/coverletter/letter-tex/covlet.tex
@@
 %* The affiliation and contact information of your corresponding author
-I have the pleasure of sending you the manuscript entitled "Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities" authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
+I have the pleasure of sending you the manuscript entitled 
+"Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities"
+authored by Miguel Xochicale and Chris Baber from the University of Birmingham to be considered for publication as a research article in your prestigious journal.
 

### 2021-06-29 docs/manuscript/main.tex
#15 minor changes to main.tex to build CITEX
#15 polishes acknowledgements CITEX

@@
 \section*{Acknowledgements}
-This work was funded by the University of Birmingham and the Mexican National Council of Science and Technology as part of Miguel Xochicale's PhD degree.
-I would like to thank to Dr.
-Dolores Columba Perez Flores and Prof.
-Martin J Russell for their helpful comments to polish the use of the language of mathematics and to Constantino Antonio Garcia Martinez et al.
-for developing the R package nonlinearTseries that help to accelerate the analysis of the nonlinear time series in this work.
+This work was partially funded by both the University of Birmingham and the Mexican National Council of Science and Technology as part of Miguel Xochicale's PhD degree.      
+I would like to thank to Dr. Dolores Columba Perez Flores and Prof. Martin J Russell for their helpful comments to polish the use of the language of Mathematics.    
+Last but not least, thanks to Constantino Antonio Garcia Martinez et al. for developing the R package nonlinearTseries that help to accelerate the analysis of the nonlinear time series in this work.
 
